### 2014-07-22 MS.tex
typesetting

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{hyperref}

% line numbers
\usepackage{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
%\usepackage{rotating}

% If you wish to include algorithms, please use one of the packages below. Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
%\usepackage{algorithmic}
%\usepackage{algorithmicx}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

\author{}
\date{}
\pagestyle{myheadings}

\begin{document}
% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Diversity dynamics in Nymphalidae butterflies: Effect of
phylogenetic uncertainty on diversification rate shift estimates}
}
% Insert Author names, affiliations and corresponding author email.
\\
Carlos Pe\~na$^{1,\ast}$, 
Marianne Espeland$^{2}$
\\
\bf{1} Carlos Pe\~na Laboratory of Genetics, Department of Biology, University of Turku, Turku
\\
\bf{2} Museum of Comparative Zoology and Department of Organismic and
Evolutionary Biology, Harvard University, Cambridge, USA. Email:
\href{mailto:marianne.espeland@gmail.com}{marianne.espeland@gmail.com}.
Current address: McGuire Center for Lepidoptera and Biodiversity,
Florida Museum of Natural History, University of Florida, Gainesville,
FL, US
\\
$\ast$ E-mail: Corresponding mycalesis@gmail.com
\end{flushleft}

\subsection{Abstract}\label{abstract}

The species rich butterfly family Nymphalidae has been used to study
evolutionary interactions between plants and insects. Theories of
insect-hostplant dynamics predict accelerated diversification due to key
innovations. We investigated whether phylogenetic uncertainty affects a
commonly used method (MEDUSA, modelling evolutionary diversity using
stepwise AIC) for estimating shifts in diversification rates in
lineages, by extending the method across a random sample of trees from
the posterior distribution of a Bayesian run. We found that phylogenetic
uncertainty greatly affects diversification rate estimates. Different
trees produced diversification rates ranging from high values to almost
zero for the same clade, and both significant rate increase and decrease
in some clades. Only four out of 18 significant shifts found on the
maximum clade credibility tree were consistent across most of the
sampled trees. Among these, we found accelerated diversification for
Ithomiini butterflies. We used the binary speciation and extinction
model (BiSSE) and found that a hostplant shift to Solanaceae is
correlated with increased net diversification rates in Ithomiini,
congruent with the diffuse cospeciation hypothesis. Our results show
that taking phylogenetic uncertainty into account when estimating net
diversification rate shifts is of great importance, and relying on the
maximum clade credibility tree alone can potentially give erroneous
results.

\textbf{Keywords:} diversification analysis, MEDUSA, BiSSE, speciation
rate, insect-hostplant dynamics

\subsection{Introduction}\label{introduction}

Hostplant shifts have been invoked to be responsible for a great part of
the biodiversity of herbivorous insects {[}1{]}. The study of the
evolution of hostplant use has spawned several theories explaining the
evolutionary interactions between plants and insects {[}2{]}. The
``escape-and-radiate'' hypothesis {[}3{]}, the ``oscillation
hypothesis'' {[}4,5{]} or ``diffuse cospeciation'' {[}2{]}.

The butterfly family Nymphalidae has been an important taxon for
developing some of the mentioned hypotheses. Nymphalidae contains around
6000 species {[}6{]}, and is the largest family within the true
butterflies. The family most likely originated around 94 MYA in the mid
Cretaceous. Diversification of the group began in the Late Cretaceous
and most major radiations (current subfamilies) appeared shortly after
the Cretaceous-Paleogene (K-Pg) boundary {[}7{]}. Several studies have
used time-calibrated phylogenies and diversification models to
reconstruct the evolutionary history of the group to identify patterns
of accelerated or decelerated diversification of some Nymphalidae clades
{[}7--10{]}. For example, it has been suggested that climate change in
the Oligocene and the subsequent diversification of grasses has led to
diversification of the subfamily Satyrinae {[}11{]} due to the abundance
of grasses over extensive geographic areas (``resource
abundance-dependent diversity dynamics'' hypothesis). Reference 9 found
increased net diversification rates in some Nymphalidae lineages after a
major hostplant shift, which appears to be in agreement with Reference 3
``escape-and-radiate'' model of diversification.

Although it has been suggested that part of the great diversity of
Nymphalidae butterflies is a result of hostplant-insect dynamics, it is
necessary to use modern techniques to investigate whether the
diversification patterns of Nymphalidae are in agreement with the
theoretical predictions. It is necessary to test whether the overall
diversification pattern of Nymphalidae is congruent with events of
sudden diversification bursts due to hostplant shift, climatic events or
shifts to closely related hostplants {[}5,12{]}.

In this study, we used a time-calibrated genus-level phylogenetic
hypothesis for Nymphalidae butterflies {[}10{]} to investigate patterns
of diversification. We applied MEDUSA {[}13,14{]}, a recently developed
statistical method, to study the diversification pattern of Nymphalidae
butterflies. MEDUSA fits likelihood models of diversification into a
time-calibrated tree and tests whether allowing increases or decreases
in speciation and extinction rates within the tree produces better fit
of the models. MEDUSA is able to take into account unsampled extant
species diversity during model fitting and it is normally applied to the
maximum clade credibility phylogenetic tree. Particularly, we wanted to
study the effects of phylogenetic uncertainty and by using the extended
MEDUSA method called MultiMEDUSA {[}13{]}. We also tested whether
hostplant association dynamics can explain the diversification patterns
of component Nymphalidae lineages by testing whether character states of
hostplant use affected the diversification pattern of those lineages
employing the method BiSSE as implemented in the R package
\texttt{diversitree} {[}15{]}.

\subsection{Methods}\label{methods}

\subsubsection{Data}\label{data}

For analyses, we used the phylogenetic trees from the study of Reference
10 that were generated using DNA sequence data from 10 gene regions for
398 of the 540 valid genera in Nymphalidae. We employed Reference 10
maximum clade credibility tree (MCC tree) (Fig. 1) as well as a random
sample of 1000 trees from their BEAST run after burnin. Their original
BEAST run was for 40 million generations. We used a burnin of 25 million
generations and took a random sample of 1000 trees using Burntrees
v.0.1.9 \url{http://www.abc.se/~nylander/} (supp. mat. 5) in order to
correct for phylogenetic uncertainty when performing the diversification
analyses.

Species richness data for Nymphalidae genera were compiled from several
sources including the specialist-curated lists on
\url{http://tolweb.org}, Reference 16 and curated lists of Global
Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}). We
assigned the species numbers of genera not included in the phylogeny to
the closest related genus that was included in Reference 10 study
according to available phylogenetic studies {[}17--28{]}, taxonomical
classification and morphological resemblance when no phylogenies were
available (supp. mat. 3).

Hostplant data for Nymphalidae species were compiled from several
sources including Reference 29, HOSTS database
(\url{http://bit.ly/YI7nwW}), Reference 30, Reference 31, Reference 32
and others (supp. mat. 15--16) for a total of 6586 hostplant records,
including 428 Nymphalidae genera and 143 plant families and 1070 plant
genera. It was not possible to find any hostplant data for 35 butterfly
genera.

\subsubsection{Analyses of
Diversification}\label{analyses-of-diversification}

We used the statistical software R version 3.0.1 {[}33{]} in combination
with the APE {[}34{]}, GEIGER {[}35{]}, MEDUSA {[}13{]} and
\texttt{diversitree} {[}15{]} packages along with our own scripts to
perform the analyses (included as supplementary materials 2, 6, 10). All
analyses were run on the 1000 random trees from Reference 10, on the
maximum clade credibility tree derived from these and the MCC tree from
Reference 10.

\subsubsection{Detecting diversification shifts on phylogenetic
trees}\label{detecting-diversification-shifts-on-phylogenetic-trees}

Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
version 093.4.33 {[}13{]} on the maximum clade credibility tree from
Reference 10. MEDUSA fits alternative birth-death likelihood models to a
phylogenetic tree in order to estimate changes in net diversification
rates along branches. MEDUSA estimates likelihood and AIC scores for the
simplest birth-death model, with two parameters, the rates \texttt{r}:
net diversification \texttt{=} speciation (\texttt{b}) \texttt{-}
extinction (\texttt{d}) and $\varepsilon$: relative extinction
\texttt{= d/b}. The AIC scores of the two-parameter model are then
compared with incrementally more complex models until the addition of
parameters do not improve the AIC scores beyond a cutoff value. MEDUSA
finds the likelihood of the models after taking into account branch
lengths and number of species per lineage {[}13{]}. Most studies using
MEDUSA only run the method on a single tree, usually the maximum clade
credibility tree, which makes the assumption that this tree is correct.
We wanted to study the effects of phylogenetic uncertainty on estimation
of net diversification rate shifts and therefore selected 1000 random
genus-level trees from the posterior distribution (MultiMEDUSA, supp.
mat. 5). We calculated a new MCC tree from the selection of trees, ran
MEDUSA and MultiMEDUSA on the selection of 1000 trees and summarized the
estimated changes in net diversification rates for nodes across all
trees. Patterns of change in net diversification rates are considered
significant if they are found at the same node in at least 90\% of the
trees. We also expected to find similar values of net diversification
(\texttt{r = b - d}) and relative extinction rate ($\varepsilon$
\texttt{= d/b}) values across the 1000 trees for the nodes where changes
in diversification tempo occurs. We used the AICc threshold of 7.8
units, as estimated by MEDUSA, as the limit for a significantly better
fit to select among increasingly complex alternative models.

\subsubsection{Estimation of trait-dependent speciation
rates}\label{estimation-of-trait-dependent-speciation-rates}

As the MEDUSA and MultiMEDUSA approaches estimated an increase in net
diversification in the clade Ithomiini, we tested whether this pattern
can be explained by increase in birth-rate due to hostplant use and
performed analyses with the ``binary state speciation and extinction''
{[}36{]} Bayesian approach as implemented in the R package
\texttt{diversitree} {[}15{]}. MuSSE {[}15{]} is designed to examine the
joint effects of two or more traits on speciation. Because most of
Nymphalidae butterflies are restricted to use one plant family as
hostplant, the character states can be coded as presence/absence, for
which the BiSSE analysis is better suited. BiSSE was designed to test
whether a binary character state has had any effect on increased net
diversification rate for a clade {[}36{]}. We used our compiled data of
hostplant use to produce a binary dataset for the character ``feeding on
the plant family Solanaceae'' with two states (absence = 0; presence =
1) (supp. mat. 17), which is the main hostplants of the diverse
Ithomiini butterflies and closest relatives {[}37{]}. Other hostplant
shifts were not tested for effect on net diversification rates due to
the low support for diversification shifts found in the MEDUSA analyses.
We analyzed the data using BiSSE employing the Markov Chain Monte Carlo
algorithm on the maximum clade credibility tree, taking into account
missing taxa by using the parameter ``sampling factor''
(\texttt{sampling.f}) in \texttt{diversitree}. We also used constrained
analyses forcing no effect of hostplant use on diversification and used
likelihood ratio tests to find out whether the hypothesis of effect on
diversification has a significantly better likelihood than the null
hypothesis (no effect). The analyses were run across a sample of 250
trees from the posterior distribution. The records of \emph{Vanessa} and
\emph{Hypanartia} feeding on Solanaceae {[}30,38{]} might be incorrect
as it is unlikely that these species can be feeding on this plant
family. We tested whether coding these two genera as not feeding on
Solanaceae affected our results.

\subsection{Results}\label{results}

\subsubsection{Detecting diversification shifts on the maximum clade
credibility
tree}\label{detecting-diversification-shifts-on-the-maximum-clade-credibility-tree}

The MEDUSA analysis on the MCC tree in combination with richness data
estimated 18 changes in the tempo of diversification in Nymphalidae
history (Fig. 1; Table 1). The corrected AICc acceptance threshold for
adding subsequent piecewise birth-death processes to the overall model
was set to 7.8 units, as prescribed by MEDUSA. In all MEDUSA analyses,
the maximum number of inferred diversification splits in all trees was
26. The background net diversification rate for Nymphalidae was
estimated as \texttt{r = 0.092} lineages per Million of years and the
AICc score for the best fit model was \texttt{5090.5} (Table 1). MEDUSA
also estimated that the basic constant birth--death model was not a
better explanation for our data (AICc \texttt{= 5449.3}).

Some of the 18 changes in diversification correspond to rate increases
in very species-rich genera: \emph{Ypthima} (\texttt{r = 0.311}),
\emph{Charaxes} (\texttt{r = 0.291}), \emph{Callicore} +
\emph{Diaethria} (\texttt{r = 0.220}), \emph{Pedaliodes}
(\texttt{r = 0.196}) and \emph{Taenaris} (\texttt{r = 0.234}). We found
rate increases for other clades as well such as: Mycalesina
(\texttt{r = 0.191}), Oleriina + Ithomiina + Napeogenina + Dircennina +
Godyridina (\texttt{r = 0.187}), Satyrini (\texttt{r = 0.116}),
Phyciodina in part (\texttt{r = 0.241}) and Satyrina
(\texttt{r = 0.221}), \emph{Coenonympha} (\texttt{r = 0.209}),
\emph{Caeruleuptychia} + \emph{Magneuptychia} (\texttt{r = 0.312}) and
\emph{Taenaris} (\texttt{r = 0.312}). We also found decreases in net
diversification rates for Limenitidinae + Heliconiinae
(\texttt{r = 0.0541}), part of Danaini (\texttt{r = 0.0423}),
Pseudergolinae (\texttt{r = 0.024}) and Coenonymphina
(\texttt{r = 0.065}) (Table 1).

\subsubsection{Phylogenetic uncertainty in the MultiMEDUSA
approach}\label{phylogenetic-uncertainty-in-the-multimedusa-approach}

We tested MEDUSA to find out whether taking into account the
phylogenetic signal from a random sample of 1000 trees from the
posterior distribution can return similar estimates of diversification
to the values obtained from the MCC tree. We run MultiMEDUSA on a random
sample of 1000 trees (supp. mat. 5) from the posterior distribution and
compared the results with a MEDUSA analysis on the MCC tree derived from
this sample (supp. mat. 9).

We found that the analysis by MultiMEDUSA on the 1000 trees estimated
lower median net diversification rates for the diversification shifts
found by MEDUSA on the MCC tree from Reference 10 (Table 2). Although
the diversification pattern found by MEDUSA and MultiMEDUSA was the
same, the latter consistently estimated lower rates. Thus, the shifts
recovered with low net diversification rate on the MCC were recovered
with negative net diversification rate by MultiMEDUSA. The background
diversification and all shifts found by MEDUSA on the 1000 trees are
provided as an R object in supp. mat. 7).

We also compared the results from MultiMEDUSA (derived from the sample
of 1000 trees) with the splits found by MEDUSA on the MCC tree derived
from this random sample. In the summary statistics, MultiMEDUSA reports
the frequency of the diversification shifts found in the trees
(parameter \texttt{sum.prop}). Thus, if a node is found in only half of
the 1000 trees, but the phylogenetic signal was strong enough to be
picked up by MEDUSA and a node shift was found most of the time, then
the \texttt{sum.prop} should be close to 1. For example the node 625
consists of the clade \texttt{Charaxes} + \texttt{Polyura} and it was
found in only 256 trees, however MultiMEDUSA was consistenly able to
find a diversification shift for that node and the \texttt{sum.prop}
value is 0.996.

We found different rates for diversification shifts even for the nodes
that appear in most of the sample of 1000 trees (Table 2) as many of the
nodes also found in the MCC tree had a probability of being recovered
higher than 0.90.

There were four diversification shifts found with more than 90\% of
probability in the trees from the random sample (Table 2): (i) net
diversification rate increase in the genus \emph{Ypthima}; (ii) net
diversification rate increase in the genus \emph{Charaxes}; (iii) rate
increase in Ithomiini subtribes Oleriina + Ithomiina + Napeogenina +
Dircennina + Godyridina; and (iv) rate increase in \emph{Callicore} +
\emph{Diaethria}.

MultiMEDUSA provided mean and standard deviation statistics for the
diversification values found on the shifts on the 1000 trees (supp. mat.
13--14), and found that some of the changes in net diversification rate
values had great variation across the posterior distribution of trees. A
boxplot of the net diversification rate values estimated for the clades
that appear in the MCC tree shows that some shifts are estimated as
increased or slowed diversification pace depending on the tree used for
analysis (Fig. 2). This variation is especially wide for the clade
formed by the genera \emph{Magneuptychia} and \emph{Caeruleuptychia}
because MEDUSA estimated diversification values from six times the
background net diversification rate (\texttt{r = 0.5234}) to almost zero
(\texttt{r = 1.22e-01}). The rates for \emph{Taenaris} were between 0.14
and 0.44 (mean value 0.25). Similar degrees of variation were found in
the nodes for \emph{Ypthima}, \emph{Charaxes} and \emph{Coenonympha}
(Fig. 2). The net diversification rates estimates for the clades
(Oleriina + Ithomiina + Napeogenina + Dircennina + Godyridina),
Limenitidinae + Heliconiinae and Pseudergolinae are relatively
consistent across the 1000 trees (Fig. 2).

It is also evident that not all the diversification shifts estimated on
the MCC tree are consistently recovered in most of the 1000 trees. Some
of the splits in the MCC tree are recovered in very few trees, for
example the split for the clade Satyrini (Euptychiina + Pronophilina +
Satyrina + Maniolina) is recovered with a probability of 18\% (Fig. 3).

\subsubsection{Estimation of trait-dependent speciation
rates}\label{estimation-of-trait-dependent-speciation-rates-1}

The MEDUSA analyses taking into account phylogenetic uncertainty
estimated a net diversification rate increase in part of the clade
Ithomiini across more than 95\% of the trees. Our BiSSE analysis found a
positive effect of the character state ``feeding on Solanaceae'' on the
net net diversification rate on part of Ithomiini (Oleriina + Ithomiina
+ Napeogenina + Dircennina + Godyridina) (Fig. 4). The Markov Chain
Monte Carlo algorithm was run for 10000 generations discarding the first
7500 as burnin. The estimated mean net diversification rate for taxa
that do not feed on Solanaceae was \texttt{r = 0.11} while the net
diversification rate for the Solanaceae feeders was \texttt{r = 0.16}
(see Fig. S1 for a boxplot of speciation and extinction values for the
95\% credibility intervals). The same analysis considering
\emph{Vanessa} and \emph{Hypanartia} as non-Solanaceae feeders due to
dubious records produced the same pattern and net diversification rates
(Fig. S2). Therefore, the rest of the analyses were performed assuming
these two genera as Solanaceae feeders. We constrained the BiSSE
likelihood model to force equal rates of speciation for both character
states in order to test whether the model of different speciation rates
is a significantly better explanation for the data. A likelihood ratio
test found that the model for increased net diversification rate for
nymphalids feeding on Solanaceae is a significantly better explanation
than this character state having no effect on diversification
(\texttt{p \textless{} 0.001}) (Table 3; character states available in
supp. mat. 17, code in supp. mat. 18, and mcmc run in supp. mat. 19). We
combined the post-burnin mcmc generations from running BiSSE on 250
trees from the posterior distribution and found the same pattern as the
BiSSE analysis on the maximum clade credibility tree (combined mcmc run
in supp. mat. 20; profiles plot of speciation rates in Fig. S3; boxplot
of 95\% credibility intervals in Fig. S4). A BiSSE analysis to test
whether the trait ``feeding on Apocynaceae'' had any effect on increased
net diversification rates found similar speciation rates for lineages
feeding on Apocynaceae and other plants (Fig. S5). It has been shown
that BiSSE performs poorly under certain conditions {[}39{]}. However,
our data has adequate number of taxa under analysis (more than 300
tips), adequate speciation bias (between 1.5x and 2.0x), character state
bias (around 8x) and extinction bias (around 4x) for the analysis of
Solanaceae hostplants. Thus, BiSSE is expected to produce robust results
{[}39{]}.

\subsection{Discussion}\label{discussion}

\subsubsection{Effects of phylogenetic uncertainty on the performance of
MEDUSA}\label{effects-of-phylogenetic-uncertainty-on-the-performance-of-medusa}

The MEDUSA method has been used to infer changes in net diversification
rates along a phylogenetic tree. Since its publication {[}13{]} the
results of using MEDUSA on a single tree, the maximum clade credibility
tree, have been used for generation of hypotheses and discussion
{[}7,40,41{]}. However, we found that MEDUSA estimated different
diversification shifts and different rates of diversification for
certain lineages when phylogenetic uncertainty was taken into account by
using MEDUSA on a random sample of trees from the posterior distribution
of a Bayesian run. We found that some diversification splits, estimated
on the Nymphalidae maximum clade credibility tree, were found with a
very low probability in a random sample of 1000 trees from the posterior
distribution (Fig. 3, Table 2). We also found that, even though MEDUSA
could estimate the same diversification splits on two or more trees, the
estimated net diversification rates could vary widely (Fig. 2). For
example, in our Nymphalidae trees, we found that the split for
\emph{Magneuptychia} and \emph{Caeruleuptychia} had a variation from
\texttt{r = 0.5234}, higher than the background net diversification
rate, to almost zero. This means that observed patterns and conclusions
can be completely contradictory depending on tree choice.

In this study, the effect of phylogenetic uncertainty on the inferred
diversification splits by MEDUSA is amplified because some Nymphalidae
taxa appear to be strongly affected by long-branch attraction artifacts
{[}27{]}. Thus, the Bayesian runs are expected to recover alternative
topologies on the posterior distribution of trees, resulting in low
support and posterior probability values for the nodes. For example,
posterior probability values for clades in Satyrini are very low
{[}10{]}. As a result, MEDUSA inferred a net diversification rate
increase for part of Satyrini in the maximum clade credibility tree, but
this was recovered only in 13\% of probabilities in the MultiMEDUSA
analysis on the random sample of trees.

If there is strong phylogenetic signal for increases or decreases in net
diversification rates for a node, it is expected that these splits would
be inferred by MEDUSA in most of the posterior distribution of trees.
However, weak phylogenetic signal for some nodes can cause some clades
to be absent in some trees and MEDUSA will be unable to estimate any
diversification shift (due to a non-existent node). This is the reason
why MEDUSA estimated net diversification rate splits with more than 90\%
of probabilities in the sample of trees for only four splits: the genus
\emph{Charaxes}, the genus \emph{Ypthima}, part of Ithomiini and the
clade \emph{Callicore} + \emph{Diaethria} (Fig. 3), while estimating
splits for other lineages with much lower probability.

The clade Ithomiini and the non-basal danaids are well supported by high
posterior probability values in Reference 10. Therefore our MEDUSA
analyses recovered an increase in net diversification rate with
probability higher than 90\% in the posterior distribution of trees
(Fig. 3).

\subsubsection{Hostplant use and diversification in
Nymphalidae}\label{hostplant-use-and-diversification-in-nymphalidae}

\paragraph{Ithomiini}\label{ithomiini}

Keith Brown suggested that feeding on Solanaceae was an important event
in the diversification of Ithomiini butterflies {[}42{]}. Ithomiini
butterflies are exclusively Neotropical and most species feed on
Solanaceae hostplants during larval stage {[}37{]}. Optimizations of the
evolution of hostplant use on phylogenies evidence a probable shift from
Apocynaceae to Solanaceae in the ancestor of the tribe {[}37,43{]}.
Reference 9 found that the Gamma statistics, a LTT plot of an Ithomiini
phylogeny and the fit of the density-dependent model of diversification
are consistent with a burst of diversification in Ithomiini following
the shift from Apocynaceae to Solanaceae.

We investigated whether the strong signal for an increase in net
diversification rate for Ithomiini (found by MEDUSA) can be explained
due to the use of Solanaceae plants as hosts during larval stage. For
this, we used a Bayesian approach {[}44{]} to test whether the trait
``feeding on Solanaceae'' had any effect on the diversification of the
group.

Our BiSSE analysis, extended to take into account missing taxa and
phylogenetic uncertainty, shows a significantly higher net
diversification rate for Ithomiini taxa, which can be attributed to the
trait ``feeding on Solanaceae hostplants'' (Fig. 4). This is in
agreement with the findings of Reference 9 using other statistical
methods. Due to the fact that Ithomiini are virtually the only
nymphalids using Solanaceae as hostplants, it is possible that the trait
responsible for a higher diversification of Ithomiini might not be the
hostplant character. As noted by Reference 36, the responsible trait
might be a codistributed character such as a trait related to the
ability to digest secondary metabolites.

Solanaceae plants contain chemical compounds and it has been suggested
that the high diversity of Ithomiini is consistent with the
``escape-and-radiate scenario'' due to a shift onto Solanaceae {[}9{]}
and radiation scenarios among chemically different lineages of
Solanaceae plants {[}37,42{]}. According to this theory, the shift from
Apocynaceae to Solanaceae allowed Ithomiini to invade newly available
resources due to a possible key innovation that allowed them cope with
secondary metabolites of the new hosts. Additional studies are needed to
identify the actual enzymes that Ithomiini species might be using for
detoxification of ingested food as they have been found in other
butterfly groups {[}45{]}.

The increase in diversification inferred by MEDUSA ocurred after the
probable shift from Apocynaceae to Solanaceae, as the Solanaceae feeders
in the subtribes Melinaeina and Mechanitina are not included in the
diversification shift (shift number 5 in Fig. 1). The apparent
conflicting results from MEDUSA and BiSSE can be explained by the low
species-richness of the subtribes Melinaeina and Mechanitina compared to
the other subtribes included in the shift (52 versus 272 species). It
can be that MEDUSA is more conservative than BiSSE and is not including
Melinaeina and Mechanitina in the shift due to low species numbers.

Although the Solanaceae genera used by the Ithomiini clades are well
known {[}37{]}, we do not have any understanding on the physiological
routes involved in the detoxification of Solanaceae compounds by the
several lineages of Ithomiini. We can speculate that older lineages
exploiting a novel toxic resource {[}10,37{]} might not be too efficient
in metabolizing plant toxins and that younger lineages are able to deal
with toxins more efficiently, so that host switching events within
Solanaceae are possible, which can lead to higher diversification.
Studies in \emph{Papilio} species have reported that detoxification
enzymes can become more efficient in metabolizing toxins than ancestral
configurations of the proteins, providing more opportunities for
hostplant switch {[}46{]}. This might be the reason why the basal
Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and
restricted to few Solanaceae hosts {[}37{]}, while recent subtribes are
species-rich and have expanded their host range into several Solanaceae
lineages {[}37{]}. It might be that the switch to feeding in Solanaceae
was an important event in the evolutionary history of Ithomiini, but the
actual radiation occurred after critical physiological changes (a
probable key innovation) allowed efficient detoxification of Solanaceae
toxins.

The diffuse cospeciation hypothesis predicts almost identical ages of
insects and their hostplants, while the ``resource abundance-dependent
diversity'' and the ``escape-and-radiate'' hypotheses state that insects
diversify after their hostplants {[}2--4{]}. Reference 45 found strong
evidence for a model of speciation congruent with Ehrlich and Raven's
hypothesis in Pieridae butterflies due to, in addition to the
identification of a key innovation, a burst of diversification in
glucosinolate-feeding taxa shortly afterwards (with a lag of
\textasciitilde{}10 MY). According to a recent dated phylogeny of the
Angiosperms {[}47{]}, the family Solanaceae split from its sister group
about 59 (49-68) MYA and diversification started (crown group age)
around 37 (29-47) MYA. Reference 10 give the corresponding ages for
Ithomiini as 45 (39-53) and 37 (32-43) MYA, respectively. Thus, current
evidence shows that Solanaceae and Ithomiini might have diversified
around the same time, during the Late Eocene and Oligocene, and this
would be congruent with the diffuse cospeciation hypothesis.

The uplift of the Andes was a major tectonic event that underwent higher
activity during the Oligocene {[}48{]}. This caused climatic changes in
the region that affected the flora and fauna of the time, which
coincides with the diversification of modern montane plant and animal
taxa {[}49{]} including Ithomiini butterflies and Solanaceae hostplants.
Moreover, all Solanaceae clades currently present in New World
originated in South America {[}50{]} as well as Ithomiini butterflies
{[}10{]}. This suggests a process of ``diffuse cospeciation'' of
Ithomiini and hostplants.

\paragraph{Danaini}\label{danaini}

Our MultiMEDUSA approach gave a significant slowdown in net
diversification rate in the subtribe Danaina of the Danini. Both Danaina
and the sister clade Euploeina feed mainly on Apocynaceae and thus a
hostplant shift should not be responsible for the observed slowdown of
diversification in the Danaina. As expected, our BiSSE analysis of
Apocynaceae feeders shows that there is no effect of feeding on this
plant family on the net diversification rates of Nymphalidae lineages.
Many of the Danaina are large, strong fliers, highly migratory and
involved in mimicry rings. Among them is for example the monarch
(\emph{Danaus plexippus}), probably the most well known of all migratory
butterflies. The causes for a lower net diversification rate in the
Danaina remains to be investigated, but their great dispersal power
might be involved in preventing allopatric speciation. It has been found
in highly vagile species in the nymphalid genus \emph{Vanessa} that
dispersal has homogenized populations due to gene flow, as old and
vagile species seem to be genetically homogeneous while younger
widespread species show higher genetic differentiation in their
populations {[}51{]}.

\paragraph{\emph{Charaxes}}\label{charaxes}

The genus \emph{Charaxes} contain 193 species distributed in the Old
World with highest diversity in the Afrotropical region. These
butterflies are also very strong fliers, but contrary to Danaina, which
are specialized Apocynaceae feeders, \emph{Charaxes} are known to feed
on at least 28 plant families in 18 orders {[}29{]} and some species
appear to be polyphagous {[}52{]}. Reference 53 showed that most of the
diversification of the genus occurred from the late Oligocene to the
Miocene when there were drastic global climatic fluctuations, indicating
that the diversity mainly is driven by climate change. Reference 52
found that climatic changes during the Pliocene, Pleistocene as well as
dispersal and vicariance might have been responsible for the high
diversification of the genus.

\paragraph{Satyrini}\label{satyrini}

The diverse tribe Satyrini radiated simultaneously with the radiation of
their main hostplant, grasses, during the climatic cooling in the
Oligocene {[}11{]}. Thus, it is somewhat surprising that part of
Satyrini (the subtribes Euptychiina, Satyrina and Pronophilina) were
found to have accelerated diversification in only 13\% of probabilities
in the trees from the posterior distribution. Although this can be
attributed to low phylogenetic signal {[}10{]}, the clade Satyrini is
very robust {[}10{]} and MEDUSA failed to identify any significant
accelerated net diversification rate for Satyrini. It appears that the
radiation of Satyrini as a whole was not remarkably fast and therefore
not picked up by MEDUSA.

The origin of the tribe Satyrini is not completely clear (originated
either in the Neotropical or Eastern Palaearctic, Oriental and/or
Indo-Australian regions {[}27{]} and their radiation involved colonizing
almost all continents starting from their place of origin. For
butterflies, a prerequisite for colonizing new areas is that suitable
hostplants are already present. Therefore, it should be expected that
the diversification of Satyrinae occurred in a stepwise manner, with
pulses or bursts of diversification for certain lineages but unlikely
for the tribe Satyrini as a whole.

We found that even though MEDUSA estimated several diversification
shifts in the maximum clade credibility tree of Nymphalidae, only a few
of these splits were found in more than 90\% of the trees from the
posterior distribution. In the literature, it is common practice that
conclusions are based on the splits estimated on the maximum clade
credibility tree. However, by using a MultiMEDUSA approach, we found
that some of this splits might be greatly affected by phylogenetic
uncertainty. Moreover, some of these splits can be recovered either as
increases or decreases in net diversification rate depending on the tree
from the posterior distribution that was used for analysis. This means
that contradictory conclusions would be made if only the maximum clade
credibility tree was used for analysis. MEDUSA appears to be sensitive
to the number of nodes with high posterior probability and width of age
confidence intervals. For our data, it would be necessary to obtain a
posterior distribution of trees with no conflicting topology, and very
similar estimated ages for nodes in order to consistently recover most
of the diversification splits on the posterior distribution of trees
that were inferred by MEDUSA on the MCC tree.

Our MultiMEDUSA approach to perform analyses on the posterior
distribution of trees found strong support for an increase in net
diversification rate in the tribe Ithomiini and the genus
\emph{Charaxes}, and for a decrease in net diversification rate in the
subtribe Danaina. Due to phylogenetic uncertainty, we did not obtain
strong support for other diversification splits in Nymphalidae. Our
BiSSE analysis corroborated other studies in that the trait ``feeding on
Solanaceae'', or a codistributed character, was important in the
diversification of Ithomiini butterflies. However, by applying MEDUSA we
found that a critical character in the radiation of the group might have
appeared after the shift from Apocynaceae to Solanaceae. We also found
that the trait ``feeding on Apocynaceae'' is not responsible for the
slowdown of diversification in Danaina. Ithomiini and Solanaceae
diversified near simultaneously, which is in agreement with the diffuse
co-speciation hypothesis {[}2,4{]}.

\subsection{Acknowledgments}\label{acknowledgments}

We are thankful to Mark Cornwall for help with the script to extend
MEDUSA to include phylogenetic uncertainty, Niklas Wahlberg for
commenting on the manuscript and giving us the posterior distribution of
trees, Luke Harmon for commenting on the manuscript and anonymous
reviewers for their comments, which greatly improved the manuscript,
Jessica Slove Davidson and Niklas Janz for access to their hostplant
data. The study was supported by a Kone Foundation grant (awarded to
Niklas Wahlberg), Finland (C. Peña) and the Research Council of Norway
(grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
Science Ltd. (Finland) for the allocation of computational resources.

\subsection{Figure legends}\label{figure-legends}

\texttt{Figure 1.} Results of the MEDUSA analysis run on the maximum
clade credibility tree from Reference 10. Rate shifts were estimated for
the following nodes (besides the background rate): 2) Limenitidinae +
Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
part, 6) Satyrini, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
\emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
Satyrina, 14) Mycalesinaa, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae.

\texttt{Figure 2.} Boxplot of the range of diversification values for
tips or clades estimated by MEDUSA on the 1000 random trees from the
posterior distribution of the Nymphalidae phylogeny. The tips or clades
shown are those present on the maximum clade credibility tree from
Reference 10.

\texttt{Figure 3.} Results of the MultiMEDUSA analysis on 1000 randomly
sampled trees from the posterior distribution of the Nymphalidae
phylogeny. Bars show the probability for nodes of being recovered as
significant increases or decreases in net diversification rates by
MultiMEDUSA.

\texttt{Figure 4.} BiSSE analysis of diversification of nymphalids due
to feeding on Solanaceae hostplants. Speciation and net diversification
rates are significantly higher in Solanaceae feeders (\$lambda\$1, r1).

\subsection{Tables}\label{tables}

\texttt{Table 1.} Significant net diversification rate shifts found in
the MEDUSA analysis of the Nymphalid maximum clade credibility tree.
Split.Node = node number, Model = preferred diversification model by
MEDUSA, r = net diversification rate, LnLik.part = log likelihood value.

\texttt{Table 2.} Differences in rates estimated by MEDUSA on the MCC
tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
from the posterior distribution.

\texttt{Table 3.} Likelihood ratio test between the model of increased
diversification of nymphalids feeding on Solanaceae against a model
forcing equal speciation rates (no effect on diversification).

\section{Supp. mat.}\label{supp.-mat.}

\textbf{\texttt{supp\_mat\_01\_genus.nex}}: MCC Nymphalidae tree from
Reference 10.

\textbf{\texttt{supp\_mat\_02\_run\_medusa\_on\_mct.R}}: Run MEDUSA on
MCC tree from Reference 10.

\textbf{\texttt{supp\_mat\_03\_richness.nex}}: Species richness for
lineages in Nymphalidae.

\textbf{\texttt{supp\_mat\_04\_MEDUSA\_results.txt}}: Results of running
MEDUSA on the MCC tree.

\textbf{\texttt{supp\_mat\_05\_1000\_random\_trees\_no\_outgroups.nex}}:
1000 random trees from Reference 10.

\textbf{\texttt{supp\_mat\_06\_run\_multimedusa.R}}: Run MultiMEDUSA on
1000 random trees from Reference 10.

\textbf{\texttt{supp\_mat\_07\_multimedusa\_on\_1000\_trees.txt}}: Raw
results from the MultiMEDUSA run on the random sample of trees from the
posterior distribution.

\textbf{\texttt{supp\_mat\_08\_multimedusa\_result01.pdf}}: Summary of
MulitMEDUSA analysis on the 1000 trees from the posterior distribution.

\textbf{\texttt{supp\_mat\_09\_mcc\_tree\_from\_random\_sample\_of\_1000.tree}}:
MCC tree from the 1000 random trees selected from the posterior
distribution.

\textbf{\texttt{supp\_mat\_10\_run\_medusa\_on\_mct\_from\_1000\_trees.R}}:
Run MEDUSA on MCC tree from the 1000 random trees selected from the
posterior distribution.

\textbf{\texttt{supp\_mat\_11\_results\_of\_MEDUSA\_run\_on\_mcc\_from\_1000\_trees.txt}}:
Summary results from a MEDUSA run on the MCC tree from the 1000 random
trees selected from the posterior distribution.

\textbf{\texttt{supp\_mat\_12\_MEDUSA\_mcc\_tree\_from\_random\_sample\_of\_1000\_trees.pdf}}:
Figure for MEDUSA run on MCC tree from random 1000 trees.

\textbf{\texttt{supp\_mat\_13\_results\_of\_Multi-MEDUSA\_run\_on\_1000\_trees.txt}}:
Summary results from a MultiMEDUSA run on the 1000 random trees selected
from the posterior distribution.

\textbf{\texttt{supp\_mat\_14\_multimedusa\_result\_prob\_nodes.pdf}}:
Probability of nodes with estimated rates from a MultiMEDUSA run on the
1000 random trees selected from the posterior distribution.

\textbf{\texttt{supp\_mat\_15\_host\_plant\_data.csv}}: Hostplants of
Nymphalidae butterflies recorded from the literature.

\textbf{\texttt{supp\_mat\_16\_host\_plant\_data\_references.csv}}:
References for hostplants data.

\textbf{\texttt{supp\_mat\_17\_hostplant\_states.csv}}: Data matrix with
character states for hosplant use.

\textbf{\texttt{supp\_mat\_18\_bisse\_source.R}}: R code for running the
BiSSE analysis.

\textbf{\texttt{supp\_mat\_19\_bisse\_mcmc\_run.csv}}: Raw results for
the BiSSE analysis.

\textbf{\texttt{supp\_mat\_20\_combined\_bisse.csv}}: Raw results for
the combined BiSSE analysis.

\texttt{Figure S1.} Boxplot of speciation (lambda) and extinction (mu)
values for the 95\% credibility intervals of values estimated by BiSSE
analysis of diversification due to feeding on Solanaceae plants.

\texttt{Figure S2.} BiSSE analysis of diversification of nymphalids due
to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
recovered, speciation and net diversification rates are significantly
higher for Solanaceae feeders (\$lambda\$1, r1).

\texttt{Figure S3.} Net diversification rates of nymphalids feeding on
Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
the 1000 trees from the posterior distribution.

\texttt{Figure S4.} Boxplot of speciation and extinction values for the
95\% credibility intervals of values estimated by BiSSE analysis of
diversification due to feeding on Solanaceae plants on the combined
post-burnin runs on 1000 trees from the posterior distribution.

\texttt{Figure S5.} BiSSE analysis of diversification of nymphalids due
to feeding on Apocynaceae hostplants. Speciation and net diversification
rates are similar.

\subsection{References}\label{references}

1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
adaptive zones: has phytophagy promoted insect diversification? Am Nat
132: 107--128.
doi:\href{http://dx.doi.org/10.1086/284840}{10.1086/284840}.

2. Nyman T, Linder HP, Peña C, Malm T, Wahlberg N (2012) Climate-driven
diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
889--898. Available:
\url{http://doi.wiley.com/10.1111/j.1461-0248.2012.01782.x}. Accessed 29
March 2013.

3. Ehrlich PR, Raven PH (1964) Butterflies and plants: a study in
coevolution. Evolution 18: 586--608.

4. Janz N (2011) Ehrlich and raven revisited: mechanisms underlying
codiversification of plants and enemies. Annu Rev Ecol Evol S 42:
71--89. Available:
\url{http://www.annualreviews.org/doi/abs/10.1146/annurev-ecolsys-102710-145024}.

5. Nylin S, Slove J, Janz N (2013) Host plant utilization, host range
oscillations and diversification in nymphalid butterflies: a
phylogenetic investigation. Evolution: in press.

6. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
al. (2011) Order lepidoptera. In:. Animal biodiversity: an outline of
higher-level classification and survey of taxonomic richness. Zootaxa.
pp. 212--221.

7. Heikkilä M, Kaila L, Mutanen M, Peña C, Wahlberg N (2012) Cretaceous
origin and repeated tertiary diversification of the redefined
butterflies. P Roy Soc B 279: 1093--1099. Available:
\href{http://rspb.royalsocietypublishing.org/cgi/doi/10.1098/rspb.2011.1430 http://www.ncbi.nlm.nih.gov/pubmed/21920981}{http://rspb.royalsocietypublishing.org/cgi/doi/10.1098/rspb.2011.1430
http://www.ncbi.nlm.nih.gov/pubmed/21920981}.

8. Elias M, Joron M, Willmott KR, Silva-Brandão KL, Kaiser V, et al.
(2009) Out of the andes: patterns of diversification in clearwing
butterflies. Mol Ecol 18: 1716--1729. Available:
\url{http://www.ncbi.nlm.nih.gov/pubmed/19386035}.

9. Fordyce JA (2010) Host shifts and evolutionary radiations of
butterflies. P Roy Soc B 277: 3735--3743. Available:
\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=2992698\&tool=pmcentrez\&rendertype=abstract}.

10. Wahlberg N, Leneveu J, Kodandaramaiah U, Peña C, Nylin S, et al.
(2009) Nymphalid butterflies diversify following near demise at the
Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302. Available:
\url{http://rspb.royalsocietypublishing.org/cgi/doi/10.1098/rspb.2009.1303}.
Accessed 29 March 2013.

11. Peña C, Wahlberg N (2008) Prehistorical climate change increased
diversification of a group of butterflies. Biol Lett 4: 274--278.
Available:
\url{http://rsbl.royalsocietypublishing.org/cgi/doi/10.1098/rsbl.2008.0062}.
Accessed 29 March 2013.

12. Ferrer-Paris JR, Sánchez-Mercado A, Viloria ÁL, Donaldson J (2013)
Congruence and diversity of butterfly-host plant associations at higher
taxonomic levels. PloS one 8: e63570.

13. Alfaro ME, Santini F, Brock C, Alamillo H, Dornburg A, et al. (2009)
Nine exceptional radiations plus high turnover explain species diversity
in jawed vertebrates. P Natl Acad Sci USA 106: 13410--13414. Available:
\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=2715324\$/backslash\$\&tool=pmcentrez\$/backslash\$\&rendertype=abstract}.

14. Harmon LJ, Rabosky DL, FitzJohn RG, Brown JW (2011) turboMEDUSA:
MEDUSA: modeling evolutionary diversification using stepwise AIC.

15. FitzJohn RG (2012) Diversitree: comparative phylogenetic analyses of
diversification in r. Method Ecol Evol 3: 1084--1092. Available:
\url{http://doi.wiley.com/10.1111/j.2041-210X.2012.00234.x}.

16. Lamas G (2004) Checklist: part 4A. hesperioidea-papilionoidea.
Gainesville.

17. Matos-Maraví PF, Peña C, Willmott KR, Freitas AV, Wahlberg N (2013)
Systematics and evolutionary history of butterflies in the ``Taygetis
clade'' (nymphalidae: satyrinae: euptychiina): towards a better
understanding of neotropical biogeography. Mol Phylogenet Evol 66:
54--68. Available:
\url{http://linkinghub.elsevier.com/retrieve/pii/S1055790312003624}.
Accessed 29 March 2013.

18. Brower AVZ, Wahlberg N, Ogawa JR, Boppré M, Vane-Wright RI (2010)
Phylogenetic relationships among genera of danaine butterflies
(lepidoptera: nymphalidae) as implied by morphology and DNA sequences.
Syst Biodivers 8: 75--89.
doi:\href{http://dx.doi.org/10.1080/14772001003626814}{10.1080/14772001003626814}.

19. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
(2010) Phylogenetics and biogeography of a spectacular old world
radiation of butterflies: the subtribe mycalesina (lepidoptera:
nymphalidae: satyrini). BMC Evol Biol 10: 172. Available:
\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=2898688\&tool=pmcentrez\&rendertype=abstract}.

20. Kodandaramaiah U, Peña C, Braby MF, Grund R, Müller CJ, et al.
(2010) Phylogenetics of coenonymphina (nymphalidae: satyrinae) and the
problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
Available:
\url{http://linkinghub.elsevier.com/retrieve/pii/S1055790309003273}.
Accessed 29 March 2013.

21. Ortiz-Acevedo E, Willmott KR (2013) Molecular systematics of the
butterfly tribe preponini (nymphalidae: charaxinae). Syst Entomol:
no--o. Available: \url{http://doi.wiley.com/10.1111/syen.12008}.

22. De-Silva DL, Day JJ, Elias M, Willmott KR, Whinnett A, et al. (2010)
Molecular phylogenetics of the neotropical butterfly subtribe oleriina
(nymphalidae: danainae: ithomiini). Mol Phylogenet Evol 55: 1032--1041.
Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/20079859}.

23. Freitas AVL, Brown Jr. KS (2004) Phylogeny of the nymphalidae
(lepidoptera). Syst Biol 53: 363--383. Available:
\url{http://sysbio.oxfordjournals.org/cgi/doi/10.1080/10635150490445670}.

24. Peña C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
(2006) Higher level phylogeny of satyrinae butterflies (lepidoptera:
nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/16563805}.

25. Penz CM (1999) Higher level phylogeny for the passion-vine
butterflies (nymphalidae, heliconiinae) based on early stage and adult
morphology. Zool J Linn Soc 127: 277--344. Available:
\url{http://doi.wiley.com/10.1111/j.1096-3642.1999.tb00680.x}. Accessed
03 September 2013.

26. Silva-Brandão KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
tribe acraeini (lepidoptera, nymphalidae, heliconiinae) and the
evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/18191590}.

27. Peña C, Nylin S, Wahlberg N (2011) The radiation of satyrini
butterflies (nymphalidae: satyrinae): a challenge for phylogenetic
methods. Zool J Linn Soc 161: 64--87. Available:
\url{http://doi.wiley.com/10.1111/j.1096-3642.2009.00627.x}. Accessed 29
March 2013.

28. Peña C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
history of the butterfly subtribe euptychiina (lepidoptera, nymphalidae,
satyrinae). Zool Scr 39: 243--258. Available:
\url{http://blackwell-synergy.com/doi/abs/10.1111/j.1463-6409.2010.00421.x}.

29. Ackery PR (1988) Hostplants and classification: a review of
nymphalid butterflies. Biol J Linn Soc 33: 95--203.

30. Dyer LA, Gentry GL (2002) Caterpillars and parasitoids of a tropical
lowland wet forest. Available: \url{http://www.caterpillars.org}.
Accessed 17 April 2013.

31. Beccaloni G, Viloria Á, Hall S, Robinson G (2008) Catalogue of the
hostplants of the neotropical butterflies. London: The Natural History
Museum.

32. Janzen DH, Hallwachs W (2009) Dynamic database for an inventory of
the macrocaterpillar fauna, and its food plants and parasitoids, of area
de conservacion guanacaste (aCG), northwestern costa rica. Available:
\url{http://janzen.sas.upenn.edu}. Accessed 17 April 2013.

33. Core Team R (2013) R: a language and environment for statistical
computing. Vienna, Austria: R Foundation for Statistical Computing.
Available: \url{http://www.r-project.org}.

34. Popescu A-A, Huber KT, Paradis E (2012) ape 3.0: new tools for
distance-based phylogenetics and evolutionary analysis in r.
Bioinformatics 28: 1536--1537.

35. Harmon LJ, Weir JT, Brock CD, Glor RE, Challenger W (2008) GEIGER:
investigating evolutionary radiations. Bioinformatics 24: 129--131.
Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/18006550}.

36. Maddison WP, Midford PE, Otto SP (2007) Estimating a binary
character's effect on speciation and extinction. Syst Biol 56: 701--710.
Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/17849325}.

37. Willmott KR, Freitas AVL (2006) Higher-level phylogeny of the
ithomiinae (lepidoptera: nymphalidae): classification, patterns of
larval hostplant colonization and diversification. Cladistics 22:
297--368. Available:
\url{http://doi.wiley.com/10.1111/j.1096-0031.2006.00108.x}.

38. Scott JA (1986) The butterflies of north america. Stanford
University Press.

39. Davis MP, Midford PE, Maddison W (2013) Exploring power and
parameter estimation of the BiSSE method for analyzing species
diversification. BMC Evol Biol 13: 38. Available:
\url{http://www.biomedcentral.com/1471-2148/13/38}.

40. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
bees cut leaves? new insights into the early evolution of bees. P Roy
Soc B 278: 3593--3600. Available:
\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3189370\$/backslash\$\&tool=pmcentrez\$/backslash\$\&rendertype=abstract}.

41. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
clades of agaricales. P Roy Soc B 279: 2003--2011. Available:
\url{http://www.ncbi.nlm.nih.gov/pubmed/22171078}.

42. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
interface. Ann Mo Bot Gard 74: 359--397.

43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brandão KL, Whinnett A, et
al. (2006) Phylogenetic relationships among the ithomiini (lepidoptera:
nymphalidae) inferred from one mitochondrial and two nuclear gene
regions. Syst Entomol 31: 288--301. Available:
\url{http://doi.wiley.com/10.1111/j.1365-3113.2006.00321.x}.

44. FitzJohn RG, Maddison WP, Otto SP (2009) Estimating trait-dependent
speciation and extinction rates from incompletely resolved phylogenies.
Syst Biol 58: 595--611. Available:
\url{http://www.ncbi.nlm.nih.gov/pubmed/20525612}.

45. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
The genetic basis of a plant-insect coevolutionary key innovation. P
Natl Acad Sci USA 104: 20427--20431.
doi:\href{http://dx.doi.org/10.1073/pnas.0706229104}{10.1073/pnas.0706229104}.

46. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
furanocoumarin-metabolizing cytochrome p450 monooxygenases in two
papilionids: specificity and substrate encounter rate. P Natl Acad Sci
USA 100: 14593--14598. Available:
\url{http://www.pnas.org/content/100/suppl.2/14593.abstract}.

47. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
the angiosperms re-revisited. Am J Bot 97: 1296--1303. Available:
\url{http://www.ncbi.nlm.nih.gov/pubmed/21616882}.

48. Somoza R (1998) Updated nazca (farallon) - south america relative
motions during the last 40 my: implications for mountain building in the
central andean region. J S Am Earth Sci 11: 211--215.

49. Hoorn C, Wesselingh FP, Steege H ter, Bermudez MA, Mora A, et al.
(2010) Amazonia through time: andean uplift, climate change, landscape
evolution, and biodiversity. Science 330: 927--931. Available:
\url{http://www.ncbi.nlm.nih.gov/pubmed/21071659}.

50. Olmstead RG (2013) Phylogeny and biogeography in solanaceae,
verbenaceae and bignoniaceae: a comparison of continental and
intercontinental diversification patterns. Bot J Linn Soc 171: 80--102.
Available: \url{http://doi.wiley.com/10.1111/j.1095-8339.2012.01306.x}.

51. Wahlberg N, Rubinoff D (2011) Vagility across vanessa (lepidoptera:
nymphalidae): mobility in butterfly species does not inhibit the
formation and persistence of isolated sister taxa. Syst Entomol 36:
362--370. Available:
\url{http://doi.wiley.com/10.1111/j.1365-3113.2010.00566.x}. Accessed 16
April 2013.

52. Müller CJ, Wahlberg N, Beheregaray LB (2010) `After africa': the
evolutionary history and systematics of the genus charaxes ochsenheimer
(lepidoptera: nymphalidae) in the indo-pacific region. Biol J Linn Soc
100: 457--481. Available:
\url{http://doi.wiley.com/10.1111/j.1095-8312.2010.01426.x}.

53. Aduse-Poku K, Vingerhoedt E, Wahlberg N (2009) Out-of-africa again:
a phylogenetic hypothesis of the genus charaxes (lepidoptera:
nymphalidae) based on 5 gene regions. Mol Phylogenet Evol 53: 463--478.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.06.021}{10.1016/j.ympev.2009.06.021}.

\end{document}

### 2014-07-22 MS.tex
sections and subsections

@@
 Evolutionary Biology, Harvard University, Cambridge, USA. Email:
-\href{mailto:marianne.espeland@gmail.com}{marianne.espeland@gmail.com}.
+marianne.espeland@gmail.com.
 Current address: McGuire Center for Lepidoptera and Biodiversity,
@@
 
-\subsection{Abstract}\label{abstract}
+\section*{Abstract}
 
@@
 
-\subsection{Introduction}\label{introduction}
+\section*{Introduction}
 
@@
 
-\subsection{Methods}\label{methods}
-
-\subsubsection{Data}\label{data}
+\section*{Materials and Methods}
+
+\subsection*{Data}
 
@@
 
-\subsubsection{Analyses of
-Diversification}\label{analyses-of-diversification}
+\subsection*{Analyses of Diversification}
 
@@
 
-\subsubsection{Detecting diversification shifts on phylogenetic
-trees}\label{detecting-diversification-shifts-on-phylogenetic-trees}
+\subsection*{Detecting diversification shifts on phylogenetic trees}
 
@@
 
-\subsubsection{Estimation of trait-dependent speciation
-rates}\label{estimation-of-trait-dependent-speciation-rates}
+\subsection*{Estimation of trait-dependent speciation rates}
 
@@
 
-\subsection{Results}\label{results}
-
-\subsubsection{Detecting diversification shifts on the maximum clade
-credibility
-tree}\label{detecting-diversification-shifts-on-the-maximum-clade-credibility-tree}
+\section*{Results}
+
+\subsection*{Detecting diversification shifts on the maximum clade credibility tree}
 
@@
 
-\subsubsection{Phylogenetic uncertainty in the MultiMEDUSA
-approach}\label{phylogenetic-uncertainty-in-the-multimedusa-approach}
+\subsection*{Phylogenetic uncertainty in the MultiMEDUSA approach}
 
@@
 
-\subsubsection{Estimation of trait-dependent speciation
-rates}\label{estimation-of-trait-dependent-speciation-rates-1}
+\subsection*{Estimation of trait-dependent speciation rates}
 
@@
 
-\subsection{Discussion}\label{discussion}
-
-\subsubsection{Effects of phylogenetic uncertainty on the performance of
-MEDUSA}\label{effects-of-phylogenetic-uncertainty-on-the-performance-of-medusa}
+\section*{Discussion}
+
+\subsection*{Effects of phylogenetic uncertainty on the performance of MEDUSA}
 
@@
 
-\subsubsection{Hostplant use and diversification in
-Nymphalidae}\label{hostplant-use-and-diversification-in-nymphalidae}
-
-\paragraph{Ithomiini}\label{ithomiini}
+\subsection*{Hostplant use and diversification in Nymphalidae}
+
+\subsubsection*{Ithomiini}
 
@@
 
-\paragraph{Danaini}\label{danaini}
+\subsubsection*{Danaini}
 
@@
 
-\paragraph{\emph{Charaxes}}\label{charaxes}
+\subsubsection*{\emph{Charaxes}}
 
@@
 
-\paragraph{Satyrini}\label{satyrini}
+\subsubsection*{Satyrini}
 
@@
 
-\subsection{Acknowledgments}\label{acknowledgments}
+\section*{Acknowledgments}
 
@@
 
-\subsection{Figure legends}\label{figure-legends}
+\section*{Figure legends}
 
@@
 
-\subsection{Tables}\label{tables}
+\section*{Tables}
 
@@
 
-\section{Supp. mat.}\label{supp.-mat.}
+\section*{Supp. mat.}
 
@@
 
-\subsection{References}\label{references}
+\section*{References}
 

### 2014-07-22 MS.tex
sanchez

@@
 \section*{References}
-
 1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
@@
 diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
-889--898. Available:
-\url{http://doi.wiley.com/10.1111/j.1461-0248.2012.01782.x}. Accessed 29
-March 2013.
-
-3. Ehrlich PR, Raven PH (1964) Butterflies and plants: a study in
+889--898.
+doi:\href{http://dx.doi.org/10.1111/j.1461-0248.2012.01782.x}{10.1111/j.1461-0248.2012.01782.x}.
+
+3. Ehrlich PR, Raven PH (1964) Butterflies and plants: A study in
 coevolution. Evolution 18: 586--608.
-
-4. Janz N (2011) Ehrlich and raven revisited: mechanisms underlying
-codiversification of plants and enemies. Annu Rev Ecol Evol S 42:
-71--89. Available:
-\url{http://www.annualreviews.org/doi/abs/10.1146/annurev-ecolsys-102710-145024}.
-
-5. Nylin S, Slove J, Janz N (2013) Host plant utilization, host range
-oscillations and diversification in nymphalid butterflies: a
-phylogenetic investigation. Evolution: in press.
+doi:\href{http://dx.doi.org/10.2307/2406212}{10.2307/2406212}.
+
+4. Janz N (2011) Ehrlich and Raven Revisited: Mechanisms Underlying
+Codiversification of Plants and Enemies. Annu Rev Ecol Evol S 42:
+71--89.
+doi:\href{http://dx.doi.org/10.1146/annurev-ecolsys-102710-145024}{10.1146/annurev-ecolsys-102710-145024}.
+
+5. Nylin S, Slove J, Janz N (2014) Host plant utilization, host range
+oscillations and diversification in Nymphalid butterflies: A
+phylogenetic investigation. Evolution 68: 105--124.
+doi:\href{http://dx.doi.org/10.1111/evo.12227}{10.1111/evo.12227}.
 
 6. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
-al. (2011) Order lepidoptera. In:. Animal biodiversity: an outline of
+al. (2011) Order Lepidoptera. In:. Animal biodiversity: an outline of
 higher-level classification and survey of taxonomic richness. Zootaxa.
@@
 origin and repeated tertiary diversification of the redefined
-butterflies. P Roy Soc B 279: 1093--1099. Available:
-\href{http://rspb.royalsocietypublishing.org/cgi/doi/10.1098/rspb.2011.1430 http://www.ncbi.nlm.nih.gov/pubmed/21920981}{http://rspb.royalsocietypublishing.org/cgi/doi/10.1098/rspb.2011.1430
-http://www.ncbi.nlm.nih.gov/pubmed/21920981}.
+butterflies. P Roy Soc B 279: 1093--1099.
+doi:\href{http://dx.doi.org/10.1098/rspb.2011.1430}{10.1098/rspb.2011.1430}.
 
 8. Elias M, Joron M, Willmott KR, Silva-Brandão KL, Kaiser V, et al.
-(2009) Out of the andes: patterns of diversification in clearwing
-butterflies. Mol Ecol 18: 1716--1729. Available:
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19386035}.
+(2009) Out of the Andes: Patterns of diversification in clearwing
+butterflies. Mol Ecol 18: 1716--1729.
+doi:\href{http://dx.doi.org/10.1111/j.1365-294X.2009.04149.x}{10.1111/j.1365-294X.2009.04149.x}.
 
 9. Fordyce JA (2010) Host shifts and evolutionary radiations of
-butterflies. P Roy Soc B 277: 3735--3743. Available:
-\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=2992698\&tool=pmcentrez\&rendertype=abstract}.
+butterflies. P Roy Soc B 277: 3735--3743.
+doi:\href{http://dx.doi.org/10.1098/rspb.2010.0211}{10.1098/rspb.2010.0211}.
 
@@
 (2009) Nymphalid butterflies diversify following near demise at the
-Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302. Available:
-\url{http://rspb.royalsocietypublishing.org/cgi/doi/10.1098/rspb.2009.1303}.
-Accessed 29 March 2013.
+Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302.
+doi:\href{http://dx.doi.org/10.1098/rspb.2009.1303}{10.1098/rspb.2009.1303}.
 
@@
 diversification of a group of butterflies. Biol Lett 4: 274--278.
-Available:
-\url{http://rsbl.royalsocietypublishing.org/cgi/doi/10.1098/rsbl.2008.0062}.
-Accessed 29 March 2013.
-
-12. Ferrer-Paris JR, Sánchez-Mercado A, Viloria ÁL, Donaldson J (2013)
-Congruence and diversity of butterfly-host plant associations at higher
-taxonomic levels. PloS one 8: e63570.
+doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.
+
+12. Ferrer-Paris JR, S\'anchez-Mercado A, Viloria ÁL, Donaldson J (2013)
+Congruence and Diversity of Butterfly-Host Plant Associations at Higher
+Taxonomic Levels. PloS one 8: e63570.
+doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.
 
@@
 Nine exceptional radiations plus high turnover explain species diversity
-in jawed vertebrates. P Natl Acad Sci USA 106: 13410--13414. Available:
-\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=2715324\$/backslash\$\&tool=pmcentrez\$/backslash\$\&rendertype=abstract}.
+in jawed vertebrates. P Natl Acad Sci USA 106: 13410--13414.
+doi:\href{http://dx.doi.org/10.1073/pnas.0811087106}{10.1073/pnas.0811087106}.
 
 14. Harmon LJ, Rabosky DL, FitzJohn RG, Brown JW (2011) turboMEDUSA:
-MEDUSA: modeling evolutionary diversification using stepwise AIC.
+MEDUSA: Modeling Evolutionary Diversification Using Stepwise AIC.
 
 15. FitzJohn RG (2012) Diversitree: comparative phylogenetic analyses of
-diversification in r. Method Ecol Evol 3: 1084--1092. Available:
-\url{http://doi.wiley.com/10.1111/j.2041-210X.2012.00234.x}.
-
-16. Lamas G (2004) Checklist: part 4A. hesperioidea-papilionoidea.
+diversification in R. Method Ecol Evol 3: 1084--1092.
+doi:\href{http://dx.doi.org/10.1111/j.2041-210X.2012.00234.x}{10.1111/j.2041-210X.2012.00234.x}.
+
+16. Lamas G (2004) Checklist: Part 4A. Hesperioidea-Papilionoidea.
 Gainesville.
@@
 Systematics and evolutionary history of butterflies in the ``Taygetis
-clade'' (nymphalidae: satyrinae: euptychiina): towards a better
-understanding of neotropical biogeography. Mol Phylogenet Evol 66:
-54--68. Available:
-\url{http://linkinghub.elsevier.com/retrieve/pii/S1055790312003624}.
-Accessed 29 March 2013.
+clade'' (Nymphalidae: Satyrinae: Euptychiina): Towards a better
+understanding of Neotropical biogeography. Mol Phylogenet Evol 66:
+54--68.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2012.09.005}{10.1016/j.ympev.2012.09.005}.
 
@@
 Phylogenetic relationships among genera of danaine butterflies
-(lepidoptera: nymphalidae) as implied by morphology and DNA sequences.
+(Lepidoptera: Nymphalidae) as implied by morphology and DNA sequences.
 Syst Biodivers 8: 75--89.
@@
 19. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
-(2010) Phylogenetics and biogeography of a spectacular old world
-radiation of butterflies: the subtribe mycalesina (lepidoptera:
-nymphalidae: satyrini). BMC Evol Biol 10: 172. Available:
+(2010) Phylogenetics and biogeography of a spectacular Old World
+radiation of butterflies: the subtribe Mycalesina (Lepidoptera:
+Nymphalidae: Satyrini). BMC Evol Biol 10: 172. Available:
 \url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=2898688\&tool=pmcentrez\&rendertype=abstract}.
@@
 20. Kodandaramaiah U, Peña C, Braby MF, Grund R, Müller CJ, et al.
-(2010) Phylogenetics of coenonymphina (nymphalidae: satyrinae) and the
+(2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
 problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
-Available:
-\url{http://linkinghub.elsevier.com/retrieve/pii/S1055790309003273}.
-Accessed 29 March 2013.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.08.012}{10.1016/j.ympev.2009.08.012}.
 
 21. Ortiz-Acevedo E, Willmott KR (2013) Molecular systematics of the
-butterfly tribe preponini (nymphalidae: charaxinae). Syst Entomol:
-no--o. Available: \url{http://doi.wiley.com/10.1111/syen.12008}.
+butterfly tribe Preponini (Nymphalidae: Charaxinae). Syst Entomol:
+no--o.
+doi:\href{http://dx.doi.org/10.1111/syen.12008}{10.1111/syen.12008}.
 
 22. De-Silva DL, Day JJ, Elias M, Willmott KR, Whinnett A, et al. (2010)
-Molecular phylogenetics of the neotropical butterfly subtribe oleriina
-(nymphalidae: danainae: ithomiini). Mol Phylogenet Evol 55: 1032--1041.
-Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/20079859}.
-
-23. Freitas AVL, Brown Jr. KS (2004) Phylogeny of the nymphalidae
-(lepidoptera). Syst Biol 53: 363--383. Available:
-\url{http://sysbio.oxfordjournals.org/cgi/doi/10.1080/10635150490445670}.
+Molecular phylogenetics of the Neotropical butterfly subtribe Oleriina
+(Nymphalidae: Danainae: Ithomiini). Mol Phylogenet Evol 55: 1032--1041.
+doi:\href{http://dx.doi.org/http://dx.doi.org/10.1016/j.ympev.2010.01.010}{http://dx.doi.org/10.1016/j.ympev.2010.01.010}.
+
+23. Freitas AVL, Brown Jr. KS (2004) Phylogeny of the Nymphalidae
+(Lepidoptera). Syst Biol 53: 363--383.
+doi:\href{http://dx.doi.org/10.1080/10635150490445670}{10.1080/10635150490445670}.
 
 24. Peña C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
-(2006) Higher level phylogeny of satyrinae butterflies (lepidoptera:
-nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
-Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/16563805}.
+(2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
+Nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2006.02.007}{10.1016/j.ympev.2006.02.007}.
 
 25. Penz CM (1999) Higher level phylogeny for the passion-vine
-butterflies (nymphalidae, heliconiinae) based on early stage and adult
-morphology. Zool J Linn Soc 127: 277--344. Available:
-\url{http://doi.wiley.com/10.1111/j.1096-3642.1999.tb00680.x}. Accessed
-03 September 2013.
+butterflies (Nymphalidae, Heliconiinae) based on early stage and adult
+morphology. Zool J Linn Soc 127: 277--344.
+doi:\href{http://dx.doi.org/10.1111/j.1096-3642.1999.tb00680.x}{10.1111/j.1096-3642.1999.tb00680.x}.
 
@@
 Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
-tribe acraeini (lepidoptera, nymphalidae, heliconiinae) and the
+tribe Acraeini (Lepidoptera, Nymphalidae, Heliconiinae) and the
 evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
-Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/18191590}.
-
-27. Peña C, Nylin S, Wahlberg N (2011) The radiation of satyrini
-butterflies (nymphalidae: satyrinae): a challenge for phylogenetic
-methods. Zool J Linn Soc 161: 64--87. Available:
-\url{http://doi.wiley.com/10.1111/j.1096-3642.2009.00627.x}. Accessed 29
-March 2013.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2007.11.024}{10.1016/j.ympev.2007.11.024}.
+
+27. Peña C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
+butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
+methods. Zool J Linn Soc 161: 64--87.
+doi:\href{http://dx.doi.org/10.1111/j.1096-3642.2009.00627.x}{10.1111/j.1096-3642.2009.00627.x}.
 
 28. Peña C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
-history of the butterfly subtribe euptychiina (lepidoptera, nymphalidae,
-satyrinae). Zool Scr 39: 243--258. Available:
-\url{http://blackwell-synergy.com/doi/abs/10.1111/j.1463-6409.2010.00421.x}.
+history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
+Satyrinae). Zool Scr 39: 243--258.
+doi:\href{http://dx.doi.org/10.1111/j.1463-6409.2010.00421.x}{10.1111/j.1463-6409.2010.00421.x}.
 
@@
 nymphalid butterflies. Biol J Linn Soc 33: 95--203.
+doi:\href{http://dx.doi.org/10.1111/j.1095-8312.1988.tb00446.x}{10.1111/j.1095-8312.1988.tb00446.x}.
 
@@
 31. Beccaloni G, Viloria Á, Hall S, Robinson G (2008) Catalogue of the
-hostplants of the neotropical butterflies. London: The Natural History
+hostplants of the Neotropical butterflies. London: The Natural History
 Museum.
@@
 32. Janzen DH, Hallwachs W (2009) Dynamic database for an inventory of
-the macrocaterpillar fauna, and its food plants and parasitoids, of area
-de conservacion guanacaste (aCG), northwestern costa rica. Available:
+the macrocaterpillar fauna, and its food plants and parasitoids, of Area
+de Conservacion Guanacaste (ACG), northwestern Costa Rica. Available:
 \url{http://janzen.sas.upenn.edu}. Accessed 17 April 2013.
 
-33. Core Team R (2013) R: a language and environment for statistical
-computing. Vienna, Austria: R Foundation for Statistical Computing.
+33. Core Team R (2013) R: A Language and Environment for Statistical
+Computing. Vienna, Austria: R Foundation for Statistical Computing.
 Available: \url{http://www.r-project.org}.
 
-34. Popescu A-A, Huber KT, Paradis E (2012) ape 3.0: new tools for
-distance-based phylogenetics and evolutionary analysis in r.
+34. Popescu A-A, Huber KT, Paradis E (2012) ape 3.0: New tools for
+distance-based phylogenetics and evolutionary analysis in R.
 Bioinformatics 28: 1536--1537.
+doi:\href{http://dx.doi.org/10.1093/bioinformatics/bts184}{10.1093/bioinformatics/bts184}.
 
@@
 investigating evolutionary radiations. Bioinformatics 24: 129--131.
-Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/18006550}.
+doi:\href{http://dx.doi.org/10.1093/bioinformatics/btm538}{10.1093/bioinformatics/btm538}.
 
@@
 character's effect on speciation and extinction. Syst Biol 56: 701--710.
-Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/17849325}.
+doi:\href{http://dx.doi.org/10.1080/10635150701607033}{10.1080/10635150701607033}.
 
 37. Willmott KR, Freitas AVL (2006) Higher-level phylogeny of the
-ithomiinae (lepidoptera: nymphalidae): classification, patterns of
+Ithomiinae (Lepidoptera: Nymphalidae): classification, patterns of
 larval hostplant colonization and diversification. Cladistics 22:
-297--368. Available:
-\url{http://doi.wiley.com/10.1111/j.1096-0031.2006.00108.x}.
-
-38. Scott JA (1986) The butterflies of north america. Stanford
+297--368.
+doi:\href{http://dx.doi.org/10.1111/j.1096-0031.2006.00108.x}{10.1111/j.1096-0031.2006.00108.x}.
+
+38. Scott JA (1986) The Butterflies of North America. Stanford
 University Press.
@@
 parameter estimation of the BiSSE method for analyzing species
-diversification. BMC Evol Biol 13: 38. Available:
-\url{http://www.biomedcentral.com/1471-2148/13/38}.
+diversification. BMC Evol Biol 13: 38.
+doi:\href{http://dx.doi.org/10.1186/1471-2148-13-38}{10.1186/1471-2148-13-38}.
 
 40. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
-bees cut leaves? new insights into the early evolution of bees. P Roy
-Soc B 278: 3593--3600. Available:
-\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3189370\$/backslash\$\&tool=pmcentrez\$/backslash\$\&rendertype=abstract}.
+bees cut leaves? New insights into the early evolution of bees. P Roy
+Soc B 278: 3593--3600.
+doi:\href{http://dx.doi.org/10.1098/rspb.2011.0365}{10.1098/rspb.2011.0365}.
 
 41. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
-clades of agaricales. P Roy Soc B 279: 2003--2011. Available:
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22171078}.
+clades of Agaricales. P Roy Soc B 279: 2003--2011.
+doi:\href{http://dx.doi.org/10.1098/rspb.2011.2428}{10.1098/rspb.2011.2428}.
 
 42. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
-interface. Ann Mo Bot Gard 74: 359--397.
+Interface. Ann Mo Bot Gard 74: 359--397.
 
 43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brandão KL, Whinnett A, et
-al. (2006) Phylogenetic relationships among the ithomiini (lepidoptera:
-nymphalidae) inferred from one mitochondrial and two nuclear gene
-regions. Syst Entomol 31: 288--301. Available:
-\url{http://doi.wiley.com/10.1111/j.1365-3113.2006.00321.x}.
+al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
+Nymphalidae) inferred from one mitochondrial and two nuclear gene
+regions. Syst Entomol 31: 288--301.
+doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2006.00321.x}{10.1111/j.1365-3113.2006.00321.x}.
 
@@
 46. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
-furanocoumarin-metabolizing cytochrome p450 monooxygenases in two
-papilionids: specificity and substrate encounter rate. P Natl Acad Sci
-USA 100: 14593--14598. Available:
-\url{http://www.pnas.org/content/100/suppl.2/14593.abstract}.
+furanocoumarin-metabolizing cytochrome P450 monooxygenases in two
+papilionids: Specificity and substrate encounter rate. P Natl Acad Sci
+USA 100: 14593--14598.
+doi:\href{http://dx.doi.org/10.1073/pnas.1934643100}{10.1073/pnas.1934643100}.
 
 47. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
-the angiosperms re-revisited. Am J Bot 97: 1296--1303. Available:
-\url{http://www.ncbi.nlm.nih.gov/pubmed/21616882}.
-
-48. Somoza R (1998) Updated nazca (farallon) - south america relative
-motions during the last 40 my: implications for mountain building in the
-central andean region. J S Am Earth Sci 11: 211--215.
+the angiosperms re-revisited. Am J Bot 97: 1296--1303.
+doi:\href{http://dx.doi.org/10.3732/ajb.0900346}{10.3732/ajb.0900346}.
+
+48. Somoza R (1998) Updated Nazca (Farallon) - South America relative
+motions during the last 40 My: implications for mountain building in the
+central Andean region. J S Am Earth Sci 11: 211--215.
+doi:\href{http://dx.doi.org/10.1016/S0895-9811(98)00012-1}{10.1016/S0895-9811(98)00012-1}.
 
 49. Hoorn C, Wesselingh FP, Steege H ter, Bermudez MA, Mora A, et al.
-(2010) Amazonia through time: andean uplift, climate change, landscape
-evolution, and biodiversity. Science 330: 927--931. Available:
-\url{http://www.ncbi.nlm.nih.gov/pubmed/21071659}.
-
-50. Olmstead RG (2013) Phylogeny and biogeography in solanaceae,
-verbenaceae and bignoniaceae: a comparison of continental and
+(2010) Amazonia through time: Andean uplift, climate change, landscape
+evolution, and biodiversity. Science 330: 927--931.
+doi:\href{http://dx.doi.org/10.1126/science.1194585}{10.1126/science.1194585}.
+
+50. Olmstead RG (2013) Phylogeny and biogeography in Solanaceae,
+Verbenaceae and Bignoniaceae: a comparison of continental and
 intercontinental diversification patterns. Bot J Linn Soc 171: 80--102.
-Available: \url{http://doi.wiley.com/10.1111/j.1095-8339.2012.01306.x}.
-
-51. Wahlberg N, Rubinoff D (2011) Vagility across vanessa (lepidoptera:
-nymphalidae): mobility in butterfly species does not inhibit the
+doi:\href{http://dx.doi.org/10.1111/j.1095-8339.2012.01306.x}{10.1111/j.1095-8339.2012.01306.x}.
+
+51. Wahlberg N, Rubinoff D (2011) Vagility across Vanessa (Lepidoptera:
+Nymphalidae): mobility in butterfly species does not inhibit the
 formation and persistence of isolated sister taxa. Syst Entomol 36:
-362--370. Available:
-\url{http://doi.wiley.com/10.1111/j.1365-3113.2010.00566.x}. Accessed 16
-April 2013.
-
-52. Müller CJ, Wahlberg N, Beheregaray LB (2010) `After africa': the
-evolutionary history and systematics of the genus charaxes ochsenheimer
-(lepidoptera: nymphalidae) in the indo-pacific region. Biol J Linn Soc
-100: 457--481. Available:
-\url{http://doi.wiley.com/10.1111/j.1095-8312.2010.01426.x}.
-
-53. Aduse-Poku K, Vingerhoedt E, Wahlberg N (2009) Out-of-africa again:
-a phylogenetic hypothesis of the genus charaxes (lepidoptera:
-nymphalidae) based on 5 gene regions. Mol Phylogenet Evol 53: 463--478.
+362--370.
+doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2010.00566.x}{10.1111/j.1365-3113.2010.00566.x}.
+
+52. Müller CJ, Wahlberg N, Beheregaray LB (2010) `After Africa': the
+evolutionary history and systematics of the genus Charaxes Ochsenheimer
+(Lepidoptera: Nymphalidae) in the Indo-Pacific region. Biol J Linn Soc
+100: 457--481.
+doi:\href{http://dx.doi.org/10.1111/j.1095-8312.2010.01426.x}{10.1111/j.1095-8312.2010.01426.x}.
+
+53. Aduse-Poku K, Vingerhoedt E, Wahlberg N (2009) Out-of-Africa again:
+a phylogenetic hypothesis of the genus Charaxes (Lepidoptera:
+Nymphalidae) based on 5 gene regions. Mol Phylogenet Evol 53: 463--478.
 doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.06.021}{10.1016/j.ympev.2009.06.021}.

### 2014-07-22 MS.tex
~n

@@
 data. The study was supported by a Kone Foundation grant (awarded to
-Niklas Wahlberg), Finland (C. Peña) and the Research Council of Norway
+Niklas Wahlberg), Finland (C. Pe\~na) and the Research Council of Norway
 (grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
@@
 
-2. Nyman T, Linder HP, Peña C, Malm T, Wahlberg N (2012) Climate-driven
+2. Nyman T, Linder HP, Pe\~na C, Malm T, Wahlberg N (2012) Climate-driven
 diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
@@
 
-7. Heikkilä M, Kaila L, Mutanen M, Peña C, Wahlberg N (2012) Cretaceous
+7. Heikkilä M, Kaila L, Mutanen M, Pe\~na C, Wahlberg N (2012) Cretaceous
 origin and repeated tertiary diversification of the redefined
@@
 
-10. Wahlberg N, Leneveu J, Kodandaramaiah U, Peña C, Nylin S, et al.
+10. Wahlberg N, Leneveu J, Kodandaramaiah U, Pe\~na C, Nylin S, et al.
 (2009) Nymphalid butterflies diversify following near demise at the
@@
 
-11. Peña C, Wahlberg N (2008) Prehistorical climate change increased
+11. Pe\~na C, Wahlberg N (2008) Prehistorical climate change increased
 diversification of a group of butterflies. Biol Lett 4: 274--278.
@@
 
-12. Ferrer-Paris JR, S\'anchez-Mercado A, Viloria ÁL, Donaldson J (2013)
+12. Ferrer-Paris JR, S\'anchez-Mercado A, Viloria \'AL, Donaldson J (2013)
 Congruence and Diversity of Butterfly-Host Plant Associations at Higher
@@
 
-17. Matos-Maraví PF, Peña C, Willmott KR, Freitas AV, Wahlberg N (2013)
+17. Matos-Marav\'i PF, Pe\~na C, Willmott KR, Freitas AV, Wahlberg N (2013)
 Systematics and evolutionary history of butterflies in the ``Taygetis
@@
 
-18. Brower AVZ, Wahlberg N, Ogawa JR, Boppré M, Vane-Wright RI (2010)
+18. Brower AVZ, Wahlberg N, Ogawa JR, Boppr\'e M, Vane-Wright RI (2010)
 Phylogenetic relationships among genera of danaine butterflies
@@
 
-20. Kodandaramaiah U, Peña C, Braby MF, Grund R, Müller CJ, et al.
+20. Kodandaramaiah U, Pe\~na C, Braby MF, Grund R, Müller CJ, et al.
 (2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
@@
 
-24. Peña C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
+24. Pe\~na C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
 (2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
@@
 
-27. Peña C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
+27. Pe\~na C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
 butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
@@
 
-28. Peña C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
+28. Pe\~na C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
 history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
@@
 
-31. Beccaloni G, Viloria Á, Hall S, Robinson G (2008) Catalogue of the
+31. Beccaloni G, Viloria \'A, Hall S, Robinson G (2008) Catalogue of the
 hostplants of the Neotropical butterflies. London: The Natural History

### 2014-07-22 MS.tex
ä
ã

@@
 
-7. Heikkilä M, Kaila L, Mutanen M, Pe\~na C, Wahlberg N (2012) Cretaceous
+7. Heikkil\"a M, Kaila L, Mutanen M, Pe\~na C, Wahlberg N (2012) Cretaceous
 origin and repeated tertiary diversification of the redefined
@@
 
-8. Elias M, Joron M, Willmott KR, Silva-Brandão KL, Kaiser V, et al.
+8. Elias M, Joron M, Willmott KR, Silva-Brand\~ao KL, Kaiser V, et al.
 (2009) Out of the Andes: Patterns of diversification in clearwing
@@
 
-26. Silva-Brandão KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
+26. Silva-Brand\~ao KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
 Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
@@
 
-43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brandão KL, Whinnett A, et
+43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand\~ao KL, Whinnett A, et
 al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:

### 2014-07-22 MS.tex
latex ms ready

@@
 6. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
-al. (2011) Order Lepidoptera. In:. Animal biodiversity: an outline of
+al. (2011) Order Lepidoptera. In: Animal biodiversity: an outline of
 higher-level classification and survey of taxonomic richness. Zootaxa.
@@
 Taxonomic Levels. PloS one 8: e63570.
-doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.
+doi:\href{http://dx.doi.org/10.1371/journal.pone.0063570}{10.1371/journal.pone.0063570}.
 
@@
 
-19. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
+19. Kodandaramaiah U, Lees DC, M\"uller CJ, Torres E, Karanth KP, et al.
 (2010) Phylogenetics and biogeography of a spectacular Old World
 radiation of butterflies: the subtribe Mycalesina (Lepidoptera:
-Nymphalidae: Satyrini). BMC Evol Biol 10: 172. Available:
-\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=2898688\&tool=pmcentrez\&rendertype=abstract}.
-
-20. Kodandaramaiah U, Pe\~na C, Braby MF, Grund R, Müller CJ, et al.
+Nymphalidae: Satyrini). BMC Evol Biol 10: 172. 
+doi:\href{http://dx.doi.org/10.1186/1471-2148-10-172}{10.1186/1471-2148-10-172}.
+
+20. Kodandaramaiah U, Pe\~na C, Braby MF, Grund R, M\"uller CJ, et al.
 (2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
@@
 21. Ortiz-Acevedo E, Willmott KR (2013) Molecular systematics of the
-butterfly tribe Preponini (Nymphalidae: Charaxinae). Syst Entomol:
-no--o.
+butterfly tribe Preponini (Nymphalidae: Charaxinae). Syst Entomol 38:
+440--449.
 doi:\href{http://dx.doi.org/10.1111/syen.12008}{10.1111/syen.12008}.
@@
 (Nymphalidae: Danainae: Ithomiini). Mol Phylogenet Evol 55: 1032--1041.
-doi:\href{http://dx.doi.org/http://dx.doi.org/10.1016/j.ympev.2010.01.010}{http://dx.doi.org/10.1016/j.ympev.2010.01.010}.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2010.01.010}{10.1016/j.ympev.2010.01.010}.
 
@@
 Syst Biol 58: 595--611. Available:
-\url{http://www.ncbi.nlm.nih.gov/pubmed/20525612}.
+doi:\href{http://dx.doi.org/10.1093/sysbio/syp067}{10.1093/sysbio/syp067}.
 
@@
 
-52. Müller CJ, Wahlberg N, Beheregaray LB (2010) `After Africa': the
+52. M\"uller CJ, Wahlberg N, Beheregaray LB (2010) `After Africa': the
 evolutionary history and systematics of the genus Charaxes Ochsenheimer

### 2014-07-23 MS.tex
italics for names in the ref list

@@
 
-51. Wahlberg N, Rubinoff D (2011) Vagility across Vanessa (Lepidoptera:
+51. Wahlberg N, Rubinoff D (2011) Vagility across \textit{Vanessa} (Lepidoptera:
 Nymphalidae): mobility in butterfly species does not inhibit the
@@
 52. M\"uller CJ, Wahlberg N, Beheregaray LB (2010) `After Africa': the
-evolutionary history and systematics of the genus Charaxes Ochsenheimer
+evolutionary history and systematics of the genus \textit{Charaxes} Ochsenheimer
 (Lepidoptera: Nymphalidae) in the Indo-Pacific region. Biol J Linn Soc
@@
 53. Aduse-Poku K, Vingerhoedt E, Wahlberg N (2009) Out-of-Africa again:
-a phylogenetic hypothesis of the genus Charaxes (Lepidoptera:
+a phylogenetic hypothesis of the genus \textit{Charaxes} (Lepidoptera:
 Nymphalidae) based on 5 gene regions. Mol Phylogenet Evol 53: 463--478.

### 2014-07-23 MS.tex
removing table legends from MS

@@
 
-\section*{Tables}
-
-\texttt{Table 1.} Significant net diversification rate shifts found in
-the MEDUSA analysis of the Nymphalid maximum clade credibility tree.
-Split.Node = node number, Model = preferred diversification model by
-MEDUSA, r = net diversification rate, LnLik.part = log likelihood value.
-
-\texttt{Table 2.} Differences in rates estimated by MEDUSA on the MCC
-tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
-from the posterior distribution.
-
-\texttt{Table 3.} Likelihood ratio test between the model of increased
-diversification of nymphalids feeding on Solanaceae against a model
-forcing equal speciation rates (no effect on diversification).
-
 \section*{Supp. mat.}

### 2014-07-23 MS.tex
legend fig 1

@@
 \texttt{Figure 1.} Results of the MEDUSA analysis run on the maximum
-clade credibility tree from Reference 10. Rate shifts were estimated for
-the following nodes (besides the background rate): 2) Limenitidinae +
-Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
-part, 6) Satyrini, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
-part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
-\emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
-Satyrina, 14) Mycalesinaa, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
-\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae.
+clade credibility tree from Reference 10. Estimated rate shifts in Table 1.
 

### 2014-07-23 MS.tex
legend figs 1,2,3

@@
 
-\texttt{Figure 1.} Results of the MEDUSA analysis run on the maximum
-clade credibility tree from Reference 10. Estimated rate shifts in Table 1.
-
-\texttt{Figure 2.} Boxplot of the range of diversification values for
-tips or clades estimated by MEDUSA on the 1000 random trees from the
-posterior distribution of the Nymphalidae phylogeny. The tips or clades
-shown are those present on the maximum clade credibility tree from
-Reference 10.
-
-\texttt{Figure 3.} Results of the MultiMEDUSA analysis on 1000 randomly
-sampled trees from the posterior distribution of the Nymphalidae
-phylogeny. Bars show the probability for nodes of being recovered as
-significant increases or decreases in net diversification rates by
-MultiMEDUSA.
+\texttt{Figure 1.} MEDUSA analysis run on the MCT 
+from Reference 10. Estimated rate shifts in Table 1.
+
+\texttt{Figure 2.} Range of diversification values for
+tips or clades estimated by MEDUSA on the 1000 random trees.
+
+\texttt{Figure 3.} Results of the MultiMEDUSA analysis showing
+the probability for nodes of being recovered as diversification events.
 

### 2014-07-23 MS.tex
legend fig 4

@@
 
-\texttt{Figure 4.} BiSSE analysis of diversification of nymphalids due
-to feeding on Solanaceae hostplants. Speciation and net diversification
-rates are significantly higher in Solanaceae feeders (\$lambda\$1, r1).
+\texttt{Figure 4.} BiSSE analysis of diversification of Solanaceae feeding 
+nymphalids shows higher speciation and net diversification
+rates in Solanaceae feeders.
 

### 2014-07-23 ancillary/supp_mat_S06_run_multimedusa.R
renaming supp mat

@@ -1,0 +1,147 @@
+library("MEDUSA")
+library("multicore")
+
+mct <- read.nexus("ancillary/supp_mat_09_mcc_tree_from_random_sample_of_1000.tree")
+phy <- read.nexus("ancillary/supp_mat_05_1000_random_trees_no_outgroups.nex")
+
+richness <- read.csv("ancillary/supp_mat_03_richness.csv")
+
+res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=27, mc=TRUE)
+
+
+pdf(file="ancillary/supp_mat_08_multimedusa_result01.pdf", width=9, height=22)
+summ <- multiMedusaSummary(res, conTree=mct, cex=0.3)
+dev.off()
+
+source("code/alt_plotMultimedusa.R")
+pdf(file="ancillary/supp_mat_14_multimedusa_result_rates_nodes.pdf", width=9, height=22)
+plotMultiMedusa_alt(summ)
+dev.off()
+
+save(res, file="ancillary/supp_mat_07_multimedusa_on_1000_trees.txt", ascii=TRUE)
+
+
+# plot the summary statistics usig a boxplot Fig. 02
+# save as supp. mat. 15
+library("ggplot2")
+load("ancillary/supp_mat_07_multimedusa_on_1000_trees.txt")
+summ <- multiMedusaSummary(res, conTree=mct, cex=0.3)
+data <- summ$shift.summary
+data <- as.data.frame(data)
+
+names <- c("Charaxes", "Ypthima", "Ithomiini", "Callicore + Diaethria", 
+           "Danaini", "Limenitidinae + Heliconiinae", "Caeruleuptychia + Magneuptychia", 
+           "Pedaliodes", "Taenaris", "Dryas + Dryadula", "Coenonympha", 
+           "Anaeomorpha + Hypna", "Mycalesina", "Pseudergolinae", 
+           "Phyciodina", "Satyrina", "Coenonymphina", 
+           "Satyrini")
+
+# delete rows that are not too important
+data <- data[c(1:33), 1:7]
+data <- data[-32,]
+data <- data[-31,]
+data <- data[-30,]
+data <- data[-29,]
+data <- data[-28,]
+data <- data[-25,]
+data <- data[-24,]
+data <- data[-22,]
+data <- data[-21,]
+data <- data[-17,]
+data <- data[-16,]
+data <- data[-15,]
+data <- data[-14,]
+data <- data[-13,]
+data <- data[-11,]
+
[…]

### 2014-07-23 MS.tex
sup inf legends

@@
 
-\section*{Supp. mat.}
-
-\textbf{\texttt{supp\_mat\_01\_genus.nex}}: MCC Nymphalidae tree from
-Reference 10.
-
-\textbf{\texttt{supp\_mat\_02\_run\_medusa\_on\_mct.R}}: Run MEDUSA on
-MCC tree from Reference 10.
-
-\textbf{\texttt{supp\_mat\_03\_richness.nex}}: Species richness for
-lineages in Nymphalidae.
-
-\textbf{\texttt{supp\_mat\_04\_MEDUSA\_results.txt}}: Results of running
-MEDUSA on the MCC tree.
-
-\textbf{\texttt{supp\_mat\_05\_1000\_random\_trees\_no\_outgroups.nex}}:
-1000 random trees from Reference 10.
-
-\textbf{\texttt{supp\_mat\_06\_run\_multimedusa.R}}: Run MultiMEDUSA on
-1000 random trees from Reference 10.
-
-\textbf{\texttt{supp\_mat\_07\_multimedusa\_on\_1000\_trees.txt}}: Raw
-results from the MultiMEDUSA run on the random sample of trees from the
-posterior distribution.
-
-\textbf{\texttt{supp\_mat\_08\_multimedusa\_result01.pdf}}: Summary of
-MulitMEDUSA analysis on the 1000 trees from the posterior distribution.
-
-\textbf{\texttt{supp\_mat\_09\_mcc\_tree\_from\_random\_sample\_of\_1000.tree}}:
-MCC tree from the 1000 random trees selected from the posterior
-distribution.
-
-\textbf{\texttt{supp\_mat\_10\_run\_medusa\_on\_mct\_from\_1000\_trees.R}}:
-Run MEDUSA on MCC tree from the 1000 random trees selected from the
-posterior distribution.
-
-\textbf{\texttt{supp\_mat\_11\_results\_of\_MEDUSA\_run\_on\_mcc\_from\_1000\_trees.txt}}:
-Summary results from a MEDUSA run on the MCC tree from the 1000 random
-trees selected from the posterior distribution.
-
-\textbf{\texttt{supp\_mat\_12\_MEDUSA\_mcc\_tree\_from\_random\_sample\_of\_1000\_trees.pdf}}:
-Figure for MEDUSA run on MCC tree from random 1000 trees.
-
-\textbf{\texttt{supp\_mat\_13\_results\_of\_Multi-MEDUSA\_run\_on\_1000\_trees.txt}}:
-Summary results from a MultiMEDUSA run on the 1000 random trees selected
-from the posterior distribution.
-
-\textbf{\texttt{supp\_mat\_14\_multimedusa\_result\_prob\_nodes.pdf}}:
-Probability of nodes with estimated rates from a MultiMEDUSA run on the
-1000 random trees selected from the posterior distribution.
-
-\textbf{\texttt{supp\_mat\_15\_host\_plant\_data.csv}}: Hostplants of
-Nymphalidae butterflies recorded from the literature.
-
-\textbf{\texttt{supp\_mat\_16\_host\_plant\_data\_references.csv}}:
-References for hostplants data.
-
-\textbf{\texttt{supp\_mat\_17\_hostplant\_states.csv}}: Data matrix with
-character states for hosplant use.
-
-\textbf{\texttt{supp\_mat\_18\_bisse\_source.R}}: R code for running the
-BiSSE analysis.
-
-\textbf{\texttt{supp\_mat\_19\_bisse\_mcmc\_run.csv}}: Raw results for
-the BiSSE analysis.
-
-\textbf{\texttt{supp\_mat\_20\_combined\_bisse.csv}}: Raw results for
-the combined BiSSE analysis.
-
-\texttt{Figure S1.} Boxplot of speciation (lambda) and extinction (mu)
-values for the 95\% credibility intervals of values estimated by BiSSE
-analysis of diversification due to feeding on Solanaceae plants.
-
-\texttt{Figure S2.} BiSSE analysis of diversification of nymphalids due
-to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
-\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
-recovered, speciation and net diversification rates are significantly
-higher for Solanaceae feeders (\$lambda\$1, r1).
-
-\texttt{Figure S3.} Net diversification rates of nymphalids feeding on
-Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
-the 1000 trees from the posterior distribution.
-
-\texttt{Figure S4.} Boxplot of speciation and extinction values for the
-95\% credibility intervals of values estimated by BiSSE analysis of
-diversification due to feeding on Solanaceae plants on the combined
-post-burnin runs on 1000 trees from the posterior distribution.
-
-\texttt{Figure S5.} BiSSE analysis of diversification of nymphalids due
-to feeding on Apocynaceae hostplants. Speciation and net diversification
-rates are similar.
-
 \section*{References}
@@
 
+\section*{Supp. mat.}
+
+\textbf{\texttt{Supporting_Information_S1.nex}}: \textbf{Genus}. MCC Nymphalidae tree from
+Reference 10.
+
+\textbf{\texttt{Supporting_Information_S2.R}}: \textbf{Run_medusa_on_mct}. Run MEDUSA on
+MCC tree from Reference 10.
+
+\textbf{\texttt{Supporting\_Information\_S3.csv}}: Species richness for
+lineages in Nymphalidae.
+
+\textbf{\texttt{Supporting\_Information\_S4.txt}}: Results of running
+MEDUSA on the MCC tree.
+
+\textbf{\texttt{Supporting\_Information\_S5.zip}}: 1000 random trees from Reference 10.
+
+\textbf{\texttt{Supporting\_Information\_S6.R}}: Run MultiMEDUSA on
+1000 random trees from Reference 10.
+
+\textbf{\texttt{Supporting\_Information\_S7.zip}}: Raw
+results from the MultiMEDUSA run on the random sample of trees from the
+posterior distribution.
+
+\textbf{\texttt{Supporting\_Information\_S8.pdf}}: Summary of
+MulitMEDUSA analysis on the 1000 trees from the posterior distribution.
+
+\textbf{\texttt{Supporting\_Information\_S9.tree}}:
+MCC tree from the 1000 random trees selected from the posterior
+distribution.
+
+\textbf{\texttt{Supporting_Information_S10.R}}:
+Run MEDUSA on MCC tree from the 1000 random trees selected from the
+posterior distribution.
+
+\textbf{\texttt{Supporting_Information_S11.txt}}:
+Summary results from a MEDUSA run on the MCC tree from the 1000 random
+trees selected from the posterior distribution.
+
+\textbf{\texttt{Supporting_Information_S12.pdf}}:
+Figure for MEDUSA run on MCC tree from random 1000 trees.
+
+\textbf{\texttt{Supporting_Information_S13.txt}}:
+Summary results from a MultiMEDUSA run on the 1000 random trees selected
+from the posterior distribution.
+
+\textbf{\texttt{Supporting_Information_S14.pdf}}:
+Probability of nodes with estimated rates from a MultiMEDUSA run on the
+1000 random trees selected from the posterior distribution.
+
+\textbf{\texttt{Supporting_Information_S15.csv}}: Hostplants of
+Nymphalidae butterflies recorded from the literature.
+
+\textbf{\texttt{Supporting_Information_S16.csv}}:
+References for hostplants data.
+
+\textbf{\texttt{Supporting_Information_S17.csv}}: Data matrix with
+character states for hosplant use.
+
+\textbf{\texttt{Supporting_Information_S18.R}}: R code for running the
+BiSSE analysis.
+
+\textbf{\texttt{Supporting_Information_S19.csv}}: Raw results for
+the BiSSE analysis.
+
+\textbf{\texttt{Supporting_Information_S20.csv}}: Raw results for
+the combined BiSSE analysis.
+
 \end{document}

### 2014-07-24 ancillary/Supporting_Information_S6_run_multimedusa.R
start removing big files

@@ -1,147 +1,0 @@
-library("MEDUSA")
-library("multicore")
-
-mct <- read.nexus("ancillary/supp_mat_09_mcc_tree_from_random_sample_of_1000.tree")
-phy <- read.nexus("ancillary/supp_mat_05_1000_random_trees_no_outgroups.nex")
-
-richness <- read.csv("ancillary/supp_mat_03_richness.csv")
-
-res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=27, mc=TRUE)
-
-
-pdf(file="ancillary/supp_mat_08_multimedusa_result01.pdf", width=9, height=22)
-summ <- multiMedusaSummary(res, conTree=mct, cex=0.3)
-dev.off()
-
-source("code/alt_plotMultimedusa.R")
-pdf(file="ancillary/supp_mat_14_multimedusa_result_rates_nodes.pdf", width=9, height=22)
-plotMultiMedusa_alt(summ)
-dev.off()
-
-save(res, file="ancillary/supp_mat_07_multimedusa_on_1000_trees.txt", ascii=TRUE)
-
-
-# plot the summary statistics usig a boxplot Fig. 02
-# save as supp. mat. 15
-library("ggplot2")
-load("ancillary/supp_mat_07_multimedusa_on_1000_trees.txt")
-summ <- multiMedusaSummary(res, conTree=mct, cex=0.3)
-data <- summ$shift.summary
-data <- as.data.frame(data)
-
-names <- c("Charaxes", "Ypthima", "Ithomiini", "Callicore + Diaethria", 
-           "Danaini", "Limenitidinae + Heliconiinae", "Caeruleuptychia + Magneuptychia", 
-           "Pedaliodes", "Taenaris", "Dryas + Dryadula", "Coenonympha", 
-           "Anaeomorpha + Hypna", "Mycalesina", "Pseudergolinae", 
-           "Phyciodina", "Satyrina", "Coenonymphina", 
-           "Satyrini")
-
-# delete rows that are not too important
-data <- data[c(1:33), 1:7]
-data <- data[-32,]
-data <- data[-31,]
-data <- data[-30,]
-data <- data[-29,]
-data <- data[-28,]
-data <- data[-25,]
-data <- data[-24,]
-data <- data[-22,]
-data <- data[-21,]
-data <- data[-17,]
-data <- data[-16,]
-data <- data[-15,]
-data <- data[-14,]
-data <- data[-13,]
-data <- data[-11,]
-
[…]

### 2014-07-24 ancillary/supp_mat_18_bisse_source.R

@@ -1,158 +1,0 @@
-library(diversitree);
-library(RSvgDevice);
-
-add.alpha <- function(col, alpha=.5) {
-    tmp <- col2rgb(col)/255
-    rgb(tmp[1,], tmp[2,], tmp[3,], alpha=alpha)
-}
-
-phy <- read.nexus("ancillary/supp_mat_01_genus.nex");
-tips <- c("Achlyodes", "Graphium", "Parnassius", "Baronia", "Troides", "Papilio1", "Papilio2", "Pieris", "Aporia", "Styx", "Hamearis", "Euselasia", "Nymphidium", "Emesis", "Crocozona", "Riodina", "Amarynthis", "Baliochila", "Poritia", "Miletus", "Liphyra", "Lycaena", "Celastrina", "Thecla", "Lucia", "Curetis", "Eurema", "Colias", "Leptidea", "Pseudopontia", "Libyt");
-phy  <- drop.tip(phy, tips)
-new.phy <- multi2di(phy);
-phy <- new.phy;
-
-# Process data on using **Solanaceae** as hostplant
-data <- read.csv("ancillary/supp_mat_09_states.csv", sep=",", header=T, row.names=1);
-data <- data[order(order(phy$tip.label)),]
-
-# Column 2 is the data for **Solanaceae**
-# Column 1 is for Apocynaceae (data not used in this analysis)
-data.v <- data[,2]
-
-names(data.v) <- row.names(data)
-
-
-
-phy$tip.state<-data.v
-
-##################
-# get sampling values. Propotion of extant species in our tree that have state 0 and state 1
-# in relation with the total 
-richness <- read.csv("ancillary/supp_mat_03_richness.csv")
-richness <- richness[order(order(phy$tip.label)),]
-
-richness_cero <- c();
-richness_one <- c();
-for(i in 1:length(names(data.v))) {
-  x <- match(names(data.v)[i], richness$taxon)
-  x.state <- data.v[i]
-  
-  if( is.na(x.state) == FALSE) {
-    if( x.state == 0) {
-      richness_cero <- c(richness_cero, richness$n.taxa[i])
-    }
-    if( x.state == 1) {
-      richness_one <- c(richness_one, richness$n.taxa[i])
-    }
-  }
-}
-
-total_richness_state_cero <- sum(richness_cero);
-total_richness_state_one <- sum(richness_one);
-
[…]

### 2014-07-24 ancillary/supp_mat_S02_run_medusa_on_mct.R

@@ -1,24 +1,0 @@
-# run MEDUSA 0.93-4-33
-
-library("MEDUSA")
-library("multicore")
-phy <- read.nexus("ancillary/supp_mat_01_genus.nex");
-phy <- ladderize(phy)
-
-# we have to remove the tips
-tips <- c("Achlyodes", "Graphium", "Parnassius", "Baronia", "Troides", "Papilio1", "Papilio2", "Pieris", "Aporia", "Styx", "Hamearis", "Euselasia", "Nymphidium", "Emesis", "Crocozona", "Riodina", "Amarynthis", "Baliochila", "Poritia", "Miletus", "Liphyra", "Lycaena", "Celastrina", "Thecla", "Lucia", "Curetis", "Eurema", "Colias", "Leptidea", "Pseudopontia", "Libyt")
-phy <- drop.tip(phy, tip=tips)
-
-
-richness <- read.csv("ancillary/supp_mat_03_richness.csv")
-
-res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25, mc=TRUE)
-
-
-pdf(file="output/medusa_on_mct.pdf", width=9, height=19)
-summ <- medusaSummary(res, cex=0.3)
-dev.off()
-
-
-save(res, file="ancillary/supp_mat_04_MEDUSA_results.txt", ascii=TRUE)
-

### 2014-07-24 ancillary/supp_mat_S10_run_medusa_on_mct_from_1000_trees.R

@@ -1,17 +1,0 @@
-# run MEDUSA 0.93-4-33
-
-library("MEDUSA")
-library("multicore")
-phy <- read.nexus("ancillary/supp_mat_09_mcc_tree_from_random_sample_of_1000.tree");
-phy <- ladderize(phy)
-
-
-
-richness <- read.csv("ancillary/supp_mat_03_richness.csv")
-
-res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25, mc=TRUE)
-
-
-pdf(file="ancillary/supp_mat_12_MEDUSA_mcc_tree_from_random_sample_of_1000_trees.pdf", width=9, height=23)
-summ <- medusaSummary(res, cex=0.3)
-dev.off()

### 2014-07-24 MS.tex
legends for supp figures

@@
 
-\section*{Supp. mat.}
-
-\textbf{\texttt{Supporting_Information_S1.nex}}: \textbf{Genus}. MCC Nymphalidae tree from
+\section*{Supporting Information}
+\texttt{Figure S1.} Boxplot of speciation (lambda) and extinction (mu)
+values for the 95\% credibility intervals of values estimated by BiSSE
+analysis of diversification due to feeding on Solanaceae plants.
+
+\texttt{Figure S2.} BiSSE analysis of diversification of nymphalids due
+to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
+\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
+recovered, speciation and net diversification rates are significantly
+higher for Solanaceae feeders (\$lambda\$1, r1).
+
+\texttt{Figure S3.} Net diversification rates of nymphalids feeding on
+Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
+the 1000 trees from the posterior distribution.
+
+\texttt{Figure S4.} Boxplot of speciation and extinction values for the
+95\% credibility intervals of values estimated by BiSSE analysis of
+diversification due to feeding on Solanaceae plants on the combined
+post-burnin runs on 1000 trees from the posterior distribution.
+
+\texttt{Figure S5.} BiSSE analysis of diversification of nymphalids due
+to feeding on Apocynaceae hostplants. Speciation and net diversification
+rates are similar.
+
+\textbf{\texttt{Supporting\_Information\_S1.nex}}: \textbf{Genus}. MCC Nymphalidae tree from
 Reference 10.
 
-\textbf{\texttt{Supporting_Information_S2.R}}: \textbf{Run_medusa_on_mct}. Run MEDUSA on
+\textbf{\texttt{Supporting\_Information\_S2.R}}: \textbf{Run medusa on mct}. Run MEDUSA on
 MCC tree from Reference 10.
@@
 
-\textbf{\texttt{Supporting_Information_S10.R}}:
+\textbf{\texttt{Supporting\_Information\_S10.R}}:
 Run MEDUSA on MCC tree from the 1000 random trees selected from the
@@
 
-\textbf{\texttt{Supporting_Information_S11.txt}}:
+\textbf{\texttt{Supporting\_Information\_S11.txt}}:
 Summary results from a MEDUSA run on the MCC tree from the 1000 random
@@
 
-\textbf{\texttt{Supporting_Information_S12.pdf}}:
+\textbf{\texttt{Supporting\_Information\_S12.pdf}}:
 Figure for MEDUSA run on MCC tree from random 1000 trees.
 
-\textbf{\texttt{Supporting_Information_S13.txt}}:
+\textbf{\texttt{Supporting\_Information\_S13.txt}}:
 Summary results from a MultiMEDUSA run on the 1000 random trees selected
@@
 
-\textbf{\texttt{Supporting_Information_S14.pdf}}:
+\textbf{\texttt{Supporting\_Information\_S14.pdf}}:
 Probability of nodes with estimated rates from a MultiMEDUSA run on the
@@
 
-\textbf{\texttt{Supporting_Information_S15.csv}}: Hostplants of
+\textbf{\texttt{Supporting\_Information\_S15.csv}}: Hostplants of
 Nymphalidae butterflies recorded from the literature.
 
-\textbf{\texttt{Supporting_Information_S16.csv}}:
+\textbf{\texttt{Supporting\_Information\_S16.csv}}:
 References for hostplants data.
 
-\textbf{\texttt{Supporting_Information_S17.csv}}: Data matrix with
+\textbf{\texttt{Supporting\_Information\_S17.csv}}: Data matrix with
 character states for hosplant use.
 
-\textbf{\texttt{Supporting_Information_S18.R}}: R code for running the
+\textbf{\texttt{Supporting\_Information\_S18.R}}: R code for running the
 BiSSE analysis.
 
-\textbf{\texttt{Supporting_Information_S19.csv}}: Raw results for
+\textbf{\texttt{Supporting\_Information\_S19.csv}}: Raw results for
 the BiSSE analysis.
 
-\textbf{\texttt{Supporting_Information_S20.csv}}: Raw results for
+\textbf{\texttt{Supporting\_Information\_S20.csv}}: Raw results for
 the combined BiSSE analysis.

### 2014-07-24 ancillary/Supporting_Information_S10.R
ready to submit

@@ -1,0 +1,17 @@
+# run MEDUSA 0.93-4-33
+
+library("MEDUSA")
+library("multicore")
+phy <- read.nexus("ancillary/supp_mat_09_mcc_tree_from_random_sample_of_1000.tree");
+phy <- ladderize(phy)
+
+
+
+richness <- read.csv("ancillary/supp_mat_03_richness.csv")
+
+res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25, mc=TRUE)
+
+
+pdf(file="ancillary/supp_mat_12_MEDUSA_mcc_tree_from_random_sample_of_1000_trees.pdf", width=9, height=23)
+summ <- medusaSummary(res, cex=0.3)
+dev.off()

### 2014-07-24 ancillary/Supporting_Information_S2.R

@@ -1,0 +1,24 @@
+# run MEDUSA 0.93-4-33
+
+library("MEDUSA")
+library("multicore")
+phy <- read.nexus("ancillary/supp_mat_01_genus.nex");
+phy <- ladderize(phy)
+
+# we have to remove the tips
+tips <- c("Achlyodes", "Graphium", "Parnassius", "Baronia", "Troides", "Papilio1", "Papilio2", "Pieris", "Aporia", "Styx", "Hamearis", "Euselasia", "Nymphidium", "Emesis", "Crocozona", "Riodina", "Amarynthis", "Baliochila", "Poritia", "Miletus", "Liphyra", "Lycaena", "Celastrina", "Thecla", "Lucia", "Curetis", "Eurema", "Colias", "Leptidea", "Pseudopontia", "Libyt")
+phy <- drop.tip(phy, tip=tips)
+
+
+richness <- read.csv("ancillary/supp_mat_03_richness.csv")
+
+res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25, mc=TRUE)
+
+
+pdf(file="output/medusa_on_mct.pdf", width=9, height=19)
+summ <- medusaSummary(res, cex=0.3)
+dev.off()
+
+
+save(res, file="ancillary/supp_mat_04_MEDUSA_results.txt", ascii=TRUE)
+

### 2014-10-06 MS_revised.tex
start revision

--- 2014-07-24 MS.tex
@@
 
-\textbf{Keywords:} diversification analysis, MEDUSA, BiSSE, speciation
-rate, insect-hostplant dynamics
+\textbf{Keywords:} diversification analysis; MEDUSA; BiSSE; speciation
+rate; insect-hostplant dynamics
 
@@
 Jessica Slove Davidson and Niklas Janz for access to their hostplant
-data. The study was supported by a Kone Foundation grant (awarded to
-Niklas Wahlberg), Finland (C. Pe\~na) and the Research Council of Norway
-(grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
+data. We acknowledge CSC--IT Center for
 Science Ltd. (Finland) for the allocation of computational resources.

### 2014-10-08 response_to_reviewers.md
punto 2

# Response to reviewers

## Reviewer 1

**1. Reviewer's Comment:**

Introduction
1. Second paragraph, second sentence: The fact that the Nymphalids are the “largest family” of butterflies conveys an exaggerated importance of their diversity based on idiosyncrasy of Linnean taxonomy (the sister group to Nymphalidae, the clade that includes riodinids and lycaenids, has over 6500 species). Consider revising the sentence, omitting superlatives (“biggest”, “largest”, “most”, etc.).

**1. Response:**

> We have rewritten the sentence to:
> "Nymphalidae contains around 6000 species [6], and several members are
considered model organisms in evolutionary biology [7–9]."

**2. Reviewer's Comment:**

The authors should mention that the ideal approach would accommodate uncertainty by sampling from a distribution of trees which reflected the variation in confidence in various parts of the tree (such as a Bayesian posterior distribution or a bootstrap sample). See Holder,M. and Lewis,P.O. (2003) Phylogeny estimation: traditional and Bayesian approaches. Nat. Rev. Genet., 4, 275–284

> We have specified that the sample of trees was taken from the Bayesian run in
Wahlberg et al., 2009.

### 2014-10-08 response_to_reviewers.tex
punto 2

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in,a4paper]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\usepackage{color}
\date{}

\begin{document}

\section{Response to reviewers}\label{response-to-reviewers}

\subsection{Reviewer 1}\label{reviewer-1}

\textbf{1. Reviewer's Comment:}

Introduction 1. Second paragraph, second sentence: The fact that the
Nymphalids are the ``largest family'' of butterflies conveys an
exaggerated importance of their diversity based on idiosyncrasy of
Linnean taxonomy (the sister group to Nymphalidae, the clade that
includes riodinids and lycaenids, has over 6500 species). Consider
revising the sentence, omitting superlatives (``biggest'', ``largest'',
``most'', etc.).

\textbf{1. Response:}

\begin{quote}
\color{blue}
We have rewritten the sentence to: ``Nymphalidae contains around 6000
species {[}6{]}, and several members are considered model organisms in
evolutionary biology {[}7--9{]}.''
\end{quote}

\textbf{2. Reviewer's Comment:}

The authors should mention that the ideal approach would accommodate
uncertainty by sampling from a distribution of trees which reflected the
variation in confidence in various parts of the tree (such as a Bayesian
posterior distribution or a bootstrap sample). See Holder,M. and
Lewis,P.O. (2003) Phylogeny estimation: traditional and Bayesian
approaches. Nat. Rev.~Genet., 4, 275--284

\end{document}

### 2014-10-08 response_to_reviewers.md
punto 3

@@
 
+**2. Response:**
+
 > We have specified that the sample of trees was taken from the Bayesian run in
 Wahlberg et al., 2009.
+
+**3. Reviewer's Comment:**
+
+Middle of paragraph (and throughout the manuscript): revise the tree sampling to “...therefore randomly sampled 1000 trees from the posterior distribution”; in the remainder of the manuscript, refer to these as the “1000 sampled trees”, not “1000 random trees”, which can take on a very different meaning.
+
+**3. Response:**
+
+> We have corrected some sentences in the manuscript to specify that we are
+referring to the same set of 1000 trees.

### 2014-10-08 response_to_reviewers.md
punto 4
punto 5

@@
 referring to the same set of 1000 trees.
+
+**4. Reviewer's Comment:**
+
+Define what 'MuSSE' refers to (Multiple State Speciation and Extinction?).
+
+**4. Response:**
+
+> We have included the definition of MuSSE in the manuscript.
+
+**5. Reviewer's Comment:**
+
+A reference for the statement “most of Nymphalidae butterflies are restricted to use one plant family as hostplant” is warranted.
+
+**5. Response:**
+
+> We have included the numbers of single-family feeders based on calculations
+of our own data on hostplant use across Nymphalidae genera. The text has been
+modified accordingly:
+
+> "Because most of Nymphalidae butterflies are restricted to use one plant
+family as hostplant (232 genera use only one plant family versus 176 that use
+        more than one; based on our data in supp. mat. 15), the ..."

### 2014-10-09 response_to_reviewers.md
punto 6

@@
         more than one; based on our data in supp. mat. 15), the ..."
+
+**6. Reviewer's Comment:**
+
+At the end of this paragraph, the authors indicated they tested the effect of character coding on their analyses.
+A more explicit description of the test would be useful, along with predictions of how results would be effected (i.e.
+how would a large effect vs.
+no effect of coding Hypanartia and Vanessa as Solanaceae feeders manifest in the results?).
+
+**6. Response:**
+
+> We included the following sentence to clarify the issue: **"We expected that
+coding these two genera as either absence or presence would not have a
+significant effect on our results."**

### 2014-10-10 response_to_reviewers.md
puntos 7,8,9
punto10
punto 11

@@
 significant effect on our results."**
+
+**7, 8, 9, 10. Reviewer's Comment:**
+
+7. First paragraph, second sentence: change the tense of “run” to “ran”.
+8. Second paragraph, third sentence: replace “Thus” with “Furthermore”.
+9. There appears to be an orphaned right-parenthesis at the end of the second paragraph.
+10. Third paragraph, last sentence: The node number is not really relevant, so beginning of the sentence could be revised to “For example, the Charaxes + Polyura clade was found in only 256 trees...”
+
+**7, 8, 9, 10. Response:**
+
+> We made the suggested corrections.
+
+**11. Reviewer's Comment:**
+
+Fourth paragraph: The point of the paragraph is not clear. What rates are being
+compared? MEDUSA vs. MultiMEDUSA? Consider revising to make the point clearer
+or omit the sentence.
+
+**11. Response:**
+
+> We have rewritten the sentence and now is more clear: **"For the diversification shifts found in both the MCC tree and most of the
+samples of 1000 trees (frequency more than 90%; Table 2), the MultiMEDUSA
+approach recovered different rates of diversification than
+those found using when the MCC tree alone."**.

### 2014-10-10 response_to_reviewers.md
punto 12

@@
 those found using when the MCC tree alone."**.
+
+**12. Reviewer's Comment:**
+
+Fifth paragraph, first sentence (and potentially elsewhere): Consider using a consistent means of presenting probability values.
+In this sentence, the authors refer to “90% of probability in the trees”, but in the preceding paragraph, the authors present probabilities as a decimal (“probability of being recovered higher than 0.90”).
+Granted, these two probabilities refer to different events, but perhaps it would be best to refer to them in a consistent manner, such as “0.90 posterior probability”.
+
+**12. Response:**
+
+> We have corrected this and normalized the probability values to decimals in the manuscript.

### 2014-10-10 response_to_reviewers.md
punto 13
punto 14

@@
 > We have corrected this and normalized the probability values to decimals in the manuscript.
+
+**13. Reviewer's Comment:**
+The third sentence (length of analysis and burnin details) should be moved to the methods section.
+
+**13. Response:**
+
+> We made the suggested corrections.
+
+**14. Reviewer's Comment:**
+
+Mid-way through paragraph, discussion of likelihood ratio test: The source of this P-value needs to be described (chi-squared approximation with 1 degree of freedom?), here or in the methods section (see p.
+617 in Pagel 1999 Syst.
+Biol.
+48:612-622).
+
+**14. Response:**
+
+> We have include the statistics suggested by the reviewer: 
+"**$\chi^2 = 12.3; 1 df; p < 0.001$**".

### 2014-10-10 ancillary/Supporting_Information_S18.R
need to fix code for right filenames

@@ -6,14 +6,14 @@ add.alpha <- function(col, alpha=.5) {
     rgb(tmp[1,], tmp[2,], tmp[3,], alpha=alpha)
 }
 
-phy <- read.nexus("ancillary/supp_mat_01_genus.nex");
+phy <- read.nexus("ancillary/Supporting_Information_S1.nex");
 tips <- c("Achlyodes", "Graphium", "Parnassius", "Baronia", "Troides", "Papilio1", "Papilio2", "Pieris", "Aporia", "Styx", "Hamearis", "Euselasia", "Nymphidium", "Emesis", "Crocozona", "Riodina", "Amarynthis", "Baliochila", "Poritia", "Miletus", "Liphyra", "Lycaena", "Celastrina", "Thecla", "Lucia", "Curetis", "Eurema", "Colias", "Leptidea", "Pseudopontia", "Libyt");
 phy  <- drop.tip(phy, tips)
 new.phy <- multi2di(phy);
 phy <- new.phy;
 
 # Process data on using **Solanaceae** as hostplant
-data <- read.csv("ancillary/supp_mat_09_states.csv", sep=",", header=T, row.names=1);
+data <- read.csv("ancillary/Supporting_Information_S17.csv", sep=",", header=T, row.names=1);
 data <- data[order(order(phy$tip.label)),]
 
 # Column 2 is the data for **Solanaceae**
@@ -29,7 +29,7 @@ phy$tip.state<-data.v
 ##################
 # get sampling values. Propotion of extant species in our tree that have state 0 and state 1
 # in relation with the total 
-richness <- read.csv("ancillary/supp_mat_03_richness.csv")
+richness <- read.csv("ancillary/Supporting_Information_S3.csv")
 richness <- richness[order(order(phy$tip.label)),]
 
 richness_cero <- c();
@@ -78,10 +78,10 @@ w <- diff(sapply(tmp[2:7], quantile, c(0.05, 0.95)));
 
 # this runs the markov chain
 samples <- mcmc(lik, fit$par, nsteps=10000, w=w, lower=0, 
-                prior=prior, save.every=1000, save.file="bisse_mcmc_run.csv");
+                prior=prior, save.every=1000, save.file="ancillary/Supporting_Information_S19.csv");
 
 # read markov chain from saved data file
-#samples <- read.csv("bisse_mcmc_run.csv")
[…]

### 2014-10-13 ancillary/Supporting_Information_S2.R
fixed code S2

@@ -2,7 +2,7 @@
 
 library("MEDUSA")
 library("multicore")
-phy <- read.nexus("ancillary/supp_mat_01_genus.nex");
+phy <- read.nexus("ancillary/Supporting_Information_S1.nex");
 phy <- ladderize(phy)
 
 # we have to remove the tips
@@ -10,7 +10,7 @@ tips <- c("Achlyodes", "Graphium", "Parnassius", "Baronia", "Troides", "Papilio1
 phy <- drop.tip(phy, tip=tips)
 
 
-richness <- read.csv("ancillary/supp_mat_03_richness.csv")
+richness <- read.csv("ancillary/Supporting_Information_S3.csv")
 
 res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25, mc=TRUE)
 
@@ -20,5 +20,5 @@ summ <- medusaSummary(res, cex=0.3)
 dev.off()
 
 
-save(res, file="ancillary/supp_mat_04_MEDUSA_results.txt", ascii=TRUE)
+# save(res, file="ancillary/supp_mat_04_MEDUSA_results.txt", ascii=TRUE)
 

### 2014-10-13 ancillary/Supporting_Information_S6.R
fixed S6

@@ -1,10 +1,10 @@
 library("MEDUSA")
 library("multicore")
 
-mct <- read.nexus("ancillary/supp_mat_09_mcc_tree_from_random_sample_of_1000.tree")
-phy <- read.nexus("ancillary/supp_mat_05_1000_random_trees_no_outgroups.nex")
+mct <- read.nexus("ancillary/Supporting_Information_S9.tree")
+phy <- read.nexus("ancillary/Supporting_Information_S5_1000_random_trees_no_outgroups.nex")
 
-richness <- read.csv("ancillary/supp_mat_03_richness.csv")
+richness <- read.csv("ancillary/Supporting_Information_S3.csv")
 
 res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=27, mc=TRUE)
 
@@ -102,8 +102,8 @@ dev.off()
 library("phytools")
 
 # for a given tree, get its descendant tip names for a given node
-mct <- read.nexus("ancillary/supp_mat_09_mcc_tree_from_random_sample_of_1000.tree")
-phy <- read.nexus("ancillary/supp_mat_05_1000_random_trees_no_outgroups.nex")
+mct <- read.nexus("ancillary/Supporting_Information_S9.tree")
+phy <- read.nexus("ancillary/Supporting_Information_S5_1000_random_trees_no_outgroups.nex")
 
 get_descendant_tips <- function(phy, node) {
   tips <- c()

### 2014-10-13 ancillary/Supporting_Information_S10.R
S10

@@ -2,12 +2,12 @@
 
 library("MEDUSA")
 library("multicore")
-phy <- read.nexus("ancillary/supp_mat_09_mcc_tree_from_random_sample_of_1000.tree");
+phy <- read.nexus("ancillary/Supporting_Information_S9.tree");
 phy <- ladderize(phy)
 
 
 
-richness <- read.csv("ancillary/supp_mat_03_richness.csv")
+richness <- read.csv("ancillary/Supporting_Information_S3.csv")
 
 res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25, mc=TRUE)
 

### 2014-10-13 response_to_reviewers.md
punto 15
punto 16

@@
 "**$\chi^2 = 12.3; 1 df; p < 0.001$**".
+
+**15. Reviewer's Comment:**
+
+The closing three sentences (about BiSSE performance) should be moved to the methods section.
+
+**15. Response:**
+
+> We made the suggested corrections.
+
+**16. Reviewer's Comment:**
+
+First paragraph: The authors' assessment of analyzing diversification shifts using a single tree, without accommodating our level of certainty in relationships, is justified; however, MEDUSA itself, that is, the approach using stepwise AIC to identify shifts in diversification rate, should not be the target of criticism.
+After all, MultiMEDUSA is doing the same thing, but on a sample of trees.
+Revise this paragraph with a more appropriate target for the criticism (i.e.
+analyses performed on single trees).
+
+**16. Response:**
+
+> We have reworded the paragraph to avoid blaming the method:
+
+> "The MEDUSA method has been used to infer changes in net diversification rates in a phylogenetic
+tree. Since its publication [16] the results of using MEDUSA on a single tree, the maximum clade
+credibility tree, have been used for generation of hypotheses and discussion [10,42,43]. However, different
+diversification shifts and different rates of diversification are found for certain lineages when phylogenetic
+uncertainty was taken into account by using MEDUSA on a random sample of trees from the posterior
+distribution of a Bayesian run. We found that some diversification splits, estimated on the Nymphalidae
+maximum clade credibility tree, were found with very low probability in the random sample of 1000 trees
+from the posterior distribution (Fig. 3, Table 2). We also found that, even though the analyses estimated
+the same diversification splits on two or more trees, the estimated net diversification rates could vary
+widely (Fig. 2)."

### 2014-10-13 response_to_reviewers.md
punto 17

@@
 widely (Fig. 2)."
+
+**17. Reviewer's Comment:**
+
+Seventh paragraph, second to last sentence: Two ages for Ithomiini are given – are they the origin and diversification ages?
+If so, revise the sentence to include this information (rather than requiring readers to refer to the preceding sentence): “Reference 10 gives the ages of origin and diversification for Ithomiini at 45 (39-53) and 37 (32-43) MYA, respectively”.
+
+**17. Response:**
+
+> We have followed the reviewer's suggestion.

### 2014-10-13 response_to_reviewers.md
punto 18
punto 19

@@
 > We have followed the reviewer's suggestion.
+
+**18. Reviewer's Comment:**
+
+Eighth paragraph: This paragraph discussing biogeography does not fit and could be omitted without detriment to the manuscript.
+
+**18. Response:**
+
+> We have followed the reviewer's suggestion and removed this paragraph.
+
+**19. Reviewer's Comment:**
+
+First paragraph, first sentence: replace “gave” with “showed”.
+
+**19. Response:**
+
+> We corrected the text following the reviewer's suggestion.

### 2014-10-13 response_to_reviewers.md
punto 20
punto 21
punto 22

@@
 > We corrected the text following the reviewer's suggestion.
+
+**20. Reviewer's Comment:**
+
+This section is predominantly speculative. Consider a more rigorous discussion (if appropriate) or omit.
+
+**20. Response:**
+
+> We have followed the reviewer's suggestion and removed this paragraph.
+
+**21. Reviewer's Comment:**
+
+First paragraph, second sentence: revise “13% of probabilities of the trees from the posterior distribution” to “13% of the trees from the posterior distribution”.
+
+**21. Response:**
+
+> We corrected the text following the reviewer's suggestion.
+
+**22. Reviewer's Comment:**
+
+Second paragraph: this paragraph is, for the most part, unnecessary.
+The last sentence of the paragraph should be revised and appended to the end of the preceding (first) paragraph: “...not picked up by MEDUSA.
+This would be expected if the diversification of Satyrinae occurred in a stepwise manner, with pulses or bursts of diversification for certain lineages but unlikely for the tribe Satyrini as a whole”.
+
+**22. Response:**
+
+> We corrected the text following the reviewer's suggestion.

### 2014-10-13 response_to_reviewers.md
punto 23

@@
 > We corrected the text following the reviewer's suggestion.
+
+**23. Reviewer's Comment:**
+
+Final two paragraphs: There needs to be a section break, to indicate a change in the discussion topic.
+“Conclusions” would be appropriate to indicate to readers that these paragraphs represent a summary of the major points of the work.
+The final paragraph is a good summary of the work, and it would be a disservice to have readers miss it because it appears under the 'Satyrini' section.
+
+**23. Response:**
+
+> As suggested by the reviewer, he have created a **Conclusions** section for
+the last two paragraphs of the text.

### 2014-10-13 response_to_reviewers.md
punto 24

@@
 the last two paragraphs of the text.
+
+**24. Reviewer's Comment:**
+
+Figure 1: The legend needs to indicate what the node labels (the circled integers) and the decimal values in the tree indicate.
+
+**24. Response:**
+
+> We have appended the following text to the legend of figure 1.
+**"Circles on nodes indicate the diversification shift number as found by MEDUSA.
+Numbers next to circles indicat the posterior probability values for
+such nodes."**

### 2014-10-13 response_to_reviewers.md
punto 25
punto 26

@@
 such nodes."**
+
+**25. Reviewer's Comment:**
+
+Figure 2: Omit the legend in the figure (“Boxplot of diversification...”) and revise the legend to read: “Diversification rates for taxa estimated by MEDUSA on the sample of 1000 trees.”
+
+**25. Response:**
+
+> We removed the legend from the figure file and rewrote the legend in the MS
+as suggested by the reviewer.
+
+**26. Reviewer's Comment:**
+
+Figure 3: Omit the legend in the figure (“Probability of finding nodes...”) and revise the legend to read: “Results of MultiMEDUSA analysis showing the probability of specific nodes being characterized by significant shifts in diversification rate.”
+
+**26. Response:**
+
+> We removed the legend from the figure file and rewrote the legend in the MS
+as suggested by the reviewer.

### 2014-10-14 ancillary/Supporting_Information_S02.R
revised filenames for supp mat

@@ -1,0 +1,24 @@
+# run MEDUSA 0.93-4-33
+
+library("MEDUSA")
+library("multicore")
+phy <- read.nexus("ancillary/Supporting_Information_S01.nex");
+phy <- ladderize(phy)
+
+# we have to remove the tips
+tips <- c("Achlyodes", "Graphium", "Parnassius", "Baronia", "Troides", "Papilio1", "Papilio2", "Pieris", "Aporia", "Styx", "Hamearis", "Euselasia", "Nymphidium", "Emesis", "Crocozona", "Riodina", "Amarynthis", "Baliochila", "Poritia", "Miletus", "Liphyra", "Lycaena", "Celastrina", "Thecla", "Lucia", "Curetis", "Eurema", "Colias", "Leptidea", "Pseudopontia", "Libyt")
+phy <- drop.tip(phy, tip=tips)
+
+
+richness <- read.csv("ancillary/Supporting_Information_S03.csv")
+
+res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25, mc=TRUE)
+
+
+pdf(file="output/medusa_on_mct.pdf", width=9, height=19)
+summ <- medusaSummary(res, cex=0.3)
+dev.off()
+
+
+# save(res, file="ancillary/supp_mat_04_MEDUSA_results.txt", ascii=TRUE)
+

### 2014-10-14 ancillary/Supporting_Information_S10.R

@@ -2,12 +2,12 @@
 
 library("MEDUSA")
 library("multicore")
-phy <- read.nexus("ancillary/Supporting_Information_S9.tree");
+phy <- read.nexus("ancillary/Supporting_Information_S09.tree");
 phy <- ladderize(phy)
 
 
 
-richness <- read.csv("ancillary/Supporting_Information_S3.csv")
+richness <- read.csv("ancillary/Supporting_Information_S03.csv")
 
 res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25, mc=TRUE)
 

### 2014-10-14 ancillary/Supporting_Information_S18.R

@@ -6,7 +6,7 @@ add.alpha <- function(col, alpha=.5) {
     rgb(tmp[1,], tmp[2,], tmp[3,], alpha=alpha)
 }
 
-phy <- read.nexus("ancillary/Supporting_Information_S1.nex");
+phy <- read.nexus("ancillary/Supporting_Information_S01.nex");
 tips <- c("Achlyodes", "Graphium", "Parnassius", "Baronia", "Troides", "Papilio1", "Papilio2", "Pieris", "Aporia", "Styx", "Hamearis", "Euselasia", "Nymphidium", "Emesis", "Crocozona", "Riodina", "Amarynthis", "Baliochila", "Poritia", "Miletus", "Liphyra", "Lycaena", "Celastrina", "Thecla", "Lucia", "Curetis", "Eurema", "Colias", "Leptidea", "Pseudopontia", "Libyt");
 phy  <- drop.tip(phy, tips)
 new.phy <- multi2di(phy);
@@ -29,7 +29,7 @@ phy$tip.state<-data.v
 ##################
 # get sampling values. Propotion of extant species in our tree that have state 0 and state 1
 # in relation with the total 
-richness <- read.csv("ancillary/Supporting_Information_S3.csv")
+richness <- read.csv("ancillary/Supporting_Information_S03.csv")
 richness <- richness[order(order(phy$tip.label)),]
 
 richness_cero <- c();

### 2014-10-14 ancillary/Supporting_Information_S2.R

@@ -1,24 +1,0 @@
-# run MEDUSA 0.93-4-33
-
-library("MEDUSA")
-library("multicore")
-phy <- read.nexus("ancillary/Supporting_Information_S1.nex");
-phy <- ladderize(phy)
-
-# we have to remove the tips
-tips <- c("Achlyodes", "Graphium", "Parnassius", "Baronia", "Troides", "Papilio1", "Papilio2", "Pieris", "Aporia", "Styx", "Hamearis", "Euselasia", "Nymphidium", "Emesis", "Crocozona", "Riodina", "Amarynthis", "Baliochila", "Poritia", "Miletus", "Liphyra", "Lycaena", "Celastrina", "Thecla", "Lucia", "Curetis", "Eurema", "Colias", "Leptidea", "Pseudopontia", "Libyt")
-phy <- drop.tip(phy, tip=tips)
-
-
-richness <- read.csv("ancillary/Supporting_Information_S3.csv")
-
-res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25, mc=TRUE)
-
-
-pdf(file="output/medusa_on_mct.pdf", width=9, height=19)
-summ <- medusaSummary(res, cex=0.3)
-dev.off()
-
-
-# save(res, file="ancillary/supp_mat_04_MEDUSA_results.txt", ascii=TRUE)
-

### 2014-10-14 ancillary/Supporting_Information_S6.R

@@ -1,147 +1,0 @@
-library("MEDUSA")
-library("multicore")
-
-mct <- read.nexus("ancillary/Supporting_Information_S9.tree")
-phy <- read.nexus("ancillary/Supporting_Information_S5_1000_random_trees_no_outgroups.nex")
-
-richness <- read.csv("ancillary/Supporting_Information_S3.csv")
-
-res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=27, mc=TRUE)
-
-
-pdf(file="ancillary/supp_mat_08_multimedusa_result01.pdf", width=9, height=22)
-summ <- multiMedusaSummary(res, conTree=mct, cex=0.3)
-dev.off()
-
-source("code/alt_plotMultimedusa.R")
-pdf(file="ancillary/supp_mat_14_multimedusa_result_rates_nodes.pdf", width=9, height=22)
-plotMultiMedusa_alt(summ)
-dev.off()
-
-save(res, file="ancillary/supp_mat_07_multimedusa_on_1000_trees.txt", ascii=TRUE)
-
-
-# plot the summary statistics usig a boxplot Fig. 02
-# save as supp. mat. 15
-library("ggplot2")
-load("ancillary/supp_mat_07_multimedusa_on_1000_trees.txt")
-summ <- multiMedusaSummary(res, conTree=mct, cex=0.3)
-data <- summ$shift.summary
-data <- as.data.frame(data)
-
-names <- c("Charaxes", "Ypthima", "Ithomiini", "Callicore + Diaethria", 
-           "Danaini", "Limenitidinae + Heliconiinae", "Caeruleuptychia + Magneuptychia", 
-           "Pedaliodes", "Taenaris", "Dryas + Dryadula", "Coenonympha", 
-           "Anaeomorpha + Hypna", "Mycalesina", "Pseudergolinae", 
-           "Phyciodina", "Satyrina", "Coenonymphina", 
-           "Satyrini")
-
-# delete rows that are not too important
-data <- data[c(1:33), 1:7]
-data <- data[-32,]
-data <- data[-31,]
-data <- data[-30,]
-data <- data[-29,]
-data <- data[-28,]
-data <- data[-25,]
-data <- data[-24,]
-data <- data[-22,]
-data <- data[-21,]
-data <- data[-17,]
-data <- data[-16,]
-data <- data[-15,]
-data <- data[-14,]
-data <- data[-13,]
-data <- data[-11,]
-
[…]

### 2014-10-14 ancillary/Supporting_Information_S02.R
punto 28

@@ -1,3 +1,6 @@
+# Script to run MEDUSA on the MCC Nymphalidae tree from Wahlberg et al., 2009
+# This script removes the outgroup taxa and loads the richness data for the tree terminals.
+
 # run MEDUSA 0.93-4-33
 
 library("MEDUSA")

### 2014-10-14 response_to_reviewers.md
punto 27
punto 29

@@
 as suggested by the reviewer.
+
+**27. Reviewer's Comment:**
+
+Figure 4: Describe what lambda (λ) and r represent in the figure.
+
+**27. Response:**
+
+> We included the following clarification in the legend: **"(speciation rate =
+$\lambda1$, net diversification rate = $r1$)."**
+
+**28, 29. Reviewer's Comment:**
+
+Supporting_Information_S1.nex: remove bolded “Genus.”
+
+**28, 29. Response:**
+
+> We followed the reviewer's suggestions and included better descriptions for
+the supplementary materials.

### 2014-10-14 response_to_reviewers.md
rev2 punto1

@@
 
-## Reviewer 1
+# Reviewer 1
 
@@
 
-> We included the following sentence to clarify the issue: **"We expected that
+> We included the following sentence to clarify the issue:  
+**"We expected that
 coding these two genera as either absence or presence would not have a
@@
 
-> We have rewritten the sentence and now is more clear: **"For the diversification shifts found in both the MCC tree and most of the
+> We have rewritten the sentence and now is more clear:  
+**"For the diversification shifts found in both the MCC tree and most of the
 samples of 1000 trees (frequency more than 90%; Table 2), the MultiMEDUSA
@@
 
-> We have appended the following text to the legend of figure 1.
+> We have appended the following text to the legend of figure 1:  
 **"Circles on nodes indicate the diversification shift number as found by MEDUSA.
@@
 
-Figure 4: Describe what lambda (λ) and r represent in the figure.
+Figure 4: Describe what lambda ($\lambda$) and $r$ represent in the figure.
 
@@
 the supplementary materials.
+
+# Reviewer 2
+
+**1. Reviewer's Comment:**
+
+The authors need to provide more information on what cutoff they used and why.
+
+**1. Response:**
+
+> Our manuscript already includes information about the AIC value used for
+choosing the best model. The following text appears in our original submission
+in the Methods-> Detecting diversification shifts on phylogenetic trees->
+last paragraph:  
+>**"We used the AICc threshold of 7.8 units, as estimated by MEDUSA,
+as the limit for a significantly better fit to select among increasingly
+complex alternative models."**

### 2014-10-14 response_to_reviewers.md
rev2 punto2 a

@@
 complex alternative models."**
+
+**2. Reviewer's Comment:**
+
+I’m not sure discussing the limitations of MEDUSA, or the tendency for MEDUSA
+to give inconsistent results across post burn-in trees, is really valid based
+on one dataset. That is, the unit of replication is the alignment – therefore
+the authors have effectively one data point and zero degrees of freedom to say
+much of anything about the power of MEDUSA across the post burnin sample. I
+think if the authors want to explore the power of MEDUSA, they need many
+different alignments (real or simulated). As it is now, I don’t think they
+really have the data to say anything about MEDUSA in general, and only about
+MEDUSA in the context of the Nymphalid tree. They need to at least point out
+there inability to generalize in the manuscript.
+
+**2. Response:**
+
+> As reviewer 1 suggested, our text gave the impression that we were
+criticizing the method MEDUSA. We have rewritten the following sections to
+clarify our criticism **that relying on the inferences made on only one tree
+(the MCC tree) is risky as analyses of equally good hyphoteses (other trees from the
+posterior distribution) cand potentially return different results.
+
+> * abstract
+  *

### 2014-10-15 response_to_reviewers.md
punto 2

@@
 
-> As reviewer 1 suggested, our text gave the impression that we were
+> * As reviewer 1 suggested, our text gave the impression that we were
 criticizing the method MEDUSA. We have rewritten the following sections to
@@
 (the MCC tree) is risky as analyses of equally good hyphoteses (other trees from the
-posterior distribution) cand potentially return different results.
+posterior distribution) cand potentially return different results**:
 
-> * abstract
-  *
+> * Abstract  
+"In evolutionary biology, analysis of maximum credibility trees in the software
+MEDUSA (modelling evolutionary diversity using stepwise AIC) is a popular
+method for estimation of shifts in diversification rates.
+We investigated whether phylogenetic uncertainty can produce different results
+by extending the method across a 
+random sample of trees from the posterior distribution of a Bayesian run."
+
+> * Results -> Phylogenetic uncertainty in the MultiMEDUSA approach
+We changed the sentence "We tested MEDUSA..." to "We used MEDUSA to find out...".
+
+> * Also, as suggested by the reviewer 1 (point 16 lines above) we improved the
+discussion.
+
+> * We have added the following in the Conclusions to acknowledge that the
+reported behaviour might also occur in analyses of other taxa:  
+> " However, by using a
+MultiMEDUSA approach, we found that for this Nymphalidae dataset some of these
+splits might be greatly
+affected by phylogenetic uncertainty.
+> We recommend that all datasets should be analyzed using both approaches, MEDUSA
+and MultiMEDUSA, in order to test whether the results are robust when
+phylogenetic uncertainty is taken into account."

### 2014-10-15 response_to_reviewers.md
punto 3

@@
 phylogenetic uncertainty is taken into account."
+
+**3. Reviewer's Comment:**
+
+“Our results show that taking phylogenetic uncertainty into account when estimating diversification rate shifts is of great importance, and relying on the maximum credibility tree alone potentially can give erroneous results”.
+The results are not really in error.
+The results are the correct results for the model.
+The real question is whether it is the appropriate model.
+
+**3. Response:**
+
+> We have rewritten that sentence in the abstract as follows:  
+"Our results show that taking phylogenetic uncertainty into account when
+estimating net diversification rate shifts is of great importance, as very
+different results can be obtained when using the maximum clade credibility tree
+and other trees from the posterior distribution."

### 2014-10-24 response_to_reviewers.md
punto4

@@
 and other trees from the posterior distribution."
+
+**4. Reviewer's Comment:**
+
+Why are they restricting their analysis to only one method (a method that they point out might not be reliable)?
+There are alternative methods available, and it might be interesting to see if consistent results are observed (eg., RC, PRC, BAMM).
+There is very little development on other aspects of plant – herbivore interactions and how they might be predicted to affect diversification rate.
+
+**4. Response:**
+
+[**>>>> Marianne**, can you write whey we are not using the RC and PRC methods? "The
+parametric rate comparison and relative rates test that he suggests both
+require incomplete taxon sampling to be nonrandom, which ours definitely isn't
+so we can easily explain why we didn't use these."**<<<<<<**]
+
+The method that we used, BiSSE, might not be reliable when the number of taxa
+is small, and there are biases in the number of species that have each of the
+character states. However, we stated in the manuscript that according to the
+evaluation of BiSSE by @davis2013, our data should not be affected for such
+biases.
+
+The method BAMM can be used to explore the evolution of phenotypic traits. 
+However the trait characters
+[**must be continuous**](http://bamm-project.org/configuration.html#id5).
+In our case the "feeding on Solanaceae plants" trait is categorical 
+(absence/presence) so it cannot be analyzed with BAMM.
+
+BAMM also measures the diversification rates of
+particular clades. We run an analysis to test whether we obtained congruent
+results. In the Fig. 1, shown below, we can see that BAMM was able to pick up two
+increases in diversification for Ithomiini butterflies, one for the origin of
+the clade and, most importantly, another increase during the recent
+diversification of the group. This is in agreement with our results from using
+BiSSE, that the ithomiines using Solanaceae underwent an increase in
+diversification rates.
+
+![Bamm analysis](response_to_reviewers_fig1.pdf)
+
+# References

### 2014-10-28 response_to_reviewers.md
puntos 5-6

@@
 
+**5. Reviewer's Comment:**
+
+The authors should ‘play’ with an AIC cutoff, simply to explore how it might affect there inferences.
+
+**5. Response:**
+
+> As specified lines above and in the manuscript, we used the AIC cutoff value
+7.8 as suggested by MEDUSA (this value is regarded as optimal due to the number
+of taxa in our analyses).
+Changing the cutoff value would allow fewer or additional diversification
+shifts to be picked up by MEDUSA. Hovewer the initial diversification shifts
+will remain the same unless the AIC cutoff used is so hight that no single
+diversification shifts are recovered.
+Thus playing with the AIC cutoff values does not affect most of the
+diversification shifts discussed in our manuscript. 
+We rerun all our MEDUSA and MultiMEDUSA analysis with differente AIC cutoff
+values to see any effect on the results. We used the AIC values 4 and 11.
+
+> The results in Fig.2 (also attached as PDF file) shows our MEDUSA results from
+the manuscript (cutoff 7.8) along with results with AIC cutoff values 4 and 11.
+We see that the first 19 diversification shitfs are the same for cutoff values
+7.8 and 4. However when using AIC cutoff 4 we recovered additional
+diversification shifts as the software kept trying to fit additional models
+until the differente in AIC values was less than 4. As expected MultiMEDUSA
+also found additinal diversification shifts across the random sample of 1000
+trees.
+
+> When we used AIC cutoff 11, we obtained fewer diversification shifts. We
+recovered only 16, which are the same first 16 that we found in our analysis
+with AIC 7.8.
+
+![MEDUSA analyses under three AIC values](multimedusa_7.8_vs_4_vs_11_AIC_cutoff.pdf)
+
+**6. Reviewer's Comment:**
+
+In addition, the authors should be cautious about interpreting AIC in a
+frequentist manner (e.g., “Only three out of 13 significant shifts found on the
+maximum credibility tree were consistent across more than 95% of the
+trees from the posterior”) – that is, AIC is a model selection tool,
+which should act as a guide for selecting the ‘best’ family of
+models. Delta AIC should really not be thought of as alpha = 0.05.
+
+**6. Response:**
+
+> We agree that AIC values should not be treated in a frequentist manner.
+However, we use the term **"significant shifts"**, in the same way as the authors of the
+method MEDUSA [@alfaro2009], to pinpoint to specific nodes that show a higher
+net diversification rate than ancestral or descendant nodes.
+
+> We also used the AIC value in the correct way. We used the cutoff 7.8 for all
+MEDUSA analyses. Then we counted the number of significant shifts infered in
+each tree and tried to pinpoint those that were consistently recovered across
+the sample of 1000 random trees (and are also present in the MCC tree).
+Since no single significant shift was found in all 1000 trees we wanted to show
+those are were most commonly recovered. This is way we considered consistent
+diversification shifts those that were found in most of the trees (in more than
+95% of them).
+
+**7. Reviewer's Comment:**
+
+Why are authors referred to as references, rather than by authors.
+
+As noted by Reference 36, the responsible trait might….
+
+**7. Response:**
+
 # References

### 2014-10-29 response_to_reviewers.md
punto7

@@
 
-![MEDUSA analyses under three AIC values](multimedusa_7.8_vs_4_vs_11_AIC_cutoff.pdf)
+![MEDUSA analyses under three AIC values](multimedusa_7_8_vs_4_vs_11_AIC_cutoff.pdf)
 
@@
 
-As noted by Reference 36, the responsible trait might….
+As noted by Reference 36, the responsible trait might...
 
@@
 
+> We have fixed those references as suggested by the reviewer.
+
 # References

### 2014-10-29 MS_revised.tex
table 1

--- 2014-07-24 MS.tex
@@
-\documentclass[10pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-\usepackage{hyperref}
-
-% line numbers
-\usepackage{lineno}
-
-% ligatures disabled
-\usepackage{microtype}
-\DisableLigatures[f]{encoding = *, family = * }
-
-% rotating package for sideways tables
-%\usepackage{rotating}
-
-% If you wish to include algorithms, please use one of the packages below.
-Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
-%\usepackage{algorithmic}
-%\usepackage{algorithmicx}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
+\documentclass[]{article}
+\usepackage{setspace}
+\usepackage[]{lineno}
+\def\linenumberfont{\normalfont\small\sffamily}
+\doublespacing
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{tabu}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage[none]{hyphenat}
+\usepackage{fixltx2e} % provides \textsubscript
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+\usepackage[margin=1in,a4paper]{geometry}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={},
+            colorlinks=true,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
+
+\linenumbers
 
@@
 \date{}
-\pagestyle{myheadings}
 
 \begin{document}
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{Diversity dynamics in Nymphalidae butterflies: Effect of
+
+\textbf{Title: Diversity dynamics in Nymphalidae butterflies: Effect of
 phylogenetic uncertainty on diversification rate shift estimates}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Carlos Pe\~na$^{1,\ast}$, 
-Marianne Espeland$^{2}$
-\\
-\bf{1} Carlos Pe\~na Laboratory of Genetics, Department of Biology, University of Turku, Turku
-\\
-\bf{2} Museum of Comparative Zoology and Department of Organismic and
-Evolutionary Biology, Harvard University, Cambridge, USA. Email:
-marianne.espeland@gmail.com.
+
+\subsubsection{Authors: Carlos Peña\textsuperscript{1} and Marianne
+Espeland\textsuperscript{2}}\label{authors-carlos-peuxf1a1-and-marianne-espeland2}
+
+\subsubsection{Running title: Diversity dynamics in Nymphalidae
+butterflies}\label{running-title-diversity-dynamics-in-nymphalidae-butterflies}
+
+\emph{\textsuperscript{1} Laboratory of Genetics, Department of Biology,
+University of Turku, Turku, Finland. Email:
+\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}}
+
+\emph{\textsuperscript{2} Museum of Comparative Zoology and Department
+of Organismic and Evolutionary Biology, Harvard University, Cambridge,
+USA. Email:
+\href{mailto:marianne.espeland@gmail.com}{\nolinkurl{marianne.espeland@gmail.com}}.
 Current address: McGuire Center for Lepidoptera and Biodiversity,
 Florida Museum of Natural History, University of Florida, Gainesville,
-FL, US
-\\
-$\ast$ E-mail: Corresponding mycalesis@gmail.com
-\end{flushleft}
-
-\section*{Abstract}
+FL, US}.
+
+\textbf{Corresponding author:}
+\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}
+Telephone: +358 417063065 Fax: +358 2 333 6680
+
+\textbf{Comments:} 5814 words, 7 figures, 2 tables and 25 supplementary
+material files.
+
+\textbf{Additional Supplementary materials:}
+\url{http://dx.doi.org/10.6084/m9.figshare.639208}
+
+\subsection{Abstract}\label{abstract}
 
@@
 insect-hostplant dynamics predict accelerated diversification due to key
-innovations. We investigated whether phylogenetic uncertainty affects a
-commonly used method (MEDUSA, modelling evolutionary diversity using
-stepwise AIC) for estimating shifts in diversification rates in
-lineages, by extending the method across a random sample of trees from
-the posterior distribution of a Bayesian run. We found that phylogenetic
-uncertainty greatly affects diversification rate estimates. Different
-trees produced diversification rates ranging from high values to almost
-zero for the same clade, and both significant rate increase and decrease
-in some clades. Only four out of 18 significant shifts found on the
-maximum clade credibility tree were consistent across most of the
-sampled trees. Among these, we found accelerated diversification for
-Ithomiini butterflies. We used the binary speciation and extinction
-model (BiSSE) and found that a hostplant shift to Solanaceae is
-correlated with increased net diversification rates in Ithomiini,
-congruent with the diffuse cospeciation hypothesis. Our results show
-that taking phylogenetic uncertainty into account when estimating net
-diversification rate shifts is of great importance, and relying on the
-maximum clade credibility tree alone can potentially give erroneous
-results.
+innovations. In evolutionary biology, analysis of maximum credibility
+trees in the software MEDUSA (modelling evolutionary diversity using
+stepwise AIC) is a popular method for estimation of shifts in
+diversification rates. We investigated whether phylogenetic uncertainty
+can produce different results by extending the method across a random
+sample of trees from the posterior distribution of a Bayesian run. Using
+the MultiMEDUSA approach, we found that phylogenetic uncertainty greatly
+affects diversification rate estimates. Different trees produced
+diversification rates ranging from high values to almost zero for the
+same clade, and both significant rate increase and decrease in some
+clades. Only four out of 18 significant shifts found on the maximum
+clade credibility tree were consistent across most of the sampled trees.
+Among these, we found accelerated diversification for Ithomiini
+butterflies. We used the binary speciation and extinction model (BiSSE)
+and found that a hostplant shift to Solanaceae is correlated with
+increased net diversification rates in Ithomiini, congruent with the
+diffuse cospeciation hypothesis. Our results show that taking
+phylogenetic uncertainty into account when estimating net
+diversification rate shifts is of great importance, as very different
+results can be obtained when using the maximum clade credibility tree
+and other trees from the posterior distribution.
 
@@
 
-\section*{Introduction}
+\subsection{Introduction}\label{introduction}
 
@@
 developing some of the mentioned hypotheses. Nymphalidae contains around
-6000 species {[}6{]}, and is the largest family within the true
-butterflies. The family most likely originated around 94 MYA in the mid
-Cretaceous. Diversification of the group began in the Late Cretaceous
-and most major radiations (current subfamilies) appeared shortly after
-the Cretaceous-Paleogene (K-Pg) boundary {[}7{]}. Several studies have
-used time-calibrated phylogenies and diversification models to
-reconstruct the evolutionary history of the group to identify patterns
-of accelerated or decelerated diversification of some Nymphalidae clades
-{[}7--10{]}. For example, it has been suggested that climate change in
-the Oligocene and the subsequent diversification of grasses has led to
-diversification of the subfamily Satyrinae {[}11{]} due to the abundance
-of grasses over extensive geographic areas (``resource
-abundance-dependent diversity dynamics'' hypothesis). Reference 9 found
-increased net diversification rates in some Nymphalidae lineages after a
-major hostplant shift, which appears to be in agreement with Reference 3
-``escape-and-radiate'' model of diversification.
+6000 species {[}6{]}, and several members are considered model organisms
+in evolutionary biology {[}7--9{]}. The family most likely originated
+around 94 MYA in the mid Cretaceous. Diversification of the group began
+in the Late Cretaceous and most major radiations (current subfamilies)
+appeared shortly after the Cretaceous-Paleogene (K-Pg) boundary
+{[}10{]}. Several studies have used time-calibrated phylogenies and
+diversification models to reconstruct the evolutionary history of the
+group to identify patterns of accelerated or decelerated diversification
+of some Nymphalidae clades {[}10--13{]}. For example, it has been
+suggested that climate change in the Oligocene and the subsequent
+diversification of grasses has led to diversification of the subfamily
+Satyrinae {[}14{]} due to the abundance of grasses over extensive
+geographic areas (``resource abundance-dependent diversity dynamics''
+hypothesis). Fordyce (2010) {[}12{]} found increased net diversification
+rates in some Nymphalidae lineages after a major hostplant shift, which
+appears to be in agreement with the ``escape-and-radiate'' model of
+diversification {[}3{]}.
 
@@
 sudden diversification bursts due to hostplant shift, climatic events or
-shifts to closely related hostplants {[}5,12{]}.
+shifts to closely related hostplants {[}5,15{]}.
 
 In this study, we used a time-calibrated genus-level phylogenetic
-hypothesis for Nymphalidae butterflies {[}10{]} to investigate patterns
-of diversification. We applied MEDUSA {[}13,14{]}, a recently developed
+hypothesis for Nymphalidae butterflies {[}13{]} to investigate patterns
+of diversification. We applied MEDUSA {[}16,17{]}, a recently developed
 statistical method, to study the diversification pattern of Nymphalidae
@@
 study the effects of phylogenetic uncertainty and by using the extended
-MEDUSA method called MultiMEDUSA {[}13{]}. We also tested whether
+MEDUSA method called MultiMEDUSA {[}16{]}. We also tested whether
 hostplant association dynamics can explain the diversification patterns
@@
 employing the method BiSSE as implemented in the R package
-\texttt{diversitree} {[}15{]}.
-
-\section*{Materials and Methods}
-
-\subsection*{Data}
-
-For analyses, we used the phylogenetic trees from the study of Reference
-10 that were generated using DNA sequence data from 10 gene regions for
-398 of the 540 valid genera in Nymphalidae. We employed Reference 10
-maximum clade credibility tree (MCC tree) (Fig. 1) as well as a random
-sample of 1000 trees from their BEAST run after burnin. Their original
-BEAST run was for 40 million generations. We used a burnin of 25 million
-generations and took a random sample of 1000 trees using Burntrees
-v.0.1.9 \url{http://www.abc.se/~nylander/} (supp. mat. 5) in order to
-correct for phylogenetic uncertainty when performing the diversification
-analyses.
+\texttt{diversitree} {[}18{]}.
+
+\subsection{Methods}\label{methods}
+
+\subsubsection{Data}\label{data}
+
+For analyses, we used the phylogenetic trees from the study of Wahlberg
+et al. (2009) {[}13{]} that were generated using DNA sequence data from
+10 gene regions for 398 of the 540 valid genera in Nymphalidae. We
+employed the maximum clade credibility tree (MCC tree) (Fig. 1) as well
+as a random sample of 1000 trees from their BEAST run after burnin
+{[}13{]}. Their original BEAST run was for 40 million generations. We
+used a burnin of 25 million generations and took a random sample of 1000
+trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (supp.
+mat. 5) in order to correct for phylogenetic uncertainty when performing
+the diversification analyses.
 
@@
 sources including the specialist-curated lists on
-\url{http://tolweb.org}, Reference 16 and curated lists of Global
-Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}). We
-assigned the species numbers of genera not included in the phylogeny to
-the closest related genus that was included in Reference 10 study
-according to available phylogenetic studies {[}17--28{]}, taxonomical
-classification and morphological resemblance when no phylogenies were
-available (supp. mat. 3).
+\url{http://tolweb.org}, Lamas (2004) {[}19{]} and curated lists of
+Global Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}).
+We assigned the species numbers of genera not included in the phylogeny
+to the closest related genus that was included in Wahlberg et al. (2009)
+{[}13{]} study according to available phylogenetic studies {[}20--31{]},
+taxonomical classification and morphological resemblance when no
+phylogenies were available (supp. mat. 3).
 
 Hostplant data for Nymphalidae species were compiled from several
-sources including Reference 29, HOSTS database
-(\url{http://bit.ly/YI7nwW}), Reference 30, Reference 31, Reference 32
-and others (supp. mat. 15--16) for a total of 6586 hostplant records,
-including 428 Nymphalidae genera and 143 plant families and 1070 plant
-genera. It was not possible to find any hostplant data for 35 butterfly
-genera.
-
-\subsection*{Analyses of Diversification}
-
-We used the statistical software R version 3.0.1 {[}33{]} in combination
-with the APE {[}34{]}, GEIGER {[}35{]}, MEDUSA {[}13{]} and
-\texttt{diversitree} {[}15{]} packages along with our own scripts to
+sources including {[}32{]}, HOSTS database (\url{http://bit.ly/YI7nwW}),
+{[}33--35{]} and others (supp. mat. 15--16) for a total of 6586
+hostplant records, including 428 Nymphalidae genera and 143 plant
+families and 1070 plant genera. It was not possible to find any
+hostplant data for 35 butterfly genera.
+
+\subsubsection{Analyses of
+Diversification}\label{analyses-of-diversification}
+
+We used the statistical software R version 3.0.1 {[}36{]} in combination
+with the APE {[}37{]}, GEIGER {[}38{]}, MEDUSA {[}16{]} and
+\texttt{diversitree} {[}18{]} packages along with our own scripts to
 perform the analyses (included as supplementary materials 2, 6, 10). All
-analyses were run on the 1000 random trees from Reference 10, on the
-maximum clade credibility tree derived from these and the MCC tree from
-Reference 10.
-
-\subsection*{Detecting diversification shifts on phylogenetic trees}
+analyses were run on the 1000 random trees from Wahlberg et al. (2009)
+{[}13{]}, on the maximum clade credibility tree derived from these and
+the MCC tree {[}13{]}.
+
+\subsubsection{Detecting diversification shifts on phylogenetic
+trees}\label{detecting-diversification-shifts-on-phylogenetic-trees}
 
 Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
-version 093.4.33 {[}13{]} on the maximum clade credibility tree from
-Reference 10. MEDUSA fits alternative birth-death likelihood models to a
-phylogenetic tree in order to estimate changes in net diversification
-rates along branches. MEDUSA estimates likelihood and AIC scores for the
-simplest birth-death model, with two parameters, the rates \texttt{r}:
-net diversification \texttt{=} speciation (\texttt{b}) \texttt{-}
-extinction (\texttt{d}) and $\varepsilon$: relative extinction
-\texttt{= d/b}. The AIC scores of the two-parameter model are then
-compared with incrementally more complex models until the addition of
-parameters do not improve the AIC scores beyond a cutoff value. MEDUSA
-finds the likelihood of the models after taking into account branch
-lengths and number of species per lineage {[}13{]}. Most studies using
-MEDUSA only run the method on a single tree, usually the maximum clade
-credibility tree, which makes the assumption that this tree is correct.
-We wanted to study the effects of phylogenetic uncertainty on estimation
-of net diversification rate shifts and therefore selected 1000 random
-genus-level trees from the posterior distribution (MultiMEDUSA, supp.
-mat. 5). We calculated a new MCC tree from the selection of trees, ran
-MEDUSA and MultiMEDUSA on the selection of 1000 trees and summarized the
-estimated changes in net diversification rates for nodes across all
-trees. Patterns of change in net diversification rates are considered
-significant if they are found at the same node in at least 90\% of the
-trees. We also expected to find similar values of net diversification
-(\texttt{r = b - d}) and relative extinction rate ($\varepsilon$
-\texttt{= d/b}) values across the 1000 trees for the nodes where changes
-in diversification tempo occurs. We used the AICc threshold of 7.8
-units, as estimated by MEDUSA, as the limit for a significantly better
-fit to select among increasingly complex alternative models.
-
-\subsection*{Estimation of trait-dependent speciation rates}
+version 093.4.33 {[}16{]} on the maximum clade credibility tree from
+Wahlberg et al. (2009) {[}13{]}. MEDUSA fits alternative birth-death
+likelihood models to a phylogenetic tree in order to estimate changes in
+net diversification rates along branches. MEDUSA estimates likelihood
+and AIC scores for the simplest birth-death model, with two parameters,
+the rates \texttt{r}: net diversification \texttt{=} speciation
+(\texttt{b}) \texttt{-} extinction (\texttt{d}) and \(\varepsilon\):
+relative extinction \texttt{= d/b}. The AIC scores of the two-parameter
+model are then compared with incrementally more complex models until the
+addition of parameters do not improve the AIC scores beyond a cutoff
+value. MEDUSA finds the likelihood of the models after taking into
+account branch lengths and number of species per lineage {[}16{]}. Most
+studies using MEDUSA only run the method on a single tree, usually the
+maximum clade credibility tree, which makes the assumption that this
+tree is correct. We wanted to study the effects of phylogenetic
+uncertainty on estimation of net diversification rate shifts and
+therefore selected 1000 random genus-level trees from the posterior
+distribution of the Bayesian run from Wahlberg et al., 2009 {[}13{]}
+(MultiMEDUSA, supp. mat. 5). We calculated a new MCC tree from the
+selection of trees, ran MEDUSA and MultiMEDUSA on the selection of 1000
+trees and summarized the estimated changes in net diversification rates
+for nodes across all trees. Patterns of change in net diversification
+rates are considered significant if they are found at the same node in
+at least 90\% of the trees. We also expected to find similar values of
+net diversification (\texttt{r = b - d}) and relative extinction rate
+(\(\varepsilon\) \texttt{= d/b}) values across the 1000 trees for the
+nodes where changes in diversification tempo occurs. We used the AICc
+threshold of 7.8 units, as estimated by MEDUSA, as the limit for a
+significantly better fit to select among increasingly complex
+alternative models.
+
+\subsubsection{Estimation of trait-dependent speciation
+rates}\label{estimation-of-trait-dependent-speciation-rates}
 
@@
 diversification in the clade Ithomiini, we tested whether this pattern
-can be explained by increase in birth-rate due to hostplant use and
-performed analyses with the ``binary state speciation and extinction''
-{[}36{]} Bayesian approach as implemented in the R package
-\texttt{diversitree} {[}15{]}. MuSSE {[}15{]} is designed to examine the
-joint effects of two or more traits on speciation. Because most of
-Nymphalidae butterflies are restricted to use one plant family as
-hostplant, the character states can be coded as presence/absence, for
-which the BiSSE analysis is better suited. BiSSE was designed to test
-whether a binary character state has had any effect on increased net
-diversification rate for a clade {[}36{]}. We used our compiled data of
+can be explained by an increase in the birth-rate due to hostplant use
+and performed analyses using the ``binary state speciation and
+extinction'' {[}39{]} Bayesian approach as implemented in the R package
+\texttt{diversitree} {[}18{]}. The Markov Chain Monte Carlo algorithm
+was run for 10000 generations discarding the first 7500 as burnin. The
+Multiple State Speciation and Extinction approach (MuSSE) {[}18{]} is
+designed to examine the joint effects of two or more traits on
+speciation. Because most of Nymphalidae butterflies are restricted to
+use one plant family as hostplant (232 genera use only one plant family
+versus 176 that use more than one; based on our data in supp. mat. 15),
+the character states can be coded as presence/absence, for which the
+BiSSE analysis is better suited. BiSSE was designed to test whether a
+binary character state has had any effect on increased net
+diversification rate for a clade {[}39{]}. We used our compiled data of
 hostplant use to produce a binary dataset for the character ``feeding on
 the plant family Solanaceae'' with two states (absence = 0; presence =
-1) (supp. mat. 17), which is the main hostplants of the diverse
-Ithomiini butterflies and closest relatives {[}37{]}. Other hostplant
-shifts were not tested for effect on net diversification rates due to
-the low support for diversification shifts found in the MEDUSA analyses.
-We analyzed the data using BiSSE employing the Markov Chain Monte Carlo
-algorithm on the maximum clade credibility tree, taking into account
-missing taxa by using the parameter ``sampling factor''
+1) (supp. mat. 17). The family Solanaceae include the hostplants of the
+diverse Ithomiini butterflies and closest relatives {[}8{]}. Other
+hostplant shifts were not tested for effect on net diversification rates
+due to the low support for diversification shifts found in the MEDUSA
+analyses. We analyzed the data using BiSSE employing the Markov Chain
+Monte Carlo algorithm on the maximum clade credibility tree, taking into
+account missing taxa by using the parameter ``sampling factor''
 (\texttt{sampling.f}) in \texttt{diversitree}. We also used constrained
@@
 hypothesis (no effect). The analyses were run across a sample of 250
-trees from the posterior distribution. The records of \emph{Vanessa} and
-\emph{Hypanartia} feeding on Solanaceae {[}30,38{]} might be incorrect
-as it is unlikely that these species can be feeding on this plant
-family. We tested whether coding these two genera as not feeding on
-Solanaceae affected our results.
-
-\section*{Results}
-
-\subsection*{Detecting diversification shifts on the maximum clade credibility tree}
+trees from the selected 1000 random trees from the posterior
+distribution. The records of \emph{Vanessa} and \emph{Hypanartia}
+feeding on Solanaceae {[}33,40{]} might be incorrect as it is unlikely
+that these species can be feeding on this plant family. We tested
+whether coding these two genera as not feeding on Solanaceae affected
+our results. We expected that coding these two genera as either absence
+or presence would not have a significant effect on our results. It has
+been shown that BiSSE performs poorly under certain conditions {[}41{]}.
+However, our data has adequate number of taxa under analysis (more than
+300 tips), adequate speciation bias (between 1.5x and 2.0x), character
+state bias (around 8x) and extinction bias (around 4x) for the analysis
+of Solanaceae hostplants. Thus, BiSSE is expected to produce robust
+results {[}41{]}.
+
+\subsection{Results}\label{results}
+
+\subsubsection{Detecting diversification shifts on the maximum clade
+credibility
+tree}\label{detecting-diversification-shifts-on-the-maximum-clade-credibility-tree}
 
@@
 AICc score for the best fit model was \texttt{5090.5} (Table 1). MEDUSA
-also estimated that the basic constant birth--death model was not a
+also estimated that the basic constant birth-death model was not a
 better explanation for our data (AICc \texttt{= 5449.3}).
@@
 
-\subsection*{Phylogenetic uncertainty in the MultiMEDUSA approach}
-
-We tested MEDUSA to find out whether taking into account the
-phylogenetic signal from a random sample of 1000 trees from the
-posterior distribution can return similar estimates of diversification
-to the values obtained from the MCC tree. We run MultiMEDUSA on a random
+\subsubsection{Phylogenetic uncertainty in the MultiMEDUSA
+approach}\label{phylogenetic-uncertainty-in-the-multimedusa-approach}
+
+We used MEDUSA to find out whether taking into account the phylogenetic
+signal from the random sample of 1000 trees from the posterior
+distribution can return similar estimates of diversification to the
+values obtained from the MCC tree. We ran MultiMEDUSA on the random
 sample of 1000 trees (supp. mat. 5) from the posterior distribution and
@@
 lower median net diversification rates for the diversification shifts
-found by MEDUSA on the MCC tree from Reference 10 (Table 2). Although
-the diversification pattern found by MEDUSA and MultiMEDUSA was the
-same, the latter consistently estimated lower rates. Thus, the shifts
+found by MEDUSA on the MCC tree from {[}13{]} (Table 2). Although the
+diversification pattern found by MEDUSA and MultiMEDUSA was the same,
+the latter consistently estimated lower rates. Furthermore, the shifts
 recovered with low net diversification rate on the MCC were recovered
@@
 diversification and all shifts found by MEDUSA on the 1000 trees are
-provided as an R object in supp. mat. 7).
+provided as an R object in supp. mat. 7.
 
@@
 picked up by MEDUSA and a node shift was found most of the time, then
-the \texttt{sum.prop} should be close to 1. For example the node 625
-consists of the clade \texttt{Charaxes} + \texttt{Polyura} and it was
-found in only 256 trees, however MultiMEDUSA was consistenly able to
-find a diversification shift for that node and the \texttt{sum.prop}
-value is 0.996.
-
-We found different rates for diversification shifts even for the nodes
-that appear in most of the sample of 1000 trees (Table 2) as many of the
-nodes also found in the MCC tree had a probability of being recovered
-higher than 0.90.
-
-There were four diversification shifts found with more than 90\% of
+the \texttt{sum.prop} should be close to 1. For example, the
+\texttt{Charaxes} + \texttt{Polyura} clade was found in only 256 trees,
+however MultiMEDUSA was consistently able to find a diversification
+shift for that node and the \texttt{sum.prop} value is 0.996.
+
+For the diversification shifts found in both the MCC tree and most of
+the samples of 1000 trees (frequency more than 90\%; Table 2), the
+MultiMEDUSA approach recovered different rates of diversification than
+those found using when the MCC tree alone.
+
+There were four diversification shifts found with more than 0.90 of
 probability in the trees from the random sample (Table 2): (i) net
@@
 example the split for the clade Satyrini (Euptychiina + Pronophilina +
-Satyrina + Maniolina) is recovered with a probability of 18\% (Fig. 3).
-
-\subsection*{Estimation of trait-dependent speciation rates}
-
-The MEDUSA analyses taking into account phylogenetic uncertainty
+Satyrina + Maniolina) is recovered with a probability of 0.18 (Fig. 3).
+
+\subsubsection{Estimation of trait-dependent speciation
+rates}\label{estimation-of-trait-dependent-speciation-rates-1}
+
+The MEDUSA analyses, taking into account phylogenetic uncertainty,
 estimated a net diversification rate increase in part of the clade
@@
 net net diversification rate on part of Ithomiini (Oleriina + Ithomiina
-+ Napeogenina + Dircennina + Godyridina) (Fig. 4). The Markov Chain
-Monte Carlo algorithm was run for 10000 generations discarding the first
-7500 as burnin. The estimated mean net diversification rate for taxa
-that do not feed on Solanaceae was \texttt{r = 0.11} while the net
-diversification rate for the Solanaceae feeders was \texttt{r = 0.16}
-(see Fig. S1 for a boxplot of speciation and extinction values for the
-95\% credibility intervals). The same analysis considering
-\emph{Vanessa} and \emph{Hypanartia} as non-Solanaceae feeders due to
-dubious records produced the same pattern and net diversification rates
-(Fig. S2). Therefore, the rest of the analyses were performed assuming
-these two genera as Solanaceae feeders. We constrained the BiSSE
-likelihood model to force equal rates of speciation for both character
-states in order to test whether the model of different speciation rates
-is a significantly better explanation for the data. A likelihood ratio
-test found that the model for increased net diversification rate for
-nymphalids feeding on Solanaceae is a significantly better explanation
-than this character state having no effect on diversification
-(\texttt{p \textless{} 0.001}) (Table 3; character states available in
-supp. mat. 17, code in supp. mat. 18, and mcmc run in supp. mat. 19). We
-combined the post-burnin mcmc generations from running BiSSE on 250
-trees from the posterior distribution and found the same pattern as the
-BiSSE analysis on the maximum clade credibility tree (combined mcmc run
-in supp. mat. 20; profiles plot of speciation rates in Fig. S3; boxplot
-of 95\% credibility intervals in Fig. S4). A BiSSE analysis to test
-whether the trait ``feeding on Apocynaceae'' had any effect on increased
-net diversification rates found similar speciation rates for lineages
-feeding on Apocynaceae and other plants (Fig. S5). It has been shown
-that BiSSE performs poorly under certain conditions {[}39{]}. However,
-our data has adequate number of taxa under analysis (more than 300
-tips), adequate speciation bias (between 1.5x and 2.0x), character state
-bias (around 8x) and extinction bias (around 4x) for the analysis of
-Solanaceae hostplants. Thus, BiSSE is expected to produce robust results
-{[}39{]}.
-
-\section*{Discussion}
-
-\subsection*{Effects of phylogenetic uncertainty on the performance of MEDUSA}
++ Napeogenina + Dircennina + Godyridina) (Fig. 4). The estimated mean
+net diversification rate for taxa that do not feed on Solanaceae was
+\texttt{r = 0.11} while the net diversification rate for the Solanaceae
+feeders was \texttt{r = 0.16} (see Fig. S1 for a boxplot of speciation
+and extinction values for the 95\% credibility intervals). The same
+analysis considering \emph{Vanessa} and \emph{Hypanartia} as
+non-Solanaceae feeders due to dubious records produced the same pattern
+and net diversification rates (Fig. S2). Therefore, the rest of the
+analyses were performed assuming these two genera as Solanaceae feeders.
+We constrained the BiSSE likelihood model to force equal rates of
+speciation for both character states in order to test whether the model
+of different speciation rates is a significantly better explanation for
+the data. A likelihood ratio test found that the model for increased net
+diversification rate for nymphalids feeding on Solanaceae is a
+significantly better explanation than this character state having no
+effect on diversification (\(\chi^2 = 12.3; 1 df; p < 0.001\)) (Table 3;
+character states available in supp. mat. 17, code in supp. mat. 18, and
+mcmc run in supp. mat. 19). We combined the post-burnin mcmc generations
+from running BiSSE on 250 trees from the random sample of 1000 trees and
+found the same pattern as the BiSSE analysis on the maximum clade
+credibility tree (combined mcmc run in supp. mat. 20; profiles plot of
+speciation rates in Fig. S3; boxplot of 95\% credibility intervals in
+Fig. S4). A BiSSE analysis to test whether the trait ``feeding on
+Apocynaceae'' had any effect on increased net diversification rates
+found similar speciation rates for lineages feeding on Apocynaceae and
+other plants (Fig. S5).
+
+\subsection{Discussion}\label{discussion}
+
+\subsubsection{Effects of phylogenetic uncertainty on the performance of
+MEDUSA}\label{effects-of-phylogenetic-uncertainty-on-the-performance-of-medusa}
 
 The MEDUSA method has been used to infer changes in net diversification
-rates along a phylogenetic tree. Since its publication {[}13{]} the
-results of using MEDUSA on a single tree, the maximum clade credibility
-tree, have been used for generation of hypotheses and discussion
-{[}7,40,41{]}. However, we found that MEDUSA estimated different
-diversification shifts and different rates of diversification for
-certain lineages when phylogenetic uncertainty was taken into account by
-using MEDUSA on a random sample of trees from the posterior distribution
-of a Bayesian run. We found that some diversification splits, estimated
-on the Nymphalidae maximum clade credibility tree, were found with a
-very low probability in a random sample of 1000 trees from the posterior
-distribution (Fig. 3, Table 2). We also found that, even though MEDUSA
-could estimate the same diversification splits on two or more trees, the
-estimated net diversification rates could vary widely (Fig. 2). For
+rates in a phylogenetic tree. Since its publication {[}16{]} the results
+of using MEDUSA on a single tree, the maximum clade credibility tree,
+have been used for generation of hypotheses and discussion
+{[}10,42,43{]}. However, different diversification shifts and different
+rates of diversification are found for certain lineages when
+phylogenetic uncertainty was taken into account by using MEDUSA on a
+random sample of trees from the posterior distribution of a Bayesian
+run. We found that some diversification splits, estimated on the
+Nymphalidae maximum clade credibility tree, were found with very low
+probability in the random sample of 1000 trees from the posterior
+distribution (Fig. 3, Table 2). We also found that, even though the
+analyses estimated the same diversification splits on two or more trees,
+the estimated net diversification rates could vary widely (Fig. 2). For
 example, in our Nymphalidae trees, we found that the split for
@@
 taxa appear to be strongly affected by long-branch attraction artifacts
-{[}27{]}. Thus, the Bayesian runs are expected to recover alternative
+{[}30{]}. Thus, the Bayesian runs are expected to recover alternative
 topologies on the posterior distribution of trees, resulting in low
@@
 posterior probability values for clades in Satyrini are very low
-{[}10{]}. As a result, MEDUSA inferred a net diversification rate
+{[}13{]}. As a result, MEDUSA inferred a net diversification rate
 increase for part of Satyrini in the maximum clade credibility tree, but
-this was recovered only in 13\% of probabilities in the MultiMEDUSA
-analysis on the random sample of trees.
+this was recovered in the MultiMEDUSA analysis in only in 13\% of the
+random sample of trees.
 
@@
 diversification shift (due to a non-existent node). This is the reason
-why MEDUSA estimated net diversification rate splits with more than 90\%
-of probabilities in the sample of trees for only four splits: the genus
+why MEDUSA estimated net diversification rate splits with a probability
+higher than 0.90 in the sample of trees for only four splits: the genus
 \emph{Charaxes}, the genus \emph{Ypthima}, part of Ithomiini and the
@@
 The clade Ithomiini and the non-basal danaids are well supported by high
-posterior probability values in Reference 10. Therefore our MEDUSA
-analyses recovered an increase in net diversification rate with
-probability higher than 90\% in the posterior distribution of trees
-(Fig. 3).
-
-\subsection*{Hostplant use and diversification in Nymphalidae}
-
-\subsubsection*{Ithomiini}
+posterior probability values in Wahlberg et al. (2009) {[}13{]}.
+Therefore our MEDUSA analyses recovered an increase in net
+diversification rate with probability higher than 0.90 in the posterior
+distribution of trees (Fig. 3).
+
+\subsubsection{Hostplant use and diversification in
+Nymphalidae}\label{hostplant-use-and-diversification-in-nymphalidae}
+
+\paragraph{Ithomiini}\label{ithomiini}
 
 Keith Brown suggested that feeding on Solanaceae was an important event
-in the diversification of Ithomiini butterflies {[}42{]}. Ithomiini
+in the diversification of Ithomiini butterflies {[}44{]}. Ithomiini
 butterflies are exclusively Neotropical and most species feed on
-Solanaceae hostplants during larval stage {[}37{]}. Optimizations of the
+Solanaceae hostplants during larval stage {[}8{]}. Optimizations of the
 evolution of hostplant use on phylogenies evidence a probable shift from
-Apocynaceae to Solanaceae in the ancestor of the tribe {[}37,43{]}.
-Reference 9 found that the Gamma statistics, a LTT plot of an Ithomiini
-phylogeny and the fit of the density-dependent model of diversification
-are consistent with a burst of diversification in Ithomiini following
-the shift from Apocynaceae to Solanaceae.
+Apocynaceae to Solanaceae in the ancestor of the tribe {[}8,45{]}.
+Fordyce (2010) {[}12{]} found that the Gamma statistics, a LTT plot of
+an Ithomiini phylogeny and the fit of the density-dependent model of
+diversification are consistent with a burst of diversification in
+Ithomiini following the shift from Apocynaceae to Solanaceae.
 
@@
 due to the use of Solanaceae plants as hosts during larval stage. For
-this, we used a Bayesian approach {[}44{]} to test whether the trait
+this, we used a Bayesian approach {[}46{]} to test whether the trait
 ``feeding on Solanaceae'' had any effect on the diversification of the
@@
 trait ``feeding on Solanaceae hostplants'' (Fig. 4). This is in
-agreement with the findings of Reference 9 using other statistical
-methods. Due to the fact that Ithomiini are virtually the only
+agreement with previoues findings using other statistical methods
+{[}12{]}. Due to the fact that Ithomiini are virtually the only
 nymphalids using Solanaceae as hostplants, it is possible that the trait
 responsible for a higher diversification of Ithomiini might not be the
-hostplant character. As noted by Reference 36, the responsible trait
-might be a codistributed character such as a trait related to the
-ability to digest secondary metabolites.
+hostplant character. As noted by Maddison et al. (2007) {[}39{]}, the
+responsible trait might be a codistributed character such as a trait
+related to the ability to digest secondary metabolites.
 
@@
 that the high diversity of Ithomiini is consistent with the
-``escape-and-radiate scenario'' due to a shift onto Solanaceae {[}9{]}
+``escape-and-radiate scenario'' due to a shift onto Solanaceae {[}12{]}
 and radiation scenarios among chemically different lineages of
-Solanaceae plants {[}37,42{]}. According to this theory, the shift from
+Solanaceae plants {[}8,44{]}. According to this theory, the shift from
 Apocynaceae to Solanaceae allowed Ithomiini to invade newly available
@@
 detoxification of ingested food as they have been found in other
-butterfly groups {[}45{]}.
+butterfly groups {[}47{]}.
 
@@
 Although the Solanaceae genera used by the Ithomiini clades are well
-known {[}37{]}, we do not have any understanding on the physiological
+known {[}8{]}, we do not have any understanding on the physiological
 routes involved in the detoxification of Solanaceae compounds by the
 several lineages of Ithomiini. We can speculate that older lineages
-exploiting a novel toxic resource {[}10,37{]} might not be too efficient
+exploiting a novel toxic resource {[}8,13{]} might not be too efficient
 in metabolizing plant toxins and that younger lineages are able to deal
@@
 configurations of the proteins, providing more opportunities for
-hostplant switch {[}46{]}. This might be the reason why the basal
+hostplant switch {[}48{]}. This might be the reason why the basal
 Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and
-restricted to few Solanaceae hosts {[}37{]}, while recent subtribes are
+restricted to few Solanaceae hosts {[}8{]}, while recent subtribes are
 species-rich and have expanded their host range into several Solanaceae
-lineages {[}37{]}. It might be that the switch to feeding in Solanaceae
+lineages {[}8{]}. It might be that the switch to feeding in Solanaceae
 was an important event in the evolutionary history of Ithomiini, but the
@@
 diversity'' and the ``escape-and-radiate'' hypotheses state that insects
-diversify after their hostplants {[}2--4{]}. Reference 45 found strong
-evidence for a model of speciation congruent with Ehrlich and Raven's
-hypothesis in Pieridae butterflies due to, in addition to the
-identification of a key innovation, a burst of diversification in
-glucosinolate-feeding taxa shortly afterwards (with a lag of
-\textasciitilde{}10 MY). According to a recent dated phylogeny of the
-Angiosperms {[}47{]}, the family Solanaceae split from its sister group
-about 59 (49-68) MYA and diversification started (crown group age)
-around 37 (29-47) MYA. Reference 10 give the corresponding ages for
-Ithomiini as 45 (39-53) and 37 (32-43) MYA, respectively. Thus, current
-evidence shows that Solanaceae and Ithomiini might have diversified
-around the same time, during the Late Eocene and Oligocene, and this
-would be congruent with the diffuse cospeciation hypothesis.
-
-The uplift of the Andes was a major tectonic event that underwent higher
-activity during the Oligocene {[}48{]}. This caused climatic changes in
-the region that affected the flora and fauna of the time, which
-coincides with the diversification of modern montane plant and animal
-taxa {[}49{]} including Ithomiini butterflies and Solanaceae hostplants.
-Moreover, all Solanaceae clades currently present in New World
-originated in South America {[}50{]} as well as Ithomiini butterflies
-{[}10{]}. This suggests a process of ``diffuse cospeciation'' of
-Ithomiini and hostplants.
-
-\subsubsection*{Danaini}
-
-Our MultiMEDUSA approach gave a significant slowdown in net
+diversify after their hostplants {[}2--4{]}. Wheat et al. (2007)
+{[}47{]} found strong evidence for a model of speciation congruent with
+Ehrlich and Raven's hypothesis in Pieridae butterflies due to, in
+addition to the identification of a key innovation, a burst of
+diversification in glucosinolate-feeding taxa shortly afterwards (with a
+lag of \textasciitilde{}10 MY). According to a recent dated phylogeny of
+the Angiosperms {[}49{]}, the family Solanaceae split from its sister
+group about 59 (49--68) MYA and diversification started (crown group
+age) around 37 (29--47) MYA. Wahlberg et al. (2009) {[}13{]} give the
+ages for origin and diversification for Ithomiini at 45 (39--53) and 37
+(32--43) MYA, respectively. Thus, current evidence shows that Solanaceae
+and Ithomiini might have diversified around the same time, during the
+Late Eocene and Oligocene, and this would be congruent with the diffuse
+cospeciation hypothesis.
+
+\paragraph{Danaini}\label{danaini}
+
+Our MultiMEDUSA approach showed a significant slowdown in net
 diversification rate in the subtribe Danaina of the Danini. Both Danaina
@@
 widespread species show higher genetic differentiation in their
-populations {[}51{]}.
-
-\subsubsection*{\emph{Charaxes}}
-
-The genus \emph{Charaxes} contain 193 species distributed in the Old
-World with highest diversity in the Afrotropical region. These
-butterflies are also very strong fliers, but contrary to Danaina, which
-are specialized Apocynaceae feeders, \emph{Charaxes} are known to feed
-on at least 28 plant families in 18 orders {[}29{]} and some species
-appear to be polyphagous {[}52{]}. Reference 53 showed that most of the
-diversification of the genus occurred from the late Oligocene to the
-Miocene when there were drastic global climatic fluctuations, indicating
-that the diversity mainly is driven by climate change. Reference 52
-found that climatic changes during the Pliocene, Pleistocene as well as
-dispersal and vicariance might have been responsible for the high
-diversification of the genus.
-
-\subsubsection*{Satyrini}
+populations {[}50{]}.
+
+\paragraph{Satyrini}\label{satyrini}
 
@@
 their main hostplant, grasses, during the climatic cooling in the
-Oligocene {[}11{]}. Thus, it is somewhat surprising that part of
+Oligocene {[}14{]}. Thus, it is somewhat surprising that part of
 Satyrini (the subtribes Euptychiina, Satyrina and Pronophilina) were
-found to have accelerated diversification in only 13\% of probabilities
-in the trees from the posterior distribution. Although this can be
-attributed to low phylogenetic signal {[}10{]}, the clade Satyrini is
-very robust {[}10{]} and MEDUSA failed to identify any significant
-accelerated net diversification rate for Satyrini. It appears that the
-radiation of Satyrini as a whole was not remarkably fast and therefore
-not picked up by MEDUSA.
-
-The origin of the tribe Satyrini is not completely clear (originated
-either in the Neotropical or Eastern Palaearctic, Oriental and/or
-Indo-Australian regions {[}27{]} and their radiation involved colonizing
-almost all continents starting from their place of origin. For
-butterflies, a prerequisite for colonizing new areas is that suitable
-hostplants are already present. Therefore, it should be expected that
-the diversification of Satyrinae occurred in a stepwise manner, with
-pulses or bursts of diversification for certain lineages but unlikely
-for the tribe Satyrini as a whole.
+found to have accelerated diversification in only 13\% of the trees from
+the posterior distribution. Although this can be attributed to low
+phylogenetic signal {[}13{]}, the clade Satyrini is very robust {[}13{]}
+and MEDUSA failed to identify any significant accelerated net
+diversification rate for Satyrini. It appears that the radiation of
+Satyrini as a whole was not remarkably fast and therefore not picked up
+by MEDUSA. This should be expected if the diversification of Satyrinae
+occurred in a stepwise manner, with pulses or bursts of diversification
+for certain lineages but unlikely for the tribe Satyrini as a whole.
+
+\subsection{Conclusions}\label{conclusions}
 
@@
 credibility tree. However, by using a MultiMEDUSA approach, we found
-that some of this splits might be greatly affected by phylogenetic
-uncertainty. Moreover, some of these splits can be recovered either as
-increases or decreases in net diversification rate depending on the tree
-from the posterior distribution that was used for analysis. This means
-that contradictory conclusions would be made if only the maximum clade
-credibility tree was used for analysis. MEDUSA appears to be sensitive
-to the number of nodes with high posterior probability and width of age
-confidence intervals. For our data, it would be necessary to obtain a
-posterior distribution of trees with no conflicting topology, and very
-similar estimated ages for nodes in order to consistently recover most
-of the diversification splits on the posterior distribution of trees
-that were inferred by MEDUSA on the MCC tree.
+that for this Nymphalidae dataset some of these splits might be greatly
+affected by phylogenetic uncertainty. Moreover, some of these splits can
+be recovered either as increases or decreases in net diversification
+rate depending on the tree from the posterior distribution that was used
+for analysis. This means that contradictory conclusions would be made if
+only the maximum clade credibility tree was used for analysis. We
+recommend that all datasets should be analyzed using both approaches,
+MEDUSA and MultiMEDUSA, in order to test whether the results are robust
+when phylogenetic uncertainty is taken into account.
+
+MEDUSA appears to be sensitive to the number of nodes with high
+posterior probability and width of age confidence intervals. For our
+data, it would be necessary to obtain a posterior distribution of trees
+with no conflicting topology, and very similar estimated ages for nodes
+in order to consistently recover most of the diversification splits on
+the posterior distribution of trees that were inferred by MEDUSA on the
+MCC tree.
 
@@
 
-\section*{Acknowledgments}
+\subsection{Acknowledgments}\label{acknowledgments}
 
@@
 data. The study was supported by a Kone Foundation grant (awarded to
-Niklas Wahlberg), Finland (C. Pe\~na) and the Research Council of Norway
+Niklas Wahlberg), Finland (C. Peña) and the Research Council of Norway
 (grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
@@
 
-\section*{Figure legends}
-
-\texttt{Figure 1.} MEDUSA analysis run on the MCT 
-from Reference 10. Estimated rate shifts in Table 1.
-
-\texttt{Figure 2.} Range of diversification values for
-tips or clades estimated by MEDUSA on the 1000 random trees.
-
-\texttt{Figure 3.} Results of the MultiMEDUSA analysis showing
-the probability for nodes of being recovered as diversification events.
-
-\texttt{Figure 4.} BiSSE analysis of diversification of Solanaceae feeding 
-nymphalids shows higher speciation and net diversification
-rates in Solanaceae feeders.
-
-\section*{References}
+\subsection{Figure legends}\label{figure-legends}
+
+\texttt{Figure 1.} Results of the MEDUSA analysis run on the maximum
+clade credibility tree from {[}13{]}. Rate shifts were estimated for the
+following nodes (besides the background rate): 2) Limenitidinae +
+Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
+part, 6) Satyrini, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
+part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
+\emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
+Satyrina, 14) Mycalesina, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
+\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae. Circles on
+nodes indicate the diversification shift number as found by MEDUSA.
+Numbers next to circles indicat the posterior probability values for
+such nodes.
+
+\texttt{Figure 2.} Diversification rates for taxa estimated by MEDUSA on
+the samples of 1000 random trees.
+
+\texttt{Figure 3.} Results of MultiMEDUSA analysis showing the
+probability of specific nodes being characterized by significant shifts
+in diversification rate.
+
+\texttt{Figure 4.} BiSSE analysis of diversification of nymphalids due
+to feeding on Solanaceae hostplants. Speciation and net diversification
+rates are significantly higher in Solanaceae feeders (speciation rate =
+\(\lambda1\), net diversification rate = \(r1\)).
+
+\texttt{Table 2.} Differences in rates estimated by MEDUSA on the MCC
+tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
+from the posterior distribution.
+
+\texttt{Table 3.} Likelihood ratio test between the model of increased
+diversification of nymphalids feeding on Solanaceae against a model
+forcing equal speciation rates (no effect on diversification).
+
+\section{Supp. mat.}\label{supp.-mat.}
+
+\texttt{Supporting\_Information\_S01.nex}: MCC Nymphalidae tree from
+{[}13{]}.
+
+\texttt{Supporting\_Information\_S02.R}: R script to run MEDUSA on the
+MCC Nymphalidae tree from {[}13{]}. This script removes the outgroup
+taxa and loads the richness data for the tree terminals.
+
+\texttt{Supporting\_Information\_S03.csv}: Species richness for lineages
+in Nymphalidae.
+
+\texttt{Supporting\_Information\_S04.txt}: Results of running MEDUSA on
+the MCC tree.
+
+\texttt{Supporting\_Information\_S05\_1000\_random\_trees\_no\_outgroups.zip}:
+1000 random trees from {[}13{]}.
+
+\texttt{Supporting\_Information\_S06.R}: R script to run a MultiMEDUSA
+analysis on 1000 random trees from {[}13{]}.
+
+\texttt{Supporting\_Information\_S07\_multimedusa\_on\_1000\_trees.zip}:
+Raw results from the MultiMEDUSA run on the random sample of trees from
+the posterior distribution.
+
+\texttt{Supporting\_Information\_S08.pdf}: Summary of the MulitMEDUSA
+analysis on the 1000 trees from the posterior distribution.
+
+\texttt{Supporting\_Information\_S09.tree}: MCC tree from the 1000
+random trees selected from the posterior distribution.
+
+\texttt{Supporting\_Information\_S10.R}: R script to run a MEDUSA
+analysis on the MCC tree from the 1000 random trees selected from the
+posterior distribution.
+
+\texttt{Supporting\_Information\_S11.txt}: Summary results from a MEDUSA
+run on the MCC tree from the 1000 random trees selected from the
+posterior distribution.
+
+\texttt{Supporting\_Information\_S12.pdf}: Figure for MEDUSA run on MCC
+tree from random 1000 trees.
+
+\texttt{Supporting\_Information\_S13.txt}: Summary results from a
+MultiMEDUSA run on the 1000 random trees selected from the posterior
+distribution.
+
+\texttt{Supporting\_Information\_S14.pdf}: Probability of nodes with
+estimated rates from a MultiMEDUSA run on the 1000 random trees selected
+from the posterior distribution.
+
+\texttt{Supporting\_Information\_S15.csv}: Hostplants of Nymphalidae
+butterflies recorded from the literature.
+
+\texttt{Supporting\_Information\_S16.csv}: References for hostplants
+data.
+
+\texttt{Supporting\_Information\_S17.csv}: Data matrix with character
+states for hosplant use.
+
+\texttt{Supporting\_Information\_S18.R}: R script for running the BiSSE
+analysis.
+
+\texttt{Supporting\_Information\_S19.csv}: Raw results for the BiSSE
+analysis.
+
+\texttt{Supporting\_Information\_S20.csv}: Raw results for the combined
+BiSSE analysis.
+
+\texttt{Figure S1.} Boxplot of speciation (\(\lambda\)) and extinction
+(\(\mu\)) values for the 95\% credibility intervals of values estimated
+by BiSSE analysis of diversification due to feeding on Solanaceae
+plants.
+
+\texttt{Figure S2.} BiSSE analysis of diversification of nymphalids due
+to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
+\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
+recovered, speciation and net diversification rates are significantly
+higher for Solanaceae feeders (\(\lambda1\), \(r1\)).
+
+\texttt{Figure S3.} Net diversification rates of nymphalids feeding on
+Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
+the 1000 trees from the posterior distribution.
+
+\texttt{Figure S4.} Boxplot of speciation and extinction values for the
+95\% credibility intervals of values estimated by BiSSE analysis of
+diversification due to feeding on Solanaceae plants on the combined
+post-burnin runs on 1000 trees from the posterior distribution.
+
+\texttt{Figure S5.} BiSSE analysis of diversification of nymphalids due
+to feeding on Apocynaceae hostplants. Speciation and net diversification
+rates are similar.
+
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
 1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
-adaptive zones: has phytophagy promoted insect diversification? Am Nat
+adaptive zones: Has phytophagy promoted insect diversification? Am Nat
 132: 107--128.
@@
 
-2. Nyman T, Linder HP, Pe\~na C, Malm T, Wahlberg N (2012) Climate-driven
+2. Nyman T, Linder HP, Peña C, Malm T, Wahlberg N (2012) Climate-driven
 diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
@@
 6. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
-al. (2011) Order Lepidoptera. In: Animal biodiversity: an outline of
+al. (2011) Order Lepidoptera. In: Animal biodiversity: An outline of
 higher-level classification and survey of taxonomic richness. Zootaxa.
@@
 
-7. Heikkil\"a M, Kaila L, Mutanen M, Pe\~na C, Wahlberg N (2012) Cretaceous
+7. Joron M, Jiggins C, Papanicolaou A, McMillan W (2006) Heliconius wing
+patterns: An evo-devo model for understanding phenotypic diversity.
+Heredity 97: 157--167.
+
+8. Willmott KR, Freitas AVL (2006) Higher-level phylogeny of the
+Ithomiinae (Lepidoptera: Nymphalidae): classification, patterns of
+larval hostplant colonization and diversification. Cladistics 22:
+297--368.
+doi:\href{http://dx.doi.org/10.1111/j.1096-0031.2006.00108.x}{10.1111/j.1096-0031.2006.00108.x}.
+
+9. Brakefield PM, Beldade P, Zwaan BJ (2009) The african butterfly
+bicyclus anynana: A model for evolutionary genetics and evolutionary
+developmental biology. Cold Spring Harbor Protocols 2009: pdb--emo122.
+
+10. Heikkilä M, Kaila L, Mutanen M, Peña C, Wahlberg N (2012) Cretaceous
 origin and repeated tertiary diversification of the redefined
@@
 
-8. Elias M, Joron M, Willmott KR, Silva-Brand\~ao KL, Kaiser V, et al.
+11. Elias M, Joron M, Willmott KR, Silva-Brandão KL, Kaiser V, et al.
 (2009) Out of the Andes: Patterns of diversification in clearwing
@@
 
-9. Fordyce JA (2010) Host shifts and evolutionary radiations of
+12. Fordyce JA (2010) Host shifts and evolutionary radiations of
 butterflies. P Roy Soc B 277: 3735--3743.
@@
 
-10. Wahlberg N, Leneveu J, Kodandaramaiah U, Pe\~na C, Nylin S, et al.
+13. Wahlberg N, Leneveu J, Kodandaramaiah U, Peña C, Nylin S, et al.
 (2009) Nymphalid butterflies diversify following near demise at the
@@
 
-11. Pe\~na C, Wahlberg N (2008) Prehistorical climate change increased
+14. Peña C, Wahlberg N (2008) Prehistorical climate change increased
 diversification of a group of butterflies. Biol Lett 4: 274--278.
@@
 
-12. Ferrer-Paris JR, S\'anchez-Mercado A, Viloria \'AL, Donaldson J (2013)
+15. Ferrer-Paris JR, S{á}nchez-Mercado A, Viloria ÁL, Donaldson J (2013)
 Congruence and Diversity of Butterfly-Host Plant Associations at Higher
 Taxonomic Levels. PloS one 8: e63570.
-doi:\href{http://dx.doi.org/10.1371/journal.pone.0063570}{10.1371/journal.pone.0063570}.
-
-13. Alfaro ME, Santini F, Brock C, Alamillo H, Dornburg A, et al. (2009)
+doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.
+
+16. Alfaro ME, Santini F, Brock C, Alamillo H, Dornburg A, et al. (2009)
 Nine exceptional radiations plus high turnover explain species diversity
@@
 
-14. Harmon LJ, Rabosky DL, FitzJohn RG, Brown JW (2011) turboMEDUSA:
+17. Harmon LJ, Rabosky DL, {FitzJohn} RG, Brown JW (2011) turboMEDUSA:
 MEDUSA: Modeling Evolutionary Diversification Using Stepwise AIC.
 
-15. FitzJohn RG (2012) Diversitree: comparative phylogenetic analyses of
-diversification in R. Method Ecol Evol 3: 1084--1092.
+18. {FitzJohn} RG (2012) Diversitree: comparative phylogenetic analyses
+of diversification in R. Method Ecol Evol 3: 1084--1092.
 doi:\href{http://dx.doi.org/10.1111/j.2041-210X.2012.00234.x}{10.1111/j.2041-210X.2012.00234.x}.
 
-16. Lamas G (2004) Checklist: Part 4A. Hesperioidea-Papilionoidea.
+19. Lamas G (2004) Checklist: Part 4A. Hesperioidea-Papilionoidea.
 Gainesville.
 
-17. Matos-Marav\'i PF, Pe\~na C, Willmott KR, Freitas AV, Wahlberg N (2013)
+20. Matos-Maraví PF, Peña C, Willmott KR, Freitas AV, Wahlberg N (2013)
 Systematics and evolutionary history of butterflies in the ``Taygetis
@@
 
-18. Brower AVZ, Wahlberg N, Ogawa JR, Boppr\'e M, Vane-Wright RI (2010)
+21. Brower AVZ, Wahlberg N, Ogawa JR, Boppré M, Vane-Wright RI (2010)
 Phylogenetic relationships among genera of danaine butterflies
@@
 
-19. Kodandaramaiah U, Lees DC, M\"uller CJ, Torres E, Karanth KP, et al.
+22. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
 (2010) Phylogenetics and biogeography of a spectacular Old World
@@
 
-20. Kodandaramaiah U, Pe\~na C, Braby MF, Grund R, M\"uller CJ, et al.
+23. Kodandaramaiah U, Peña C, Braby MF, Grund R, Müller CJ, et al.
 (2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
@@
 
-21. Ortiz-Acevedo E, Willmott KR (2013) Molecular systematics of the
+24. Ortiz-Acevedo E, Willmott KR (2013) Molecular systematics of the
 butterfly tribe Preponini (Nymphalidae: Charaxinae). Syst Entomol 38:
@@
 
-22. De-Silva DL, Day JJ, Elias M, Willmott KR, Whinnett A, et al. (2010)
+25. De-Silva DL, Day JJ, Elias M, Willmott KR, Whinnett A, et al. (2010)
 Molecular phylogenetics of the Neotropical butterfly subtribe Oleriina
@@
 
-23. Freitas AVL, Brown Jr. KS (2004) Phylogeny of the Nymphalidae
+26. Freitas AVL, Brown Jr. KS (2004) Phylogeny of the Nymphalidae
 (Lepidoptera). Syst Biol 53: 363--383.
@@
 
-24. Pe\~na C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
+27. Peña C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
 (2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
@@
 
-25. Penz CM (1999) Higher level phylogeny for the passion-vine
+28. Penz CM (1999) Higher level phylogeny for the passion-vine
 butterflies (Nymphalidae, Heliconiinae) based on early stage and adult
@@
 
-26. Silva-Brand\~ao KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
+29. Silva-Brandão KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
 Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
@@
 
-27. Pe\~na C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
+30. Peña C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
 butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
@@
 
-28. Pe\~na C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
+31. Peña C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
 history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
@@
 
-29. Ackery PR (1988) Hostplants and classification: a review of
+32. Ackery PR (1988) Hostplants and classification: A review of
 nymphalid butterflies. Biol J Linn Soc 33: 95--203.
@@
 
-30. Dyer LA, Gentry GL (2002) Caterpillars and parasitoids of a tropical
+33. Dyer LA, Gentry GL (2002) Caterpillars and parasitoids of a tropical
 lowland wet forest. Available: \url{http://www.caterpillars.org}.
@@
 
-31. Beccaloni G, Viloria \'A, Hall S, Robinson G (2008) Catalogue of the
+34. Beccaloni G, Viloria Á, Hall S, Robinson G (2008) Catalogue of the
 hostplants of the Neotropical butterflies. London: The Natural History
@@
 
-32. Janzen DH, Hallwachs W (2009) Dynamic database for an inventory of
+35. Janzen DH, Hallwachs W (2009) Dynamic database for an inventory of
 the macrocaterpillar fauna, and its food plants and parasitoids, of Area
@@
 
-33. Core Team R (2013) R: A Language and Environment for Statistical
+36. Core Team R (2013) R: A Language and Environment for Statistical
 Computing. Vienna, Austria: R Foundation for Statistical Computing.
@@
 
-34. Popescu A-A, Huber KT, Paradis E (2012) ape 3.0: New tools for
+37. Popescu A-A, Huber KT, Paradis E (2012) ape 3.0: New tools for
 distance-based phylogenetics and evolutionary analysis in R.
@@
 
-35. Harmon LJ, Weir JT, Brock CD, Glor RE, Challenger W (2008) GEIGER:
+38. Harmon LJ, Weir JT, Brock CD, Glor RE, Challenger W (2008) GEIGER:
 investigating evolutionary radiations. Bioinformatics 24: 129--131.
@@
 
-36. Maddison WP, Midford PE, Otto SP (2007) Estimating a binary
+39. Maddison WP, Midford PE, Otto SP (2007) Estimating a binary
 character's effect on speciation and extinction. Syst Biol 56: 701--710.
@@
 
-37. Willmott KR, Freitas AVL (2006) Higher-level phylogeny of the
-Ithomiinae (Lepidoptera: Nymphalidae): classification, patterns of
-larval hostplant colonization and diversification. Cladistics 22:
-297--368.
-doi:\href{http://dx.doi.org/10.1111/j.1096-0031.2006.00108.x}{10.1111/j.1096-0031.2006.00108.x}.
-
-38. Scott JA (1986) The Butterflies of North America. Stanford
+40. Scott JA (1986) The Butterflies of North America. Stanford
 University Press.
 
-39. Davis MP, Midford PE, Maddison W (2013) Exploring power and
+41. Davis MP, Midford PE, Maddison W (2013) Exploring power and
 parameter estimation of the BiSSE method for analyzing species
@@
 
-40. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
+42. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
 bees cut leaves? New insights into the early evolution of bees. P Roy
@@
 
-41. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
+43. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
 clades of Agaricales. P Roy Soc B 279: 2003--2011.
@@
 
-42. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
+44. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
 Interface. Ann Mo Bot Gard 74: 359--397.
 
-43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand\~ao KL, Whinnett A, et
+45. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brandão KL, Whinnett A, et
 al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
@@
 
-44. FitzJohn RG, Maddison WP, Otto SP (2009) Estimating trait-dependent
-speciation and extinction rates from incompletely resolved phylogenies.
-Syst Biol 58: 595--611. Available:
+46. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
+trait-dependent speciation and extinction rates from incompletely
+resolved phylogenies. Syst Biol 58: 595--611.
 doi:\href{http://dx.doi.org/10.1093/sysbio/syp067}{10.1093/sysbio/syp067}.
 
-45. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
+47. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
 The genetic basis of a plant-insect coevolutionary key innovation. P
@@
 
-46. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
+48. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
 furanocoumarin-metabolizing cytochrome P450 monooxygenases in two
@@
 
-47. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
+49. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
 the angiosperms re-revisited. Am J Bot 97: 1296--1303.
@@
 
-48. Somoza R (1998) Updated Nazca (Farallon) - South America relative
-motions during the last 40 My: implications for mountain building in the
-central Andean region. J S Am Earth Sci 11: 211--215.
-doi:\href{http://dx.doi.org/10.1016/S0895-9811(98)00012-1}{10.1016/S0895-9811(98)00012-1}.
-
-49. Hoorn C, Wesselingh FP, Steege H ter, Bermudez MA, Mora A, et al.
-(2010) Amazonia through time: Andean uplift, climate change, landscape
-evolution, and biodiversity. Science 330: 927--931.
-doi:\href{http://dx.doi.org/10.1126/science.1194585}{10.1126/science.1194585}.
-
-50. Olmstead RG (2013) Phylogeny and biogeography in Solanaceae,
-Verbenaceae and Bignoniaceae: a comparison of continental and
-intercontinental diversification patterns. Bot J Linn Soc 171: 80--102.
-doi:\href{http://dx.doi.org/10.1111/j.1095-8339.2012.01306.x}{10.1111/j.1095-8339.2012.01306.x}.
-
-51. Wahlberg N, Rubinoff D (2011) Vagility across \textit{Vanessa} (Lepidoptera:
+50. Wahlberg N, Rubinoff D (2011) Vagility across Vanessa (Lepidoptera:
 Nymphalidae): mobility in butterfly species does not inhibit the
@@
 
-52. M\"uller CJ, Wahlberg N, Beheregaray LB (2010) `After Africa': the
-evolutionary history and systematics of the genus \textit{Charaxes} Ochsenheimer
-(Lepidoptera: Nymphalidae) in the Indo-Pacific region. Biol J Linn Soc
-100: 457--481.
-doi:\href{http://dx.doi.org/10.1111/j.1095-8312.2010.01426.x}{10.1111/j.1095-8312.2010.01426.x}.
-
-53. Aduse-Poku K, Vingerhoedt E, Wahlberg N (2009) Out-of-Africa again:
-a phylogenetic hypothesis of the genus \textit{Charaxes} (Lepidoptera:
-Nymphalidae) based on 5 gene regions. Mol Phylogenet Evol 53: 463--478.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.06.021}{10.1016/j.ympev.2009.06.021}.
-
-\section*{Supporting Information}
-\texttt{Figure S1.} Boxplot of speciation (lambda) and extinction (mu)
-values for the 95\% credibility intervals of values estimated by BiSSE
-analysis of diversification due to feeding on Solanaceae plants.
-
-\texttt{Figure S2.} BiSSE analysis of diversification of nymphalids due
-to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
-\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
-recovered, speciation and net diversification rates are significantly
-higher for Solanaceae feeders (\$lambda\$1, r1).
-
-\texttt{Figure S3.} Net diversification rates of nymphalids feeding on
-Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
-the 1000 trees from the posterior distribution.
-
-\texttt{Figure S4.} Boxplot of speciation and extinction values for the
-95\% credibility intervals of values estimated by BiSSE analysis of
-diversification due to feeding on Solanaceae plants on the combined
-post-burnin runs on 1000 trees from the posterior distribution.
-
-\texttt{Figure S5.} BiSSE analysis of diversification of nymphalids due
-to feeding on Apocynaceae hostplants. Speciation and net diversification
-rates are similar.
-
-\textbf{\texttt{Supporting\_Information\_S1.nex}}: \textbf{Genus}. MCC Nymphalidae tree from
-Reference 10.
-
-\textbf{\texttt{Supporting\_Information\_S2.R}}: \textbf{Run medusa on mct}. Run MEDUSA on
-MCC tree from Reference 10.
-
-\textbf{\texttt{Supporting\_Information\_S3.csv}}: Species richness for
-lineages in Nymphalidae.
-
-\textbf{\texttt{Supporting\_Information\_S4.txt}}: Results of running
-MEDUSA on the MCC tree.
-
-\textbf{\texttt{Supporting\_Information\_S5.zip}}: 1000 random trees from Reference 10.
-
-\textbf{\texttt{Supporting\_Information\_S6.R}}: Run MultiMEDUSA on
-1000 random trees from Reference 10.
-
-\textbf{\texttt{Supporting\_Information\_S7.zip}}: Raw
-results from the MultiMEDUSA run on the random sample of trees from the
-posterior distribution.
-
-\textbf{\texttt{Supporting\_Information\_S8.pdf}}: Summary of
-MulitMEDUSA analysis on the 1000 trees from the posterior distribution.
-
-\textbf{\texttt{Supporting\_Information\_S9.tree}}:
-MCC tree from the 1000 random trees selected from the posterior
-distribution.
-
-\textbf{\texttt{Supporting\_Information\_S10.R}}:
-Run MEDUSA on MCC tree from the 1000 random trees selected from the
-posterior distribution.
-
-\textbf{\texttt{Supporting\_Information\_S11.txt}}:
-Summary results from a MEDUSA run on the MCC tree from the 1000 random
-trees selected from the posterior distribution.
-
-\textbf{\texttt{Supporting\_Information\_S12.pdf}}:
-Figure for MEDUSA run on MCC tree from random 1000 trees.
-
-\textbf{\texttt{Supporting\_Information\_S13.txt}}:
-Summary results from a MultiMEDUSA run on the 1000 random trees selected
-from the posterior distribution.
-
-\textbf{\texttt{Supporting\_Information\_S14.pdf}}:
-Probability of nodes with estimated rates from a MultiMEDUSA run on the
-1000 random trees selected from the posterior distribution.
-
-\textbf{\texttt{Supporting\_Information\_S15.csv}}: Hostplants of
-Nymphalidae butterflies recorded from the literature.
-
-\textbf{\texttt{Supporting\_Information\_S16.csv}}:
-References for hostplants data.
-
-\textbf{\texttt{Supporting\_Information\_S17.csv}}: Data matrix with
-character states for hosplant use.
-
-\textbf{\texttt{Supporting\_Information\_S18.R}}: R code for running the
-BiSSE analysis.
-
-\textbf{\texttt{Supporting\_Information\_S19.csv}}: Raw results for
-the BiSSE analysis.
-
-\textbf{\texttt{Supporting\_Information\_S20.csv}}: Raw results for
-the combined BiSSE analysis.
+\section*{Tables}
+% 
+% See introductory notes if you wish to include sideways tables.
+%
+% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements.
+Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
+%
+%\begin{table}[!ht]
+%\caption{
+%\bf{Table title}}
+%\begin{tabular}{|c|c|c|}
+%table information
+%\end{tabular}
+%\begin{flushleft}Table caption
+%\end{flushleft}
+%\label{tab:label}
+% \end{table}
+
+\begin{table}[!h]
+\caption{\bf{Significant net diversification rate shifts found in the MEDUSA analysis of the Nymphalid maximum clade credibility tree.}}
+\begin{tabular}{lllllll}
+Shift N$^\circ$ & Split.Node & Model & r          & LnLik.part & AICc     & Taxa                                                       \\
+1               & 399        & yule  & 0.092459   & -1055.957  &          & Nymphalidae (root)                                         \\
+2               & 691        & bd    & 0.054129   & -406.3703  &          & Limenitidinae + Heliconiinae                               \\
+3               & 299        & yule  & 0.311199   & -6.3058    &          & \emph{Ypthima}                                             \\
+4               & 224        & yule  & 0.290989   & -6.2601    &          & \emph{Charaxes}                                            \\
+5               & 750        & yule  & 0.186913   & -147.4146  &          & Oleriina + Ithomiina + Napeogenina + Dircennina + Godyrina \\
+6               & 405        & yule  & 0.116252   & -555.0276  &          & Satyrini                                                   \\
+7               & 495        & yule  & 0.064656   & -124.9143  &          & Coenonymphina                                              \\
+8               & 609        & yule  & 0.240562   & -35.0908   &          & Phyciodina in part                                         \\
+9               & 787        & bd    & 0.042332   & -43.7819   &          & Danaini in part                                            \\
+10              & 231        & yule  & 0.209416   & -4.7955    &          & \emph{Coenonympha}                                         \\
+11              & 478        & yule  & 0.311684   & -9.2218    &          & \emph{Caeruleuptychia} + \emph{Magneuptychia}              \\
+12              & 659        & yule  & 0.219253   & -11.2686   &          & \emph{Callicore} + \emph{Diaethria}                        \\
+13              & 444        & yule  & 0.220615   & -43.7812   &          & Satyrina                                                   \\
+14              & 524        & yule  & 0.190754   & -26.0651   &          & Mycalesina                                                 \\
+15              & 355        & yule  & 0.234041   & -6.2013    &          & \emph{Pedaliodes}                                          \\
+16              & 714        & yule  & 0          & 0          &          & \emph{Dryas} + \emph{Dryadula}                             \\
+17              & 377        & yule  & 0.311671   & -4.1986    &          & \emph{Taenaris}                                            \\
+18              & 688        & yule  & 0.024724   & -17.5256   &          & Pseudergolinae                                             \\
+19              & 583        & yule  & 0          & 0          & 5090.492 & \emph{Anaeomorpha} + \emph{Hypna}                                       
+\end{tabular}
+\begin{flushleft}Split.Node=node number, Model=preferred diversification model by MEDUSA, r=net diversification rate, LnLik.part=log likelihood value.
+\end{flushleft}
+\end{table}
 

### 2014-10-29 MS_revised.tex
table 2

@@
 
-\texttt{Table 2.} Differences in rates estimated by MEDUSA on the MCC
-tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
-from the posterior distribution.
-
 \texttt{Table 3.} Likelihood ratio test between the model of increased
@@
 \section*{Tables}
-% 
-% See introductory notes if you wish to include sideways tables.
-%
-% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements.
-Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
-%
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
 \begin{table}[!h]
 \caption{\bf{Significant net diversification rate shifts found in the MEDUSA analysis of the Nymphalid maximum clade credibility tree.}}
-\begin{tabular}{lllllll}
+\begin{tabular}{lccrrcl}
 Shift N$^\circ$ & Split.Node & Model & r          & LnLik.part & AICc     & Taxa                                                       \\
@@
 
+\begin{table}[!h]
+    \caption{\bf{Differences in rates estimated by MEDUSA on the MCC
+tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
+from the posterior distribution.}}
+\begin{tabular}{lcrr}
+Split.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
+1          & 0.090          & not found                  & 0.000                          \\
+2          & 0.054          & -0.030                     & 0.864                          \\
+3          & 0.311          & 0.219                      & 0.961                          \\
+4          & 0.290          & 0.212                      & 0.996                          \\
+5          & 0.187          & 0.101                      & 0.927                          \\
+6          & 0.116          & 0.039                      & 0.131                          \\
+7          & 0.064          & -0.052                     & 0.195                          \\
+8          & 0.240          & 0.166                      & 0.319                          \\
+9          & 0.042          & -0.049                     & 0.897                          \\
+10         & 0.209          & 0.149                      & 0.619                          \\
+11         & 0.311          & 0.208                      & 0.831                          \\
+12         & 0.219          & 0.135                      & 0.911                          \\
+13         & 0.220          & 0.117                      & 0.276                          \\
+14         & 0.190          & 0.115                      & 0.379                          \\
+15         & 0.234          & 0.127                      & 0.825                          \\
+16         & 0.000          & -0.005                     & 0.659                          \\
+17         & 0.311          & 0.248                      & 0.785                          \\
+18         & 0.024          & -0.064                     & 0.024                          \\
+19         & 0.000          & -0.087                     & 0.381                         
+\end{tabular}
+\end{table}
+% 
+% See introductory notes if you wish to include sideways tables.
+%
+% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements.
+Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
+%
+%\begin{table}[!ht]
+%\caption{
+%\bf{Table title}}
+%\begin{tabular}{|c|c|c|}
+%table information
+%\end{tabular}
+%\begin{flushleft}Table caption
+%\end{flushleft}
+%\label{tab:label}
+% \end{table}
+
 \end{document}

### 2014-10-29 MS_revised.tex
figures

@@
-\documentclass[]{article}
-\usepackage{setspace}
-\usepackage[]{lineno}
-\def\linenumberfont{\normalfont\small\sffamily}
-\doublespacing
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{tabu}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage[none]{hyphenat}
-\usepackage{fixltx2e} % provides \textsubscript
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-\usepackage[margin=1in,a4paper]{geometry}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={},
-            colorlinks=true,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\setcounter{secnumdepth}{0}
-
-\linenumbers
-
-\author{}
+\documentclass[10pt]{article}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{cite}
+\usepackage{hyperref}
+
+\usepackage{lineno}
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% Text layout
+\topmargin 0.0cm
+\oddsidemargin 0.5cm
+\evensidemargin 0.5cm
+\textwidth 16cm 
+\textheight 21cm
+
+% Bold the 'Figure #' in the caption and separate it with a period
+% Captions will be left justified
+\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2009}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
 \date{}
 
+\pagestyle{myheadings}
+
+%% Include all macros below. Please limit the use of macros.
+
+%% END MACROS SECTION
+
 \begin{document}
 
-\textbf{Title: Diversity dynamics in Nymphalidae butterflies: Effect of
-phylogenetic uncertainty on diversification rate shift estimates}
-
-\subsubsection{Authors: Carlos Peña\textsuperscript{1} and Marianne
-Espeland\textsuperscript{2}}\label{authors-carlos-peuxf1a1-and-marianne-espeland2}
-
-\subsubsection{Running title: Diversity dynamics in Nymphalidae
-butterflies}\label{running-title-diversity-dynamics-in-nymphalidae-butterflies}
-
-\emph{\textsuperscript{1} Laboratory of Genetics, Department of Biology,
-University of Turku, Turku, Finland. Email:
-\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}}
-
-\emph{\textsuperscript{2} Museum of Comparative Zoology and Department
+\begin{flushleft}
+{\Large
+    \title{Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates}
+}
+\maketitle
+Carlos Pe\~na$^{1,\ast}$,
+Marianne Espeland$^{2}$
+\\
+\bf{1} Carlos Pe\~na Laboratory of Genetics, Department of Biology, University of Turku, Turku, Finland. Email: \href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}
+\\
+\bf{2} Marianne Espeland Museum of Comparative Zoology and Department
 of Organismic and Evolutionary Biology, Harvard University, Cambridge,
@@
 Florida Museum of Natural History, University of Florida, Gainesville,
-FL, US}.
-
-\textbf{Corresponding author:}
+FL, US.
+\\
+$\ast$ \textbf{Corresponding author:}
 \href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}
@@
 
-\textbf{Comments:} 5814 words, 7 figures, 2 tables and 25 supplementary
-material files.
-
-\textbf{Additional Supplementary materials:}
-\url{http://dx.doi.org/10.6084/m9.figshare.639208}
-
-\subsection{Abstract}\label{abstract}
+\end{flushleft}
+
+\subsubsection*{Running title: Diversity dynamics in Nymphalidae
+butterflies}
+
+\subsection*{Abstract}
 
@@
 
-\subsection{Introduction}\label{introduction}
+\subsection*{Introduction}
 
@@
 
-\subsection{Methods}\label{methods}
-
-\subsubsection{Data}\label{data}
+\subsection*{Materialas and Methods}
+
+\subsubsection*{Data}
 
@@
 
-\subsubsection{Analyses of
-Diversification}\label{analyses-of-diversification}
+\subsubsection*{Analyses of
+Diversification}
 
@@
 
-\subsubsection{Detecting diversification shifts on phylogenetic
-trees}\label{detecting-diversification-shifts-on-phylogenetic-trees}
+\subsubsection*{Detecting diversification shifts on phylogenetic
+trees}
 
@@
 
-\subsubsection{Estimation of trait-dependent speciation
-rates}\label{estimation-of-trait-dependent-speciation-rates}
+\subsubsection*{Estimation of trait-dependent speciation
+rates}
 
@@
 
-\subsection{Results}\label{results}
-
-\subsubsection{Detecting diversification shifts on the maximum clade
+\subsection*{Results}
+
+\subsubsection*{Detecting diversification shifts on the maximum clade
 credibility
-tree}\label{detecting-diversification-shifts-on-the-maximum-clade-credibility-tree}
+tree}
 
@@
 
-\subsubsection{Phylogenetic uncertainty in the MultiMEDUSA
-approach}\label{phylogenetic-uncertainty-in-the-multimedusa-approach}
+\subsubsection*{Phylogenetic uncertainty in the MultiMEDUSA
+approach}
 
@@
 
-\subsubsection{Estimation of trait-dependent speciation
-rates}\label{estimation-of-trait-dependent-speciation-rates-1}
+\subsubsection*{Estimation of trait-dependent speciation
+rates}
 
@@
 
-\subsection{Discussion}\label{discussion}
-
-\subsubsection{Effects of phylogenetic uncertainty on the performance of
-MEDUSA}\label{effects-of-phylogenetic-uncertainty-on-the-performance-of-medusa}
+\subsection*{Discussion}
+
+\subsubsection*{Effects of phylogenetic uncertainty on the performance of
+MEDUSA}
 
@@
 
-\subsubsection{Hostplant use and diversification in
-Nymphalidae}\label{hostplant-use-and-diversification-in-nymphalidae}
-
-\paragraph{Ithomiini}\label{ithomiini}
+\subsubsection*{Hostplant use and diversification in
+Nymphalidae}
+
+\paragraph*{Ithomiini}
 
@@
 
-\paragraph{Danaini}\label{danaini}
-
+\paragraph*{Danaini}
 Our MultiMEDUSA approach showed a significant slowdown in net
@@
 
-\paragraph{Satyrini}\label{satyrini}
+\paragraph*{Satyrini}
 
@@
 
-\subsection{Conclusions}\label{conclusions}
-
+\subsection*{Conclusions}
 We found that even though MEDUSA estimated several diversification
@@
 strong support for other diversification splits in Nymphalidae. Our
-BiSSE analysis corroborated other studies in that the trait ``feeding on
-Solanaceae'', or a codistributed character, was important in the
+BiSSE analysis corroborated other studies in that the trait 
+``feeding on Solanaceae'', or a codistributed character, was important in the
 diversification of Ithomiini butterflies. However, by applying MEDUSA we
@@
 
-\subsection{Acknowledgments}\label{acknowledgments}
-
+\subsection*{Acknowledgments}
 We are thankful to Mark Cornwall for help with the script to extend
@@
 
-\subsection{Figure legends}\label{figure-legends}
-
-\texttt{Figure 1.} Results of the MEDUSA analysis run on the maximum
-clade credibility tree from {[}13{]}. Rate shifts were estimated for the
+\section*{Figure legends}
+\begin{description}
+ \item {\bf Figure 1. Results of the MEDUSA analysis run on the maximum
+clade credibility tree from Wahlberg et al. (2009)} Rate shifts were estimated for the
 following nodes (besides the background rate): 2) Limenitidinae +
@@
 
-\texttt{Figure 2.} Diversification rates for taxa estimated by MEDUSA on
-the samples of 1000 random trees.
-
-\texttt{Figure 3.} Results of MultiMEDUSA analysis showing the
+\item {\bf Figure 2. Diversification rates for taxa estimated by MEDUSA on
+the samples of 1000 random trees}
+
+\item {\bf Figure 3. Results of MultiMEDUSA analysis showing the
 probability of specific nodes being characterized by significant shifts
-in diversification rate.
-
-\texttt{Figure 4.} BiSSE analysis of diversification of nymphalids due
-to feeding on Solanaceae hostplants. Speciation and net diversification
+in diversification rate}
+
+\item {\bf Figure 4. BiSSE analysis of diversification of nymphalids due
+to feeding on Solanaceae hostplants} Speciation and net diversification
 rates are significantly higher in Solanaceae feeders (speciation rate =
-\(\lambda1\), net diversification rate = \(r1\)).
-
-\texttt{Table 3.} Likelihood ratio test between the model of increased
-diversification of nymphalids feeding on Solanaceae against a model
-forcing equal speciation rates (no effect on diversification).
-
-\section{Supp. mat.}\label{supp.-mat.}
-
-\texttt{Supporting\_Information\_S01.nex}: MCC Nymphalidae tree from
-{[}13{]}.
-
-\texttt{Supporting\_Information\_S02.R}: R script to run MEDUSA on the
+$\lambda1$, net diversification rate = $r1$.
+\end{description}
+
+\section*{Supporting Information Legends}
+\begin{description}
+ \item {\bf Supporting\_Information\_S01.nex.} MCC Nymphalidae tree from
+{[}13{]}
+\item {\bf Supporting\_Information\_S02.R} R script to run MEDUSA on the
 MCC Nymphalidae tree from {[}13{]}. This script removes the outgroup
 taxa and loads the richness data for the tree terminals.
+\end{description}
 
@@
 
-\texttt{Figure S1.} Boxplot of speciation (\(\lambda\)) and extinction
+\texttt{Figure S1.} Boxplot of speciation ($\lambda$) and extinction
 (\(\mu\)) values for the 95\% credibility intervals of values estimated
@@
 recovered, speciation and net diversification rates are significantly
-higher for Solanaceae feeders (\(\lambda1\), \(r1\)).
+higher for Solanaceae feeders ($\lambda1$), r1).
 
@@
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
-
+\section*{References}
 1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
@@
 
-14. Peña C, Wahlberg N (2008) Prehistorical climate change increased
+14. Pe\~na C, Wahlberg N (2008) Prehistorical climate change increased
 diversification of a group of butterflies. Biol Lett 4: 274--278.
@@
 
+% Table 1
 \section*{Tables}
@@
 
+% Table 2
 \begin{table}[!h]
@@
 \end{table}
-% 
-% See introductory notes if you wish to include sideways tables.
-%
-% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements.
-Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
-%
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
+
+% Table 3
+\begin{table}[!h]
+    \caption{\bf{Likelihood ratio test between the model of increased
+diversification of nymphalids feeding on Solanaceae against a model
+forcing equal speciation rates (no effect on diversification).}}
+\begin{tabular}{lcrrlr}
+Df           & lnLik & AIC     & ChiSq  & \multicolumn{1}{r}{p} &         \\
+full         & 6     & -1613.3 & 3238.5 &                       &         \\
+equal.lambda & 5     & -1619.4 & 3248.9 & 12.3                  & 0.00045
+\end{tabular}
+\end{table}
 

### 2014-10-29 MS_revised.tex
supp mat

@@
 {[}13{]}
+
  \item {\bf Supporting\_Information\_S02.R} R script to run MEDUSA on the
@@
 taxa and loads the richness data for the tree terminals.
-\end{description}
-
-\texttt{Supporting\_Information\_S03.csv}: Species richness for lineages
+
+ \item {\bf Supporting\_Information\_S03.csv} Species richness for lineages
 in Nymphalidae.
 
-\texttt{Supporting\_Information\_S04.txt}: Results of running MEDUSA on
+ \item {\bf Supporting\_Information\_S04.txt} Results of running MEDUSA on
 the MCC tree.
 
-\texttt{Supporting\_Information\_S05\_1000\_random\_trees\_no\_outgroups.zip}:
-1000 random trees from {[}13{]}.
-
-\texttt{Supporting\_Information\_S06.R}: R script to run a MultiMEDUSA
+ \item {\bf Supporting\_Information\_S05\_1000\_random\_trees\_no\_outgroups.zip} 1000 random trees from {[}13{]}.
+
+ \item {\bf Supporting\_Information\_S06.R} R script to run a MultiMEDUSA
 analysis on 1000 random trees from {[}13{]}.
 
-\texttt{Supporting\_Information\_S07\_multimedusa\_on\_1000\_trees.zip}:
-Raw results from the MultiMEDUSA run on the random sample of trees from
+ \item {\bf Supporting\_Information\_S07\_multimedusa\_on\_1000\_trees.zip} Raw results from the MultiMEDUSA run on the random sample of trees from
 the posterior distribution.
 
-\texttt{Supporting\_Information\_S08.pdf}: Summary of the MulitMEDUSA
+ \item {\bf Supporting\_Information\_S08.pdf} Summary of the MulitMEDUSA
 analysis on the 1000 trees from the posterior distribution.
 
-\texttt{Supporting\_Information\_S09.tree}: MCC tree from the 1000
+ \item {\bf Supporting\_Information\_S09.tree} MCC tree from the 1000
 random trees selected from the posterior distribution.
 
-\texttt{Supporting\_Information\_S10.R}: R script to run a MEDUSA
+ \item {\bf Supporting\_Information\_S10.R} R script to run a MEDUSA
 analysis on the MCC tree from the 1000 random trees selected from the
@@
 
-\texttt{Supporting\_Information\_S11.txt}: Summary results from a MEDUSA
+ \item {\bf Supporting\_Information\_S11.txt} Summary results from a MEDUSA
 run on the MCC tree from the 1000 random trees selected from the
@@
 
-\texttt{Supporting\_Information\_S12.pdf}: Figure for MEDUSA run on MCC
+ \item {\bf Supporting\_Information\_S12.pdf} Figure for MEDUSA run on MCC
 tree from random 1000 trees.
 
-\texttt{Supporting\_Information\_S13.txt}: Summary results from a
+ \item {\bf Supporting\_Information\_S13.txt} Summary results from a
 MultiMEDUSA run on the 1000 random trees selected from the posterior
@@
 
-\texttt{Supporting\_Information\_S14.pdf}: Probability of nodes with
+ \item {\bf Supporting\_Information\_S14.pdf} Probability of nodes with
 estimated rates from a MultiMEDUSA run on the 1000 random trees selected
@@
 
-\texttt{Supporting\_Information\_S15.csv}: Hostplants of Nymphalidae
+ \item {\bf Supporting\_Information\_S15.csv} Hostplants of Nymphalidae
 butterflies recorded from the literature.
 
-\texttt{Supporting\_Information\_S16.csv}: References for hostplants
+ \item {\bf Supporting\_Information\_S16.csv} References for hostplants
 data.
 
-\texttt{Supporting\_Information\_S17.csv}: Data matrix with character
+ \item {\bf Supporting\_Information\_S17.csv} Data matrix with character
 states for hosplant use.
 
-\texttt{Supporting\_Information\_S18.R}: R script for running the BiSSE
+ \item {\bf Supporting\_Information\_S18.R} R script for running the BiSSE
 analysis.
 
-\texttt{Supporting\_Information\_S19.csv}: Raw results for the BiSSE
+ \item {\bf Supporting\_Information\_S19.csv} Raw results for the BiSSE
 analysis.
 
-\texttt{Supporting\_Information\_S20.csv}: Raw results for the combined
+ \item {\bf Supporting\_Information\_S20.csv} Raw results for the combined
 BiSSE analysis.
 
-\texttt{Figure S1.} Boxplot of speciation ($\lambda$) and extinction
-(\(\mu\)) values for the 95\% credibility intervals of values estimated
+ \item {\bf Figure S1.} Boxplot of speciation ($\lambda$) and extinction
+($\mu$) values for the 95\% credibility intervals of values estimated
 by BiSSE analysis of diversification due to feeding on Solanaceae
@@
 
-\texttt{Figure S2.} BiSSE analysis of diversification of nymphalids due
+ \item {\bf Figure S2.} BiSSE analysis of diversification of nymphalids due
 to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
@@
 recovered, speciation and net diversification rates are significantly
-higher for Solanaceae feeders ($\lambda1$), r1).
-
-\texttt{Figure S3.} Net diversification rates of nymphalids feeding on
+higher for Solanaceae feeders $\lambda1$, r1).
+
+ \item {\bf Figure S3.} Net diversification rates of nymphalids feeding on
 Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
@@
 
-\texttt{Figure S4.} Boxplot of speciation and extinction values for the
+ \item {\bf Figure S4.} Boxplot of speciation and extinction values for the
 95\% credibility intervals of values estimated by BiSSE analysis of
@@
 
-\texttt{Figure S5.} BiSSE analysis of diversification of nymphalids due
+ \item {\bf Figure S5.} BiSSE analysis of diversification of nymphalids due
 to feeding on Apocynaceae hostplants. Speciation and net diversification
 rates are similar.
+\end{description}
 

### 2014-10-29 MS_orig.tex
work on diff

--- 2014-07-23 MS.tex
@@
-\documentclass[10pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-\usepackage{hyperref}
-
-% line numbers
-\usepackage{lineno}
-
-% ligatures disabled
-\usepackage{microtype}
-\DisableLigatures[f]{encoding = *, family = * }
-
-% rotating package for sideways tables
-%\usepackage{rotating}
-
-% If you wish to include algorithms, please use one of the packages below.
-Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
-%\usepackage{algorithmic}
-%\usepackage{algorithmicx}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
+\documentclass[]{article}
+\usepackage{setspace}
+\usepackage[]{lineno}
+\def\linenumberfont{\normalfont\small\sffamily}
+\doublespacing
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage[none]{hyphenat}
+\usepackage{fixltx2e} % provides \textsubscript
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+\usepackage[margin=1in,a4paper]{geometry}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={},
+            colorlinks=true,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
+
+\linenumbers
 
@@
 \date{}
-\pagestyle{myheadings}
 
 \begin{document}
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{Diversity dynamics in Nymphalidae butterflies: Effect of
+
+\textbf{Title: Diversity dynamics in Nymphalidae butterflies: Effect of
 phylogenetic uncertainty on diversification rate shift estimates}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Carlos Pe\~na$^{1,\ast}$, 
-Marianne Espeland$^{2}$
-\\
-\bf{1} Carlos Pe\~na Laboratory of Genetics, Department of Biology, University of Turku, Turku
-\\
-\bf{2} Museum of Comparative Zoology and Department of Organismic and
+
+\subsubsection{Authors: Carlos Peña\textsuperscript{1,*} and Marianne
+Espeland\textsuperscript{2,*}}\label{authors-carlos-peuxf1a1-and-marianne-espeland2}
+
+\subsubsection{Running title: Diversity dynamics in Nymphalidae
+butterflies}\label{running-title-diversity-dynamics-in-nymphalidae-butterflies}
+
+\emph{1 Laboratory of Genetics, Department of Biology, University of
+Turku, Turku, Finland. Email:
+\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}}
+
+\emph{2 Museum of Comparative Zoology and Department of Organismic and
 Evolutionary Biology, Harvard University, Cambridge, USA. Email:
-marianne.espeland@gmail.com.
+\href{mailto:marianne.espeland@gmail.com}{\nolinkurl{marianne.espeland@gmail.com}}.
 Current address: McGuire Center for Lepidoptera and Biodiversity,
 Florida Museum of Natural History, University of Florida, Gainesville,
-FL, US
-\\
-$\ast$ E-mail: Corresponding mycalesis@gmail.com
-\end{flushleft}
-
-\section*{Abstract}
+FL, US}.
+
+\textbf{Corresponding author:}
+\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}
+Telephone: +358 417063065 Fax: +358 2 333 6680
+
+\textbf{Comments:} 5814 words, 7 figures, 2 tables and 25 supplementary
+material files.
+
+\textbf{Additional Supplementary materials:}
+\url{http://dx.doi.org/10.6084/m9.figshare.639208}
+
+\subsection{Abstract}\label{abstract}
 
@@
 
-\section*{Materials and Methods}
-
-\subsection*{Data}
+\subsection{Methods}\label{methods}
+
+\subsubsection{Data}\label{data}
 
@@
 
-\subsection*{Analyses of Diversification}
+\subsubsection{Analyses of
+Diversification}\label{analyses-of-diversification}
 
@@
 
-\subsection*{Detecting diversification shifts on phylogenetic trees}
+\subsubsection{Detecting diversification shifts on phylogenetic
+trees}\label{detecting-diversification-shifts-on-phylogenetic-trees}
 
@@
 net diversification \texttt{=} speciation (\texttt{b}) \texttt{-}
-extinction (\texttt{d}) and $\varepsilon$: relative extinction
+extinction (\texttt{d}) and \(\varepsilon\): relative extinction
 \texttt{= d/b}. The AIC scores of the two-parameter model are then
@@
 trees. We also expected to find similar values of net diversification
-(\texttt{r = b - d}) and relative extinction rate ($\varepsilon$
+(\texttt{r = b - d}) and relative extinction rate (\(\varepsilon\)
 \texttt{= d/b}) values across the 1000 trees for the nodes where changes
@@
 
-\subsection*{Estimation of trait-dependent speciation rates}
+\subsubsection{Estimation of trait-dependent speciation
+rates}\label{estimation-of-trait-dependent-speciation-rates}
 
@@
 
-\section*{Results}
-
-\subsection*{Detecting diversification shifts on the maximum clade credibility tree}
+\subsection{Results}\label{results}
+
+\subsubsection{Detecting diversification shifts on the maximum clade
+credibility
+tree}\label{detecting-diversification-shifts-on-the-maximum-clade-credibility-tree}
 
@@
 
-\subsection*{Phylogenetic uncertainty in the MultiMEDUSA approach}
+\subsubsection{Phylogenetic uncertainty in the MultiMEDUSA
+approach}\label{phylogenetic-uncertainty-in-the-multimedusa-approach}
 
@@
 
-\subsection*{Estimation of trait-dependent speciation rates}
+\subsubsection{Estimation of trait-dependent speciation
+rates}\label{estimation-of-trait-dependent-speciation-rates-1}
 
@@
 
-\section*{Discussion}
-
-\subsection*{Effects of phylogenetic uncertainty on the performance of MEDUSA}
+\subsection{Discussion}\label{discussion}
+
+\subsubsection{Effects of phylogenetic uncertainty on the performance of
+MEDUSA}\label{effects-of-phylogenetic-uncertainty-on-the-performance-of-medusa}
 
@@
 
-\subsection*{Hostplant use and diversification in Nymphalidae}
-
-\subsubsection*{Ithomiini}
+\subsubsection{Hostplant use and diversification in
+Nymphalidae}\label{hostplant-use-and-diversification-in-nymphalidae}
+
+\paragraph{Ithomiini}\label{ithomiini}
 
@@
 
-\subsubsection*{Danaini}
+\paragraph{Danaini}\label{danaini}
 
@@
 
-\subsubsection*{\emph{Charaxes}}
+\paragraph{\emph{Charaxes}}\label{charaxes}
 
@@
 
-\subsubsection*{Satyrini}
+\paragraph{Satyrini}\label{satyrini}
 
@@
 
-\section*{Figure legends}
-
-\texttt{Figure 1.} MEDUSA analysis run on the MCT 
-from Reference 10. Estimated rate shifts in Table 1.
-
-\texttt{Figure 2.} Range of diversification values for
-tips or clades estimated by MEDUSA on the 1000 random trees.
-
-\texttt{Figure 3.} Results of the MultiMEDUSA analysis showing
-the probability for nodes of being recovered as diversification events.
-
-\texttt{Figure 4.} BiSSE analysis of diversification of Solanaceae feeding 
-nymphalids shows higher speciation and net diversification
-rates in Solanaceae feeders.
-
-\section*{Supp. mat.}
+\subsection{Figure legends}\label{figure-legends}
+
+\texttt{Figure 1.} Results of the MEDUSA analysis run on the maximum
+clade credibility tree from Reference 10. Rate shifts were estimated for
+the following nodes (besides the background rate): 2) Limenitidinae +
+Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
+part, 6) Satyrini, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
+part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
+\emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
+Satyrina, 14) Mycalesinaa, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
+\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae.
+
+\texttt{Figure 2.} Boxplot of the range of diversification values for
+tips or clades estimated by MEDUSA on the 1000 random trees from the
+posterior distribution of the Nymphalidae phylogeny. The tips or clades
+shown are those present on the maximum clade credibility tree from
+Reference 10.
+
+\texttt{Figure 3.} Results of the MultiMEDUSA analysis on 1000 randomly
+sampled trees from the posterior distribution of the Nymphalidae
+phylogeny. Bars show the probability for nodes of being recovered as
+significant increases or decreases in net diversification rates by
+MultiMEDUSA.
+
+\texttt{Figure 4.} BiSSE analysis of diversification of nymphalids due
+to feeding on Solanaceae hostplants. Speciation and net diversification
+rates are significantly higher in Solanaceae feeders (\$lambda\$1, r1).
+
+\subsection{Tables}\label{tables}
+
+\texttt{Table 1.} Significant net diversification rate shifts found in
+the MEDUSA analysis of the Nymphalid maximum clade credibility tree.
+Split.Node = node number, Model = preferred diversification model by
+MEDUSA, r = net diversification rate, LnLik.part = log likelihood value.
+
+\texttt{Table 2.} Differences in rates estimated by MEDUSA on the MCC
+tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
+from the posterior distribution.
+
+\texttt{Table 3.} Likelihood ratio test between the model of increased
+diversification of nymphalids feeding on Solanaceae against a model
+forcing equal speciation rates (no effect on diversification).
+
+\section{Supp. mat.}\label{supp.-mat.}
 
@@
 
-\section*{References}
+\subsection*{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
+
 1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
-adaptive zones: has phytophagy promoted insect diversification? Am Nat
+adaptive zones: Has phytophagy promoted insect diversification? Am Nat
 132: 107--128.
@@
 
-7. Heikkil\"a M, Kaila L, Mutanen M, Pe\~na C, Wahlberg N (2012) Cretaceous
+7. Heikkilä M, Kaila L, Mutanen M, Peña C, Wahlberg N (2012) Cretaceous
 origin and repeated tertiary diversification of the redefined
@@
 
-8. Elias M, Joron M, Willmott KR, Silva-Brand\~ao KL, Kaiser V, et al.
+8. Elias M, Joron M, Willmott KR, Silva-Brandão KL, Kaiser V, et al.
 (2009) Out of the Andes: Patterns of diversification in clearwing
@@
 
-10. Wahlberg N, Leneveu J, Kodandaramaiah U, Pe\~na C, Nylin S, et al.
+10. Wahlberg N, Leneveu J, Kodandaramaiah U, Peña C, Nylin S, et al.
 (2009) Nymphalid butterflies diversify following near demise at the
@@
 
-11. Pe\~na C, Wahlberg N (2008) Prehistorical climate change increased
+11. Peña C, Wahlberg N (2008) Prehistorical climate change increased
 diversification of a group of butterflies. Biol Lett 4: 274--278.
@@
 
-12. Ferrer-Paris JR, S\'anchez-Mercado A, Viloria \'AL, Donaldson J (2013)
+12. Ferrer-Paris JR, S{á}nchez-Mercado A, Viloria ÁL, Donaldson J (2013)
 Congruence and Diversity of Butterfly-Host Plant Associations at Higher
 Taxonomic Levels. PloS one 8: e63570.
-doi:\href{http://dx.doi.org/10.1371/journal.pone.0063570}{10.1371/journal.pone.0063570}.
+doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.
 
@@
 
-14. Harmon LJ, Rabosky DL, FitzJohn RG, Brown JW (2011) turboMEDUSA:
+14. Harmon LJ, Rabosky DL, {FitzJohn} RG, Brown JW (2011) turboMEDUSA:
 MEDUSA: Modeling Evolutionary Diversification Using Stepwise AIC.
 
-15. FitzJohn RG (2012) Diversitree: comparative phylogenetic analyses of
-diversification in R. Method Ecol Evol 3: 1084--1092.
+15. {FitzJohn} RG (2012) Diversitree: comparative phylogenetic analyses
+of diversification in R. Method Ecol Evol 3: 1084--1092.
 doi:\href{http://dx.doi.org/10.1111/j.2041-210X.2012.00234.x}{10.1111/j.2041-210X.2012.00234.x}.
@@
 
-17. Matos-Marav\'i PF, Pe\~na C, Willmott KR, Freitas AV, Wahlberg N (2013)
+17. Matos-Maraví PF, Peña C, Willmott KR, Freitas AV, Wahlberg N (2013)
 Systematics and evolutionary history of butterflies in the ``Taygetis
@@
 
-18. Brower AVZ, Wahlberg N, Ogawa JR, Boppr\'e M, Vane-Wright RI (2010)
+18. Brower AVZ, Wahlberg N, Ogawa JR, Boppré M, Vane-Wright RI (2010)
 Phylogenetic relationships among genera of danaine butterflies
@@
 
-19. Kodandaramaiah U, Lees DC, M\"uller CJ, Torres E, Karanth KP, et al.
+19. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
 (2010) Phylogenetics and biogeography of a spectacular Old World
@@
 
-20. Kodandaramaiah U, Pe\~na C, Braby MF, Grund R, M\"uller CJ, et al.
+20. Kodandaramaiah U, Peña C, Braby MF, Grund R, Müller CJ, et al.
 (2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
@@
 
-24. Pe\~na C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
+24. Peña C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
 (2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
@@
 
-26. Silva-Brand\~ao KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
+26. Silva-Brandão KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
 Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
@@
 
-27. Pe\~na C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
+27. Peña C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
 butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
@@
 
-28. Pe\~na C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
+28. Peña C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
 history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
@@
 
-29. Ackery PR (1988) Hostplants and classification: a review of
+29. Ackery PR (1988) Hostplants and classification: A review of
 nymphalid butterflies. Biol J Linn Soc 33: 95--203.
@@
 
-31. Beccaloni G, Viloria \'A, Hall S, Robinson G (2008) Catalogue of the
+31. Beccaloni G, Viloria Á, Hall S, Robinson G (2008) Catalogue of the
 hostplants of the Neotropical butterflies. London: The Natural History
@@
 
-43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand\~ao KL, Whinnett A, et
+43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brandão KL, Whinnett A, et
 al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
@@
 
-44. FitzJohn RG, Maddison WP, Otto SP (2009) Estimating trait-dependent
-speciation and extinction rates from incompletely resolved phylogenies.
-Syst Biol 58: 595--611. Available:
+44. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
+trait-dependent speciation and extinction rates from incompletely
+resolved phylogenies. Syst Biol 58: 595--611.
 doi:\href{http://dx.doi.org/10.1093/sysbio/syp067}{10.1093/sysbio/syp067}.
@@
 
-51. Wahlberg N, Rubinoff D (2011) Vagility across \textit{Vanessa} (Lepidoptera:
+51. Wahlberg N, Rubinoff D (2011) Vagility across Vanessa (Lepidoptera:
 Nymphalidae): mobility in butterfly species does not inhibit the
@@
 
-52. M\"uller CJ, Wahlberg N, Beheregaray LB (2010) `After Africa': the
-evolutionary history and systematics of the genus \textit{Charaxes} Ochsenheimer
+52. Müller CJ, Wahlberg N, Beheregaray LB (2010) `After Africa': the
+evolutionary history and systematics of the genus Charaxes Ochsenheimer
 (Lepidoptera: Nymphalidae) in the Indo-Pacific region. Biol J Linn Soc
@@
 53. Aduse-Poku K, Vingerhoedt E, Wahlberg N (2009) Out-of-Africa again:
-a phylogenetic hypothesis of the genus \textit{Charaxes} (Lepidoptera:
+a phylogenetic hypothesis of the genus Charaxes (Lepidoptera:
 Nymphalidae) based on 5 gene regions. Mol Phylogenet Evol 53: 463--478.

### 2014-10-29 response_to_reviewers.tex
work on diff

@@
 \usepackage[margin=1in,a4paper]{geometry}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \ifxetex
@@
 
-\subsection{Reviewer 1}\label{reviewer-1}
+\section{Reviewer 1}\label{reviewer-1}
 
@@
 
+\textbf{2. Response:}
+
+\begin{quote}
+\color{blue}
+We have specified that the sample of trees was taken from the Bayesian
+run in Wahlberg et al., 2009.
+\end{quote}
+
+\textbf{3. Reviewer's Comment:}
+
+Middle of paragraph (and throughout the manuscript): revise the tree
+sampling to ``\ldots{}therefore randomly sampled 1000 trees from the
+posterior distribution''; in the remainder of the manuscript, refer to
+these as the ``1000 sampled trees'', not ``1000 random trees'', which
+can take on a very different meaning.
+
+\textbf{3. Response:}
+
+\begin{quote}
+\color{blue}
+We have corrected some sentences in the manuscript to specify that we
+are referring to the same set of 1000 trees.
+\end{quote}
+
+\textbf{4. Reviewer's Comment:}
+
+Define what `MuSSE' refers to (Multiple State Speciation and
+Extinction?).
+
+\textbf{4. Response:}
+
+\begin{quote}
+\color{blue}
+We have included the definition of MuSSE in the manuscript.
+\end{quote}
+
+\textbf{5. Reviewer's Comment:}
+
+A reference for the statement ``most of Nymphalidae butterflies are
+restricted to use one plant family as hostplant'' is warranted.
+
+\textbf{5. Response:}
+
+\begin{quote}
+\color{blue}
+We have included the numbers of single-family feeders based on
+calculations of our own data on hostplant use across Nymphalidae genera.
+The text has been modified accordingly:
+\end{quote}
+
+\begin{quote}
+\color{blue}
+``Because most of Nymphalidae butterflies are restricted to use one
+plant family as hostplant (232 genera use only one plant family versus
+176 that use more than one; based on our data in supp. mat. 15), the
+\ldots{}''
+\end{quote}
+
+\textbf{6. Reviewer's Comment:}
+
+At the end of this paragraph, the authors indicated they tested the
+effect of character coding on their analyses. A more explicit
+description of the test would be useful, along with predictions of how
+results would be effected (i.e.~how would a large effect vs.~no effect
+of coding Hypanartia and Vanessa as Solanaceae feeders manifest in the
+results?).
+
+\textbf{6. Response:}
+
+\begin{quote}
+\color{blue}
+We included the following sentence to clarify the issue:\\\textbf{``We
+expected that coding these two genera as either absence or presence
+would not have a significant effect on our results.''}
+\end{quote}
+
+\textbf{7, 8, 9, 10. Reviewer's Comment:}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{6}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  First paragraph, second sentence: change the tense of ``run'' to
+  ``ran''.
+\item
+  Second paragraph, third sentence: replace ``Thus'' with
+  ``Furthermore''.
+\item
+  There appears to be an orphaned right-parenthesis at the end of the
+  second paragraph.
+\item
+  Third paragraph, last sentence: The node number is not really
+  relevant, so beginning of the sentence could be revised to ``For
+  example, the Charaxes + Polyura clade was found in only 256
+  trees\ldots{}''
+\end{enumerate}
+
+\textbf{7, 8, 9, 10. Response:}
+
+\begin{quote}
+\color{blue}
+We made the suggested corrections.
+\end{quote}
+
+\textbf{11. Reviewer's Comment:}
+
+Fourth paragraph: The point of the paragraph is not clear. What rates
+are being compared? MEDUSA vs.~MultiMEDUSA? Consider revising to make
+the point clearer or omit the sentence.
+
+\textbf{11. Response:}
+
+\begin{quote}
+\color{blue}
+We have rewritten the sentence and now is more clear:\\\textbf{``For the
+diversification shifts found in both the MCC tree and most of the
+samples of 1000 trees (frequency more than 90\%; Table 2), the
+MultiMEDUSA approach recovered different rates of diversification than
+those found using when the MCC tree alone.''}.
+\end{quote}
+
+\textbf{12. Reviewer's Comment:}
+
+Fifth paragraph, first sentence (and potentially elsewhere): Consider
+using a consistent means of presenting probability values. In this
+sentence, the authors refer to ``90\% of probability in the trees'', but
+in the preceding paragraph, the authors present probabilities as a
+decimal (``probability of being recovered higher than 0.90''). Granted,
+these two probabilities refer to different events, but perhaps it would
+be best to refer to them in a consistent manner, such as ``0.90
+posterior probability''.
+
+\textbf{12. Response:}
+
+\begin{quote}
+\color{blue}
+We have corrected this and normalized the probability values to decimals
+in the manuscript.
+\end{quote}
+
+\textbf{13. Reviewer's Comment:} The third sentence (length of analysis
+and burnin details) should be moved to the methods section.
+
+\textbf{13. Response:}
+
+\begin{quote}
+\color{blue}
+We made the suggested corrections.
+\end{quote}
+
+\textbf{14. Reviewer's Comment:}
+
+Mid-way through paragraph, discussion of likelihood ratio test: The
+source of this P-value needs to be described (chi-squared approximation
+with 1 degree of freedom?), here or in the methods section (see p.~617
+in Pagel 1999 Syst. Biol. 48:612-622).
+
+\textbf{14. Response:}
+
+\begin{quote}
+\color{blue}
+We have include the statistics suggested by the reviewer:
+``\textbf{\(\chi^2 = 12.3; 1 df; p < 0.001\)}''.
+\end{quote}
+
+\textbf{15. Reviewer's Comment:}
+
+The closing three sentences (about BiSSE performance) should be moved to
+the methods section.
+
+\textbf{15. Response:}
+
+\begin{quote}
+\color{blue}
+We made the suggested corrections.
+\end{quote}
+
+\textbf{16. Reviewer's Comment:}
+
+First paragraph: The authors' assessment of analyzing diversification
+shifts using a single tree, without accommodating our level of certainty
+in relationships, is justified; however, MEDUSA itself, that is, the
+approach using stepwise AIC to identify shifts in diversification rate,
+should not be the target of criticism. After all, MultiMEDUSA is doing
+the same thing, but on a sample of trees. Revise this paragraph with a
+more appropriate target for the criticism (i.e.~analyses performed on
+single trees).
+
+\textbf{16. Response:}
+
+\begin{quote}
+\color{blue}
+We have reworded the paragraph to avoid blaming the method:
+\end{quote}
+
+\begin{quote}
+\color{blue}
+``The MEDUSA method has been used to infer changes in net
+diversification rates in a phylogenetic tree. Since its publication
+{[}16{]} the results of using MEDUSA on a single tree, the maximum clade
+credibility tree, have been used for generation of hypotheses and
+discussion {[}10,42,43{]}. However, different diversification shifts and
+different rates of diversification are found for certain lineages when
+phylogenetic uncertainty was taken into account by using MEDUSA on a
+random sample of trees from the posterior distribution of a Bayesian
+run. We found that some diversification splits, estimated on the
+Nymphalidae maximum clade credibility tree, were found with very low
+probability in the random sample of 1000 trees from the posterior
+distribution (Fig. 3, Table 2). We also found that, even though the
+analyses estimated the same diversification splits on two or more trees,
+the estimated net diversification rates could vary widely (Fig. 2).''
+\end{quote}
+
+\textbf{17. Reviewer's Comment:}
+
+Seventh paragraph, second to last sentence: Two ages for Ithomiini are
+given -- are they the origin and diversification ages? If so, revise the
+sentence to include this information (rather than requiring readers to
+refer to the preceding sentence): ``Reference 10 gives the ages of
+origin and diversification for Ithomiini at 45 (39-53) and 37 (32-43)
+MYA, respectively''.
+
+\textbf{17. Response:}
+
+\begin{quote}
+\color{blue}
+We have followed the reviewer's suggestion.
+\end{quote}
+
+\textbf{18. Reviewer's Comment:}
+
+Eighth paragraph: This paragraph discussing biogeography does not fit
+and could be omitted without detriment to the manuscript.
+
+\textbf{18. Response:}
+
+\begin{quote}
+\color{blue}
+We have followed the reviewer's suggestion and removed this paragraph.
+\end{quote}
+
+\textbf{19. Reviewer's Comment:}
+
+First paragraph, first sentence: replace ``gave'' with ``showed''.
+
+\textbf{19. Response:}
+
+\begin{quote}
+\color{blue}
+We corrected the text following the reviewer's suggestion.
+\end{quote}
+
+\textbf{20. Reviewer's Comment:}
+
+This section is predominantly speculative. Consider a more rigorous
+discussion (if appropriate) or omit.
+
+\textbf{20. Response:}
+
+\begin{quote}
+\color{blue}
+We have followed the reviewer's suggestion and removed this paragraph.
+\end{quote}
+
+\textbf{21. Reviewer's Comment:}
+
+First paragraph, second sentence: revise ``13\% of probabilities of the
+trees from the posterior distribution'' to ``13\% of the trees from the
+posterior distribution''.
+
+\textbf{21. Response:}
+
+\begin{quote}
+\color{blue}
+We corrected the text following the reviewer's suggestion.
+\end{quote}
+
+\textbf{22. Reviewer's Comment:}
+
+Second paragraph: this paragraph is, for the most part, unnecessary. The
+last sentence of the paragraph should be revised and appended to the end
+of the preceding (first) paragraph: ``\ldots{}not picked up by MEDUSA.
+This would be expected if the diversification of Satyrinae occurred in a
+stepwise manner, with pulses or bursts of diversification for certain
+lineages but unlikely for the tribe Satyrini as a whole''.
+
+\textbf{22. Response:}
+
+\begin{quote}
+\color{blue}
+We corrected the text following the reviewer's suggestion.
+\end{quote}
+
+\textbf{23. Reviewer's Comment:}
+
+Final two paragraphs: There needs to be a section break, to indicate a
+change in the discussion topic. ``Conclusions'' would be appropriate to
+indicate to readers that these paragraphs represent a summary of the
+major points of the work. The final paragraph is a good summary of the
+work, and it would be a disservice to have readers miss it because it
+appears under the `Satyrini' section.
+
+\textbf{23. Response:}
+
+\begin{quote}
+\color{blue}
+As suggested by the reviewer, he have created a \textbf{Conclusions}
+section for the last two paragraphs of the text.
+\end{quote}
+
+\textbf{24. Reviewer's Comment:}
+
+Figure 1: The legend needs to indicate what the node labels (the circled
+integers) and the decimal values in the tree indicate.
+
+\textbf{24. Response:}
+
+\begin{quote}
+\color{blue}
+We have appended the following text to the legend of figure
+1:\\\textbf{``Circles on nodes indicate the diversification shift number
+as found by MEDUSA. Numbers next to circles indicat the posterior
+probability values for such nodes.''}
+\end{quote}
+
+\textbf{25. Reviewer's Comment:}
+
+Figure 2: Omit the legend in the figure (``Boxplot of
+diversification\ldots{}'') and revise the legend to read:
+``Diversification rates for taxa estimated by MEDUSA on the sample of
+1000 trees.''
+
+\textbf{25. Response:}
+
+\begin{quote}
+\color{blue}
+We removed the legend from the figure file and rewrote the legend in the
+MS as suggested by the reviewer.
+\end{quote}
+
+\textbf{26. Reviewer's Comment:}
+
+Figure 3: Omit the legend in the figure (``Probability of finding
+nodes\ldots{}'') and revise the legend to read: ``Results of MultiMEDUSA
+analysis showing the probability of specific nodes being characterized
+by significant shifts in diversification rate.''
+
+\textbf{26. Response:}
+
+\begin{quote}
+\color{blue}
+We removed the legend from the figure file and rewrote the legend in the
+MS as suggested by the reviewer.
+\end{quote}
+
+\textbf{27. Reviewer's Comment:}
+
+Figure 4: Describe what lambda (\(\lambda\)) and \(r\) represent in the
+figure.
+
+\textbf{27. Response:}
+
+\begin{quote}
+\color{blue}
+We included the following clarification in the legend:
+\textbf{``(speciation rate = \(\lambda1\), net diversification rate =
+\(r1\)).''}
+\end{quote}
+
+\textbf{28, 29. Reviewer's Comment:}
+
+Supporting\_Information\_S1.nex: remove bolded ``Genus.''
+
+\textbf{28, 29. Response:}
+
+\begin{quote}
+\color{blue}
+We followed the reviewer's suggestions and included better descriptions
+for the supplementary materials.
+\end{quote}
+
+\section{Reviewer 2}\label{reviewer-2}
+
+\textbf{1. Reviewer's Comment:}
+
+The authors need to provide more information on what cutoff they used
+and why.
+
+\textbf{1. Response:}
+
+\begin{quote}
+\color{blue}
+Our manuscript already includes information about the AIC value used for
+choosing the best model. The following text appears in our original
+submission in the Methods-\textgreater{} Detecting diversification
+shifts on phylogenetic trees-\textgreater{} last
+paragraph:\\\textbf{``We used the AICc threshold of 7.8 units, as
+estimated by MEDUSA, as the limit for a significantly better fit to
+select among increasingly complex alternative models.''}
+\end{quote}
+
+\textbf{2. Reviewer's Comment:}
+
+I'm not sure discussing the limitations of MEDUSA, or the tendency for
+MEDUSA to give inconsistent results across post burn-in trees, is really
+valid based on one dataset. That is, the unit of replication is the
+alignment -- therefore the authors have effectively one data point and
+zero degrees of freedom to say much of anything about the power of
+MEDUSA across the post burnin sample. I think if the authors want to
+explore the power of MEDUSA, they need many different alignments (real
+or simulated). As it is now, I don't think they really have the data to
+say anything about MEDUSA in general, and only about MEDUSA in the
+context of the Nymphalid tree. They need to at least point out there
+inability to generalize in the manuscript.
+
+\textbf{2. Response:}
+
+\begin{quote}
+\color{blue}
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  As reviewer 1 suggested, our text gave the impression that we were
+  criticizing the method MEDUSA. We have rewritten the following
+  sections to clarify our criticism \textbf{that relying on the
+  inferences made on only one tree (the MCC tree) is risky as analyses
+  of equally good hyphoteses (other trees from the posterior
+  distribution) cand potentially return different results}:
+\end{itemize}
+\end{quote}
+
+\begin{quote}
+\color{blue}
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Abstract\\``In evolutionary biology, analysis of maximum credibility
+  trees in the software MEDUSA (modelling evolutionary diversity using
+  stepwise AIC) is a popular method for estimation of shifts in
+  diversification rates. We investigated whether phylogenetic
+  uncertainty can produce different results by extending the method
+  across a random sample of trees from the posterior distribution of a
+  Bayesian run.''
+\end{itemize}
+\end{quote}
+
+\begin{quote}
+\color{blue}
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Results -\textgreater{} Phylogenetic uncertainty in the MultiMEDUSA
+  approach We changed the sentence ``We tested MEDUSA\ldots{}'' to ``We
+  used MEDUSA to find out\ldots{}''.
+\end{itemize}
+\end{quote}
+
+\begin{quote}
+\color{blue}
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Also, as suggested by the reviewer 1 (point 16 lines above) we
+  improved the discussion.
+\end{itemize}
+\end{quote}
+
+\begin{quote}
+\color{blue}
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  We have added the following in the Conclusions to acknowledge that the
+  reported behaviour might also occur in analyses of other taxa:\\"
+  However, by using a MultiMEDUSA approach, we found that for this
+  Nymphalidae dataset some of these splits might be greatly affected by
+  phylogenetic uncertainty. We recommend that all datasets should be
+  analyzed using both approaches, MEDUSA and MultiMEDUSA, in order to
+  test whether the results are robust when phylogenetic uncertainty is
+  taken into account."
+\end{itemize}
+\end{quote}
+
+\textbf{3. Reviewer's Comment:}
+
+``Our results show that taking phylogenetic uncertainty into account
+when estimating diversification rate shifts is of great importance, and
+relying on the maximum credibility tree alone potentially can give
+erroneous results''. The results are not really in error. The results
+are the correct results for the model. The real question is whether it
+is the appropriate model.
+
+\textbf{3. Response:}
+
+\begin{quote}
+\color{blue}
+We have rewritten that sentence in the abstract as follows:\\``Our
+results show that taking phylogenetic uncertainty into account when
+estimating net diversification rate shifts is of great importance, as
+very different results can be obtained when using the maximum clade
+credibility tree and other trees from the posterior distribution.''
+\end{quote}
+
+\textbf{4. Reviewer's Comment:}
+
+Why are they restricting their analysis to only one method (a method
+that they point out might not be reliable)? There are alternative
+methods available, and it might be interesting to see if consistent
+results are observed (eg., RC, PRC, BAMM). There is very little
+development on other aspects of plant -- herbivore interactions and how
+they might be predicted to affect diversification rate.
+
+\textbf{4. Response:}
+
+{[}\textbf{\textgreater{}\textgreater{}\textgreater{}\textgreater{}
+Marianne}, can you write whey we are not using the RC and PRC methods?
+``The parametric rate comparison and relative rates test that he
+suggests both require incomplete taxon sampling to be nonrandom, which
+ours definitely isn't so we can easily explain why we didn't use
+these.''\textbf{\textless{}\textless{}\textless{}\textless{}\textless{}\textless{}}{]}
+
+The method that we used, BiSSE, might not be reliable when the number of
+taxa is small, and there are biases in the number of species that have
+each of the character states. However, we stated in the manuscript that
+according to the evaluation of BiSSE by Davis, Midford, and Maddison
+(2013), our data should not be affected for such biases.
+
+The method BAMM can be used to explore the evolution of phenotypic
+traits. However the trait characters
+\href{http://bamm-project.org/configuration.html\#id5}{\textbf{must be
+continuous}}. In our case the ``feeding on Solanaceae plants'' trait is
+categorical (absence/presence) so it cannot be analyzed with BAMM.
+
+BAMM also measures the diversification rates of particular clades. We
+run an analysis to test whether we obtained congruent results. In the
+Fig. 1, shown below, we can see that BAMM was able to pick up two
+increases in diversification for Ithomiini butterflies, one for the
+origin of the clade and, most importantly, another increase during the
+recent diversification of the group. This is in agreement with our
+results from using BiSSE, that the ithomiines using Solanaceae underwent
+an increase in diversification rates.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{response_to_reviewers_fig1.pdf}
+\caption{Bamm analysis}
+\end{figure}
+
+\textbf{5. Reviewer's Comment:}
+
+The authors should `play' with an AIC cutoff, simply to explore how it
+might affect there inferences.
+
+\textbf{5. Response:}
+
+\begin{quote}
+\color{blue}
+As specified lines above and in the manuscript, we used the AIC cutoff
+value 7.8 as suggested by MEDUSA (this value is regarded as optimal due
+to the number of taxa in our analyses). Changing the cutoff value would
+allow fewer or additional diversification shifts to be picked up by
+MEDUSA. Hovewer the initial diversification shifts will remain the same
+unless the AIC cutoff used is so hight that no single diversification
+shifts are recovered. Thus playing with the AIC cutoff values does not
+affect most of the diversification shifts discussed in our manuscript.
+We rerun all our MEDUSA and MultiMEDUSA analysis with differente AIC
+cutoff values to see any effect on the results. We used the AIC values 4
+and 11.
+\end{quote}
+
+\begin{quote}
+\color{blue}
+The results in Fig.2 (also attached as PDF file) shows our MEDUSA
+results from the manuscript (cutoff 7.8) along with results with AIC
+cutoff values 4 and 11. We see that the first 19 diversification shitfs
+are the same for cutoff values 7.8 and 4. However when using AIC cutoff
+4 we recovered additional diversification shifts as the software kept
+trying to fit additional models until the differente in AIC values was
+less than 4. As expected MultiMEDUSA also found additinal
+diversification shifts across the random sample of 1000 trees.
+\end{quote}
+
+\begin{quote}
+\color{blue}
+When we used AIC cutoff 11, we obtained fewer diversification shifts. We
+recovered only 16, which are the same first 16 that we found in our
+analysis with AIC 7.8.
+\end{quote}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{multimedusa_7_8_vs_4_vs_11_AIC_cutoff.pdf}
+\caption{MEDUSA analyses under three AIC values}
+\end{figure}
+
+\textbf{6. Reviewer's Comment:}
+
+In addition, the authors should be cautious about interpreting AIC in a
+frequentist manner (e.g., ``Only three out of 13 significant shifts
+found on the maximum credibility tree were consistent across more than
+95\% of the trees from the posterior'') -- that is, AIC is a model
+selection tool, which should act as a guide for selecting the `best'
+family of models. Delta AIC should really not be thought of as alpha =
+0.05.
+
+\textbf{6. Response:}
+
+\begin{quote}
+\color{blue}
+We agree that AIC values should not be treated in a frequentist manner.
+However, we use the term \textbf{``significant shifts''}, in the same
+way as the authors of the method MEDUSA (Alfaro et al. 2009), to
+pinpoint to specific nodes that show a higher net diversification rate
+than ancestral or descendant nodes.
+\end{quote}
+
+\begin{quote}
+\color{blue}
+We also used the AIC value in the correct way. We used the cutoff 7.8
+for all MEDUSA analyses. Then we counted the number of significant
+shifts infered in each tree and tried to pinpoint those that were
+consistently recovered across the sample of 1000 random trees (and are
+also present in the MCC tree). Since no single significant shift was
+found in all 1000 trees we wanted to show those are were most commonly
+recovered. This is way we considered consistent diversification shifts
+those that were found in most of the trees (in more than 95\% of them).
+\end{quote}
+
+\textbf{7. Reviewer's Comment:}
+
+Why are authors referred to as references, rather than by authors.
+
+As noted by Reference 36, the responsible trait might\ldots{}
+
+\textbf{7. Response:}
+
+\begin{quote}
+\color{blue}
+We have fixed those references as suggested by the reviewer.
+\end{quote}
+
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
+Alfaro, Michael E, Francesco Santini, Chad Brock, Hugo Alamillo, Alex
+Dornburg, Daniel L Rabosky, Giorgio Carnevale, and Luke J Harmon. 2009.
+``Nine Exceptional Radiations Plus High Turnover Explain Species
+Diversity in Jawed Vertebrates.'' \emph{P. Natl. Acad. Sci. USA} 106
+(32): 13410--14.
+doi:\href{http://dx.doi.org/10.1073/pnas.0811087106}{10.1073/pnas.0811087106}.
+
+Davis, Matthew P, Peter E Midford, and Wayne Maddison. 2013. ``Exploring
+power and parameter estimation of the BiSSE method for analyzing species
+diversification.'' \emph{BMC Evol. Biol.} 13 (1): 38.
+doi:\href{http://dx.doi.org/10.1186/1471-2148-13-38}{10.1186/1471-2148-13-38}.
+
 \end{document}

### 2014-10-29 ancillary/Supporting_Information_S02.R

@@ -15,7 +15,8 @@ phy <- drop.tip(phy, tip=tips)
 
 richness <- read.csv("ancillary/Supporting_Information_S03.csv")
 
-res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25, mc=TRUE)
+res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25,
+              mc=TRUE, numCores=2, fixThreshold=4)
 
 
 pdf(file="output/medusa_on_mct.pdf", width=9, height=19)
@@ -23,5 +24,5 @@ summ <- medusaSummary(res, cex=0.3)
 dev.off()
 
 
-# save(res, file="ancillary/supp_mat_04_MEDUSA_results.txt", ascii=TRUE)
+save(res, file="ancillary/supp_mat_04_MEDUSA_results.txt", ascii=TRUE)
 

### 2014-10-29 MS_orig.tex
fixed special characters

@@
-\documentclass[]{article}
-\usepackage{setspace}
-\usepackage[]{lineno}
-\def\linenumberfont{\normalfont\small\sffamily}
-\doublespacing
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage[none]{hyphenat}
-\usepackage{fixltx2e} % provides \textsubscript
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-\usepackage[margin=1in,a4paper]{geometry}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={},
-            colorlinks=true,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\setcounter{secnumdepth}{0}
-
-\linenumbers
-
-\author{}
+\documentclass[10pt]{article}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{cite}
+\usepackage{hyperref}
+
+\usepackage{lineno}
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% Text layout
+\topmargin 0.0cm
+\oddsidemargin 0.5cm
+\evensidemargin 0.5cm
+\textwidth 16cm 
+\textheight 21cm
+
+% Bold the 'Figure #' in the caption and separate it with a period
+% Captions will be left justified
+\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2009}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
 \date{}
 
+\pagestyle{myheadings}
+
+%% Include all macros below. Please limit the use of macros.
+
+%% END MACROS SECTION
+
 \begin{document}
 
-\textbf{Title: Diversity dynamics in Nymphalidae butterflies: Effect of
+\begin{flushleft}
+{\Large
+\title{Diversity dynamics in Nymphalidae butterflies: Effect of
 phylogenetic uncertainty on diversification rate shift estimates}
-
-\subsubsection{Authors: Carlos Peña\textsuperscript{1,*} and Marianne
-Espeland\textsuperscript{2,*}}\label{authors-carlos-peuxf1a1-and-marianne-espeland2}
-
-\subsubsection{Running title: Diversity dynamics in Nymphalidae
-butterflies}\label{running-title-diversity-dynamics-in-nymphalidae-butterflies}
-
-\emph{1 Laboratory of Genetics, Department of Biology, University of
-Turku, Turku, Finland. Email:
-\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}}
-
-\emph{2 Museum of Comparative Zoology and Department of Organismic and
-Evolutionary Biology, Harvard University, Cambridge, USA. Email:
+}
+\maketitle
+Carlos Pe\~na$^{1,\ast}$,
+Marianne Espeland$^{2}$
+\\
+\bf{1} Carlos Pe\~na Laboratory of Genetics, Department of Biology, University of Turku, Turku, Finland. Email: \href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}
+\\
+\bf{2} Marianne Espeland Museum of Comparative Zoology and Department
+of Organismic and Evolutionary Biology, Harvard University, Cambridge,
+USA. Email:
 \href{mailto:marianne.espeland@gmail.com}{\nolinkurl{marianne.espeland@gmail.com}}.
@@
 \url{http://dx.doi.org/10.6084/m9.figshare.639208}
-
-\subsection{Abstract}\label{abstract}
+\end{flushleft}
+
+\subsubsection*{Running title: Diversity dynamics in Nymphalidae
+butterflies}
+
+\section*{Abstract}
 
@@
 
-\subsection{Methods}\label{methods}
-
-\subsubsection{Data}\label{data}
+\section*{Methods}
+
+\subsection*{Data}
 
@@
 
-\subsubsection{Analyses of
-Diversification}\label{analyses-of-diversification}
+\subsection*{Analyses of
+Diversification}
 
@@
 
-\subsubsection{Detecting diversification shifts on phylogenetic
-trees}\label{detecting-diversification-shifts-on-phylogenetic-trees}
+\subsection*{Detecting diversification shifts on phylogenetic
+trees}
 
@@
 
-\subsubsection{Estimation of trait-dependent speciation
-rates}\label{estimation-of-trait-dependent-speciation-rates}
+\subsection*{Estimation of trait-dependent speciation
+rates}
 
@@
 
-\subsection{Results}\label{results}
-
-\subsubsection{Detecting diversification shifts on the maximum clade
-credibility
-tree}\label{detecting-diversification-shifts-on-the-maximum-clade-credibility-tree}
+\section*{Results}
+
+\subsection*{Detecting diversification shifts on the maximum clade
+credibility tree}
 
@@
 
-\subsubsection{Phylogenetic uncertainty in the MultiMEDUSA
-approach}\label{phylogenetic-uncertainty-in-the-multimedusa-approach}
+\subsection*{Phylogenetic uncertainty in the MultiMEDUSA
+approach}
 
@@
 
-\subsubsection{Estimation of trait-dependent speciation
-rates}\label{estimation-of-trait-dependent-speciation-rates-1}
+\subsection*{Estimation of trait-dependent speciation
+rates}
 
@@
 
-\subsection{Discussion}\label{discussion}
-
-\subsubsection{Effects of phylogenetic uncertainty on the performance of
-MEDUSA}\label{effects-of-phylogenetic-uncertainty-on-the-performance-of-medusa}
+\section*{Discussion}
+
+\subsection*{Effects of phylogenetic uncertainty on the performance of
+MEDUSA}
 
@@
 
-\subsubsection{Hostplant use and diversification in
-Nymphalidae}\label{hostplant-use-and-diversification-in-nymphalidae}
-
-\paragraph{Ithomiini}\label{ithomiini}
+\subsection*{Hostplant use and diversification in
+Nymphalidae}
+
+\paragraph*{Ithomiini}
 
@@
 
-\paragraph{Danaini}\label{danaini}
+\paragraph{Danaini}
 
@@
 
-\paragraph{\emph{Charaxes}}\label{charaxes}
+\paragraph{\emph{Charaxes}}
 
@@
 
-\subsection{Acknowledgments}\label{acknowledgments}
+\subsection*{Acknowledgments}\label{acknowledgments}
 
@@
 
-\subsection{Figure legends}\label{figure-legends}
+\subsection*{Figure legends}\label{figure-legends}
 
@@
 
-\subsection{Tables}\label{tables}
+\subsection*{Tables}\label{tables}
 
@@
 
-\section{Supp. mat.}\label{supp.-mat.}
-
-\textbf{\texttt{supp\_mat\_01\_genus.nex}}: MCC Nymphalidae tree from
+\section*{Supp. mat.}\label{supp.-mat.}
+
+\textbf{\texttt{supp mat 01 genus.nex}}: MCC Nymphalidae tree from
 Reference 10.
 
-\textbf{\texttt{supp\_mat\_02\_run\_medusa\_on\_mct.R}}: Run MEDUSA on
+\textbf{\texttt{supp mat 02 run medusa on mct.R}}: Run MEDUSA on
 MCC tree from Reference 10.
 
-\textbf{\texttt{supp\_mat\_03\_richness.nex}}: Species richness for
+\textbf{\texttt{supp mat 03 richness.nex}}: Species richness for
 lineages in Nymphalidae.
 
-\textbf{\texttt{supp\_mat\_04\_MEDUSA\_results.txt}}: Results of running
+\textbf{\texttt{supp mat 04 MEDUSA results.txt}}: Results of running
 MEDUSA on the MCC tree.
 
-\textbf{\texttt{supp\_mat\_05\_1000\_random\_trees\_no\_outgroups.nex}}:
+\textbf{\texttt{supp mat 05 1000 random trees no outgroups.nex}}:
 1000 random trees from Reference 10.
 
-\textbf{\texttt{supp\_mat\_06\_run\_multimedusa.R}}: Run MultiMEDUSA on
+\textbf{\texttt{supp mat 06 run multimedusa.R}}: Run MultiMEDUSA on
 1000 random trees from Reference 10.
 
-\textbf{\texttt{supp\_mat\_07\_multimedusa\_on\_1000\_trees.txt}}: Raw
+\textbf{\texttt{supp mat 07 multimedusa on 1000 trees.txt}}: Raw
 results from the MultiMEDUSA run on the random sample of trees from the
@@
 
-\textbf{\texttt{supp\_mat\_08\_multimedusa\_result01.pdf}}: Summary of
+\textbf{\texttt{supp mat 08 multimedusa result01.pdf}}: Summary of
 MulitMEDUSA analysis on the 1000 trees from the posterior distribution.
 
-\textbf{\texttt{supp\_mat\_09\_mcc\_tree\_from\_random\_sample\_of\_1000.tree}}:
+\textbf{\texttt{supp mat 09 mcc tree from random sample of 1000.tree}}:
 MCC tree from the 1000 random trees selected from the posterior
@@
 
-\textbf{\texttt{supp\_mat\_10\_run\_medusa\_on\_mct\_from\_1000\_trees.R}}:
+\textbf{\texttt{supp mat 10 run medusa on mct from 1000 trees.R}}:
 Run MEDUSA on MCC tree from the 1000 random trees selected from the
@@
 
-\textbf{\texttt{supp\_mat\_11\_results\_of\_MEDUSA\_run\_on\_mcc\_from\_1000\_trees.txt}}:
+\textbf{\texttt{supp mat 11 results of MEDUSA run on mcc from 1000 trees.txt}}:
 Summary results from a MEDUSA run on the MCC tree from the 1000 random
@@
 
-\textbf{\texttt{supp\_mat\_12\_MEDUSA\_mcc\_tree\_from\_random\_sample\_of\_1000\_trees.pdf}}:
+\textbf{\texttt{supp mat 12 MEDUSA mcc tree from random sample of 1000 trees.pdf}}:
 Figure for MEDUSA run on MCC tree from random 1000 trees.
 
-\textbf{\texttt{supp\_mat\_13\_results\_of\_Multi-MEDUSA\_run\_on\_1000\_trees.txt}}:
+\textbf{\texttt{supp mat 13 results of Multi-MEDUSA run on 1000 trees.txt}}:
 Summary results from a MultiMEDUSA run on the 1000 random trees selected
@@
 
-\textbf{\texttt{supp\_mat\_14\_multimedusa\_result\_prob\_nodes.pdf}}:
+\textbf{\texttt{supp mat 14 multimedusa result prob nodes.pdf}}:
 Probability of nodes with estimated rates from a MultiMEDUSA run on the
@@
 
-\textbf{\texttt{supp\_mat\_15\_host\_plant\_data.csv}}: Hostplants of
+\textbf{\texttt{supp mat 15 host plant data.csv}}: Hostplants of
 Nymphalidae butterflies recorded from the literature.
 
-\textbf{\texttt{supp\_mat\_16\_host\_plant\_data\_references.csv}}:
+\textbf{\texttt{supp mat 16 host plant data references.csv}}:
 References for hostplants data.
 
-\textbf{\texttt{supp\_mat\_17\_hostplant\_states.csv}}: Data matrix with
+\textbf{\texttt{supp mat 17 hostplant states.csv}}: Data matrix with
 character states for hosplant use.
 
-\textbf{\texttt{supp\_mat\_18\_bisse\_source.R}}: R code for running the
+\textbf{\texttt{supp mat 18 bisse source.R}}: R code for running the
 BiSSE analysis.
 
-\textbf{\texttt{supp\_mat\_19\_bisse\_mcmc\_run.csv}}: Raw results for
+\textbf{\texttt{supp mat 19 bisse mcmc run.csv}}: Raw results for
 the BiSSE analysis.
 
-\textbf{\texttt{supp\_mat\_20\_combined\_bisse.csv}}: Raw results for
+\textbf{\texttt{supp mat 20 combined bisse.csv}}: Raw results for
 the combined BiSSE analysis.

### 2014-10-29 MS_revised.tex
fixed special characters

@@
 
-\subsection*{Abstract}
+\section*{Abstract}
 
@@
 
-\subsection*{Introduction}
+\section*{Introduction}
 
@@
 
-\subsection*{Materialas and Methods}
-
-\subsubsection*{Data}
+\section*{Materials and Methods}
+
+\subsection*{Data}
 
@@
 
-\subsubsection*{Analyses of
+\subsection*{Analyses of
 Diversification}
@@
 
-\subsubsection*{Detecting diversification shifts on phylogenetic
+\subsection*{Detecting diversification shifts on phylogenetic
 trees}
@@
 
-\subsubsection*{Estimation of trait-dependent speciation
+\subsection*{Estimation of trait-dependent speciation
 rates}
@@
 
-\subsection*{Results}
-
-\subsubsection*{Detecting diversification shifts on the maximum clade
-credibility
-tree}
+\section*{Results}
+
+\subsection*{Detecting diversification shifts on the maximum clade
+credibility tree}
 
@@
 
-\subsubsection*{Phylogenetic uncertainty in the MultiMEDUSA
+\subsection*{Phylogenetic uncertainty in the MultiMEDUSA
 approach}
@@
 
-\subsection*{Discussion}
-
-\subsubsection*{Effects of phylogenetic uncertainty on the performance of
+\section*{Discussion}
+
+\subsection*{Effects of phylogenetic uncertainty on the performance of
 MEDUSA}
@@
 
-\subsubsection*{Hostplant use and diversification in
+\subsection*{Hostplant use and diversification in
 Nymphalidae}
 
-\paragraph*{Ithomiini}
+\subsubsection*{Ithomiini}
 
@@
 diversification in glucosinolate-feeding taxa shortly afterwards (with a
-lag of \textasciitilde{}10 MY). According to a recent dated phylogeny of
+lag of circa 10 MY). According to a recent dated phylogeny of
 the Angiosperms {[}49{]}, the family Solanaceae split from its sister
@@
 
-\paragraph*{Danaini}
+\subsubsection*{Danaini}
 Our MultiMEDUSA approach showed a significant slowdown in net
@@
 
-\paragraph*{Satyrini}
+\subsubsection*{Satyrini}
 
@@
 
-\subsection*{Conclusions}
+\section*{Conclusions}
 We found that even though MEDUSA estimated several diversification
@@
 
-\subsection*{Acknowledgments}
+\section*{Acknowledgments}
 We are thankful to Mark Cornwall for help with the script to extend
@@
 \begin{description}
- \item {\bf Supporting\_Information\_S01.nex.} MCC Nymphalidae tree from
+ \item {\bf Supporting Information S01.nex.} MCC Nymphalidae tree from
 {[}13{]}
 
- \item {\bf Supporting\_Information\_S02.R} R script to run MEDUSA on the
+ \item {\bf Supporting Information S02.R} R script to run MEDUSA on the
 MCC Nymphalidae tree from {[}13{]}. This script removes the outgroup
@@
 
- \item {\bf Supporting\_Information\_S03.csv} Species richness for lineages
+ \item {\bf Supporting Information S03.csv} Species richness for lineages
 in Nymphalidae.
 
- \item {\bf Supporting\_Information\_S04.txt} Results of running MEDUSA on
+ \item {\bf Supporting Information S04.txt} Results of running MEDUSA on
 the MCC tree.
 
- \item {\bf Supporting\_Information\_S05\_1000\_random\_trees\_no\_outgroups.zip} 1000 random trees from {[}13{]}.
-
- \item {\bf Supporting\_Information\_S06.R} R script to run a MultiMEDUSA
+ \item {\bf Supporting Information S05 1000 random trees no outgroups.zip} 1000 random trees from {[}13{]}.
+
+ \item {\bf Supporting Information S06.R} R script to run a MultiMEDUSA
 analysis on 1000 random trees from {[}13{]}.
 
- \item {\bf Supporting\_Information\_S07\_multimedusa\_on\_1000\_trees.zip} Raw results from the MultiMEDUSA run on the random sample of trees from
+ \item {\bf Supporting Information S07 multimedusa on 1000 trees.zip} Raw results from the MultiMEDUSA run on the random sample of trees from
 the posterior distribution.
 
- \item {\bf Supporting\_Information\_S08.pdf} Summary of the MulitMEDUSA
+ \item {\bf Supporting Information S08.pdf} Summary of the MulitMEDUSA
 analysis on the 1000 trees from the posterior distribution.
 
- \item {\bf Supporting\_Information\_S09.tree} MCC tree from the 1000
+ \item {\bf Supporting Information S09.tree} MCC tree from the 1000
 random trees selected from the posterior distribution.
 
- \item {\bf Supporting\_Information\_S10.R} R script to run a MEDUSA
+ \item {\bf Supporting Information S10.R} R script to run a MEDUSA
 analysis on the MCC tree from the 1000 random trees selected from the
@@
 
- \item {\bf Supporting\_Information\_S11.txt} Summary results from a MEDUSA
+ \item {\bf Supporting Information S11.txt} Summary results from a MEDUSA
 run on the MCC tree from the 1000 random trees selected from the
@@
 
- \item {\bf Supporting\_Information\_S12.pdf} Figure for MEDUSA run on MCC
+ \item {\bf Supporting Information S12.pdf} Figure for MEDUSA run on MCC
 tree from random 1000 trees.
 
- \item {\bf Supporting\_Information\_S13.txt} Summary results from a
+ \item {\bf Supporting Information S13.txt} Summary results from a
 MultiMEDUSA run on the 1000 random trees selected from the posterior
@@
 
- \item {\bf Supporting\_Information\_S14.pdf} Probability of nodes with
+ \item {\bf Supporting Information S14.pdf} Probability of nodes with
 estimated rates from a MultiMEDUSA run on the 1000 random trees selected
@@
 
- \item {\bf Supporting\_Information\_S15.csv} Hostplants of Nymphalidae
+ \item {\bf Supporting Information S15.csv} Hostplants of Nymphalidae
 butterflies recorded from the literature.
 
- \item {\bf Supporting\_Information\_S16.csv} References for hostplants
+ \item {\bf Supporting Information S16.csv} References for hostplants
 data.
 
- \item {\bf Supporting\_Information\_S17.csv} Data matrix with character
+ \item {\bf Supporting Information S17.csv} Data matrix with character
 states for hosplant use.
 
- \item {\bf Supporting\_Information\_S18.R} R script for running the BiSSE
+ \item {\bf Supporting Information S18.R} R script for running the BiSSE
 analysis.
 
- \item {\bf Supporting\_Information\_S19.csv} Raw results for the BiSSE
+ \item {\bf Supporting Information S19.csv} Raw results for the BiSSE
 analysis.
 
- \item {\bf Supporting\_Information\_S20.csv} Raw results for the combined
+ \item {\bf Supporting Information S20.csv} Raw results for the combined
 BiSSE analysis.
@@
 
-2. Nyman T, Linder HP, Peña C, Malm T, Wahlberg N (2012) Climate-driven
+2. Nyman T, Linder HP, Pe{\~n}a C, Malm T, Wahlberg N (2012) Climate-driven
 diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
@@
 
-10. Heikkilä M, Kaila L, Mutanen M, Peña C, Wahlberg N (2012) Cretaceous
+10. Heikkil{\"a} M, Kaila L, Mutanen M, Pe{\~n}a C, Wahlberg N (2012) Cretaceous
 origin and repeated tertiary diversification of the redefined
@@
 
-11. Elias M, Joron M, Willmott KR, Silva-Brandão KL, Kaiser V, et al.
+11. Elias M, Joron M, Willmott KR, Silva-Brand{\~a}o KL, Kaiser V, et al.
 (2009) Out of the Andes: Patterns of diversification in clearwing
@@
 
-13. Wahlberg N, Leneveu J, Kodandaramaiah U, Peña C, Nylin S, et al.
+13. Wahlberg N, Leneveu J, Kodandaramaiah U, Pe{\~n}a C, Nylin S, et al.
 (2009) Nymphalid butterflies diversify following near demise at the
@@
 
-14. Pe\~na C, Wahlberg N (2008) Prehistorical climate change increased
+14. Pe{\~n}a C, Wahlberg N (2008) Prehistorical climate change increased
 diversification of a group of butterflies. Biol Lett 4: 274--278.
@@
 
-15. Ferrer-Paris JR, S{á}nchez-Mercado A, Viloria ÁL, Donaldson J (2013)
+15. Ferrer-Paris JR, S{\'a}nchez-Mercado A, Viloria {\'A}L, Donaldson J (2013)
 Congruence and Diversity of Butterfly-Host Plant Associations at Higher
@@
 
-20. Matos-Maraví PF, Peña C, Willmott KR, Freitas AV, Wahlberg N (2013)
+20. Matos-Marav{\'i} PF, Pe{\~n}a C, Willmott KR, Freitas AV, Wahlberg N (2013)
 Systematics and evolutionary history of butterflies in the ``Taygetis
@@
 
-21. Brower AVZ, Wahlberg N, Ogawa JR, Boppré M, Vane-Wright RI (2010)
+21. Brower AVZ, Wahlberg N, Ogawa JR, Boppr{\'e} M, Vane-Wright RI (2010)
 Phylogenetic relationships among genera of danaine butterflies
@@
 
-22. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
+22. Kodandaramaiah U, Lees DC, M{\"u}ller CJ, Torres E, Karanth KP, et al.
 (2010) Phylogenetics and biogeography of a spectacular Old World
@@
 
-23. Kodandaramaiah U, Peña C, Braby MF, Grund R, Müller CJ, et al.
+23. Kodandaramaiah U, Pe{\~n}a C, Braby MF, Grund R, M{\"u}ller CJ, et al.
 (2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
@@
 
-27. Peña C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
+27. Pe{\~n}a C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
 (2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
@@
 
-29. Silva-Brandão KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
+29. Silva-Brand{\~a}o KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
 Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
@@
 
-30. Peña C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
+30. Pe{\~n}a C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
 butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
@@
 
-31. Peña C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
+31. Pe{\~n}a C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
 history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
@@
 
-34. Beccaloni G, Viloria Á, Hall S, Robinson G (2008) Catalogue of the
+34. Beccaloni G, Viloria {\'A}, Hall S, Robinson G (2008) Catalogue of the
 hostplants of the Neotropical butterflies. London: The Natural History
@@
 
-45. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brandão KL, Whinnett A, et
+45. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand{\~a}o KL, Whinnett A, et
 al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
@@
 
-50. Wahlberg N, Rubinoff D (2011) Vagility across Vanessa (Lepidoptera:
+50. Wahlberg N, Rubinoff D (2011) Vagility across \emph{Vanessa} (Lepidoptera:
 Nymphalidae): mobility in butterfly species does not inhibit the

### 2014-11-04 response_to_reviewers.tex
we ran BAMM to reviewer 2

@@
 
-{[}\textbf{\textgreater{}\textgreater{}\textgreater{}\textgreater{}
-Marianne}, can you write whey we are not using the RC and PRC methods?
-``The parametric rate comparison and relative rates test that he
-suggests both require incomplete taxon sampling to be nonrandom, which
-ours definitely isn't so we can easily explain why we didn't use
-these.''\textbf{\textless{}\textless{}\textless{}\textless{}\textless{}\textless{}}{]}
+\begin{quote}
+\color{blue}
+The methods RC and PRC (parametric rate comparison and relative rates tests) 
+fit several distribution models to the trees in order to
+infer branch lengths. This requires random sampling of taxa to avoid biases.
+This assumption is not satisfied with our data as the sampling is non-random.
+This is one of the characteristics of butterfly datasets where many geographic
+areas are not sampled leaving out major lineages.
 
@@
 an increase in diversification rates.
+
+We added to the revised version of the MS that we also ran BAMM in order to
+test congruence of the results with this recent menthods.
+\end{quote}
 

### 2014-11-04 bamm/analyze_bamm_output.R
bamm run as traits

@@ -1,0 +1,52 @@
+library(BAMMtools)
+tree <- read.tree("bamm/mcc.tre")
+edata <- getEventData(tree, eventdata = "bamm/event_data.txt", burnin=0.95, type="trait")
+mcmcout <- read.csv("bamm/mcmc_out.txt", header=TRUE)
+plot(mcmcout$logLik ~ mcmcout$generation, type="l")
+
+# 50 million generations were run
+burnstart <- floor(0.95 * nrow(mcmcout))
+postburn <- mcmcout[burnstart:nrow(mcmcout),]
+plot(postburn$logLik ~ postburn$generation, type="l")
+
+library(coda)
+effectiveSize(postburn$N_shifts)
+effectiveSize(postburn$logLik)
+
+post_probs <- table(postburn$N_shifts) / nrow(postburn)
+
+shift_probs <- summary(edata)
+
+# compare with simulated priors
+postfile <- "bamm/mcmc_out.txt"
+priorfile <- "bamm/prior_probs.txt"
+pprior <- getBranchShiftPriors(tree, priorfile)
+css_nymp <- credibleShiftSet(edata, pprior)
+bmat <- computeBayesFactors(postfile, priorfile, burnin=0.1)
+bmat # most likely 8 shifts
+summary(edata)
+plot.bammdata(edata, lwd=2, legend=TRUE)
+plot.credibleshiftset(css_nymp, lwd=1.7, plotmax=4)
+
+# find best configuration
+
+
+allrates <- getCladeRates(edata)
+ithomiiinae <- getCladeRates(edata, node=749)
+
+st <- max(branching.times(tree))
+plot.new()
+par(mfrow=c(2,3))
+plotRateThroughTime(edata, intervalCol = "red", avgCol = "red", start.time=st, ylim=c(0,1), cex.axis=1.5, 
+                    cex.lab=1.5)
+text(x=80,y=0.8, label="All butterflies", font=3, cex=2.0, pos=4)
+plotRateThroughTime(edata, intervalCol = "blue", avgCol = "blue", start.time=st, ylim=c(0,1), cex.axis=1.5, 
+                    cex.lab=1.5, node=749)
+text(x=80,y=0.8, label="Ithomiinae only", font=3, cex=2.0, pos=4)
+plotRateThroughTime(edata, intervalCol = "darkgreen", avgCol = "darkgreen", start.time=st, ylim=c(0,1), cex.axis=1.5, 
[…]

### 2014-11-04 bamm/run_bamm.R

@@ -1,0 +1,14 @@
+library(ape)
+phy <- read.nexus("ancillary/Supporting_Information_S01.nex");
+phy <- ladderize(phy)
+
+# we have to remove the tips
+tips <- c("Achlyodes", "Graphium", "Parnassius", "Baronia", "Troides", "Papilio1", "Papilio2", "Pieris", "Aporia", "Styx", "Hamearis", "Euselasia", "Nymphidium", "Emesis", "Crocozona", "Riodina", "Amarynthis", "Baliochila", "Poritia", "Miletus", "Liphyra", "Lycaena", "Celastrina", "Thecla", "Lucia", "Curetis", "Eurema", "Colias", "Leptidea", "Pseudopontia", "Libyt")
+phy <- drop.tip(phy, tip=tips)
+
+is.ultrametric(phy)
+is.binary.tree(phy)
+# Now to check min branch length:
+min(phy$edge.length)
+
+write.tree(phy, file="bamm/mcc.tre")

### 2014-11-04 MS_revised.tex
2nd run bamm

@@
 
+We contrasted these results with those obtained from analysing the data in
+the software BAMM. We found 
+that the best model (best configuration) estimated 8 diversification shifts in
+our MCC tree (posterior probability = 0.65). Bayes factor values strongly 
+favored this model in comparison with the second best model (7 diversification
+shifts, BF > 37,000). <<<<CHECK AGAIN AFTER 2nd BAMM Run >>>
+
 \subsection*{Phylogenetic uncertainty in the MultiMEDUSA

### 2014-11-05 MS_revised.tex
add ref for BAMM

--- 2014-10-29 MS_revised.tex
@@
 
+We contrasted these results with those obtained from analysing the data in
+the software BAMM [@rabosky2014]. We found 
+that the best model (best configuration) estimated 2 diversification shifts in
+our MCC tree (posterior probability = 0.79) for the Ithomiini
+ and Satyrini. Bayes factor (BF) values strongly 
+favored this model in comparison with a constant rate model (no diversification
+shifts, BF $>$ 1,884).
+
 \subsection*{Phylogenetic uncertainty in the MultiMEDUSA

### 2014-11-05 MS_revised.tex
ref 49

@@
 lag of circa 10 MY). According to a recent dated phylogeny of
-the Angiosperms {[}49{]}, the family Solanaceae split from its sister
+the Angiosperms {[}50{]}, the family Solanaceae split from its sister
 group about 59 (49--68) MYA and diversification started (crown group
@@
 widespread species show higher genetic differentiation in their
-populations {[}50{]}.
+populations {[}51{]}.
 

### 2014-11-05 MS_revised.tex
ref 48

@@
 configurations of the proteins, providing more opportunities for
-hostplant switch {[}48{]}. This might be the reason why the basal
+hostplant switch {[}49{]}. This might be the reason why the basal
 Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and

### 2014-11-05 MS_revised.tex
ref 47

@@
 detoxification of ingested food as they have been found in other
-butterfly groups {[}47{]}.
+butterfly groups {[}48{]}.
 
@@
 diversify after their hostplants {[}2--4{]}. Wheat et al. (2007)
-{[}47{]} found strong evidence for a model of speciation congruent with
+{[}48{]} found strong evidence for a model of speciation congruent with
 Ehrlich and Raven's hypothesis in Pieridae butterflies due to, in

### 2014-11-05 MS_revised.tex
ref 46

@@
 due to the use of Solanaceae plants as hosts during larval stage. For
-this, we used a Bayesian approach {[}46{]} to test whether the trait
+this, we used a Bayesian approach {[}47{]} to test whether the trait
 ``feeding on Solanaceae'' had any effect on the diversification of the

### 2014-11-05 MS_revised.tex
ref 45

@@
 evolution of hostplant use on phylogenies evidence a probable shift from
-Apocynaceae to Solanaceae in the ancestor of the tribe {[}8,45{]}.
+Apocynaceae to Solanaceae in the ancestor of the tribe {[}8,46{]}.
 Fordyce (2010) {[}12{]} found that the Gamma statistics, a LTT plot of

### 2014-11-05 MS_revised.tex
ref 44

@@
 Keith Brown suggested that feeding on Solanaceae was an important event
-in the diversification of Ithomiini butterflies {[}44{]}. Ithomiini
+in the diversification of Ithomiini butterflies {[}45{]}. Ithomiini
 butterflies are exclusively Neotropical and most species feed on
@@
 and radiation scenarios among chemically different lineages of
-Solanaceae plants {[}8,44{]}. According to this theory, the shift from
+Solanaceae plants {[}8,45{]}. According to this theory, the shift from
 Apocynaceae to Solanaceae allowed Ithomiini to invade newly available

### 2014-11-05 MS_revised.tex
ref 43

@@
 have been used for generation of hypotheses and discussion
-{[}10,42,43{]}. However, different diversification shifts and different
+{[}10,42,44{]}. However, different diversification shifts and different
 rates of diversification are found for certain lineages when

### 2014-11-05 MS_revised.tex
ref 42

@@
 have been used for generation of hypotheses and discussion
-{[}10,42,44{]}. However, different diversification shifts and different
+{[}10,43,44{]}. However, different diversification shifts and different
 rates of diversification are found for certain lineages when

### 2014-11-05 MS_revised.tex
added new ref 42 rabosky2014
added rabosky to list of refs

@@
 We contrasted these results with those obtained from analysing the data in
-the software BAMM [@rabosky2014]. We found 
+the software BAMM {[}42{]}. We found 
 that the best model (best configuration) estimated 2 diversification shifts in
@@
 
-42. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
+42. Rabosky DL (2014) Automatic Detection of Key Innovations, Rate Shifts, and
+Diversity-Dependence on Phylogenetic Trees. PLoS ONE 9: e89543. doi:
+\href{http://dx.doi.org/10.1371/journal.pone.0089543}{10.1371/journal.pone.0089543}.
+
+43. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
 bees cut leaves? New insights into the early evolution of bees. P Roy
@@
 
-43. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
+44. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
 clades of Agaricales. P Roy Soc B 279: 2003--2011.
@@
 
-44. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
+45. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
 Interface. Ann Mo Bot Gard 74: 359--397.
 
-45. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand{\~a}o KL, Whinnett A, et
+46. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand{\~a}o KL, Whinnett A, et
 al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
@@
 
-46. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
+47. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
 trait-dependent speciation and extinction rates from incompletely
@@
 
-47. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
+48. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
 The genetic basis of a plant-insect coevolutionary key innovation. P
@@
 
-48. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
+49. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
 furanocoumarin-metabolizing cytochrome P450 monooxygenases in two
@@
 
-49. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
+50. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
 the angiosperms re-revisited. Am J Bot 97: 1296--1303.
@@
 
-50. Wahlberg N, Rubinoff D (2011) Vagility across \emph{Vanessa} (Lepidoptera:
+51. Wahlberg N, Rubinoff D (2011) Vagility across \emph{Vanessa} (Lepidoptera:
 Nymphalidae): mobility in butterfly species does not inhibit the

### 2014-11-10 MS_revised.tex
remove lineno to avoid utf-8 error
add refs to TeX file

@@
 \usepackage{hyperref}
-
-\usepackage{lineno}
 \usepackage{microtype}
@@
 \usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2009}
 
@@
 Hostplant shifts have been invoked to be responsible for a great part of
-the biodiversity of herbivorous insects {[}1{]}. The study of the
+the biodiversity of herbivorous insects \cite{mitter1988}. The study of the
 evolution of hostplant use has spawned several theories explaining the
-evolutionary interactions between plants and insects {[}2{]}. The
-``escape-and-radiate'' hypothesis {[}3{]}, the ``oscillation
-hypothesis'' {[}4,5{]} or ``diffuse cospeciation'' {[}2{]}.
+evolutionary interactions between plants and insects \cite{nyman2012}. The
+``escape-and-radiate'' hypothesis \cite{ehrlich1964}, the 
+``oscillation hypothesis''
+\cite{janz2011,nylin2014} or ``diffuse cospeciation'' \cite{nyman2012}.
 
@@
 developing some of the mentioned hypotheses. Nymphalidae contains around
-6000 species {[}6{]}, and several members are considered model organisms
-in evolutionary biology {[}7--9{]}. The family most likely originated
+6000 species \cite{van_nieukerken2011},
+and several members are considered model organisms
+in evolutionary biology \cite{joron2006,willmott2006,brakefield2009}.
+The family most likely originated
 around 94 MYA in the mid Cretaceous. Diversification of the group began
@@
 appeared shortly after the Cretaceous-Paleogene (K-Pg) boundary
-{[}10{]}. Several studies have used time-calibrated phylogenies and
+\cite{heikkila2012}. Several studies have used time-calibrated phylogenies and
 diversification models to reconstruct the evolutionary history of the
 group to identify patterns of accelerated or decelerated diversification
-of some Nymphalidae clades {[}10--13{]}. For example, it has been
+of some Nymphalidae clades
+\cite{heikkila2012,elias2009,fordyce2010,wahlberg2009}.
+For example, it has been
 suggested that climate change in the Oligocene and the subsequent
 diversification of grasses has led to diversification of the subfamily
-Satyrinae {[}14{]} due to the abundance of grasses over extensive
+Satyrinae \cite{pena2008} due to the abundance of grasses over extensive
 geographic areas (``resource abundance-dependent diversity dynamics''
-hypothesis). Fordyce (2010) {[}12{]} found increased net diversification
+hypothesis). Fordyce (2010) \cite{fordyce2010} found increased net diversification
 rates in some Nymphalidae lineages after a major hostplant shift, which
 appears to be in agreement with the ``escape-and-radiate'' model of
-diversification {[}3{]}.
+diversification \cite{ehrlich1964}.
 
@@
 sudden diversification bursts due to hostplant shift, climatic events or
-shifts to closely related hostplants {[}5,15{]}.
+shifts to closely related hostplants \cite{nylin2014,ferrer2013}.
 
@@
 
+% Use the PLoS provided BiBTeX style
+\bibliography{refs}{}
+\bibliographystyle{plos2009}
+
 \section*{References}

### 2014-11-10 MS_revised.tex
ref 35

@@
 In this study, we used a time-calibrated genus-level phylogenetic
-hypothesis for Nymphalidae butterflies {[}13{]} to investigate patterns
-of diversification. We applied MEDUSA {[}16,17{]}, a recently developed
+hypothesis for Nymphalidae butterflies \cite{wahlberg2009} to
+investigate patterns
+of diversification. We applied MEDUSA \cite{alfaro2009,harmon2011},
+a recently developed
 statistical method, to study the diversification pattern of Nymphalidae
@@
 study the effects of phylogenetic uncertainty and by using the extended
-MEDUSA method called MultiMEDUSA {[}16{]}. We also tested whether
+MEDUSA method called MultiMEDUSA \cite{alfaro2009}. We also tested whether
 hostplant association dynamics can explain the diversification patterns
@@
 employing the method BiSSE as implemented in the R package
-\texttt{diversitree} {[}18{]}.
+\texttt{diversitree} \cite{fitzjohn2012}.
 
@@
 For analyses, we used the phylogenetic trees from the study of Wahlberg
-et al. (2009) {[}13{]} that were generated using DNA sequence data from
+et al. (2009) \cite{wahlberg2009} that were generated using DNA sequence data from
 10 gene regions for 398 of the 540 valid genera in Nymphalidae. We
@@
 as a random sample of 1000 trees from their BEAST run after burnin
-{[}13{]}. Their original BEAST run was for 40 million generations. We
+\cite{wahlberg2009}. Their original BEAST run was for 40 million generations. We
 used a burnin of 25 million generations and took a random sample of 1000
@@
 sources including the specialist-curated lists on
-\url{http://tolweb.org}, Lamas (2004) {[}19{]} and curated lists of
+\url{http://tolweb.org}, Lamas (2004) \cite{lamas2004} and curated lists of
 Global Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}).
@@
 to the closest related genus that was included in Wahlberg et al. (2009)
-{[}13{]} study according to available phylogenetic studies {[}20--31{]},
+\cite{wahlberg2009} study according to available phylogenetic studies 
+\cite{matos2013,brower2010,kodandaramaiah2010,kodandaramaiah2010a,ortiz2013,desilva2010,freitas2004,pena2006,penz1999,silva2008,pena2011,pena2010},
 taxonomical classification and morphological resemblance when no
@@
 Hostplant data for Nymphalidae species were compiled from several
-sources including {[}32{]}, HOSTS database (\url{http://bit.ly/YI7nwW}),
-{[}33--35{]} and others (supp. mat. 15--16) for a total of 6586
+sources including \cite{ackery1988}, HOSTS database (\url{http://bit.ly/YI7nwW}),
+\cite{dyer2002,beccaloni2008,janzen2009} and others (supp. mat. 15--16)
+for a total of 6586
 hostplant records, including 428 Nymphalidae genera and 143 plant

### 2014-11-10 MS_revised.tex
ref 41

@@
 
-We used the statistical software R version 3.0.1 {[}36{]} in combination
-with the APE {[}37{]}, GEIGER {[}38{]}, MEDUSA {[}16{]} and
-\texttt{diversitree} {[}18{]} packages along with our own scripts to
+We used the statistical software R version 3.0.1 \cite{r2013} in combination
+with the APE \cite{popescu2012}, GEIGER \cite{harmon2008}, MEDUSA 
+\cite{alfaro2009} and
+\texttt{diversitree} \cite{fitzjohn2012} packages along with our own scripts to
 perform the analyses (included as supplementary materials 2, 6, 10). All
 analyses were run on the 1000 random trees from Wahlberg et al. (2009)
-{[}13{]}, on the maximum clade credibility tree derived from these and
-the MCC tree {[}13{]}.
+\cite{wahlberg2009}, on the maximum clade credibility tree derived from these and
+the MCC tree \cite{wahlberg2009}.
 
@@
 Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
-version 093.4.33 {[}16{]} on the maximum clade credibility tree from
-Wahlberg et al. (2009) {[}13{]}. MEDUSA fits alternative birth-death
+version 093.4.33 \cite{alfaro2009} on the maximum clade credibility tree from
+Wahlberg et al. (2009) \cite{wahlberg2009}. MEDUSA fits alternative birth-death
 likelihood models to a phylogenetic tree in order to estimate changes in
@@
 value. MEDUSA finds the likelihood of the models after taking into
-account branch lengths and number of species per lineage {[}16{]}. Most
+account branch lengths and number of species per lineage 
+\cite{alfaro2009}. Most
 studies using MEDUSA only run the method on a single tree, usually the
@@
 therefore selected 1000 random genus-level trees from the posterior
-distribution of the Bayesian run from Wahlberg et al., 2009 {[}13{]}
+distribution of the Bayesian run from Wahlberg et al., 2009 \cite{wahlberg2009}
 (MultiMEDUSA, supp. mat. 5). We calculated a new MCC tree from the
@@
 net diversification (\texttt{r = b - d}) and relative extinction rate
-(\(\varepsilon\) \texttt{= d/b}) values across the 1000 trees for the
+($\varepsilon$ \texttt{= d/b}) values across the 1000 trees for the
 nodes where changes in diversification tempo occurs. We used the AICc
@@
 and performed analyses using the ``binary state speciation and
-extinction'' {[}39{]} Bayesian approach as implemented in the R package
-\texttt{diversitree} {[}18{]}. The Markov Chain Monte Carlo algorithm
+extinction'' \cite{maddison2007} Bayesian approach as implemented in the R package
+\texttt{diversitree} \cite{fitzjohn2012}. The Markov Chain Monte Carlo algorithm
 was run for 10000 generations discarding the first 7500 as burnin. The
-Multiple State Speciation and Extinction approach (MuSSE) {[}18{]} is
+Multiple State Speciation and Extinction approach (MuSSE) 
+\cite{fitzjohn2012} is
 designed to examine the joint effects of two or more traits on
@@
 binary character state has had any effect on increased net
-diversification rate for a clade {[}39{]}. We used our compiled data of
+diversification rate for a clade \cite{maddison2007}. We used our compiled data of
 hostplant use to produce a binary dataset for the character ``feeding on
@@
 1) (supp. mat. 17). The family Solanaceae include the hostplants of the
-diverse Ithomiini butterflies and closest relatives {[}8{]}. Other
+diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. Other
 hostplant shifts were not tested for effect on net diversification rates
@@
 distribution. The records of \emph{Vanessa} and \emph{Hypanartia}
-feeding on Solanaceae {[}33,40{]} might be incorrect as it is unlikely
+feeding on Solanaceae \cite{beccaloni2008,scott1986} might be incorrect as it
+is unlikely
 that these species can be feeding on this plant family. We tested
@@
 or presence would not have a significant effect on our results. It has
-been shown that BiSSE performs poorly under certain conditions {[}41{]}.
+been shown that BiSSE performs poorly under certain conditions 
+\cite{davis2013}.
 However, our data has adequate number of taxa under analysis (more than
@@
 of Solanaceae hostplants. Thus, BiSSE is expected to produce robust
-results {[}41{]}.
+results \cite{davis2013}.
 

### 2014-11-10 MS_revised.tex
all refs in

@@
 We contrasted these results with those obtained from analysing the data in
-the software BAMM {[}42{]}. We found 
+the software BAMM \cite{rabosky2014}. We found 
 that the best model (best configuration) estimated 2 diversification shifts in
@@
 lower median net diversification rates for the diversification shifts
-found by MEDUSA on the MCC tree from {[}13{]} (Table 2). Although the
+found by MEDUSA on the MCC tree from \cite{wahlberg2009} (Table 2). Although the
 diversification pattern found by MEDUSA and MultiMEDUSA was the same,
@@
 The MEDUSA method has been used to infer changes in net diversification
-rates in a phylogenetic tree. Since its publication {[}16{]} the results
+rates in a phylogenetic tree. Since its publication \cite{alfaro2009} the results
 of using MEDUSA on a single tree, the maximum clade credibility tree,
 have been used for generation of hypotheses and discussion
-{[}10,43,44{]}. However, different diversification shifts and different
+\cite{heikkila2012,litman2011,ryberg2012}. However, different diversification
+shifts and different
 rates of diversification are found for certain lineages when
@@
 taxa appear to be strongly affected by long-branch attraction artifacts
-{[}30{]}. Thus, the Bayesian runs are expected to recover alternative
+\cite{pena2011}. Thus, the Bayesian runs are expected to recover alternative
 topologies on the posterior distribution of trees, resulting in low
@@
 posterior probability values for clades in Satyrini are very low
-{[}13{]}. As a result, MEDUSA inferred a net diversification rate
+\cite{wahlberg2009}. As a result, MEDUSA inferred a net diversification rate
 increase for part of Satyrini in the maximum clade credibility tree, but
@@
 The clade Ithomiini and the non-basal danaids are well supported by high
-posterior probability values in Wahlberg et al. (2009) {[}13{]}.
+posterior probability values in Wahlberg et al. (2009) \cite{wahlberg2009}.
 Therefore our MEDUSA analyses recovered an increase in net
@@
 Keith Brown suggested that feeding on Solanaceae was an important event
-in the diversification of Ithomiini butterflies {[}45{]}. Ithomiini
+in the diversification of Ithomiini butterflies \cite{brown1987}. Ithomiini
 butterflies are exclusively Neotropical and most species feed on
-Solanaceae hostplants during larval stage {[}8{]}. Optimizations of the
+Solanaceae hostplants during larval stage \cite{willmott2006}.
+Optimizations of the
 evolution of hostplant use on phylogenies evidence a probable shift from
-Apocynaceae to Solanaceae in the ancestor of the tribe {[}8,46{]}.
-Fordyce (2010) {[}12{]} found that the Gamma statistics, a LTT plot of
+Apocynaceae to Solanaceae in the ancestor of the tribe 
+\cite{willmott2006,brower2006}.
+Fordyce (2010) \cite{fordyce2010} found that the Gamma statistics, a LTT plot of
 an Ithomiini phylogeny and the fit of the density-dependent model of
@@
 due to the use of Solanaceae plants as hosts during larval stage. For
-this, we used a Bayesian approach {[}47{]} to test whether the trait
+this, we used a Bayesian approach \cite{fitzjohn2009} to test whether the trait
 ``feeding on Solanaceae'' had any effect on the diversification of the
@@
 agreement with previoues findings using other statistical methods
-{[}12{]}. Due to the fact that Ithomiini are virtually the only
+\cite{fordyce2010}. Due to the fact that Ithomiini are virtually the only
 nymphalids using Solanaceae as hostplants, it is possible that the trait
 responsible for a higher diversification of Ithomiini might not be the
-hostplant character. As noted by Maddison et al. (2007) {[}39{]}, the
-responsible trait might be a codistributed character such as a trait
+hostplant character. As noted by Maddison et al. (2007) \cite{maddison2007},
+the responsible trait might be a codistributed character such as a trait
 related to the ability to digest secondary metabolites.
@@
 that the high diversity of Ithomiini is consistent with the
-``escape-and-radiate scenario'' due to a shift onto Solanaceae {[}12{]}
+``escape-and-radiate scenario'' due to a shift onto Solanaceae 
+\cite{fordyce2010}
 and radiation scenarios among chemically different lineages of
-Solanaceae plants {[}8,45{]}. According to this theory, the shift from
+Solanaceae plants \cite{willmott2006,brown1987}.
+According to this theory, the shift from
 Apocynaceae to Solanaceae allowed Ithomiini to invade newly available
@@
 detoxification of ingested food as they have been found in other
-butterfly groups {[}48{]}.
+butterfly groups \cite{wheat2007}.
 
@@
 Although the Solanaceae genera used by the Ithomiini clades are well
-known {[}8{]}, we do not have any understanding on the physiological
+known \cite{willmott2006}, we do not have any understanding on the physiological
 routes involved in the detoxification of Solanaceae compounds by the
 several lineages of Ithomiini. We can speculate that older lineages
-exploiting a novel toxic resource {[}8,13{]} might not be too efficient
+exploiting a novel toxic resource \cite{willmott2006,wahlberg2009}
+might not be too efficient
 in metabolizing plant toxins and that younger lineages are able to deal
@@
 configurations of the proteins, providing more opportunities for
-hostplant switch {[}49{]}. This might be the reason why the basal
+hostplant switch \cite{li2003}. This might be the reason why the basal
 Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and
-restricted to few Solanaceae hosts {[}8{]}, while recent subtribes are
+restricted to few Solanaceae hosts \cite{willmott2006}, while recent subtribes are
 species-rich and have expanded their host range into several Solanaceae
-lineages {[}8{]}. It might be that the switch to feeding in Solanaceae
+lineages \cite{willmott2006}. It might be that the switch to feeding in Solanaceae
 was an important event in the evolutionary history of Ithomiini, but the
@@
 diversity'' and the ``escape-and-radiate'' hypotheses state that insects
-diversify after their hostplants {[}2--4{]}. Wheat et al. (2007)
-{[}48{]} found strong evidence for a model of speciation congruent with
+diversify after their hostplants \cite{nyman2012,ehrlich1964,janz2011}.
+Wheat et al. (2007)
+\cite{wheat2007} found strong evidence for a model of speciation congruent with
 Ehrlich and Raven's hypothesis in Pieridae butterflies due to, in
@@
 lag of circa 10 MY). According to a recent dated phylogeny of
-the Angiosperms {[}50{]}, the family Solanaceae split from its sister
+the Angiosperms \cite{bell2010}, the family Solanaceae split from its sister
 group about 59 (49--68) MYA and diversification started (crown group
-age) around 37 (29--47) MYA. Wahlberg et al. (2009) {[}13{]} give the
+age) around 37 (29--47) MYA. Wahlberg et al. (2009) \cite{wahlberg2009} give the
 ages for origin and diversification for Ithomiini at 45 (39--53) and 37
@@
 widespread species show higher genetic differentiation in their
-populations {[}51{]}.
+populations \cite{wahlberg2011}.
 
@@
 their main hostplant, grasses, during the climatic cooling in the
-Oligocene {[}14{]}. Thus, it is somewhat surprising that part of
+Oligocene \cite{pena2008}. Thus, it is somewhat surprising that part of
 Satyrini (the subtribes Euptychiina, Satyrina and Pronophilina) were
@@
 the posterior distribution. Although this can be attributed to low
-phylogenetic signal {[}13{]}, the clade Satyrini is very robust {[}13{]}
+phylogenetic signal \cite{wahlberg2009}, the clade Satyrini is very robust
+\cite{wahlberg2009}
 and MEDUSA failed to identify any significant accelerated net
@@
 diversified near simultaneously, which is in agreement with the diffuse
-co-speciation hypothesis {[}2,4{]}.
+co-speciation hypothesis \cite{nyman2012,janz2011}.
 
@@
  \item {\bf Supporting Information S02.R} R script to run MEDUSA on the
-MCC Nymphalidae tree from {[}13{]}. This script removes the outgroup
+MCC Nymphalidae tree from \cite{wahlberg2009}. This script removes the outgroup
 taxa and loads the richness data for the tree terminals.
@@
 
- \item {\bf Supporting Information S05 1000 random trees no outgroups.zip} 1000 random trees from {[}13{]}.
+ \item {\bf Supporting Information S05 1000 random trees no outgroups.zip} 1000 random trees from \cite{wahlberg2009}.
 
  \item {\bf Supporting Information S06.R} R script to run a MultiMEDUSA
-analysis on 1000 random trees from {[}13{]}.
+analysis on 1000 random trees from \cite{wahlberg2009}.
 

### 2014-11-10 MS_revised.tex
remove all biblio lines

@@
 
-\section*{References}
-1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
-adaptive zones: Has phytophagy promoted insect diversification? Am Nat
-132: 107--128.
-doi:\href{http://dx.doi.org/10.1086/284840}{10.1086/284840}.
-
-2. Nyman T, Linder HP, Pe{\~n}a C, Malm T, Wahlberg N (2012) Climate-driven
-diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
-889--898.
-doi:\href{http://dx.doi.org/10.1111/j.1461-0248.2012.01782.x}{10.1111/j.1461-0248.2012.01782.x}.
-
-3. Ehrlich PR, Raven PH (1964) Butterflies and plants: A study in
-coevolution. Evolution 18: 586--608.
-doi:\href{http://dx.doi.org/10.2307/2406212}{10.2307/2406212}.
-
-4. Janz N (2011) Ehrlich and Raven Revisited: Mechanisms Underlying
-Codiversification of Plants and Enemies. Annu Rev Ecol Evol S 42:
-71--89.
-doi:\href{http://dx.doi.org/10.1146/annurev-ecolsys-102710-145024}{10.1146/annurev-ecolsys-102710-145024}.
-
-5. Nylin S, Slove J, Janz N (2014) Host plant utilization, host range
-oscillations and diversification in Nymphalid butterflies: A
-phylogenetic investigation. Evolution 68: 105--124.
-doi:\href{http://dx.doi.org/10.1111/evo.12227}{10.1111/evo.12227}.
-
-6. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
-al. (2011) Order Lepidoptera. In: Animal biodiversity: An outline of
-higher-level classification and survey of taxonomic richness. Zootaxa.
-pp. 212--221.
-
-7. Joron M, Jiggins C, Papanicolaou A, McMillan W (2006) Heliconius wing
-patterns: An evo-devo model for understanding phenotypic diversity.
-Heredity 97: 157--167.
-
-8. Willmott KR, Freitas AVL (2006) Higher-level phylogeny of the
-Ithomiinae (Lepidoptera: Nymphalidae): classification, patterns of
-larval hostplant colonization and diversification. Cladistics 22:
-297--368.
-doi:\href{http://dx.doi.org/10.1111/j.1096-0031.2006.00108.x}{10.1111/j.1096-0031.2006.00108.x}.
-
-9. Brakefield PM, Beldade P, Zwaan BJ (2009) The african butterfly
-bicyclus anynana: A model for evolutionary genetics and evolutionary
-developmental biology. Cold Spring Harbor Protocols 2009: pdb--emo122.
-
-10. Heikkil{\"a} M, Kaila L, Mutanen M, Pe{\~n}a C, Wahlberg N (2012) Cretaceous
-origin and repeated tertiary diversification of the redefined
-butterflies. P Roy Soc B 279: 1093--1099.
-doi:\href{http://dx.doi.org/10.1098/rspb.2011.1430}{10.1098/rspb.2011.1430}.
-
-11. Elias M, Joron M, Willmott KR, Silva-Brand{\~a}o KL, Kaiser V, et al.
-(2009) Out of the Andes: Patterns of diversification in clearwing
-butterflies. Mol Ecol 18: 1716--1729.
-doi:\href{http://dx.doi.org/10.1111/j.1365-294X.2009.04149.x}{10.1111/j.1365-294X.2009.04149.x}.
-
-12. Fordyce JA (2010) Host shifts and evolutionary radiations of
-butterflies. P Roy Soc B 277: 3735--3743.
-doi:\href{http://dx.doi.org/10.1098/rspb.2010.0211}{10.1098/rspb.2010.0211}.
-
-13. Wahlberg N, Leneveu J, Kodandaramaiah U, Pe{\~n}a C, Nylin S, et al.
-(2009) Nymphalid butterflies diversify following near demise at the
-Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302.
-doi:\href{http://dx.doi.org/10.1098/rspb.2009.1303}{10.1098/rspb.2009.1303}.
-
-14. Pe{\~n}a C, Wahlberg N (2008) Prehistorical climate change increased
-diversification of a group of butterflies. Biol Lett 4: 274--278.
-doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.
-
-15. Ferrer-Paris JR, S{\'a}nchez-Mercado A, Viloria {\'A}L, Donaldson J (2013)
-Congruence and Diversity of Butterfly-Host Plant Associations at Higher
-Taxonomic Levels. PloS one 8: e63570.
-doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.
-
-16. Alfaro ME, Santini F, Brock C, Alamillo H, Dornburg A, et al. (2009)
-Nine exceptional radiations plus high turnover explain species diversity
-in jawed vertebrates. P Natl Acad Sci USA 106: 13410--13414.
-doi:\href{http://dx.doi.org/10.1073/pnas.0811087106}{10.1073/pnas.0811087106}.
-
-17. Harmon LJ, Rabosky DL, {FitzJohn} RG, Brown JW (2011) turboMEDUSA:
-MEDUSA: Modeling Evolutionary Diversification Using Stepwise AIC.
-
-18. {FitzJohn} RG (2012) Diversitree: comparative phylogenetic analyses
-of diversification in R. Method Ecol Evol 3: 1084--1092.
-doi:\href{http://dx.doi.org/10.1111/j.2041-210X.2012.00234.x}{10.1111/j.2041-210X.2012.00234.x}.
-
-19. Lamas G (2004) Checklist: Part 4A. Hesperioidea-Papilionoidea.
-Gainesville.
-
-20. Matos-Marav{\'i} PF, Pe{\~n}a C, Willmott KR, Freitas AV, Wahlberg N (2013)
-Systematics and evolutionary history of butterflies in the ``Taygetis
-clade'' (Nymphalidae: Satyrinae: Euptychiina): Towards a better
-understanding of Neotropical biogeography. Mol Phylogenet Evol 66:
-54--68.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2012.09.005}{10.1016/j.ympev.2012.09.005}.
-
-21. Brower AVZ, Wahlberg N, Ogawa JR, Boppr{\'e} M, Vane-Wright RI (2010)
-Phylogenetic relationships among genera of danaine butterflies
-(Lepidoptera: Nymphalidae) as implied by morphology and DNA sequences.
-Syst Biodivers 8: 75--89.
-doi:\href{http://dx.doi.org/10.1080/14772001003626814}{10.1080/14772001003626814}.
-
-22. Kodandaramaiah U, Lees DC, M{\"u}ller CJ, Torres E, Karanth KP, et al.
-(2010) Phylogenetics and biogeography of a spectacular Old World
-radiation of butterflies: the subtribe Mycalesina (Lepidoptera:
-Nymphalidae: Satyrini). BMC Evol Biol 10: 172.
-doi:\href{http://dx.doi.org/10.1186/1471-2148-10-172}{10.1186/1471-2148-10-172}.
-
-23. Kodandaramaiah U, Pe{\~n}a C, Braby MF, Grund R, M{\"u}ller CJ, et al.
-(2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
-problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.08.012}{10.1016/j.ympev.2009.08.012}.
-
-24. Ortiz-Acevedo E, Willmott KR (2013) Molecular systematics of the
-butterfly tribe Preponini (Nymphalidae: Charaxinae). Syst Entomol 38:
-440--449.
-doi:\href{http://dx.doi.org/10.1111/syen.12008}{10.1111/syen.12008}.
-
-25. De-Silva DL, Day JJ, Elias M, Willmott KR, Whinnett A, et al. (2010)
-Molecular phylogenetics of the Neotropical butterfly subtribe Oleriina
-(Nymphalidae: Danainae: Ithomiini). Mol Phylogenet Evol 55: 1032--1041.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2010.01.010}{10.1016/j.ympev.2010.01.010}.
-
-26. Freitas AVL, Brown Jr. KS (2004) Phylogeny of the Nymphalidae
-(Lepidoptera). Syst Biol 53: 363--383.
-doi:\href{http://dx.doi.org/10.1080/10635150490445670}{10.1080/10635150490445670}.
-
-27. Pe{\~n}a C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
-(2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
-Nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2006.02.007}{10.1016/j.ympev.2006.02.007}.
-
-28. Penz CM (1999) Higher level phylogeny for the passion-vine
-butterflies (Nymphalidae, Heliconiinae) based on early stage and adult
-morphology. Zool J Linn Soc 127: 277--344.
-doi:\href{http://dx.doi.org/10.1111/j.1096-3642.1999.tb00680.x}{10.1111/j.1096-3642.1999.tb00680.x}.
-
-29. Silva-Brand{\~a}o KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
-Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
-tribe Acraeini (Lepidoptera, Nymphalidae, Heliconiinae) and the
-evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2007.11.024}{10.1016/j.ympev.2007.11.024}.
-
-30. Pe{\~n}a C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
-butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
-methods. Zool J Linn Soc 161: 64--87.
-doi:\href{http://dx.doi.org/10.1111/j.1096-3642.2009.00627.x}{10.1111/j.1096-3642.2009.00627.x}.
-
-31. Pe{\~n}a C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
-history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
-Satyrinae). Zool Scr 39: 243--258.
-doi:\href{http://dx.doi.org/10.1111/j.1463-6409.2010.00421.x}{10.1111/j.1463-6409.2010.00421.x}.
-
-32. Ackery PR (1988) Hostplants and classification: A review of
-nymphalid butterflies. Biol J Linn Soc 33: 95--203.
-doi:\href{http://dx.doi.org/10.1111/j.1095-8312.1988.tb00446.x}{10.1111/j.1095-8312.1988.tb00446.x}.
-
-33. Dyer LA, Gentry GL (2002) Caterpillars and parasitoids of a tropical
-lowland wet forest. Available: \url{http://www.caterpillars.org}.
-Accessed 17 April 2013.
-
-34. Beccaloni G, Viloria {\'A}, Hall S, Robinson G (2008) Catalogue of the
-hostplants of the Neotropical butterflies. London: The Natural History
-Museum.
-
-35. Janzen DH, Hallwachs W (2009) Dynamic database for an inventory of
-the macrocaterpillar fauna, and its food plants and parasitoids, of Area
-de Conservacion Guanacaste (ACG), northwestern Costa Rica. Available:
-\url{http://janzen.sas.upenn.edu}. Accessed 17 April 2013.
-
-36. Core Team R (2013) R: A Language and Environment for Statistical
-Computing. Vienna, Austria: R Foundation for Statistical Computing.
-Available: \url{http://www.r-project.org}.
-
-37. Popescu A-A, Huber KT, Paradis E (2012) ape 3.0: New tools for
-distance-based phylogenetics and evolutionary analysis in R.
-Bioinformatics 28: 1536--1537.
-doi:\href{http://dx.doi.org/10.1093/bioinformatics/bts184}{10.1093/bioinformatics/bts184}.
-
-38. Harmon LJ, Weir JT, Brock CD, Glor RE, Challenger W (2008) GEIGER:
-investigating evolutionary radiations. Bioinformatics 24: 129--131.
-doi:\href{http://dx.doi.org/10.1093/bioinformatics/btm538}{10.1093/bioinformatics/btm538}.
-
-39. Maddison WP, Midford PE, Otto SP (2007) Estimating a binary
-character's effect on speciation and extinction. Syst Biol 56: 701--710.
-doi:\href{http://dx.doi.org/10.1080/10635150701607033}{10.1080/10635150701607033}.
-
-40. Scott JA (1986) The Butterflies of North America. Stanford
-University Press.
-
-41. Davis MP, Midford PE, Maddison W (2013) Exploring power and
-parameter estimation of the BiSSE method for analyzing species
-diversification. BMC Evol Biol 13: 38.
-doi:\href{http://dx.doi.org/10.1186/1471-2148-13-38}{10.1186/1471-2148-13-38}.
-
-42. Rabosky DL (2014) Automatic Detection of Key Innovations, Rate Shifts, and
-Diversity-Dependence on Phylogenetic Trees. PLoS ONE 9: e89543. doi:
-\href{http://dx.doi.org/10.1371/journal.pone.0089543}{10.1371/journal.pone.0089543}.
-
-43. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
-bees cut leaves? New insights into the early evolution of bees. P Roy
-Soc B 278: 3593--3600.
-doi:\href{http://dx.doi.org/10.1098/rspb.2011.0365}{10.1098/rspb.2011.0365}.
-
-44. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
-clades of Agaricales. P Roy Soc B 279: 2003--2011.
-doi:\href{http://dx.doi.org/10.1098/rspb.2011.2428}{10.1098/rspb.2011.2428}.
-
-45. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
-Interface. Ann Mo Bot Gard 74: 359--397.
-
-46. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand{\~a}o KL, Whinnett A, et
-al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
-Nymphalidae) inferred from one mitochondrial and two nuclear gene
-regions. Syst Entomol 31: 288--301.
-doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2006.00321.x}{10.1111/j.1365-3113.2006.00321.x}.
-
-47. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
-trait-dependent speciation and extinction rates from incompletely
-resolved phylogenies. Syst Biol 58: 595--611.
-doi:\href{http://dx.doi.org/10.1093/sysbio/syp067}{10.1093/sysbio/syp067}.
-
-48. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
-The genetic basis of a plant-insect coevolutionary key innovation. P
-Natl Acad Sci USA 104: 20427--20431.
-doi:\href{http://dx.doi.org/10.1073/pnas.0706229104}{10.1073/pnas.0706229104}.
-
-49. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
-furanocoumarin-metabolizing cytochrome P450 monooxygenases in two
-papilionids: Specificity and substrate encounter rate. P Natl Acad Sci
-USA 100: 14593--14598.
-doi:\href{http://dx.doi.org/10.1073/pnas.1934643100}{10.1073/pnas.1934643100}.
-
-50. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
-the angiosperms re-revisited. Am J Bot 97: 1296--1303.
-doi:\href{http://dx.doi.org/10.3732/ajb.0900346}{10.3732/ajb.0900346}.
-
-51. Wahlberg N, Rubinoff D (2011) Vagility across \emph{Vanessa} (Lepidoptera:
-Nymphalidae): mobility in butterfly species does not inhibit the
-formation and persistence of isolated sister taxa. Syst Entomol 36:
-362--370.
-doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2010.00566.x}{10.1111/j.1365-3113.2010.00566.x}.
-
 % Table 1

### 2014-11-10 MS_revised.tex
fixed table 2

@@
 ``oscillation hypothesis''
-\cite{janz2011,nylin2014} or ``diffuse cospeciation'' \cite{nyman2012}.
+\cite{janz2011,nylin2014} or ``diffuse cospeciation'' \cite{nyman2012} and the
+``musical chairs hypothesis'' \cite{hardy2014}.
 
@@
 diversification pattern of Nymphalidae is congruent with events of
-sudden diversification bursts due to hostplant shift, climatic events or
-shifts to closely related hostplants \cite{nylin2014,ferrer2013}.
+sudden diversification bursts due to hostplant shift or climatic events
+\cite{nylin2014,ferrer2013}.
 
@@
 investigate patterns
-of diversification. We applied MEDUSA \cite{alfaro2009,harmon2011},
-a recently developed
-statistical method, to study the diversification pattern of Nymphalidae
+of diversification. We applied the statistical method
+MEDUSA \cite{alfaro2009,harmon2011},
+to study the diversification pattern of Nymphalidae
 butterflies. MEDUSA fits likelihood models of diversification into a
@@
 
-\subsection*{Analyses of
-Diversification}
+\subsection*{Analyses of Diversification}
 
@@
 
-\subsection*{Detecting diversification shifts on phylogenetic
-trees}
+We analyzed our MCC tree in the Bayesian software BAMM in order to contrast
+the results
+found by MEDUSA and MultiMEDUSA.
+BAMM (Bayesian analysis of macroevolutionary mixtures)
+\cite{rabosky2014} is a
+software that fits several models of speciation and extinction. The best model
+can be selected by comparison of Bayes Factors.
+
+\subsection*{Detecting diversification shifts on phylogenetic trees}
 
@@
 the rates \texttt{r}: net diversification \texttt{=} speciation
-(\texttt{b}) \texttt{-} extinction (\texttt{d}) and \(\varepsilon\):
+(\texttt{b}) \texttt{-} extinction (\texttt{d}) and ($\varepsilon$:
 relative extinction \texttt{= d/b}. The AIC scores of the two-parameter
@@
 The MEDUSA analysis on the MCC tree in combination with richness data
-estimated 18 changes in the tempo of diversification in Nymphalidae
+estimated 18 significant changes in the tempo of diversification in Nymphalidae
 history (Fig. 1; Table 1). The corrected AICc acceptance threshold for
@@
 MultiMEDUSA approach recovered different rates of diversification than
-those found using when the MCC tree alone.
+those found when the MCC tree alone.
 
-There were four diversification shifts found with more than 0.90 of
-probability in the trees from the random sample (Table 2): (i) net
-diversification rate increase in the genus \emph{Ypthima}; (ii) net
-diversification rate increase in the genus \emph{Charaxes}; (iii) rate
-increase in Ithomiini subtribes Oleriina + Ithomiina + Napeogenina +
-Dircennina + Godyridina; and (iv) rate increase in \emph{Callicore} +
-\emph{Diaethria}.
+There were four diversification shifts found in the trees from the
+random sample (Table 2): (i) net
+diversification rate increase in the genus \emph{Ypthima} ($r = 0.22$);
+(ii) net diversification rate increase in the genus \emph{Charaxes}
+($r = 0.21$); (iii) rate increase in Ithomiini subtribes Oleriina +
+Ithomiina + Napeogenina + Dircennina + Godyridina ($r = 0.10$);
+and (iv) rate increase in \emph{Callicore} + \emph{Diaethria} ($r = 0.135$).
 
 MultiMEDUSA provided mean and standard deviation statistics for the
-diversification values found on the shifts on the 1000 trees (supp. mat.
+diversification values on the shifts on the 1000 trees (supp. mat.
 13--14), and found that some of the changes in net diversification rate
@@
 when phylogenetic uncertainty is taken into account.
+Moreover, we also found that the results from BAMM, without taking into account
+phylogenetic uncertainty, correspond better to those of MultiMEDUSA than MEDUSA.
 
@@
 
- \item {\bf Supporting Information S08.pdf} Summary of the MulitMEDUSA
+ \item {\bf Supporting Information S08.pdf} Summary of the MultiMEDUSA
 analysis on the 1000 trees from the posterior distribution.
@@
     \caption{\bf{Differences in rates estimated by MEDUSA on the MCC
-tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
-from the posterior distribution.}}
+tree from the sample of trees from the posterior distribution and the
+MultiMEDUSA approach. Shift consistently recovered across the
+sample of trees in bold face.}}
 \begin{tabular}{lcrr}
 Split.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
-1          & 0.090          & not found                  & 0.000                          \\
-2          & 0.054          & -0.030                     & 0.864                          \\
-3          & 0.311          & 0.219                      & 0.961                          \\
-4          & 0.290          & 0.212                      & 0.996                          \\
-5          & 0.187          & 0.101                      & 0.927                          \\
-6          & 0.116          & 0.039                      & 0.131                          \\
-7          & 0.064          & -0.052                     & 0.195                          \\
-8          & 0.240          & 0.166                      & 0.319                          \\
+1          & 0.092          & not found                  & 0.000                          \\
+2          & 0.055          & -0.030                     & 0.864                          \\
+\bf{3}     & 0.184          & \bf{0.219}                 & \bf{0.961}                     \\
+\bf{4}     & 0.166          & \bf{0.212}                 & \bf{0.996}                     \\
+\bf{5}     & 0.111          & \bf{0.101}                 & \bf{0.927}                     \\
+6          & 0.119          & 0.039                      & 0.131                          \\
+7          & 0.066          & -0.052                     & 0.195                          \\
+8          & 0.232          & 0.166                      & 0.319                          \\
 9          & 0.042          & -0.049                     & 0.897                          \\
-10         & 0.209          & 0.149                      & 0.619                          \\
+10         & 0.058          & 0.149                      & 0.619                          \\
 11         & 0.311          & 0.208                      & 0.831                          \\
-12         & 0.219          & 0.135                      & 0.911                          \\
-13         & 0.220          & 0.117                      & 0.276                          \\
-14         & 0.190          & 0.115                      & 0.379                          \\
-15         & 0.234          & 0.127                      & 0.825                          \\
-16         & 0.000          & -0.005                     & 0.659                          \\
-17         & 0.311          & 0.248                      & 0.785                          \\
-18         & 0.024          & -0.064                     & 0.024                          \\
-19         & 0.000          & -0.087                     & 0.381                         
+\bf{12}    & 0.219          & \bf{0.135}                 & \bf{0.911}                     \\
+13         & 0.082          & 0.117                      & 0.276                          \\
+14         & 0.099          & 0.115                      & 0.379                          \\
+15         & 0.113          & 0.127                      & 0.825                          \\
+16         & 0.222          & -0.005                     & 0.659                          \\
+17         & 0.243          & 0.248                      & 0.785                          \\
+18         & 0.192          & -0.064                     & 0.024                          \\
+19         & 0.064          & -0.087                     & 0.381                         
 \end{tabular}

### 2014-11-10 MS_revised.tex
mcc tree derived from the random sample

@@
 lower median net diversification rates for the diversification shifts
-found by MEDUSA on the MCC tree from \cite{wahlberg2009} (Table 2). Although the
+found by MEDUSA on the MCC tree derived from the random sample of trees
+(Table 2). Although the
 diversification pattern found by MEDUSA and MultiMEDUSA was the same,

### 2014-11-10 MS_revised.tex
fixing several typos

@@
 
-We contrasted these results with those obtained from analysing the data in
+We contrasted these results with those obtained from analyzing the data in
 the software BAMM \cite{rabosky2014}. We found 
@@
 trait ``feeding on Solanaceae hostplants'' (Fig. 4). This is in
-agreement with previoues findings using other statistical methods
+agreement with previous findings using other statistical methods
 \cite{fordyce2010}. Due to the fact that Ithomiini are virtually the only
@@
 
-The increase in diversification inferred by MEDUSA ocurred after the
+The increase in diversification inferred by MEDUSA occurred after the
 probable shift from Apocynaceae to Solanaceae, as the Solanaceae feeders
@@
 diversified near simultaneously, which is in agreement with the diffuse
-co-speciation hypothesis \cite{nyman2012,janz2011}.
+cospeciation hypothesis \cite{nyman2012,janz2011}.
 
@@
 nodes indicate the diversification shift number as found by MEDUSA.
-Numbers next to circles indicat the posterior probability values for
+Numbers next to circles indicate the posterior probability values for
 such nodes.

### 2014-11-11 MS_revised.tex
rm in .+ in

@@
 \usepackage{microtype}
+\usepackage{lineno}
+\linenumbers
 \DisableLigatures[f]{encoding = *, family = * }
@@
 increase for part of Satyrini in the maximum clade credibility tree, but
-this was recovered in the MultiMEDUSA analysis in only in 13\% of the
+this was recovered in the MultiMEDUSA analysis in only 13\% of the
 random sample of trees.
@@
 when phylogenetic uncertainty is taken into account.
-Moreover, we also found that the results from BAMM, without taking into account
-phylogenetic uncertainty, correspond better to those of MultiMEDUSA than MEDUSA.
+The results from BAMM recovered only two significant shifts in our MCC tree
+as this method is more conservative than the MEDUSA approach. One of these 
+shifts is the Satyrini clade, which was recovered by MEDUSA in a few trees 
+from the posterior distribution. Thus, it might be important to
+take into account phylogenetic uncertainty in BAMM as well.
 

### 2014-11-11 MS_revised.tex
moved BAMM text to discussion
Fix Satyrini, it should be Satyrinae in many cases

@@
 (\texttt{r = 0.191}), Oleriina + Ithomiina + Napeogenina + Dircennina +
-Godyridina (\texttt{r = 0.187}), Satyrini (\texttt{r = 0.116}),
+Godyridina (\texttt{r = 0.187}), Satyrinae (\texttt{r = 0.116}),
 Phyciodina in part (\texttt{r = 0.241}) and Satyrina
@@
 of the splits in the MCC tree are recovered in very few trees, for
-example the split for the clade Satyrini (Euptychiina + Pronophilina +
-Satyrina + Maniolina) is recovered with a probability of 0.18 (Fig. 3).
+example the split for the clade Satyrinae is recovered with a probability
+of 0.18 (Fig. 3).
 
@@
 support and posterior probability values for the nodes. For example,
-posterior probability values for clades in Satyrini are very low
+posterior probability values for clades in Satyrinae are very low
 \cite{wahlberg2009}. As a result, MEDUSA inferred a net diversification rate
-increase for part of Satyrini in the maximum clade credibility tree, but
+increase for the Satyrinae (which includes Satyrini and its sister clade)
+in the maximum clade credibility tree, but
 this was recovered in the MultiMEDUSA analysis in only 13\% of the
@@
 distribution of trees (Fig. 3).
+
+The results from BAMM recovered only two significant shifts in our MCC tree
+as this method seems to be much more conservative than the MEDUSA approach.
+One of these 
+shifts is the Satyrini clade. MEDUSA estimated its parent node as
+significant shift in a few trees from the posterior distribution. 
+Thus, it might be important to
+take into account phylogenetic uncertainty in BAMM as well.
 
@@
 
-\subsubsection*{Satyrini}
+\subsubsection*{Satyrinae}
 
-The diverse tribe Satyrini radiated simultaneously with the radiation of
+Lineages in the diverse family Satyrinae radiated simultaneously with the
+radiation of
 their main hostplant, grasses, during the climatic cooling in the
 Oligocene \cite{pena2008}. Thus, it is somewhat surprising that part of
-Satyrini (the subtribes Euptychiina, Satyrina and Pronophilina) were
+Satyrinae were
 found to have accelerated diversification in only 13\% of the trees from
@@
 Satyrini as a whole was not remarkably fast and therefore not picked up
-by MEDUSA. This should be expected if the diversification of Satyrinae
+by MEDUSA, although it estimated a diversification shift for Satyrini + its
+sister clade. This should be expected if the diversification of Satyrini
 occurred in a stepwise manner, with pulses or bursts of diversification
@@
 when phylogenetic uncertainty is taken into account.
-The results from BAMM recovered only two significant shifts in our MCC tree
-as this method is more conservative than the MEDUSA approach. One of these 
-shifts is the Satyrini clade, which was recovered by MEDUSA in a few trees 
-from the posterior distribution. Thus, it might be important to
-take into account phylogenetic uncertainty in BAMM as well.
 
@@
 Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
-part, 6) Satyrini, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
+part, 6) Satyrinae, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
 part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
@@
 5               & 750        & yule  & 0.186913   & -147.4146  &          & Oleriina + Ithomiina + Napeogenina + Dircennina + Godyrina \\
-6               & 405        & yule  & 0.116252   & -555.0276  &          & Satyrini                                                   \\
+6               & 405        & yule  & 0.116252   & -555.0276  &          & Satyrinae                                                   \\
 7               & 495        & yule  & 0.064656   & -124.9143  &          & Coenonymphina                                              \\

### 2014-11-11 MS_revised.tex
adding BAMM results as fig S01

@@
 our MCC tree (posterior probability = 0.79) for the Ithomiini
- and Satyrini. Bayes factor (BF) values strongly 
+and Satyrini (Fig. S1). Bayes factor (BF) values strongly 
 favored this model in comparison with a constant rate model (no diversification
@@
 \texttt{r = 0.11} while the net diversification rate for the Solanaceae
-feeders was \texttt{r = 0.16} (see Fig. S1 for a boxplot of speciation
+feeders was \texttt{r = 0.16} (see Fig. S2 for a boxplot of speciation
 and extinction values for the 95\% credibility intervals). The same
@@
 non-Solanaceae feeders due to dubious records produced the same pattern
-and net diversification rates (Fig. S2). Therefore, the rest of the
+and net diversification rates (Fig. S3). Therefore, the rest of the
 analyses were performed assuming these two genera as Solanaceae feeders.
@@
 credibility tree (combined mcmc run in supp. mat. 20; profiles plot of
-speciation rates in Fig. S3; boxplot of 95\% credibility intervals in
-Fig. S4). A BiSSE analysis to test whether the trait ``feeding on
+speciation rates in Fig. S4; boxplot of 95\% credibility intervals in
+Fig. S5). A BiSSE analysis to test whether the trait ``feeding on
 Apocynaceae'' had any effect on increased net diversification rates
 found similar speciation rates for lineages feeding on Apocynaceae and
-other plants (Fig. S5).
+other plants (Fig. S6).
 
@@
 
- \item {\bf Figure S1.} Boxplot of speciation ($\lambda$) and extinction
+ \item {\bf Figure S1.} Best diversification model as estimated by BAMM. Two
+diversification shifts were recovered (Ithomiini and Satyrini). The MEDUSA 
+analysis found a diversification shifts in Satyrinae (the parent node of 
+Satyini).
+
+ \item {\bf Figure S2.} Boxplot of speciation ($\lambda$) and extinction
 ($\mu$) values for the 95\% credibility intervals of values estimated
@@
 
- \item {\bf Figure S2.} BiSSE analysis of diversification of nymphalids due
+ \item {\bf Figure S3.} BiSSE analysis of diversification of nymphalids due
 to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
@@
 
- \item {\bf Figure S3.} Net diversification rates of nymphalids feeding on
+ \item {\bf Figure S4.} Net diversification rates of nymphalids feeding on
 Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
@@
 
- \item {\bf Figure S4.} Boxplot of speciation and extinction values for the
+ \item {\bf Figure S5.} Boxplot of speciation and extinction values for the
 95\% credibility intervals of values estimated by BiSSE analysis of
@@
 
- \item {\bf Figure S5.} BiSSE analysis of diversification of nymphalids due
+ \item {\bf Figure S6.} BiSSE analysis of diversification of nymphalids due
 to feeding on Apocynaceae hostplants. Speciation and net diversification

### 2014-11-12 MS_revised.tex
in only a few
ready for resubmittion
one ref was missing

@@
 \usepackage{hyperref}
+\usepackage{lineno}
 \usepackage{microtype}
-\usepackage{lineno}
-\linenumbers
 \DisableLigatures[f]{encoding = *, family = * }
@@
 shifts is the Satyrini clade. MEDUSA estimated its parent node as
-significant shift in a few trees from the posterior distribution. 
+significant shift in only a few trees from the posterior distribution. 
 Thus, it might be important to
@@
  \item {\bf Supporting Information S01.nex.} MCC Nymphalidae tree from
-{[}13{]}
+\cite{wahlberg2009}
 

### 2014-11-12 MS_revised.tex
edited refs so they show up in PLOS resubmittion system

@@
 \usepackage{hyperref}
-\usepackage{lineno}
 \usepackage{microtype}
@@
 Hostplant shifts have been invoked to be responsible for a great part of
-the biodiversity of herbivorous insects \cite{mitter1988}. The study of the
+the biodiversity of herbivorous insects {[}1{]}. The study of the
 evolution of hostplant use has spawned several theories explaining the
-evolutionary interactions between plants and insects \cite{nyman2012}. The
-``escape-and-radiate'' hypothesis \cite{ehrlich1964}, the 
-``oscillation hypothesis''
-\cite{janz2011,nylin2014} or ``diffuse cospeciation'' \cite{nyman2012} and the
-``musical chairs hypothesis'' \cite{hardy2014}.
+evolutionary interactions between plants and insects {[}2{]}. The
+``escape-and-radiate'' hypothesis {[}3{]}, the ``oscillation
+hypothesis'' {[}4,5{]} or ``diffuse cospeciation'' {[}2{]} and the
+``musical chairs hypothesis'' {[}6{]}.
 
@@
 developing some of the mentioned hypotheses. Nymphalidae contains around
-6000 species \cite{van_nieukerken2011},
-and several members are considered model organisms
-in evolutionary biology \cite{joron2006,willmott2006,brakefield2009}.
-The family most likely originated
+6000 species {[}7{]}, and several members are considered model organisms
+in evolutionary biology {[}8--10{]}. The family most likely originated
 around 94 MYA in the mid Cretaceous. Diversification of the group began
@@
 appeared shortly after the Cretaceous-Paleogene (K-Pg) boundary
-\cite{heikkila2012}. Several studies have used time-calibrated phylogenies and
+{[}11{]}. Several studies have used time-calibrated phylogenies and
 diversification models to reconstruct the evolutionary history of the
 group to identify patterns of accelerated or decelerated diversification
-of some Nymphalidae clades
-\cite{heikkila2012,elias2009,fordyce2010,wahlberg2009}.
-For example, it has been
+of some Nymphalidae clades {[}11--14{]}. For example, it has been
 suggested that climate change in the Oligocene and the subsequent
 diversification of grasses has led to diversification of the subfamily
-Satyrinae \cite{pena2008} due to the abundance of grasses over extensive
+Satyrinae {[}15{]} due to the abundance of grasses over extensive
 geographic areas (``resource abundance-dependent diversity dynamics''
-hypothesis). Fordyce (2010) \cite{fordyce2010} found increased net diversification
+hypothesis). Fordyce (2010) {[}13{]} found increased net diversification
 rates in some Nymphalidae lineages after a major hostplant shift, which
 appears to be in agreement with the ``escape-and-radiate'' model of
-diversification \cite{ehrlich1964}.
+diversification {[}3{]}.
 
@@
 sudden diversification bursts due to hostplant shift or climatic events
-\cite{nylin2014,ferrer2013}.
+{[}5,16{]}.
 
 In this study, we used a time-calibrated genus-level phylogenetic
-hypothesis for Nymphalidae butterflies \cite{wahlberg2009} to
-investigate patterns
-of diversification. We applied the statistical method
-MEDUSA \cite{alfaro2009,harmon2011},
-to study the diversification pattern of Nymphalidae
+hypothesis for Nymphalidae butterflies {[}14{]} to investigate patterns
+of diversification. We applied the statistical method MEDUSA
+{[}17,18{]}, to study the diversification pattern of Nymphalidae
 butterflies. MEDUSA fits likelihood models of diversification into a
@@
 study the effects of phylogenetic uncertainty and by using the extended
-MEDUSA method called MultiMEDUSA \cite{alfaro2009}. We also tested whether
+MEDUSA method called MultiMEDUSA {[}17{]}. We also tested whether
 hostplant association dynamics can explain the diversification patterns
@@
 employing the method BiSSE as implemented in the R package
-\texttt{diversitree} \cite{fitzjohn2012}.
+\texttt{diversitree} {[}19{]}.
 
@@
 For analyses, we used the phylogenetic trees from the study of Wahlberg
-et al. (2009) \cite{wahlberg2009} that were generated using DNA sequence data from
+et al. (2009) {[}14{]} that were generated using DNA sequence data from
 10 gene regions for 398 of the 540 valid genera in Nymphalidae. We
@@
 as a random sample of 1000 trees from their BEAST run after burnin
-\cite{wahlberg2009}. Their original BEAST run was for 40 million generations. We
+{[}14{]}. Their original BEAST run was for 40 million generations. We
 used a burnin of 25 million generations and took a random sample of 1000
@@
 sources including the specialist-curated lists on
-\url{http://tolweb.org}, Lamas (2004) \cite{lamas2004} and curated lists of
+\url{http://tolweb.org}, Lamas (2004) {[}20{]} and curated lists of
 Global Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}).
@@
 to the closest related genus that was included in Wahlberg et al. (2009)
-\cite{wahlberg2009} study according to available phylogenetic studies 
-\cite{matos2013,brower2010,kodandaramaiah2010,kodandaramaiah2010a,ortiz2013,desilva2010,freitas2004,pena2006,penz1999,silva2008,pena2011,pena2010},
+{[}14{]} study according to available phylogenetic studies {[}21--32{]},
 taxonomical classification and morphological resemblance when no
@@
 Hostplant data for Nymphalidae species were compiled from several
-sources including \cite{ackery1988}, HOSTS database (\url{http://bit.ly/YI7nwW}),
-\cite{dyer2002,beccaloni2008,janzen2009} and others (supp. mat. 15--16)
-for a total of 6586
+sources including {[}33{]}, HOSTS database (\url{http://bit.ly/YI7nwW}),
+{[}34--36{]} and others (supp. mat. 15--16) for a total of 6586
 hostplant records, including 428 Nymphalidae genera and 143 plant
@@
 
-We used the statistical software R version 3.0.1 \cite{r2013} in combination
-with the APE \cite{popescu2012}, GEIGER \cite{harmon2008}, MEDUSA 
-\cite{alfaro2009} and
-\texttt{diversitree} \cite{fitzjohn2012} packages along with our own scripts to
+We used the statistical software R version 3.0.1 {[}37{]} in combination
+with the APE {[}38{]}, GEIGER {[}39{]}, MEDUSA {[}17{]} and
+\texttt{diversitree} {[}19{]} packages along with our own scripts to
 perform the analyses (included as supplementary materials 2, 6, 10). All
 analyses were run on the 1000 random trees from Wahlberg et al. (2009)
-\cite{wahlberg2009}, on the maximum clade credibility tree derived from these and
-the MCC tree \cite{wahlberg2009}.
-
-We analyzed our MCC tree in the Bayesian software BAMM in order to contrast
-the results
-found by MEDUSA and MultiMEDUSA.
-BAMM (Bayesian analysis of macroevolutionary mixtures)
-\cite{rabosky2014} is a
-software that fits several models of speciation and extinction. The best model
-can be selected by comparison of Bayes Factors.
+{[}14{]}, on the maximum clade credibility tree derived from these and
+the MCC tree {[}14{]}.
+
+We analyzed our MCC tree in the Bayesian software BAMM in order to
+contrast the results found by MEDUSA and MultiMEDUSA. BAMM (Bayesian
+analysis of macroevolutionary mixtures) {[}40{]} is a software that fits
+several models of speciation and extinction. The best model can be
+selected by comparison of Bayes Factors.
 
@@
 Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
-version 093.4.33 \cite{alfaro2009} on the maximum clade credibility tree from
-Wahlberg et al. (2009) \cite{wahlberg2009}. MEDUSA fits alternative birth-death
+version 093.4.33 {[}17{]} on the maximum clade credibility tree from
+Wahlberg et al. (2009) {[}14{]}. MEDUSA fits alternative birth-death
 likelihood models to a phylogenetic tree in order to estimate changes in
@@
 the rates \texttt{r}: net diversification \texttt{=} speciation
-(\texttt{b}) \texttt{-} extinction (\texttt{d}) and ($\varepsilon$:
+(\texttt{b}) \texttt{-} extinction (\texttt{d}) and (\(\varepsilon\):
 relative extinction \texttt{= d/b}. The AIC scores of the two-parameter
@@
 value. MEDUSA finds the likelihood of the models after taking into
-account branch lengths and number of species per lineage 
-\cite{alfaro2009}. Most
+account branch lengths and number of species per lineage {[}17{]}. Most
 studies using MEDUSA only run the method on a single tree, usually the
@@
 therefore selected 1000 random genus-level trees from the posterior
-distribution of the Bayesian run from Wahlberg et al., 2009 \cite{wahlberg2009}
+distribution of the Bayesian run from Wahlberg et al., 2009 {[}14{]}
 (MultiMEDUSA, supp. mat. 5). We calculated a new MCC tree from the
@@
 net diversification (\texttt{r = b - d}) and relative extinction rate
-($\varepsilon$ \texttt{= d/b}) values across the 1000 trees for the
+(\(\varepsilon\) \texttt{= d/b}) values across the 1000 trees for the
 nodes where changes in diversification tempo occurs. We used the AICc
@@
 and performed analyses using the ``binary state speciation and
-extinction'' \cite{maddison2007} Bayesian approach as implemented in the R package
-\texttt{diversitree} \cite{fitzjohn2012}. The Markov Chain Monte Carlo algorithm
+extinction'' {[}41{]} Bayesian approach as implemented in the R package
+\texttt{diversitree} {[}19{]}. The Markov Chain Monte Carlo algorithm
 was run for 10000 generations discarding the first 7500 as burnin. The
-Multiple State Speciation and Extinction approach (MuSSE) 
-\cite{fitzjohn2012} is
+Multiple State Speciation and Extinction approach (MuSSE) {[}19{]} is
 designed to examine the joint effects of two or more traits on
@@
 binary character state has had any effect on increased net
-diversification rate for a clade \cite{maddison2007}. We used our compiled data of
+diversification rate for a clade {[}41{]}. We used our compiled data of
 hostplant use to produce a binary dataset for the character ``feeding on
@@
 1) (supp. mat. 17). The family Solanaceae include the hostplants of the
-diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. Other
+diverse Ithomiini butterflies and closest relatives {[}9{]}. Other
 hostplant shifts were not tested for effect on net diversification rates
@@
 distribution. The records of \emph{Vanessa} and \emph{Hypanartia}
-feeding on Solanaceae \cite{beccaloni2008,scott1986} might be incorrect as it
-is unlikely
+feeding on Solanaceae {[}35,42{]} might be incorrect as it is unlikely
 that these species can be feeding on this plant family. We tested
@@
 or presence would not have a significant effect on our results. It has
-been shown that BiSSE performs poorly under certain conditions 
-\cite{davis2013}.
+been shown that BiSSE performs poorly under certain conditions {[}43{]}.
 However, our data has adequate number of taxa under analysis (more than
@@
 of Solanaceae hostplants. Thus, BiSSE is expected to produce robust
-results \cite{davis2013}.
+results {[}43{]}.
 
@@
 The MEDUSA analysis on the MCC tree in combination with richness data
-estimated 18 significant changes in the tempo of diversification in Nymphalidae
-history (Fig. 1; Table 1). The corrected AICc acceptance threshold for
-adding subsequent piecewise birth-death processes to the overall model
-was set to 7.8 units, as prescribed by MEDUSA. In all MEDUSA analyses,
-the maximum number of inferred diversification splits in all trees was
-26. The background net diversification rate for Nymphalidae was
-estimated as \texttt{r = 0.092} lineages per Million of years and the
-AICc score for the best fit model was \texttt{5090.5} (Table 1). MEDUSA
-also estimated that the basic constant birth-death model was not a
-better explanation for our data (AICc \texttt{= 5449.3}).
+estimated 18 significant changes in the tempo of diversification in
+Nymphalidae history (Fig. 1; Table 1). The corrected AICc acceptance
+threshold for adding subsequent piecewise birth-death processes to the
+overall model was set to 7.8 units, as prescribed by MEDUSA. In all
+MEDUSA analyses, the maximum number of inferred diversification splits
+in all trees was 26. The background net diversification rate for
+Nymphalidae was estimated as \texttt{r = 0.092} lineages per Million of
+years and the AICc score for the best fit model was \texttt{5090.5}
+(Table 1). MEDUSA also estimated that the basic constant birth-death
+model was not a better explanation for our data (AICc
+\texttt{= 5449.3}).
 
@@
 
-We contrasted these results with those obtained from analyzing the data in
-the software BAMM \cite{rabosky2014}. We found 
-that the best model (best configuration) estimated 2 diversification shifts in
-our MCC tree (posterior probability = 0.79) for the Ithomiini
-and Satyrini (Fig. S1). Bayes factor (BF) values strongly 
-favored this model in comparison with a constant rate model (no diversification
-shifts, BF $>$ 1,884).
+We contrasted these results with those obtained from analyzing the data
+in the software BAMM {[}40{]}. We found that the best model (best
+configuration) estimated 2 diversification shifts in our MCC tree
+(posterior probability = 0.79) for the Ithomiini and Satyrini (Fig. S1).
+Bayes factor (BF) values strongly favored this model in comparison with
+a constant rate model (no diversification shifts, BF \(>\) 1,884).
 
@@
 found by MEDUSA on the MCC tree derived from the random sample of trees
-(Table 2). Although the
-diversification pattern found by MEDUSA and MultiMEDUSA was the same,
-the latter consistently estimated lower rates. Furthermore, the shifts
-recovered with low net diversification rate on the MCC were recovered
-with negative net diversification rate by MultiMEDUSA. The background
-diversification and all shifts found by MEDUSA on the 1000 trees are
-provided as an R object in supp. mat. 7.
+(Table 2). Although the diversification pattern found by MEDUSA and
+MultiMEDUSA was the same, the latter consistently estimated lower rates.
+Furthermore, the shifts recovered with low net diversification rate on
+the MCC were recovered with negative net diversification rate by
+MultiMEDUSA. The background diversification and all shifts found by
+MEDUSA on the 1000 trees are provided as an R object in supp. mat. 7.
 
@@
 There were four diversification shifts found in the trees from the
-random sample (Table 2): (i) net
-diversification rate increase in the genus \emph{Ypthima} ($r = 0.22$);
-(ii) net diversification rate increase in the genus \emph{Charaxes}
-($r = 0.21$); (iii) rate increase in Ithomiini subtribes Oleriina +
-Ithomiina + Napeogenina + Dircennina + Godyridina ($r = 0.10$);
-and (iv) rate increase in \emph{Callicore} + \emph{Diaethria} ($r = 0.135$).
+random sample (Table 2): (i) net diversification rate increase in the
+genus \emph{Ypthima} (\(r = 0.22\)); (ii) net diversification rate
+increase in the genus \emph{Charaxes} (\(r = 0.21\)); (iii) rate
+increase in Ithomiini subtribes Oleriina + Ithomiina + Napeogenina +
+Dircennina + Godyridina (\(r = 0.10\)); and (iv) rate increase in
+\emph{Callicore} + \emph{Diaethria} (\(r = 0.135\)).
 
@@
 of the splits in the MCC tree are recovered in very few trees, for
-example the split for the clade Satyrinae is recovered with a probability
-of 0.18 (Fig. 3).
+example the split for the clade Satyrinae is recovered with a
+probability of 0.18 (Fig. 3).
 
@@
 The MEDUSA method has been used to infer changes in net diversification
-rates in a phylogenetic tree. Since its publication \cite{alfaro2009} the results
+rates in a phylogenetic tree. Since its publication {[}17{]} the results
 of using MEDUSA on a single tree, the maximum clade credibility tree,
 have been used for generation of hypotheses and discussion
-\cite{heikkila2012,litman2011,ryberg2012}. However, different diversification
-shifts and different
+{[}11,44,45{]}. However, different diversification shifts and different
 rates of diversification are found for certain lineages when
@@
 taxa appear to be strongly affected by long-branch attraction artifacts
-\cite{pena2011}. Thus, the Bayesian runs are expected to recover alternative
+{[}31{]}. Thus, the Bayesian runs are expected to recover alternative
 topologies on the posterior distribution of trees, resulting in low
@@
 posterior probability values for clades in Satyrinae are very low
-\cite{wahlberg2009}. As a result, MEDUSA inferred a net diversification rate
-increase for the Satyrinae (which includes Satyrini and its sister clade)
-in the maximum clade credibility tree, but
-this was recovered in the MultiMEDUSA analysis in only 13\% of the
-random sample of trees.
+{[}14{]}. As a result, MEDUSA inferred a net diversification rate
+increase for the Satyrinae (which includes Satyrini and its sister
+clade) in the maximum clade credibility tree, but this was recovered in
+the MultiMEDUSA analysis in only 13\% of the random sample of trees.
 
@@
 The clade Ithomiini and the non-basal danaids are well supported by high
-posterior probability values in Wahlberg et al. (2009) \cite{wahlberg2009}.
+posterior probability values in Wahlberg et al. (2009) {[}14{]}.
 Therefore our MEDUSA analyses recovered an increase in net
@@
 
-The results from BAMM recovered only two significant shifts in our MCC tree
-as this method seems to be much more conservative than the MEDUSA approach.
-One of these 
-shifts is the Satyrini clade. MEDUSA estimated its parent node as
-significant shift in only a few trees from the posterior distribution. 
-Thus, it might be important to
-take into account phylogenetic uncertainty in BAMM as well.
+The results from BAMM recovered only two significant shifts in our MCC
+tree as this method seems to be much more conservative than the MEDUSA
+approach. One of these shifts is the Satyrini clade. MEDUSA estimated
+its parent node as significant shift in only a few trees from the
+posterior distribution. Thus, it might be important to take into account
+phylogenetic uncertainty in BAMM as well.
 
@@
 Keith Brown suggested that feeding on Solanaceae was an important event
-in the diversification of Ithomiini butterflies \cite{brown1987}. Ithomiini
+in the diversification of Ithomiini butterflies {[}46{]}. Ithomiini
 butterflies are exclusively Neotropical and most species feed on
-Solanaceae hostplants during larval stage \cite{willmott2006}.
-Optimizations of the
+Solanaceae hostplants during larval stage {[}9{]}. Optimizations of the
 evolution of hostplant use on phylogenies evidence a probable shift from
-Apocynaceae to Solanaceae in the ancestor of the tribe 
-\cite{willmott2006,brower2006}.
-Fordyce (2010) \cite{fordyce2010} found that the Gamma statistics, a LTT plot of
+Apocynaceae to Solanaceae in the ancestor of the tribe {[}9,47{]}.
+Fordyce (2010) {[}13{]} found that the Gamma statistics, a LTT plot of
 an Ithomiini phylogeny and the fit of the density-dependent model of
@@
 due to the use of Solanaceae plants as hosts during larval stage. For
-this, we used a Bayesian approach \cite{fitzjohn2009} to test whether the trait
+this, we used a Bayesian approach {[}48{]} to test whether the trait
 ``feeding on Solanaceae'' had any effect on the diversification of the
@@
 agreement with previous findings using other statistical methods
-\cite{fordyce2010}. Due to the fact that Ithomiini are virtually the only
+{[}13{]}. Due to the fact that Ithomiini are virtually the only
 nymphalids using Solanaceae as hostplants, it is possible that the trait
 responsible for a higher diversification of Ithomiini might not be the
-hostplant character. As noted by Maddison et al. (2007) \cite{maddison2007},
-the responsible trait might be a codistributed character such as a trait
+hostplant character. As noted by Maddison et al. (2007) {[}41{]}, the
+responsible trait might be a codistributed character such as a trait
 related to the ability to digest secondary metabolites.
@@
 that the high diversity of Ithomiini is consistent with the
-``escape-and-radiate scenario'' due to a shift onto Solanaceae 
-\cite{fordyce2010}
+``escape-and-radiate scenario'' due to a shift onto Solanaceae {[}13{]}
 and radiation scenarios among chemically different lineages of
-Solanaceae plants \cite{willmott2006,brown1987}.
-According to this theory, the shift from
+Solanaceae plants {[}9,46{]}. According to this theory, the shift from
 Apocynaceae to Solanaceae allowed Ithomiini to invade newly available
@@
 detoxification of ingested food as they have been found in other
-butterfly groups \cite{wheat2007}.
+butterfly groups {[}49{]}.
 
@@
 Although the Solanaceae genera used by the Ithomiini clades are well
-known \cite{willmott2006}, we do not have any understanding on the physiological
+known {[}9{]}, we do not have any understanding on the physiological
 routes involved in the detoxification of Solanaceae compounds by the
 several lineages of Ithomiini. We can speculate that older lineages
-exploiting a novel toxic resource \cite{willmott2006,wahlberg2009}
-might not be too efficient
+exploiting a novel toxic resource {[}9,14{]} might not be too efficient
 in metabolizing plant toxins and that younger lineages are able to deal
@@
 configurations of the proteins, providing more opportunities for
-hostplant switch \cite{li2003}. This might be the reason why the basal
+hostplant switch {[}50{]}. This might be the reason why the basal
 Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and
-restricted to few Solanaceae hosts \cite{willmott2006}, while recent subtribes are
+restricted to few Solanaceae hosts {[}9{]}, while recent subtribes are
 species-rich and have expanded their host range into several Solanaceae
-lineages \cite{willmott2006}. It might be that the switch to feeding in Solanaceae
+lineages {[}9{]}. It might be that the switch to feeding in Solanaceae
 was an important event in the evolutionary history of Ithomiini, but the
@@
 diversity'' and the ``escape-and-radiate'' hypotheses state that insects
-diversify after their hostplants \cite{nyman2012,ehrlich1964,janz2011}.
-Wheat et al. (2007)
-\cite{wheat2007} found strong evidence for a model of speciation congruent with
+diversify after their hostplants {[}2--4{]}. Wheat et al. (2007)
+{[}49{]} found strong evidence for a model of speciation congruent with
 Ehrlich and Raven's hypothesis in Pieridae butterflies due to, in
@@
 diversification in glucosinolate-feeding taxa shortly afterwards (with a
-lag of circa 10 MY). According to a recent dated phylogeny of
-the Angiosperms \cite{bell2010}, the family Solanaceae split from its sister
-group about 59 (49--68) MYA and diversification started (crown group
-age) around 37 (29--47) MYA. Wahlberg et al. (2009) \cite{wahlberg2009} give the
-ages for origin and diversification for Ithomiini at 45 (39--53) and 37
+lag of circa 10 MY). According to a recent dated phylogeny of the
+Angiosperms {[}51{]}, the family Solanaceae split from its sister group
+about 59 (49--68) MYA and diversification started (crown group age)
+around 37 (29--47) MYA. Wahlberg et al. (2009) {[}14{]} give the ages
+for origin and diversification for Ithomiini at 45 (39--53) and 37
 (32--43) MYA, respectively. Thus, current evidence shows that Solanaceae
@@
 widespread species show higher genetic differentiation in their
-populations \cite{wahlberg2011}.
+populations {[}52{]}.
 
@@
 
-Lineages in the diverse family Satyrinae radiated simultaneously with the
-radiation of
-their main hostplant, grasses, during the climatic cooling in the
-Oligocene \cite{pena2008}. Thus, it is somewhat surprising that part of
-Satyrinae were
-found to have accelerated diversification in only 13\% of the trees from
-the posterior distribution. Although this can be attributed to low
-phylogenetic signal \cite{wahlberg2009}, the clade Satyrini is very robust
-\cite{wahlberg2009}
-and MEDUSA failed to identify any significant accelerated net
-diversification rate for Satyrini. It appears that the radiation of
-Satyrini as a whole was not remarkably fast and therefore not picked up
-by MEDUSA, although it estimated a diversification shift for Satyrini + its
-sister clade. This should be expected if the diversification of Satyrini
-occurred in a stepwise manner, with pulses or bursts of diversification
-for certain lineages but unlikely for the tribe Satyrini as a whole.
+Lineages in the diverse family Satyrinae radiated simultaneously with
+the radiation of their main hostplant, grasses, during the climatic
+cooling in the Oligocene {[}15{]}. Thus, it is somewhat surprising that
+part of Satyrinae were found to have accelerated diversification in only
+13\% of the trees from the posterior distribution. Although this can be
+attributed to low phylogenetic signal {[}14{]}, the clade Satyrini is
+very robust {[}14{]} and MEDUSA failed to identify any significant
+accelerated net diversification rate for Satyrini. It appears that the
+radiation of Satyrini as a whole was not remarkably fast and therefore
+not picked up by MEDUSA, although it estimated a diversification shift
+for Satyrini + its sister clade. This should be expected if the
+diversification of Satyrini occurred in a stepwise manner, with pulses
+or bursts of diversification for certain lineages but unlikely for the
+tribe Satyrini as a whole.
 
@@
 strong support for other diversification splits in Nymphalidae. Our
-BiSSE analysis corroborated other studies in that the trait 
-``feeding on Solanaceae'', or a codistributed character, was important in the
+BiSSE analysis corroborated other studies in that the trait ``feeding on
+Solanaceae'', or a codistributed character, was important in the
 diversification of Ithomiini butterflies. However, by applying MEDUSA we
@@
 diversified near simultaneously, which is in agreement with the diffuse
-cospeciation hypothesis \cite{nyman2012,janz2011}.
+cospeciation hypothesis {[}2,4{]}.
 
@@
 \section*{Figure legends}
-\begin{description}
- \item {\bf Figure 1. Results of the MEDUSA analysis run on the maximum
+{\bf Figure 1. Results of the MEDUSA analysis run on the maximum
 clade credibility tree from Wahlberg et al. (2009)} Rate shifts were estimated for the
@@
 
-\item {\bf Figure 2. Diversification rates for taxa estimated by MEDUSA on
+{\bf Figure 2. Diversification rates for taxa estimated by MEDUSA on
 the samples of 1000 random trees}
 
-\item {\bf Figure 3. Results of MultiMEDUSA analysis showing the
+{\bf Figure 3. Results of MultiMEDUSA analysis showing the
 probability of specific nodes being characterized by significant shifts
@@
 
-\item {\bf Figure 4. BiSSE analysis of diversification of nymphalids due
+{\bf Figure 4. BiSSE analysis of diversification of nymphalids due
 to feeding on Solanaceae hostplants} Speciation and net diversification
@@
 $\lambda1$, net diversification rate = $r1$.
-\end{description}
 
 \section*{Supporting Information Legends}
-\begin{description}
- \item {\bf Supporting Information S01.nex.} MCC Nymphalidae tree from
-\cite{wahlberg2009}
-
- \item {\bf Supporting Information S02.R} R script to run MEDUSA on the
-MCC Nymphalidae tree from \cite{wahlberg2009}. This script removes the outgroup
+{\textbf{Supporting Information S01.nex.}} MCC Nymphalidae tree from
+{[}14{]}
+
+{\textbf{Supporting Information S02.R}} R script to run MEDUSA on the
+MCC Nymphalidae tree from {[}14{]}. This script removes the outgroup
 taxa and loads the richness data for the tree terminals.
 
- \item {\bf Supporting Information S03.csv} Species richness for lineages
+{\textbf{Supporting Information S03.csv}} Species richness for lineages
 in Nymphalidae.
 
- \item {\bf Supporting Information S04.txt} Results of running MEDUSA on
+{\textbf{Supporting Information S04.txt}} Results of running MEDUSA on
 the MCC tree.
 
- \item {\bf Supporting Information S05 1000 random trees no outgroups.zip} 1000 random trees from \cite{wahlberg2009}.
-
- \item {\bf Supporting Information S06.R} R script to run a MultiMEDUSA
-analysis on 1000 random trees from \cite{wahlberg2009}.
-
- \item {\bf Supporting Information S07 multimedusa on 1000 trees.zip} Raw results from the MultiMEDUSA run on the random sample of trees from
-the posterior distribution.
-
- \item {\bf Supporting Information S08.pdf} Summary of the MultiMEDUSA
+{\textbf{Supporting Information S05 1000 random trees no outgroups.zip}}
+1000 random trees from {[}14{]}.
+
+{\textbf{Supporting Information S06.R}}~R script to run a MultiMEDUSA
+analysis on 1000 random trees from {[}14{]}.
+
+{\textbf{Supporting Information S07 multimedusa on 1000 trees.zip}} Raw
+results from the MultiMEDUSA run on the random sample of trees from the
+posterior distribution.
+
+{\textbf{Supporting Information S08.pdf}} Summary of the MultiMEDUSA
 analysis on the 1000 trees from the posterior distribution.
 
- \item {\bf Supporting Information S09.tree} MCC tree from the 1000
-random trees selected from the posterior distribution.
-
- \item {\bf Supporting Information S10.R} R script to run a MEDUSA
+{\textbf{Supporting Information S09.tree}} MCC tree from the 1000 random
+trees selected from the posterior distribution.
+
+{\textbf{Supporting Information S10.R}} R script to run a MEDUSA
 analysis on the MCC tree from the 1000 random trees selected from the
@@
 
- \item {\bf Supporting Information S11.txt} Summary results from a MEDUSA
+{\textbf{Supporting Information S11.txt}} Summary results from a MEDUSA
 run on the MCC tree from the 1000 random trees selected from the
@@
 
- \item {\bf Supporting Information S12.pdf} Figure for MEDUSA run on MCC
+{\textbf{Supporting Information S12.pdf}} Figure for MEDUSA run on MCC
 tree from random 1000 trees.
 
- \item {\bf Supporting Information S13.txt} Summary results from a
+{\textbf{Supporting Information S13.txt}} Summary results from a
 MultiMEDUSA run on the 1000 random trees selected from the posterior
@@
 
- \item {\bf Supporting Information S14.pdf} Probability of nodes with
+{\textbf{Supporting Information S14.pdf}} Probability of nodes with
 estimated rates from a MultiMEDUSA run on the 1000 random trees selected
@@
 
- \item {\bf Supporting Information S15.csv} Hostplants of Nymphalidae
+{\textbf{Supporting Information S15.csv}} Hostplants of Nymphalidae
 butterflies recorded from the literature.
 
- \item {\bf Supporting Information S16.csv} References for hostplants
+{\textbf{Supporting Information S16.csv}} References for hostplants
 data.
 
- \item {\bf Supporting Information S17.csv} Data matrix with character
+{\textbf{Supporting Information S17.csv}} Data matrix with character
 states for hosplant use.
 
- \item {\bf Supporting Information S18.R} R script for running the BiSSE
+{\textbf{Supporting Information S18.R}} R script for running the BiSSE
 analysis.
 
- \item {\bf Supporting Information S19.csv} Raw results for the BiSSE
+{\textbf{Supporting Information S19.csv}} Raw results for the BiSSE
 analysis.
 
- \item {\bf Supporting Information S20.csv} Raw results for the combined
+{\textbf{Supporting Information S20.csv}} Raw results for the combined
 BiSSE analysis.
 
- \item {\bf Figure S1.} Best diversification model as estimated by BAMM. Two
-diversification shifts were recovered (Ithomiini and Satyrini). The MEDUSA 
-analysis found a diversification shifts in Satyrinae (the parent node of 
-Satyini).
-
- \item {\bf Figure S2.} Boxplot of speciation ($\lambda$) and extinction
-($\mu$) values for the 95\% credibility intervals of values estimated
+{\textbf{Figure S1.}} Best diversification model as estimated by BAMM.
+Two diversification shifts were recovered (Ithomiini and Satyrini). The
+MEDUSA analysis found a diversification shifts in Satyrinae (the parent
+node of Satyini).
+
+{\textbf{Figure S2.}} Boxplot of speciation (\(\lambda\)) and extinction
+(\(\mu\)) values for the 95\% credibility intervals of values estimated
 by BiSSE analysis of diversification due to feeding on Solanaceae
@@
 
- \item {\bf Figure S3.} BiSSE analysis of diversification of nymphalids due
-to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
+{\textbf{Figure S3.}} BiSSE analysis of diversification of nymphalids
+due to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
 \emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
 recovered, speciation and net diversification rates are significantly
-higher for Solanaceae feeders $\lambda1$, r1).
-
- \item {\bf Figure S4.} Net diversification rates of nymphalids feeding on
+higher for Solanaceae feeders \(\lambda1\), r1).
+
+{\textbf{Figure S4.}} Net diversification rates of nymphalids feeding on
 Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
@@
 
- \item {\bf Figure S5.} Boxplot of speciation and extinction values for the
-95\% credibility intervals of values estimated by BiSSE analysis of
+{\textbf{Figure S5.}} Boxplot of speciation and extinction values for
+the 95\% credibility intervals of values estimated by BiSSE analysis of
 diversification due to feeding on Solanaceae plants on the combined
@@
 
- \item {\bf Figure S6.} BiSSE analysis of diversification of nymphalids due
-to feeding on Apocynaceae hostplants. Speciation and net diversification
-rates are similar.
-\end{description}
-
-% Use the PLoS provided BiBTeX style
-\bibliography{refs}{}
-\bibliographystyle{plos2009}
+{\textbf{Figure S6.}} BiSSE analysis of diversification of nymphalids
+due to feeding on Apocynaceae hostplants. Speciation and net
+diversification rates are similar.
 
@@
 
+\pagebreak
+\section*{References}
+
+1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
+adaptive zones: Has phytophagy promoted insect diversification? Am Nat
+132: 107--128.
+doi:\href{http://dx.doi.org/10.1086/284840}{10.1086/284840}.
+
+2. Nyman T, Linder HP, Pe\~na C, Malm T, Wahlberg N (2012) Climate-driven
+diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
+889--898.
+doi:\href{http://dx.doi.org/10.1111/j.1461-0248.2012.01782.x}{10.1111/j.1461-0248.2012.01782.x}.
+
+3. Ehrlich PR, Raven PH (1964) Butterflies and plants: A study in
+coevolution. Evolution 18: 586--608.
+doi:\href{http://dx.doi.org/10.2307/2406212}{10.2307/2406212}.
+
+4. Janz N (2011) Ehrlich and Raven Revisited: Mechanisms Underlying
+Codiversification of Plants and Enemies. Annu Rev Ecol Evol S 42:
+71--89.
+doi:\href{http://dx.doi.org/10.1146/annurev-ecolsys-102710-145024}{10.1146/annurev-ecolsys-102710-145024}.
+
+5. Nylin S, Slove J, Janz N (2014) Host plant utilization, host range
+oscillations and diversification in Nymphalid butterflies: A
+phylogenetic investigation. Evolution 68: 105--124.
+doi:\href{http://dx.doi.org/10.1111/evo.12227}{10.1111/evo.12227}.
+
+6. Hardy NB, Otto SP (2014) Specialization and generalization in the
+diversification of phytophagous insects: Tests of the musical chairs and
+oscillation hypotheses. P Roy Soc B 281: 20132960.
+
+7. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
+al. (2011) Order Lepidoptera. In: Animal biodiversity: An outline of
+higher-level classification and survey of taxonomic richness. Zootaxa.
+pp. 212--221.
+
+8. Joron M, Jiggins C, Papanicolaou A, McMillan W (2006) Heliconius wing
+patterns: An evo-devo model for understanding phenotypic diversity.
+Heredity 97: 157--167.
+
+9. Willmott KR, Freitas AVL (2006) Higher-level phylogeny of the
+Ithomiinae (Lepidoptera: Nymphalidae): classification, patterns of
+larval hostplant colonization and diversification. Cladistics 22:
+297--368.
+doi:\href{http://dx.doi.org/10.1111/j.1096-0031.2006.00108.x}{10.1111/j.1096-0031.2006.00108.x}.
+
+10. Brakefield PM, Beldade P, Zwaan BJ (2009) The african butterfly
+Bicyclus anynana: A model for evolutionary genetics and evolutionary
+developmental biology. Cold Spring Harb Prot 2009: pdb--emo122.
+
+11. Heikkil\"a M, Kaila L, Mutanen M, Pe\~na C, Wahlberg N (2012) Cretaceous
+origin and repeated tertiary diversification of the redefined
+butterflies. P Roy Soc B 279: 1093--1099.
+doi:\href{http://dx.doi.org/10.1098/rspb.2011.1430}{10.1098/rspb.2011.1430}.
+
+12. Elias M, Joron M, Willmott KR, Silva-Brand\~ao KL, Kaiser V, et al.
+(2009) Out of the Andes: Patterns of diversification in clearwing
+butterflies. Mol Ecol 18: 1716--1729.
+doi:\href{http://dx.doi.org/10.1111/j.1365-294X.2009.04149.x}{10.1111/j.1365-294X.2009.04149.x}.
+
+13. Fordyce JA (2010) Host shifts and evolutionary radiations of
+butterflies. P Roy Soc B 277: 3735--3743.
+doi:\href{http://dx.doi.org/10.1098/rspb.2010.0211}{10.1098/rspb.2010.0211}.
+
+14. Wahlberg N, Leneveu J, Kodandaramaiah U, Pe\~na C, Nylin S, et al.
+(2009) Nymphalid butterflies diversify following near demise at the
+Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302.
+doi:\href{http://dx.doi.org/10.1098/rspb.2009.1303}{10.1098/rspb.2009.1303}.
+
+15. Pe\~na C, Wahlberg N (2008) Prehistorical climate change increased
+diversification of a group of butterflies. Biol Lett 4: 274--278.
+doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.
+
+16. Ferrer-Paris JR, S{\'a}nchez-Mercado A, Viloria \'AL, Donaldson J (2013)
+Congruence and Diversity of Butterfly-Host Plant Associations at Higher
+Taxonomic Levels. PLoS ONE 8: e63570.
+doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.
+
+17. Alfaro ME, Santini F, Brock C, Alamillo H, Dornburg A, et al. (2009)
+Nine exceptional radiations plus high turnover explain species diversity
+in jawed vertebrates. P Natl Acad Sci USA 106: 13410--13414.
+doi:\href{http://dx.doi.org/10.1073/pnas.0811087106}{10.1073/pnas.0811087106}.
+
+18. Harmon LJ, Rabosky DL, {FitzJohn} RG, Brown JW (2011) turboMEDUSA:
+MEDUSA: Modeling Evolutionary Diversification Using Stepwise AIC.
+Available: \url{https://github.com/josephwb/turboMEDUSA}.
+
+19. {FitzJohn} RG (2012) Diversitree: comparative phylogenetic analyses
+of diversification in R. Method Ecol Evol 3: 1084--1092.
+doi:\href{http://dx.doi.org/10.1111/j.2041-210X.2012.00234.x}{10.1111/j.2041-210X.2012.00234.x}.
+
+20. Lamas G (2004) Checklist: Part 4A. Hesperioidea-Papilionoidea.
+Gainesville: Association for Tropical Lepidoptera; Scientific
+Publishers.
+
+21. Matos-Marav\'i PF, Pe\~na C, Willmott KR, Freitas AV, Wahlberg N (2013)
+Systematics and evolutionary history of butterflies in the ``Taygetis
+clade'' (Nymphalidae: Satyrinae: Euptychiina): Towards a better
+understanding of Neotropical biogeography. Mol Phylogenet Evol 66:
+54--68.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2012.09.005}{10.1016/j.ympev.2012.09.005}.
+
+22. Brower AVZ, Wahlberg N, Ogawa JR, Boppr\'e M, Vane-Wright RI (2010)
+Phylogenetic relationships among genera of danaine butterflies
+(Lepidoptera: Nymphalidae) as implied by morphology and DNA sequences.
+Syst Biodivers 8: 75--89.
+doi:\href{http://dx.doi.org/10.1080/14772001003626814}{10.1080/14772001003626814}.
+
+23. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
+(2010) Phylogenetics and biogeography of a spectacular Old World
+radiation of butterflies: the subtribe Mycalesina (Lepidoptera:
+Nymphalidae: Satyrini). BMC Evol Biol 10: 172.
+doi:\href{http://dx.doi.org/10.1186/1471-2148-10-172}{10.1186/1471-2148-10-172}.
+
+24. Kodandaramaiah U, Pe\~na C, Braby MF, Grund R, M\"uller CJ, et al.
+(2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
+problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.08.012}{10.1016/j.ympev.2009.08.012}.
+
+25. Ortiz-Acevedo E, Willmott KR (2013) Molecular systematics of the
+butterfly tribe Preponini (Nymphalidae: Charaxinae). Syst Entomol 38:
+440--449.
+doi:\href{http://dx.doi.org/10.1111/syen.12008}{10.1111/syen.12008}.
+
+26. De-Silva DL, Day JJ, Elias M, Willmott KR, Whinnett A, et al. (2010)
+Molecular phylogenetics of the Neotropical butterfly subtribe Oleriina
+(Nymphalidae: Danainae: Ithomiini). Mol Phylogenet Evol 55: 1032--1041.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2010.01.010}{10.1016/j.ympev.2010.01.010}.
+
+27. Freitas AVL, Brown Jr. KS (2004) Phylogeny of the Nymphalidae
+(Lepidoptera). Syst Biol 53: 363--383.
+doi:\href{http://dx.doi.org/10.1080/10635150490445670}{10.1080/10635150490445670}.
+
+28. Pe\~na C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
+(2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
+Nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2006.02.007}{10.1016/j.ympev.2006.02.007}.
+
+29. Penz CM (1999) Higher level phylogeny for the passion-vine
+butterflies (Nymphalidae, Heliconiinae) based on early stage and adult
+morphology. Zool J Linn Soc 127: 277--344.
+doi:\href{http://dx.doi.org/10.1111/j.1096-3642.1999.tb00680.x}{10.1111/j.1096-3642.1999.tb00680.x}.
+
+30. Silva-Brand\~ao KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
+Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
+tribe Acraeini (Lepidoptera, Nymphalidae, Heliconiinae) and the
+evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2007.11.024}{10.1016/j.ympev.2007.11.024}.
+
+31. Pe\~na C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
+butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
+methods. Zool J Linn Soc 161: 64--87.
+doi:\href{http://dx.doi.org/10.1111/j.1096-3642.2009.00627.x}{10.1111/j.1096-3642.2009.00627.x}.
+
+32. Pe\~na C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
+history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
+Satyrinae). Zool Scr 39: 243--258.
+doi:\href{http://dx.doi.org/10.1111/j.1463-6409.2010.00421.x}{10.1111/j.1463-6409.2010.00421.x}.
+
+33. Ackery PR (1988) Hostplants and classification: A review of
+nymphalid butterflies. Biol J Linn Soc 33: 95--203.
+doi:\href{http://dx.doi.org/10.1111/j.1095-8312.1988.tb00446.x}{10.1111/j.1095-8312.1988.tb00446.x}.
+
+34. Dyer LA, Gentry GL (2002) Caterpillars and parasitoids of a tropical
+lowland wet forest. Available: \url{http://www.caterpillars.org}.
+Accessed 17 April 2013.
+
+35. Beccaloni G, Viloria \'A, Hall S, Robinson G (2008) Catalogue of the
+hostplants of the Neotropical butterflies. London: The Natural History
+Museum.
+
+36. Janzen DH, Hallwachs W (2009) Dynamic database for an inventory of
+the macrocaterpillar fauna, and its food plants and parasitoids, of Area
+de Conservacion Guanacaste (ACG), northwestern Costa Rica. Available:
+\url{http://janzen.sas.upenn.edu}. Accessed 17 April 2013.
+
+37. Core Team R (2013) R: A Language and Environment for Statistical
+Computing. Vienna, Austria: R Foundation for Statistical Computing.
+Available: \url{http://www.r-project.org}.
+
+38. Popescu A-A, Huber KT, Paradis E (2012) ape 3.0: New tools for
+distance-based phylogenetics and evolutionary analysis in R.
+Bioinformatics 28: 1536--1537.
+doi:\href{http://dx.doi.org/10.1093/bioinformatics/bts184}{10.1093/bioinformatics/bts184}.
+
+39. Harmon LJ, Weir JT, Brock CD, Glor RE, Challenger W (2008) GEIGER:
+investigating evolutionary radiations. Bioinformatics 24: 129--131.
+doi:\href{http://dx.doi.org/10.1093/bioinformatics/btm538}{10.1093/bioinformatics/btm538}.
+
+40. Rabosky DL (2014) Automatic detection of key innovations, rate
+shifts, and diversity-dependence on phylogenetic trees. PLoS ONE 9:
+e89543. Available:
+\url{http://dx.doi.org/10.1371\%2Fjournal.pone.0089543}.
+
+41. Maddison WP, Midford PE, Otto SP (2007) Estimating a binary
+character's effect on speciation and extinction. Syst Biol 56: 701--710.
+doi:\href{http://dx.doi.org/10.1080/10635150701607033}{10.1080/10635150701607033}.
+
+42. Scott JA (1986) The Butterflies of North America. Stanford
+University Press.
+
+43. Davis MP, Midford PE, Maddison W (2013) Exploring power and
+parameter estimation of the BiSSE method for analyzing species
+diversification. BMC Evol Biol 13: 38.
+doi:\href{http://dx.doi.org/10.1186/1471-2148-13-38}{10.1186/1471-2148-13-38}.
+
+44. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
+bees cut leaves? New insights into the early evolution of bees. P Roy
+Soc B 278: 3593--3600.
+doi:\href{http://dx.doi.org/10.1098/rspb.2011.0365}{10.1098/rspb.2011.0365}.
+
+45. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
+clades of Agaricales. P Roy Soc B 279: 2003--2011.
+doi:\href{http://dx.doi.org/10.1098/rspb.2011.2428}{10.1098/rspb.2011.2428}.
+
+46. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
+Interface. Ann Mo Bot Gard 74: 359--397.
+
+47. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand\~ao KL, Whinnett A, et
+al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
+Nymphalidae) inferred from one mitochondrial and two nuclear gene
+regions. Syst Entomol 31: 288--301.
+doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2006.00321.x}{10.1111/j.1365-3113.2006.00321.x}.
+
+48. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
+trait-dependent speciation and extinction rates from incompletely
+resolved phylogenies. Syst Biol 58: 595--611.
+doi:\href{http://dx.doi.org/10.1093/sysbio/syp067}{10.1093/sysbio/syp067}.
+
+49. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
+The genetic basis of a plant-insect coevolutionary key innovation. P
+Natl Acad Sci USA 104: 20427--20431.
+doi:\href{http://dx.doi.org/10.1073/pnas.0706229104}{10.1073/pnas.0706229104}.
+
+50. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
+furanocoumarin-metabolizing cytochrome P450 monooxygenases in two
+papilionids: Specificity and substrate encounter rate. P Natl Acad Sci
+USA 100: 14593--14598.
+doi:\href{http://dx.doi.org/10.1073/pnas.1934643100}{10.1073/pnas.1934643100}.
+
+51. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
+the angiosperms re-revisited. Am J Bot 97: 1296--1303.
+doi:\href{http://dx.doi.org/10.3732/ajb.0900346}{10.3732/ajb.0900346}.
+
+52. Wahlberg N, Rubinoff D (2011) Vagility across Vanessa (Lepidoptera:
+Nymphalidae): mobility in butterfly species does not inhibit the
+formation and persistence of isolated sister taxa. Syst Entomol 36:
+362--370.
+doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2010.00566.x}{10.1111/j.1365-3113.2010.00566.x}.
+
 \end{document}

### 2014-12-19 MS_revised.tex
typo in acknowledgments
references in

--- 2014-11-12 MS_revised.tex
@@
 evolutionary interactions between plants and insects \cite{nyman2012}. The
-``escape-and-radiate'' hypothesis \cite{ehrlich1964}, the 
-``oscillation hypothesis''
-\cite{janz2011,nylin2014} or ``diffuse cospeciation'' \cite{nyman2012} and the
+``escape-and-radiate'' hypothesis \cite{ehrlich1964}, the ``oscillation
+hypothesis'' \cite{janz2011, nylin2014} or ``diffuse cospeciation'' \cite{nyman2012} and the
 ``musical chairs hypothesis'' \cite{hardy2014}.
@@
 developing some of the mentioned hypotheses. Nymphalidae contains around
-6000 species \cite{van_nieukerken2011},
-and several members are considered model organisms
-in evolutionary biology \cite{joron2006,willmott2006,brakefield2009}.
+6000 species \cite{van_nieukerken2011}, and several members are considered model organisms
+in evolutionary biology \cite{joron2006, willmott2006, brakefield2009}.
 The family most likely originated
@@
 group to identify patterns of accelerated or decelerated diversification
-of some Nymphalidae clades
-\cite{heikkila2012,elias2009,fordyce2010,wahlberg2009}.
+of some Nymphalidae clades \cite{heikkila2012, elias2009, fordyce2010, wahlberg2009}.
 For example, it has been
@@
 geographic areas (``resource abundance-dependent diversity dynamics''
-hypothesis). Fordyce (2010) \cite{fordyce2010} found increased net diversification
+hypothesis): Fordyce (2010) \cite{fordyce2010} found increased net diversification
 rates in some Nymphalidae lineages after a major hostplant shift, which
@@
 sudden diversification bursts due to hostplant shift or climatic events
-\cite{nylin2014,ferrer2013}.
+\cite{nylin2014, ferrer2013}.
 
 In this study, we used a time-calibrated genus-level phylogenetic
-hypothesis for Nymphalidae butterflies \cite{wahlberg2009} to
-investigate patterns
-of diversification. We applied the statistical method
-MEDUSA \cite{alfaro2009,harmon2011},
-to study the diversification pattern of Nymphalidae
+hypothesis for Nymphalidae butterflies \cite{wahlberg2009} to investigate patterns
+of diversification. We applied the statistical method MEDUSA
+\cite{alfaro2009, harmon2011}, to study the diversification pattern of Nymphalidae
 butterflies. MEDUSA fits likelihood models of diversification into a
@@
 hostplant use affected the diversification pattern of those lineages
-employing the method BiSSE as implemented in the R package
-\texttt{diversitree} \cite{fitzjohn2012}.
+employing the binary speciation and extinction model (BiSSE) as
+implemented in the R package \texttt{diversitree} \cite{fitzjohn2012}.
 
@@
 trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (supp.
-mat. 5) in order to correct for phylogenetic uncertainty when performing
+mat. 5) in order to account for phylogenetic uncertainty when performing
 the diversification analyses.
@@
 \cite{wahlberg2009} study according to available phylogenetic studies 
-\cite{matos2013,brower2010,kodandaramaiah2010,kodandaramaiah2010a,ortiz2013,desilva2010,freitas2004,pena2006,penz1999,silva2008,pena2011,pena2010},
+\cite{matos2013, brower2010, kodandaramaiah2010, kodandaramaiah2010a,
+ortiz2013, desilva2010, freitas2004, pena2006, penz1999, silva2008,
+pena2011, pena2010},
 taxonomical classification and morphological resemblance when no
@@
 sources including \cite{ackery1988}, HOSTS database (\url{http://bit.ly/YI7nwW}),
-\cite{dyer2002,beccaloni2008,janzen2009} and others (supp. mat. 15--16)
-for a total of 6586
+\cite{dyer2002, beccaloni2008, janzen2009} and others (supp. mat. 15--16) for a total of 6586
 hostplant records, including 428 Nymphalidae genera and 143 plant
@@
 
-We analyzed our MCC tree in the Bayesian software BAMM in order to contrast
-the results
-found by MEDUSA and MultiMEDUSA.
-BAMM (Bayesian analysis of macroevolutionary mixtures)
-\cite{rabosky2014} is a
-software that fits several models of speciation and extinction. The best model
-can be selected by comparison of Bayes Factors.
+We analyzed our MCC tree in the Bayesian software BAMM in order to
+contrast the results found by MEDUSA and MultiMEDUSA. BAMM (Bayesian
+analysis of macroevolutionary mixtures) \cite{rabosky2014} is a software that fits
+several models of speciation and extinction. The best model can be
+selected by comparison of Bayes Factors.
 
@@
 value. MEDUSA finds the likelihood of the models after taking into
-account branch lengths and number of species per lineage 
-\cite{alfaro2009}. Most
+account branch lengths and number of species per lineage \cite{alfaro2009}. Most
 studies using MEDUSA only run the method on a single tree, usually the
@@
 was run for 10000 generations discarding the first 7500 as burnin. The
-Multiple State Speciation and Extinction approach (MuSSE) 
-\cite{fitzjohn2012} is
+Multiple State Speciation and Extinction approach (MuSSE) \cite{fitzjohn2012} is
 designed to examine the joint effects of two or more traits on
@@
 distribution. The records of \emph{Vanessa} and \emph{Hypanartia}
-feeding on Solanaceae \cite{beccaloni2008,scott1986} might be incorrect as it
-is unlikely
+feeding on Solanaceae \cite{beccaloni2008, scott1986} might be incorrect as it is unlikely
 that these species can be feeding on this plant family. We tested
@@
 or presence would not have a significant effect on our results. It has
-been shown that BiSSE performs poorly under certain conditions 
-\cite{davis2013}.
+been shown that BiSSE performs poorly under certain conditions \cite{davis2013}.
 However, our data has adequate number of taxa under analysis (more than
@@
 
-We contrasted these results with those obtained from analyzing the data in
-the software BAMM \cite{rabosky2014}. We found 
-that the best model (best configuration) estimated 2 diversification shifts in
-our MCC tree (posterior probability = 0.79) for the Ithomiini
-and Satyrini (Fig. S1). Bayes factor (BF) values strongly 
-favored this model in comparison with a constant rate model (no diversification
-shifts, BF $>$ 1,884).
+We contrasted these results with those obtained from analyzing the data
+in the software BAMM \cite{rabosky2014}. We found that the best model (best
+configuration) estimated 2 diversification shifts in our MCC tree
+(posterior probability = 0.79) for the Ithomiini and Satyrini (Fig. S1).
+Bayes factor (BF) values strongly favored this model in comparison with
+a constant rate model (no diversification shifts, BF \(>\) 1,884).
 
@@
 have been used for generation of hypotheses and discussion
-\cite{heikkila2012,litman2011,ryberg2012}. However, different diversification
-shifts and different
+\cite{heikkila2012, litman2011, ryberg2012}. However, different diversification shifts and different
 rates of diversification are found for certain lineages when
@@
 \cite{wahlberg2009}. As a result, MEDUSA inferred a net diversification rate
-increase for the Satyrinae (which includes Satyrini and its sister clade)
-in the maximum clade credibility tree, but
-this was recovered in the MultiMEDUSA analysis in only 13\% of the
-random sample of trees.
+increase for the Satyrinae (which includes Satyrini and its sister
+clade) in the maximum clade credibility tree, but this was recovered in
+the MultiMEDUSA analysis in only 13\% of the random sample of trees.
 
@@
 butterflies are exclusively Neotropical and most species feed on
-Solanaceae hostplants during larval stage \cite{willmott2006}.
-Optimizations of the
+Solanaceae hostplants during larval stage \cite{willmott2006}. Optimizations of the
 evolution of hostplant use on phylogenies evidence a probable shift from
-Apocynaceae to Solanaceae in the ancestor of the tribe 
-\cite{willmott2006,brower2006}.
+Apocynaceae to Solanaceae in the ancestor of the tribe \cite{willmott2006, brower2006}.
 Fordyce (2010) \cite{fordyce2010} found that the Gamma statistics, a LTT plot of
@@
 responsible for a higher diversification of Ithomiini might not be the
-hostplant character. As noted by Maddison et al. (2007) \cite{maddison2007},
-the responsible trait might be a codistributed character such as a trait
+hostplant character. As noted by Maddison et al. (2007) \cite{maddison2007}, the
+responsible trait might be a codistributed character such as a trait
 related to the ability to digest secondary metabolites.
@@
 and radiation scenarios among chemically different lineages of
-Solanaceae plants \cite{willmott2006,brown1987}.
-According to this theory, the shift from
+Solanaceae plants \cite{willmott2006, brown1987}. According to this theory, the shift from
 Apocynaceae to Solanaceae allowed Ithomiini to invade newly available
@@
 several lineages of Ithomiini. We can speculate that older lineages
-exploiting a novel toxic resource \cite{willmott2006,wahlberg2009}
-might not be too efficient
+exploiting a novel toxic resource \cite{willmott2006, wahlberg2009} might not be too efficient
 in metabolizing plant toxins and that younger lineages are able to deal
@@
 diversity'' and the ``escape-and-radiate'' hypotheses state that insects
-diversify after their hostplants \cite{nyman2012,ehrlich1964,janz2011}.
+diversify after their hostplants \cite{nyman2012, ehrlich1964, janz2011}.
 Wheat et al. (2007)
@@
 diversification in glucosinolate-feeding taxa shortly afterwards (with a
-lag of circa 10 MY). According to a recent dated phylogeny of
-the Angiosperms \cite{bell2010}, the family Solanaceae split from its sister
-group about 59 (49--68) MYA and diversification started (crown group
-age) around 37 (29--47) MYA. Wahlberg et al. (2009) \cite{wahlberg2009} give the
-ages for origin and diversification for Ithomiini at 45 (39--53) and 37
+lag of circa 10 MY). According to a recent dated phylogeny of the
+Angiosperms \cite{bell2010}, the family Solanaceae split from its sister group
+about 59 (49--68) MYA and diversification started (crown group age)
+around 37 (29--47) MYA. Wahlberg et al. (2009) \cite{wahlberg2009} give the ages
+for origin and diversification for Ithomiini at 45 (39--53) and 37
 (32--43) MYA, respectively. Thus, current evidence shows that Solanaceae
@@
 
-Lineages in the diverse family Satyrinae radiated simultaneously with the
-radiation of
-their main hostplant, grasses, during the climatic cooling in the
-Oligocene \cite{pena2008}. Thus, it is somewhat surprising that part of
-Satyrinae were
-found to have accelerated diversification in only 13\% of the trees from
-the posterior distribution. Although this can be attributed to low
-phylogenetic signal \cite{wahlberg2009}, the clade Satyrini is very robust
-\cite{wahlberg2009}
-and MEDUSA failed to identify any significant accelerated net
-diversification rate for Satyrini. It appears that the radiation of
-Satyrini as a whole was not remarkably fast and therefore not picked up
-by MEDUSA, although it estimated a diversification shift for Satyrini + its
-sister clade. This should be expected if the diversification of Satyrini
-occurred in a stepwise manner, with pulses or bursts of diversification
-for certain lineages but unlikely for the tribe Satyrini as a whole.
+Lineages in the diverse family Satyrinae radiated simultaneously with
+the radiation of their main hostplant, grasses, during the climatic
+cooling in the Oligocene \cite{pena2008}. Thus, it is somewhat surprising that
+part of Satyrinae were found to have accelerated diversification in only
+13\% of the trees from the posterior distribution. Although this can be
+attributed to low phylogenetic signal \cite{wahlberg2009}, the clade Satyrini is
+very robust \cite{wahlberg2009} and MEDUSA failed to identify any significant
+accelerated net diversification rate for Satyrini. It appears that the
+radiation of Satyrini as a whole was not remarkably fast and therefore
+not picked up by MEDUSA, although it estimated a diversification shift
+for Satyrini + its sister clade. This should be expected if the
+diversification of Satyrini occurred in a stepwise manner, with pulses
+or bursts of diversification for certain lineages but unlikely for the
+tribe Satyrini as a whole.
 
@@
 diversified near simultaneously, which is in agreement with the diffuse
-cospeciation hypothesis \cite{nyman2012,janz2011}.
+cospeciation hypothesis \cite{nyman2012, janz2011}.
 
@@
 $\lambda1$, net diversification rate = $r1$.
-\end{description}
 
 \section*{Supporting Information Legends}
-\begin{description}
- \item {\bf Supporting Information S01.nex.} MCC Nymphalidae tree from
+{\textbf{Supporting Information S01.nex.}} MCC Nymphalidae tree from
 \cite{wahlberg2009}
 
- \item {\bf Supporting Information S02.R} R script to run MEDUSA on the
+{\textbf{Supporting Information S02.R}} R script to run MEDUSA on the
 MCC Nymphalidae tree from \cite{wahlberg2009}. This script removes the outgroup
@@
 
- \item {\bf Supporting Information S03.csv} Species richness for lineages
+{\textbf{Supporting Information S03.csv}} Species richness for lineages
 in Nymphalidae.
 
- \item {\bf Supporting Information S04.txt} Results of running MEDUSA on
+{\textbf{Supporting Information S04.txt}} Results of running MEDUSA on
 the MCC tree.
 
- \item {\bf Supporting Information S05 1000 random trees no outgroups.zip} 1000 random trees from \cite{wahlberg2009}.
+{\textbf{Supporting Information S05 1000 random trees no outgroups.zip}}
+1000 random trees from \cite{wahlberg2009}.
 
- \item {\bf Supporting Information S06.R} R script to run a MultiMEDUSA
+{\textbf{Supporting Information S06.R}}~R script to run a MultiMEDUSA
 analysis on 1000 random trees from \cite{wahlberg2009}.
 
- \item {\bf Supporting Information S07 multimedusa on 1000 trees.zip} Raw results from the MultiMEDUSA run on the random sample of trees from
-the posterior distribution.
+{\textbf{Supporting Information S07 multimedusa on 1000 trees.zip}} Raw
+results from the MultiMEDUSA run on the random sample of trees from the
+posterior distribution.
 
- \item {\bf Supporting Information S08.pdf} Summary of the MultiMEDUSA
+{\textbf{Supporting Information S08.pdf}} Summary of the MultiMEDUSA
 analysis on the 1000 trees from the posterior distribution.
 
- \item {\bf Supporting Information S09.tree} MCC tree from the 1000
-random trees selected from the posterior distribution.
+{\textbf{Supporting Information S09.tree}} MCC tree from the 1000 random
+trees selected from the posterior distribution.
 
- \item {\bf Supporting Information S10.R} R script to run a MEDUSA
+{\textbf{Supporting Information S10.R}} R script to run a MEDUSA
 analysis on the MCC tree from the 1000 random trees selected from the
@@
 
- \item {\bf Supporting Information S15.csv} Hostplants of Nymphalidae
+{\textbf{Supporting Information S15.csv}} Hostplants of Nymphalidae
 butterflies recorded from the literature.
 
- \item {\bf Supporting Information S16.csv} References for hostplants
+{\textbf{Supporting Information S16.csv}} References for hostplants
 data.
 
- \item {\bf Supporting Information S17.csv} Data matrix with character
+{\textbf{Supporting Information S17.csv}} Data matrix with character
 states for hosplant use.
 
- \item {\bf Supporting Information S18.R} R script for running the BiSSE
+{\textbf{Supporting Information S18.R}} R script for running the BiSSE
 analysis.
 
- \item {\bf Supporting Information S19.csv} Raw results for the BiSSE
+{\textbf{Supporting Information S19.csv}} Raw results for the BiSSE
 analysis.
 
- \item {\bf Supporting Information S20.csv} Raw results for the combined
+{\textbf{Supporting Information S20.csv}} Raw results for the combined
 BiSSE analysis.
 
- \item {\bf Figure S1.} Best diversification model as estimated by BAMM. Two
-diversification shifts were recovered (Ithomiini and Satyrini). The MEDUSA 
-analysis found a diversification shifts in Satyrinae (the parent node of 
-Satyini).
+{\textbf{Figure S1.}} Best diversification model as estimated by BAMM.
+Two diversification shifts were recovered (Ithomiini and Satyrini). The
+MEDUSA analysis found a diversification shifts in Satyrinae (the parent
+node of Satyini).
 
- \item {\bf Figure S2.} Boxplot of speciation ($\lambda$) and extinction
-($\mu$) values for the 95\% credibility intervals of values estimated
+{\textbf{Figure S2.}} Boxplot of speciation (\(\lambda\)) and extinction
+(\(\mu\)) values for the 95\% credibility intervals of values estimated
 by BiSSE analysis of diversification due to feeding on Solanaceae
@@
 
- \item {\bf Figure S3.} BiSSE analysis of diversification of nymphalids due
-to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
+{\textbf{Figure S3.}} BiSSE analysis of diversification of nymphalids
+due to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
 \emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
 recovered, speciation and net diversification rates are significantly
-higher for Solanaceae feeders $\lambda1$, r1).
+higher for Solanaceae feeders \(\lambda1\), r1).
 
- \item {\bf Figure S4.} Net diversification rates of nymphalids feeding on
+{\textbf{Figure S4.}} Net diversification rates of nymphalids feeding on
 Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
@@
 
+\pagebreak
+\bibliography{refs}{}
+\bibliographystyle{plos2009}
+
 \end{document}

### 2014-12-19 MS_revised.tex
tolweb as reference
fixes comment 4

@@
 Species richness data for Nymphalidae genera were compiled from several
-sources including the specialist-curated lists on
-\url{http://tolweb.org}, Lamas (2004) \cite{lamas2004} and curated lists of
-Global Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}).
+sources including the specialist-curated lists 
+\cite{tolweb2007, lamas2004} and curated lists of the
+Global Butterfly Names project \cite{gbnp}.
 We assigned the species numbers of genera not included in the phylogeny

### 2014-12-19 MS_revised.tex
removing BAMM

@@
 
-We analyzed our MCC tree in the Bayesian software BAMM in order to
-contrast the results found by MEDUSA and MultiMEDUSA. BAMM (Bayesian
-analysis of macroevolutionary mixtures) \cite{rabosky2014} is a software that fits
-several models of speciation and extinction. The best model can be
-selected by comparison of Bayes Factors.
-
 \subsection*{Detecting diversification shifts on phylogenetic trees}
@@
 (\texttt{r = 0.065}) (Table 1).
-
-We contrasted these results with those obtained from analyzing the data
-in the software BAMM \cite{rabosky2014}. We found that the best model (best
-configuration) estimated 2 diversification shifts in our MCC tree
-(posterior probability = 0.79) for the Ithomiini and Satyrini (Fig. S1).
-Bayes factor (BF) values strongly favored this model in comparison with
-a constant rate model (no diversification shifts, BF \(>\) 1,884).
 
@@
 distribution of trees (Fig. 3).
-
-The results from BAMM recovered only two significant shifts in our MCC
-tree as this method seems to be much more conservative than the MEDUSA
-approach. One of these shifts is the Satyrini clade. MEDUSA estimated
-its parent node as significant shift in only a few trees from the
-posterior distribution. Thus, it might be important to take into account
-phylogenetic uncertainty in BAMM as well.
 
@@
 
-{\textbf{Figure S1.}} Best diversification model as estimated by BAMM.
-Two diversification shifts were recovered (Ithomiini and Satyrini). The
-MEDUSA analysis found a diversification shifts in Satyrinae (the parent
-node of Satyini).
-
-{\textbf{Figure S2.}} Boxplot of speciation (\(\lambda\)) and extinction
+{\textbf{Figure S1.}} Boxplot of speciation (\(\lambda\)) and extinction
 (\(\mu\)) values for the 95\% credibility intervals of values estimated
@@
 
-{\textbf{Figure S3.}} BiSSE analysis of diversification of nymphalids
+{\textbf{Figure S2.}} BiSSE analysis of diversification of nymphalids
 due to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
@@
 
-{\textbf{Figure S4.}} Net diversification rates of nymphalids feeding on
+{\textbf{Figure S3.}} Net diversification rates of nymphalids feeding on
 Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
@@
 
-{\textbf{Figure S5.}} Boxplot of speciation and extinction values for
+{\textbf{Figure S4.}} Boxplot of speciation and extinction values for
 the 95\% credibility intervals of values estimated by BiSSE analysis of
@@
 
-{\textbf{Figure S6.}} BiSSE analysis of diversification of nymphalids
+{\textbf{Figure S5.}} BiSSE analysis of diversification of nymphalids
 due to feeding on Apocynaceae hostplants. Speciation and net

### 2014-12-19 MS_revised.tex
comment 6

@@
 (\texttt{b}) \texttt{-} extinction (\texttt{d}) and (\(\varepsilon\):
-relative extinction \texttt{= d/b}. The AIC scores of the two-parameter
-model are then compared with incrementally more complex models until the
-addition of parameters do not improve the AIC scores beyond a cutoff
-value. MEDUSA finds the likelihood of the models after taking into
+relative extinction \texttt{= d/b}.
+The maximum-likelihood values are recalculated when each branch in the tree
+is selected to contain an additional breakpoint (additional parameters
+\texttt{b} and \texttt{d}) in net diversification rate. 
+The process is repeated by fitting additional breaking points to the tree.
+The fit of  alternative models is compared using AIC
+scores and stop when the improvement of the incrementally more complex models 
+is smaller than a cutoff value which
+depends of the number of terminals.
+MEDUSA finds the likelihood of the models after taking into
 account branch lengths and number of species per lineage \cite{alfaro2009}. Most

### 2014-12-19 MS_revised.tex
comment 7

@@
 can be explained by an increase in the birth-rate due to hostplant use
-and performed analyses using the ``binary state speciation and
-extinction'' \cite{maddison2007} Bayesian approach as implemented in the R package
+and performed analyses using the BiSSE model \cite{maddison2007} 
+as implemented in the R package
 \texttt{diversitree} \cite{fitzjohn2012}. The Markov Chain Monte Carlo algorithm

### 2014-12-19 response_to_reviewers.tex
ready for resubmittion
comment 9

@@
 \usepackage{color}
+\usepackage{color}
+\usepackage{color}
 \date{}
@@
 
-Introduction 1. Second paragraph, second sentence: The fact that the
-Nymphalids are the ``largest family'' of butterflies conveys an
-exaggerated importance of their diversity based on idiosyncrasy of
-Linnean taxonomy (the sister group to Nymphalidae, the clade that
-includes riodinids and lycaenids, has over 6500 species). Consider
-revising the sentence, omitting superlatives (``biggest'', ``largest'',
-``most'', etc.).
+ Page 3, last sentence of the first paragraph in the Introduction is a fragment; consider joining it to
+preceding sentence with a colon (:).
 
@@
 \color{blue}
-We have rewritten the sentence to: ``Nymphalidae contains around 6000
-species {[}6{]}, and several members are considered model organisms in
-evolutionary biology {[}7--9{]}.''
+We have followed the reviewer's advise.
 \end{quote}
@@
 
-The authors should mention that the ideal approach would accommodate
-uncertainty by sampling from a distribution of trees which reflected the
-variation in confidence in various parts of the tree (such as a Bayesian
-posterior distribution or a bootstrap sample). See Holder,M. and
-Lewis,P.O. (2003) Phylogeny estimation: traditional and Bayesian
-approaches. Nat. Rev.~Genet., 4, 275--284
+Page 3, last sentence of Introduction: as this is the first mention of BiSSE inbe spelled out (even though it is spelled out in the Abstract as well).
 
@@
 \color{blue}
-We have specified that the sample of trees was taken from the Bayesian
-run in Wahlberg et al., 2009.
+We have followed the reviewer's advise.
 \end{quote}
@@
 
-Middle of paragraph (and throughout the manuscript): revise the tree
-sampling to ``\ldots{}therefore randomly sampled 1000 trees from the
-posterior distribution''; in the remainder of the manuscript, refer to
-these as the ``1000 sampled trees'', not ``1000 random trees'', which
-can take on a very different meaning.
+ Page 3, first paragraph of Data section in Materials & Methods, final sentence: replace “correct for”
+with “account for”.
 
@@
 \color{blue}
-We have corrected some sentences in the manuscript to specify that we
-are referring to the same set of 1000 trees.
+We have followed the reviewer's advise.
 \end{quote}
@@
 
-Define what `MuSSE' refers to (Multiple State Speciation and
-Extinction?).
+Page 3, second paragraph of Data section in Materials & Methods, first sentence: the two web
+resources (Tree of Life Web Project and Global Butterfly Names project) should be listed in the
+References section and treated as references in the manuscript (i.e. referred to by number).
 
@@
 \color{blue}
-We have included the definition of MuSSE in the manuscript.
+We have moved the web links to the references as advised by the reviewer.
 \end{quote}
@@
 
-A reference for the statement ``most of Nymphalidae butterflies are
-restricted to use one plant family as hostplant'' is warranted.
+Page 4, Analyses of Diversification section, second paragraph: The BAMM analyses are given less
+than cursory treatment, and no real comparison is discussed in the Discussion (different
+diversification rate shifts are identified in the BAMM analyses than in the MultiMEDUSA
+analyses). At this point, either a flushed-out discussion of the potential reasons for the differences
+should be included, or the BAMM analyses should just be removed from the manuscript (the latter
+would likely be best).
 
@@
 \color{blue}
-We have included the numbers of single-family feeders based on
-calculations of our own data on hostplant use across Nymphalidae genera.
-The text has been modified accordingly:
+We have removed the BAMM analysis from the manuscript.
 \end{quote}
@@
 
-At the end of this paragraph, the authors indicated they tested the
-effect of character coding on their analyses. A more explicit
-description of the test would be useful, along with predictions of how
-results would be effected (i.e.~how would a large effect vs.~no effect
-of coding Hypanartia and Vanessa as Solanaceae feeders manifest in the
-results?).
+Page 4, Detecting diversification shifts... section: A brief (one to two sentences) description of how
+models become iteratively more complex in MEDUSA analyses is warranted here. Something
+about how one additional clade-specific diversification rate parameter is added at each step.
 
@@
 \color{blue}
-We included the following sentence to clarify the issue:\\\textbf{``We
-expected that coding these two genera as either absence or presence
-would not have a significant effect on our results.''}
+We have rewritten the paragraph and included additional sentences as requested
+by the reviewer.
 \end{quote}
 
-\textbf{7, 8, 9, 10. Reviewer's Comment:}
+\textbf{7. Reviewer's Comment:}
+Page 4, Estimation of trait-dependent... section, first sentence: replace “binary state speciation and extinction” with “BiSSE model”.
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{6}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  First paragraph, second sentence: change the tense of ``run'' to
-  ``ran''.
-\item
-  Second paragraph, third sentence: replace ``Thus'' with
-  ``Furthermore''.
-\item
-  There appears to be an orphaned right-parenthesis at the end of the
-  second paragraph.
-\item
-  Third paragraph, last sentence: The node number is not really
-  relevant, so beginning of the sentence could be revised to ``For
-  example, the Charaxes + Polyura clade was found in only 256
-  trees\ldots{}''
-\end{enumerate}
-
-\textbf{7, 8, 9, 10. Response:}
+\textbf{7. Response:}
 
@@
 \color{blue}
-We made the suggested corrections.
+We have followed the reviewer's advise.
 \end{quote}
 
-\textbf{11. Reviewer's Comment:}
+\textbf{8. Reviewer's Comment:}
 
-Fourth paragraph: The point of the paragraph is not clear. What rates
-are being compared? MEDUSA vs.~MultiMEDUSA? Consider revising to make
-the point clearer or omit the sentence.
+ Page 4, Estimation of trait-dependent...
+section, third sentence: it is not clear why the reference to MuSSE is necessary here, as the authors do not appear to use it.
+If this is the case, the sentence could be omitted.
 
-\textbf{11. Response:}
+ \textbf{8. Response:}
 
@@
 \color{blue}
-We have rewritten the sentence and now is more clear:\\\textbf{``For the
-diversification shifts found in both the MCC tree and most of the
-samples of 1000 trees (frequency more than 90\%; Table 2), the
-MultiMEDUSA approach recovered different rates of diversification than
-those found using when the MCC tree alone.''}.
+We removed that sentence.
 \end{quote}
 
-\textbf{12. Reviewer's Comment:}
+\textbf{9. Reviewer's Comment:}
+Page 5, Estimation of trait-dependent...
+section, sentence beginning “Other hostplant shifts were not tested...”: is this accurate?
+At the end of the results section, the very last sentence describes a test for the effect of Apocynaceae feeding on diversification rates...
 
-Fifth paragraph, first sentence (and potentially elsewhere): Consider
-using a consistent means of presenting probability values. In this
-sentence, the authors refer to ``90\% of probability in the trees'', but
-in the preceding paragraph, the authors present probabilities as a
-decimal (``probability of being recovered higher than 0.90''). Granted,
-these two probabilities refer to different events, but perhaps it would
-be best to refer to them in a consistent manner, such as ``0.90
-posterior probability''.
-
-\textbf{12. Response:}
+\textbf{9. Response:}
 
@@
 \color{blue}
-We have corrected this and normalized the probability values to decimals
-in the manuscript.
-\end{quote}
-
-\textbf{13. Reviewer's Comment:} The third sentence (length of analysis
-and burnin details) should be moved to the methods section.
-
-\textbf{13. Response:}
-
-\begin{quote}
-\color{blue}
-We made the suggested corrections.
+We removed that sentence.
 \end{quote}
@@
 \color{blue}
+\color{blue}
+\color{blue}
 We corrected the text following the reviewer's suggestion.
@@
 \begin{quote}
+\color{blue}
+\color{blue}
 \color{blue}
@@
 \color{blue}
+\color{blue}
+\color{blue}
 As suggested by the reviewer, he have created a \textbf{Conclusions}

### 2014-12-19 MS_revised.tex
comment 8
comment 9

@@
 \texttt{diversitree} \cite{fitzjohn2012}. The Markov Chain Monte Carlo algorithm
-was run for 10000 generations discarding the first 7500 as burnin. The
-Multiple State Speciation and Extinction approach (MuSSE) \cite{fitzjohn2012} is
-designed to examine the joint effects of two or more traits on
-speciation. Because most of Nymphalidae butterflies are restricted to
+was run for 10000 generations discarding the first 7500 as burnin.
+Because most of Nymphalidae butterflies are restricted to
 use one plant family as hostplant (232 genera use only one plant family
@@
 1) (supp. mat. 17). The family Solanaceae include the hostplants of the
-diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. Other
-hostplant shifts were not tested for effect on net diversification rates
-due to the low support for diversification shifts found in the MEDUSA
-analyses. We analyzed the data using BiSSE employing the Markov Chain
+diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. 
+We analyzed the data using BiSSE employing the Markov Chain
 Monte Carlo algorithm on the maximum clade credibility tree, taking into

### 2014-12-19 response_to_reviewers.tex
comment 10

@@
 
-\textbf{14. Reviewer's Comment:}
+\textbf{10. Reviewer's Comment:}
 
-Mid-way through paragraph, discussion of likelihood ratio test: The
-source of this P-value needs to be described (chi-squared approximation
-with 1 degree of freedom?), here or in the methods section (see p.~617
-in Pagel 1999 Syst. Biol. 48:612-622).
+Page 5, Estimation of trait-dependent...
+section, sentence beginning “We also used constrained...”: it is not immediately clear which model is the constrained model and which model is unconstrained.
+Start the sentence with: “We also used likelihood ratio tests to evaluate the influence of Solanaceae-feeding on diversification rates by comparing...” and describing the two models being compared.
 
-\textbf{14. Response:}
+\textbf{10. Response:}
 
@@
 \color{blue}
-We have include the statistics suggested by the reviewer:
-``\textbf{\(\chi^2 = 12.3; 1 df; p < 0.001\)}''.
+We corrected the sentence as suggested by the reviewer:
+
+\textbf{We also used likelihood ratio
+tests to evaluate the influence of Solanaceae-feeding on diversification rates by comparing a constrained
+versus a relaxed model of diversification. The constrained model assumes no effect of hostplant use on
+diversification by enforcing equal speciation rates across both character states. In the relaxed model, all
+speciation and extinction parameters are estimated for each character state.}
 \end{quote}
 
-\textbf{15. Reviewer's Comment:}
+\textbf{11. Reviewer's Comment:}
 
@@
 
-\textbf{15. Response:}
+\textbf{11. Response:}
 

### 2014-12-19 MS_revised.tex
comment 10

@@
 account missing taxa by using the parameter ``sampling factor''
-(\texttt{sampling.f}) in \texttt{diversitree}. We also used constrained
-analyses forcing no effect of hostplant use on diversification and used
-likelihood ratio tests to find out whether the hypothesis of effect on
-diversification has a significantly better likelihood than the null
-hypothesis (no effect). The analyses were run across a sample of 250
+(\texttt{sampling.f}) in \texttt{diversitree}. 
+We also used likelihood ratio tests to evaluate the influence of
+Solanaceae-feeding on diversification rates by comparing a constrained versus
+a relaxed model of diversification.
+The constrained model assumes no effect of hostplant use on diversification by
+enforcing equal speciation rates across both character states. In the relaxed
+model, all speciation and extinction parameters are estimated for each
+character state.
+The analyses were run across a sample of 250
 trees from the selected 1000 random trees from the posterior

### 2014-12-19 response_to_reviewers.tex
comment 11

@@
 
-The closing three sentences (about BiSSE performance) should be moved to
-the methods section.
+Page 5, Estimation of trait-dependent... section, replace “can be feeding on” with “are able to feed on members of”
 
@@
 
-\textbf{16. Reviewer's Comment:}
-
-First paragraph: The authors' assessment of analyzing diversification
-shifts using a single tree, without accommodating our level of certainty
-in relationships, is justified; however, MEDUSA itself, that is, the
-approach using stepwise AIC to identify shifts in diversification rate,
-should not be the target of criticism. After all, MultiMEDUSA is doing
-the same thing, but on a sample of trees. Revise this paragraph with a
-more appropriate target for the criticism (i.e.~analyses performed on
-single trees).
-
-\textbf{16. Response:}
-
-\begin{quote}
-\color{blue}
-We have reworded the paragraph to avoid blaming the method:
-\end{quote}
-
-\begin{quote}
-\color{blue}
-``The MEDUSA method has been used to infer changes in net
-diversification rates in a phylogenetic tree. Since its publication
-{[}16{]} the results of using MEDUSA on a single tree, the maximum clade
-credibility tree, have been used for generation of hypotheses and
-discussion {[}10,42,43{]}. However, different diversification shifts and
-different rates of diversification are found for certain lineages when
-phylogenetic uncertainty was taken into account by using MEDUSA on a
-random sample of trees from the posterior distribution of a Bayesian
-run. We found that some diversification splits, estimated on the
-Nymphalidae maximum clade credibility tree, were found with very low
-probability in the random sample of 1000 trees from the posterior
-distribution (Fig. 3, Table 2). We also found that, even though the
-analyses estimated the same diversification splits on two or more trees,
-the estimated net diversification rates could vary widely (Fig. 2).''
-\end{quote}
-
-\textbf{17. Reviewer's Comment:}
-
-Seventh paragraph, second to last sentence: Two ages for Ithomiini are
-given -- are they the origin and diversification ages? If so, revise the
-sentence to include this information (rather than requiring readers to
-refer to the preceding sentence): ``Reference 10 gives the ages of
-origin and diversification for Ithomiini at 45 (39-53) and 37 (32-43)
-MYA, respectively''.
-
-\textbf{17. Response:}
-
-\begin{quote}
-\color{blue}
-We have followed the reviewer's suggestion.
-\end{quote}
-
-\textbf{18. Reviewer's Comment:}
-
-Eighth paragraph: This paragraph discussing biogeography does not fit
-and could be omitted without detriment to the manuscript.
-
-\textbf{18. Response:}
-
-\begin{quote}
-\color{blue}
-We have followed the reviewer's suggestion and removed this paragraph.
-\end{quote}
-
-\textbf{19. Reviewer's Comment:}
-
-First paragraph, first sentence: replace ``gave'' with ``showed''.
-
-\textbf{19. Response:}
-
-\begin{quote}
-\color{blue}
-We corrected the text following the reviewer's suggestion.
-\end{quote}
-
-\textbf{20. Reviewer's Comment:}
-
-This section is predominantly speculative. Consider a more rigorous
-discussion (if appropriate) or omit.
-
-\textbf{20. Response:}
-
-\begin{quote}
-\color{blue}
-We have followed the reviewer's suggestion and removed this paragraph.
-\end{quote}
-
-\textbf{21. Reviewer's Comment:}
-
-First paragraph, second sentence: revise ``13\% of probabilities of the
-trees from the posterior distribution'' to ``13\% of the trees from the
-posterior distribution''.
-
-\textbf{21. Response:}
-
-\begin{quote}
-\color{blue}
-\color{blue}
-\color{blue}
-We corrected the text following the reviewer's suggestion.
-\end{quote}
-
-\textbf{22. Reviewer's Comment:}
-
-Second paragraph: this paragraph is, for the most part, unnecessary. The
-last sentence of the paragraph should be revised and appended to the end
-of the preceding (first) paragraph: ``\ldots{}not picked up by MEDUSA.
-This would be expected if the diversification of Satyrinae occurred in a
-stepwise manner, with pulses or bursts of diversification for certain
-lineages but unlikely for the tribe Satyrini as a whole''.
-
-\textbf{22. Response:}
-
-\begin{quote}
-\color{blue}
-\color{blue}
-\color{blue}
-We corrected the text following the reviewer's suggestion.
-\end{quote}
-
-\textbf{23. Reviewer's Comment:}
-
-Final two paragraphs: There needs to be a section break, to indicate a
-change in the discussion topic. ``Conclusions'' would be appropriate to
-indicate to readers that these paragraphs represent a summary of the
-major points of the work. The final paragraph is a good summary of the
-work, and it would be a disservice to have readers miss it because it
-appears under the `Satyrini' section.
-
-\textbf{23. Response:}
-
-\begin{quote}
-\color{blue}
-\color{blue}
-\color{blue}
-As suggested by the reviewer, he have created a \textbf{Conclusions}
-section for the last two paragraphs of the text.
-\end{quote}
-
-\textbf{24. Reviewer's Comment:}
-
-Figure 1: The legend needs to indicate what the node labels (the circled
-integers) and the decimal values in the tree indicate.
-
-\textbf{24. Response:}
-
-\begin{quote}
-\color{blue}
-\color{blue}
-\color{blue}
-We have appended the following text to the legend of figure
-1:\\\textbf{``Circles on nodes indicate the diversification shift number
-as found by MEDUSA. Numbers next to circles indicat the posterior
-probability values for such nodes.''}
-\end{quote}
-
-\textbf{25. Reviewer's Comment:}
-
-Figure 2: Omit the legend in the figure (``Boxplot of
-diversification\ldots{}'') and revise the legend to read:
-``Diversification rates for taxa estimated by MEDUSA on the sample of
-1000 trees.''
-
-\textbf{25. Response:}
-
-\begin{quote}
-\color{blue}
-\color{blue}
-\color{blue}
-We removed the legend from the figure file and rewrote the legend in the
-MS as suggested by the reviewer.
-\end{quote}
-
-\textbf{26. Reviewer's Comment:}
-
-Figure 3: Omit the legend in the figure (``Probability of finding
-nodes\ldots{}'') and revise the legend to read: ``Results of MultiMEDUSA
-analysis showing the probability of specific nodes being characterized
-by significant shifts in diversification rate.''
-
-\textbf{26. Response:}
-
-\begin{quote}
-\color{blue}
-We removed the legend from the figure file and rewrote the legend in the
-MS as suggested by the reviewer.
-\end{quote}
-
-\textbf{27. Reviewer's Comment:}
-
-Figure 4: Describe what lambda (\(\lambda\)) and \(r\) represent in the
-figure.
-
-\textbf{27. Response:}
-
-\begin{quote}
-\color{blue}
-We included the following clarification in the legend:
-\textbf{``(speciation rate = \(\lambda1\), net diversification rate =
-\(r1\)).''}
-\end{quote}
-
-\textbf{28, 29. Reviewer's Comment:}
-
-Supporting\_Information\_S1.nex: remove bolded ``Genus.''
-
-\textbf{28, 29. Response:}
-
-\begin{quote}
-\color{blue}
-We followed the reviewer's suggestions and included better descriptions
-for the supplementary materials.
-\end{quote}
-
-\section{Reviewer 2}\label{reviewer-2}
-
-\textbf{1. Reviewer's Comment:}
-
-The authors need to provide more information on what cutoff they used
-and why.
-
-\textbf{1. Response:}
-
-\begin{quote}
-\color{blue}
-Our manuscript already includes information about the AIC value used for
-choosing the best model. The following text appears in our original
-submission in the Methods-\textgreater{} Detecting diversification
-shifts on phylogenetic trees-\textgreater{} last
-paragraph:\\\textbf{``We used the AICc threshold of 7.8 units, as
-estimated by MEDUSA, as the limit for a significantly better fit to
-select among increasingly complex alternative models.''}
-\end{quote}
-
-\textbf{2. Reviewer's Comment:}
-
-I'm not sure discussing the limitations of MEDUSA, or the tendency for
-MEDUSA to give inconsistent results across post burn-in trees, is really
-valid based on one dataset. That is, the unit of replication is the
-alignment -- therefore the authors have effectively one data point and
-zero degrees of freedom to say much of anything about the power of
-MEDUSA across the post burnin sample. I think if the authors want to
-explore the power of MEDUSA, they need many different alignments (real
-or simulated). As it is now, I don't think they really have the data to
-say anything about MEDUSA in general, and only about MEDUSA in the
-context of the Nymphalid tree. They need to at least point out there
-inability to generalize in the manuscript.
-
-\textbf{2. Response:}
-
-\begin{quote}
-\color{blue}
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  As reviewer 1 suggested, our text gave the impression that we were
-  criticizing the method MEDUSA. We have rewritten the following
-  sections to clarify our criticism \textbf{that relying on the
-  inferences made on only one tree (the MCC tree) is risky as analyses
-  of equally good hyphoteses (other trees from the posterior
-  distribution) cand potentially return different results}:
-\end{itemize}
-\end{quote}
-
-\begin{quote}
-\color{blue}
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Abstract\\``In evolutionary biology, analysis of maximum credibility
-  trees in the software MEDUSA (modelling evolutionary diversity using
-  stepwise AIC) is a popular method for estimation of shifts in
-  diversification rates. We investigated whether phylogenetic
-  uncertainty can produce different results by extending the method
-  across a random sample of trees from the posterior distribution of a
-  Bayesian run.''
-\end{itemize}
-\end{quote}
-
-\begin{quote}
-\color{blue}
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Results -\textgreater{} Phylogenetic uncertainty in the MultiMEDUSA
-  approach We changed the sentence ``We tested MEDUSA\ldots{}'' to ``We
-  used MEDUSA to find out\ldots{}''.
-\end{itemize}
-\end{quote}
-
-\begin{quote}
-\color{blue}
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Also, as suggested by the reviewer 1 (point 16 lines above) we
-  improved the discussion.
-\end{itemize}
-\end{quote}
-
-\begin{quote}
-\color{blue}
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  We have added the following in the Conclusions to acknowledge that the
-  reported behaviour might also occur in analyses of other taxa:\\"
-  However, by using a MultiMEDUSA approach, we found that for this
-  Nymphalidae dataset some of these splits might be greatly affected by
-  phylogenetic uncertainty. We recommend that all datasets should be
-  analyzed using both approaches, MEDUSA and MultiMEDUSA, in order to
-  test whether the results are robust when phylogenetic uncertainty is
-  taken into account."
-\end{itemize}
-\end{quote}
-
-\textbf{3. Reviewer's Comment:}
-
-``Our results show that taking phylogenetic uncertainty into account
-when estimating diversification rate shifts is of great importance, and
-relying on the maximum credibility tree alone potentially can give
-erroneous results''. The results are not really in error. The results
-are the correct results for the model. The real question is whether it
-is the appropriate model.
-
-\textbf{3. Response:}
-
-\begin{quote}
-\color{blue}
-We have rewritten that sentence in the abstract as follows:\\``Our
-results show that taking phylogenetic uncertainty into account when
-estimating net diversification rate shifts is of great importance, as
-very different results can be obtained when using the maximum clade
-credibility tree and other trees from the posterior distribution.''
-\end{quote}
-
-\textbf{4. Reviewer's Comment:}
-
-Why are they restricting their analysis to only one method (a method
-that they point out might not be reliable)? There are alternative
-methods available, and it might be interesting to see if consistent
-results are observed (eg., RC, PRC, BAMM). There is very little
-development on other aspects of plant -- herbivore interactions and how
-they might be predicted to affect diversification rate.
-
-\textbf{4. Response:}
-
-\begin{quote}
-\color{blue}
-The methods RC and PRC (parametric rate comparison and relative rates tests) 
-fit several distribution models to the trees in order to
-infer branch lengths. This requires random sampling of taxa to avoid biases.
-This assumption is not satisfied with our data as the sampling is non-random.
-This is one of the characteristics of butterfly datasets where many geographic
-areas are not sampled leaving out major lineages.
-
-The method that we used, BiSSE, might not be reliable when the number of
-taxa is small, and there are biases in the number of species that have
-each of the character states. However, we stated in the manuscript that
-according to the evaluation of BiSSE by Davis, Midford, and Maddison
-(2013), our data should not be affected for such biases.
-
-The method BAMM can be used to explore the evolution of phenotypic
-traits. However the trait characters
-\href{http://bamm-project.org/configuration.html\#id5}{\textbf{must be
-continuous}}. In our case the ``feeding on Solanaceae plants'' trait is
-categorical (absence/presence) so it cannot be analyzed with BAMM.
-
-BAMM also measures the diversification rates of particular clades. We
-run an analysis to test whether we obtained congruent results. In the
-Fig. 1, shown below, we can see that BAMM was able to pick up two
-increases in diversification for Ithomiini butterflies, one for the
-origin of the clade and, most importantly, another increase during the
-recent diversification of the group. This is in agreement with our
-results from using BiSSE, that the ithomiines using Solanaceae underwent
-an increase in diversification rates.
-
-We added to the revised version of the MS that we also ran BAMM in order to
-test congruence of the results with this recent menthods.
-\end{quote}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{response_to_reviewers_fig1.pdf}
-\caption{Bamm analysis}
-\end{figure}
-
-\textbf{5. Reviewer's Comment:}
-
-The authors should `play' with an AIC cutoff, simply to explore how it
-might affect there inferences.
-
-\textbf{5. Response:}
-
-\begin{quote}
-\color{blue}
-As specified lines above and in the manuscript, we used the AIC cutoff
-value 7.8 as suggested by MEDUSA (this value is regarded as optimal due
-to the number of taxa in our analyses). Changing the cutoff value would
-allow fewer or additional diversification shifts to be picked up by
-MEDUSA. Hovewer the initial diversification shifts will remain the same
-unless the AIC cutoff used is so hight that no single diversification
-shifts are recovered. Thus playing with the AIC cutoff values does not
-affect most of the diversification shifts discussed in our manuscript.
-We rerun all our MEDUSA and MultiMEDUSA analysis with differente AIC
-cutoff values to see any effect on the results. We used the AIC values 4
-and 11.
-\end{quote}
-
-\begin{quote}
-\color{blue}
-The results in Fig.2 (also attached as PDF file) shows our MEDUSA
-results from the manuscript (cutoff 7.8) along with results with AIC
-cutoff values 4 and 11. We see that the first 19 diversification shitfs
-are the same for cutoff values 7.8 and 4. However when using AIC cutoff
-4 we recovered additional diversification shifts as the software kept
-trying to fit additional models until the differente in AIC values was
-less than 4. As expected MultiMEDUSA also found additinal
-diversification shifts across the random sample of 1000 trees.
-\end{quote}
-
-\begin{quote}
-\color{blue}
-When we used AIC cutoff 11, we obtained fewer diversification shifts. We
-recovered only 16, which are the same first 16 that we found in our
-analysis with AIC 7.8.
-\end{quote}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{multimedusa_7_8_vs_4_vs_11_AIC_cutoff.pdf}
-\caption{MEDUSA analyses under three AIC values}
-\end{figure}
-
-\textbf{6. Reviewer's Comment:}
-
-In addition, the authors should be cautious about interpreting AIC in a
-frequentist manner (e.g., ``Only three out of 13 significant shifts
-found on the maximum credibility tree were consistent across more than
-95\% of the trees from the posterior'') -- that is, AIC is a model
-selection tool, which should act as a guide for selecting the `best'
-family of models. Delta AIC should really not be thought of as alpha =
-0.05.
-
-\textbf{6. Response:}
-
-\begin{quote}
-\color{blue}
-We agree that AIC values should not be treated in a frequentist manner.
-However, we use the term \textbf{``significant shifts''}, in the same
-way as the authors of the method MEDUSA (Alfaro et al. 2009), to
-pinpoint to specific nodes that show a higher net diversification rate
-than ancestral or descendant nodes.
-\end{quote}
-
-\begin{quote}
-\color{blue}
-We also used the AIC value in the correct way. We used the cutoff 7.8
-for all MEDUSA analyses. Then we counted the number of significant
-shifts infered in each tree and tried to pinpoint those that were
-consistently recovered across the sample of 1000 random trees (and are
-also present in the MCC tree). Since no single significant shift was
-found in all 1000 trees we wanted to show those are were most commonly
-recovered. This is way we considered consistent diversification shifts
-those that were found in most of the trees (in more than 95\% of them).
-\end{quote}
-
-\textbf{7. Reviewer's Comment:}
-
-Why are authors referred to as references, rather than by authors.
-
-As noted by Reference 36, the responsible trait might\ldots{}
-
-\textbf{7. Response:}
-
-\begin{quote}
-\color{blue}
-We have fixed those references as suggested by the reviewer.
-\end{quote}
-
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
-
-Alfaro, Michael E, Francesco Santini, Chad Brock, Hugo Alamillo, Alex
-Dornburg, Daniel L Rabosky, Giorgio Carnevale, and Luke J Harmon. 2009.
-``Nine Exceptional Radiations Plus High Turnover Explain Species
-Diversity in Jawed Vertebrates.'' \emph{P. Natl. Acad. Sci. USA} 106
-(32): 13410--14.
-doi:\href{http://dx.doi.org/10.1073/pnas.0811087106}{10.1073/pnas.0811087106}.
-
-Davis, Matthew P, Peter E Midford, and Wayne Maddison. 2013. ``Exploring
-power and parameter estimation of the BiSSE method for analyzing species
-diversification.'' \emph{BMC Evol. Biol.} 13 (1): 38.
-doi:\href{http://dx.doi.org/10.1186/1471-2148-13-38}{10.1186/1471-2148-13-38}.
-
 \end{document}

### 2014-12-19 MS_revised.tex
comment 11

@@
 feeding on Solanaceae \cite{beccaloni2008, scott1986} might be incorrect as it is unlikely
-that these species can be feeding on this plant family. We tested
+that these species are able to feed on members of this plant family. We tested
 whether coding these two genera as not feeding on Solanaceae affected

### 2014-12-19 response_to_reviewers.tex
comment 12
comment 13

@@
 
+\textbf{12. Reviewer's Comment:}
+ Page 5, Detecting diversification shifts... section: should final number be 1.884, instead of 1,884?
+
+\textbf{12. Response:}
+
+\begin{quote}
+\color{blue}
+We have removed the whole sentence as the reviewer recommended in point 5.
+\end{quote}
+
+\textbf{13. Reviewer's Comment:}
+Page 6, sentence beginning “For the diversification shifts..”: the last phrase could use a verb; something like “when the MCC tree alone was used”.
+
+\textbf{13. Response:}
+
+\begin{quote}
+\color{blue}
+We added "was used" to the end of the sentence.
+\end{quote}
 \end{document}

### 2014-12-19 MS_revised.tex
comment 13

@@
 MultiMEDUSA approach recovered different rates of diversification than
-those found when the MCC tree alone.
+those found when the MCC tree alone was used.
 

### 2015-01-12 response_to_reviewers.tex
comment 14

@@
 
- Page 3, first paragraph of Data section in Materials & Methods, final sentence: replace “correct for”
+ Page 3, first paragraph of Data section in Materials \& Methods, final sentence: replace “correct for”
 with “account for”.
@@
 
-Page 3, second paragraph of Data section in Materials & Methods, first sentence: the two web
+Page 3, second paragraph of Data section in Materials \& Methods, first sentence: the two web
 resources (Tree of Life Web Project and Global Butterfly Names project) should be listed in the
@@
 We have removed the BAMM analysis from the manuscript.
-\end{quote}
-
-\begin{quote}
-\color{blue}
-``Because most of Nymphalidae butterflies are restricted to use one
-plant family as hostplant (232 genera use only one plant family versus
-176 that use more than one; based on our data in supp. mat. 15), the
-\ldots{}''
 \end{quote}
@@
 \end{quote}
+
+\textbf{14. Reviewer's Comment:}
+Page 6, sentence beginning “There were four...”: all shifts described are increases.
+If this is the case, the sentence can be simplified, moving the congruent part of each shift to a single instance preceding the list: “There were four net diversification rate increases found in the trees from the random sample (Table 2); the shifts occurred in: (i) the genus Ypthima (r = 0:22); (ii) the genus Charaxes (r = 0:21); (iii) Ithomiini subtribes Oleriina + Ithomiina + Napeogenina + Dircennina + Godyridina (r = 0:10); and (iv) the clade of Callicore + Diaethria (r = 0:135).”
+
+\textbf{14. Response:}
+
+\begin{quote}
+\color{blue}
+We modified the sentence as suggested by the reviewer.
+\end{quote}
+
 \end{document}

### 2015-01-12 MS_revised.tex
comment 14

@@
 
-There were four diversification shifts found in the trees from the
-random sample (Table 2): (i) net diversification rate increase in the
-genus \emph{Ypthima} (\(r = 0.22\)); (ii) net diversification rate
-increase in the genus \emph{Charaxes} (\(r = 0.21\)); (iii) rate
-increase in Ithomiini subtribes Oleriina + Ithomiina + Napeogenina +
-Dircennina + Godyridina (\(r = 0.10\)); and (iv) rate increase in
+There were four net diversification rate increases found in the trees from the
+random sample (Table 2): (i) the
+genus \emph{Ypthima} (\(r = 0.22\)); (ii) the genus \emph{Charaxes}
+(\(r = 0.21\)); (iii) Ithomiini subtribes Oleriina + Ithomiina + Napeogenina +
+Dircennina + Godyridina (\(r = 0.10\)); and (iv) the clade of 
 \emph{Callicore} + \emph{Diaethria} (\(r = 0.135\)).

### 2015-01-12 out.tex
comment 14

--- 2014-10-29 MS_orig.tex
@@
+% !TeX encoding: utf-8
 \documentclass[10pt]{article}
@@
 
-%% Include all macros below. Please limit the use of macros.
-
-%% END MACROS SECTION
-
+% Include all macros below. Please limit the use of macros.
+% END MACROS SECTION
 \begin{document}
-
 \begin{flushleft}
 {\Large
-\title{Diversity dynamics in Nymphalidae butterflies: Effect of
-phylogenetic uncertainty on diversification rate shift estimates}
+    \title{Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates}
 }
@@
 
-\textbf{Comments:} 5814 words, 7 figures, 2 tables and 25 supplementary
-material files.
-
-\textbf{Additional Supplementary materials:}
-\url{http://dx.doi.org/10.6084/m9.figshare.639208}
 \end{flushleft}
@@
 employing the method BiSSE as implemented in the R package
-\texttt{diversitree} {[}15{]}.
-
-\section*{Methods}
+\texttt{diversitree} {[}18{]}.
+
+\section*{Materials and Methods}
 
@@
 
-For analyses, we used the phylogenetic trees from the study of Reference
-10 that were generated using DNA sequence data from 10 gene regions for
-398 of the 540 valid genera in Nymphalidae. We employed Reference 10
-maximum clade credibility tree (MCC tree) (Fig. 1) as well as a random
-sample of 1000 trees from their BEAST run after burnin. Their original
-BEAST run was for 40 million generations. We used a burnin of 25 million
-generations and took a random sample of 1000 trees using Burntrees
-v.0.1.9 \url{http://www.abc.se/~nylander/} (supp. mat. 5) in order to
-correct for phylogenetic uncertainty when performing the diversification
-analyses.
+For analyses, we used the phylogenetic trees from the study of Wahlberg
+et al. (2009) {[}13{]} that were generated using DNA sequence data from
+10 gene regions for 398 of the 540 valid genera in Nymphalidae. We
+employed the maximum clade credibility tree (MCC tree) (Fig. 1) as well
+as a random sample of 1000 trees from their BEAST run after burnin
+{[}13{]}. Their original BEAST run was for 40 million generations. We
+used a burnin of 25 million generations and took a random sample of 1000
+trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (supp.
+mat. 5) in order to correct for phylogenetic uncertainty when performing
+the diversification analyses.
 
@@
 sources including the specialist-curated lists on
-\url{http://tolweb.org}, Reference 16 and curated lists of Global
-Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}). We
-assigned the species numbers of genera not included in the phylogeny to
-the closest related genus that was included in Reference 10 study
-according to available phylogenetic studies {[}17--28{]}, taxonomical
-classification and morphological resemblance when no phylogenies were
-available (supp. mat. 3).
+\url{http://tolweb.org}, Lamas (2004) {[}19{]} and curated lists of
+Global Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}).
+We assigned the species numbers of genera not included in the phylogeny
+to the closest related genus that was included in Wahlberg et al. (2009)
+{[}13{]} study according to available phylogenetic studies {[}20--31{]},
+taxonomical classification and morphological resemblance when no
+phylogenies were available (supp. mat. 3).
 
 Hostplant data for Nymphalidae species were compiled from several
-sources including Reference 29, HOSTS database
-(\url{http://bit.ly/YI7nwW}), Reference 30, Reference 31, Reference 32
-and others (supp. mat. 15--16) for a total of 6586 hostplant records,
-including 428 Nymphalidae genera and 143 plant families and 1070 plant
-genera. It was not possible to find any hostplant data for 35 butterfly
-genera.
-
-\subsection*{Analyses of
-Diversification}
-
-We used the statistical software R version 3.0.1 {[}33{]} in combination
-with the APE {[}34{]}, GEIGER {[}35{]}, MEDUSA {[}13{]} and
-\texttt{diversitree} {[}15{]} packages along with our own scripts to
+sources including {[}32{]}, HOSTS database (\url{http://bit.ly/YI7nwW}),
+{[}33--35{]} and others (supp. mat. 15--16) for a total of 6586
+hostplant records, including 428 Nymphalidae genera and 143 plant
+families and 1070 plant genera. It was not possible to find any
+hostplant data for 35 butterfly genera.
+
+\subsection*{Analyses of Diversification}
+
+We used the statistical software R version 3.0.1 {[}36{]} in combination
+with the APE {[}37{]}, GEIGER {[}38{]}, MEDUSA {[}16{]} and
+\texttt{diversitree} {[}18{]} packages along with our own scripts to
 perform the analyses (included as supplementary materials 2, 6, 10). All
-analyses were run on the 1000 random trees from Reference 10, on the
-maximum clade credibility tree derived from these and the MCC tree from
-Reference 10.
+analyses were run on the 1000 random trees from Wahlberg et al. (2009)
+{[}13{]}, on the maximum clade credibility tree derived from these and
+the MCC tree {[}13{]}.
 
@@
 Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
-version 093.4.33 {[}13{]} on the maximum clade credibility tree from
-Reference 10. MEDUSA fits alternative birth-death likelihood models to a
-phylogenetic tree in order to estimate changes in net diversification
-rates along branches. MEDUSA estimates likelihood and AIC scores for the
-simplest birth-death model, with two parameters, the rates \texttt{r}:
-net diversification \texttt{=} speciation (\texttt{b}) \texttt{-}
-extinction (\texttt{d}) and \(\varepsilon\): relative extinction
-\texttt{= d/b}. The AIC scores of the two-parameter model are then
-compared with incrementally more complex models until the addition of
-parameters do not improve the AIC scores beyond a cutoff value. MEDUSA
-finds the likelihood of the models after taking into account branch
-lengths and number of species per lineage {[}13{]}. Most studies using
-MEDUSA only run the method on a single tree, usually the maximum clade
-credibility tree, which makes the assumption that this tree is correct.
-We wanted to study the effects of phylogenetic uncertainty on estimation
-of net diversification rate shifts and therefore selected 1000 random
-genus-level trees from the posterior distribution (MultiMEDUSA, supp.
-mat. 5). We calculated a new MCC tree from the selection of trees, ran
-MEDUSA and MultiMEDUSA on the selection of 1000 trees and summarized the
-estimated changes in net diversification rates for nodes across all
-trees. Patterns of change in net diversification rates are considered
-significant if they are found at the same node in at least 90\% of the
-trees. We also expected to find similar values of net diversification
-(\texttt{r = b - d}) and relative extinction rate (\(\varepsilon\)
-\texttt{= d/b}) values across the 1000 trees for the nodes where changes
-in diversification tempo occurs. We used the AICc threshold of 7.8
-units, as estimated by MEDUSA, as the limit for a significantly better
-fit to select among increasingly complex alternative models.
-
-\subsection*{Estimation of trait-dependent speciation
-rates}
+version 093.4.33 {[}16{]} on the maximum clade credibility tree from
+Wahlberg et al. (2009) {[}13{]}. MEDUSA fits alternative birth-death
+likelihood models to a phylogenetic tree in order to estimate changes in
+net diversification rates along branches. MEDUSA estimates likelihood
+and AIC scores for the simplest birth-death model, with two parameters,
+the rates \texttt{r}: net diversification \texttt{=} speciation
+(\texttt{b}) \texttt{-} extinction (\texttt{d}) and ($\varepsilon$):
+relative extinction \texttt{= d/b}. The AIC scores of the two-parameter
+model are then compared with incrementally more complex models until the
+addition of parameters do not improve the AIC scores beyond a cutoff
+value. MEDUSA finds the likelihood of the models after taking into
+account branch lengths and number of species per lineage {[}16{]}. Most
+studies using MEDUSA only run the method on a single tree, usually the
+maximum clade credibility tree, which makes the assumption that this
+tree is correct. We wanted to study the effects of phylogenetic
+uncertainty on estimation of net diversification rate shifts and
+therefore selected 1000 random genus-level trees from the posterior
+distribution of the Bayesian run from Wahlberg et al., 2009 {[}13{]}
+(MultiMEDUSA, supp. mat. 5). We calculated a new MCC tree from the
+selection of trees, ran MEDUSA and MultiMEDUSA on the selection of 1000
+trees and summarized the estimated changes in net diversification rates
+for nodes across all trees. Patterns of change in net diversification
+rates are considered significant if they are found at the same node in
+at least 90\% of the trees. We also expected to find similar values of
+net diversification (\texttt{r = b - d}) and relative extinction rate
+($\varepsilon$) \texttt{= d/b}) values across the 1000 trees for the
+nodes where changes in diversification tempo occurs. We used the AICc
+threshold of 7.8 units, as estimated by MEDUSA, as the limit for a
+significantly better fit to select among increasingly complex
+alternative models.
+
+\subsection*{Estimation of trait-dependent speciation rates}
 
@@
 hypothesis (no effect). The analyses were run across a sample of 250
-trees from the posterior distribution. The records of \emph{Vanessa} and
-\emph{Hypanartia} feeding on Solanaceae {[}30,38{]} might be incorrect
-as it is unlikely that these species can be feeding on this plant
-family. We tested whether coding these two genera as not feeding on
-Solanaceae affected our results.
+trees from the selected 1000 random trees from the posterior
+distribution. The records of \emph{Vanessa} and \emph{Hypanartia}
+feeding on Solanaceae {[}33,40{]} might be incorrect as it is unlikely
+that these species can be feeding on this plant family. We tested
+whether coding these two genera as not feeding on Solanaceae affected
+our results. We expected that coding these two genera as either absence
+or presence would not have a significant effect on our results. It has
+been shown that BiSSE performs poorly under certain conditions {[}41{]}.
+However, our data has adequate number of taxa under analysis (more than
+300 tips), adequate speciation bias (between 1.5x and 2.0x), character
+state bias (around 8x) and extinction bias (around 4x) for the analysis
+of Solanaceae hostplants. Thus, BiSSE is expected to produce robust
+results {[}41{]}.
 
@@
 
-\subsection*{Phylogenetic uncertainty in the MultiMEDUSA
-approach}
-
-We tested MEDUSA to find out whether taking into account the
-phylogenetic signal from a random sample of 1000 trees from the
-posterior distribution can return similar estimates of diversification
-to the values obtained from the MCC tree. We run MultiMEDUSA on a random
+We contrasted these results with those obtained from analysing the data in
+the software BAMM {[}42{]}. We found
+that the best model (best configuration) estimated 2 diversification shifts in
+our MCC tree (posterior probability = 0.79) for the Ithomiini
+ and Satyrini. Bayes factor (BF) values strongly 
+favored this model in comparison with a constant rate model (no diversification
+shifts, BF $>$ 1,884).
+
+\subsection*{Phylogenetic uncertainty in the MultiMEDUSA approach}
+
+We used MEDUSA to find out whether taking into account the phylogenetic
+signal from the random sample of 1000 trees from the posterior
+distribution can return similar estimates of diversification to the
+values obtained from the MCC tree. We ran MultiMEDUSA on the random
 sample of 1000 trees (supp. mat. 5) from the posterior distribution and
@@
 example the split for the clade Satyrini (Euptychiina + Pronophilina +
-Satyrina + Maniolina) is recovered with a probability of 18\% (Fig. 3).
-
-\subsection*{Estimation of trait-dependent speciation
-rates}
-
-The MEDUSA analyses taking into account phylogenetic uncertainty
+Satyrina + Maniolina) is recovered with a probability of 0.18 (Fig. 3).
+
+\subsection*{Estimation of trait-dependent speciation rates}
+
+The MEDUSA analyses, taking into account phylogenetic uncertainty,
 estimated a net diversification rate increase in part of the clade
@@
 net net diversification rate on part of Ithomiini (Oleriina + Ithomiina
-+ Napeogenina + Dircennina + Godyridina) (Fig. 4). The Markov Chain
-Monte Carlo algorithm was run for 10000 generations discarding the first
-7500 as burnin. The estimated mean net diversification rate for taxa
-that do not feed on Solanaceae was \texttt{r = 0.11} while the net
-diversification rate for the Solanaceae feeders was \texttt{r = 0.16}
-(see Fig. S1 for a boxplot of speciation and extinction values for the
-95\% credibility intervals). The same analysis considering
-\emph{Vanessa} and \emph{Hypanartia} as non-Solanaceae feeders due to
-dubious records produced the same pattern and net diversification rates
-(Fig. S2). Therefore, the rest of the analyses were performed assuming
-these two genera as Solanaceae feeders. We constrained the BiSSE
-likelihood model to force equal rates of speciation for both character
-states in order to test whether the model of different speciation rates
-is a significantly better explanation for the data. A likelihood ratio
-test found that the model for increased net diversification rate for
-nymphalids feeding on Solanaceae is a significantly better explanation
-than this character state having no effect on diversification
-(\texttt{p \textless{} 0.001}) (Table 3; character states available in
-supp. mat. 17, code in supp. mat. 18, and mcmc run in supp. mat. 19). We
-combined the post-burnin mcmc generations from running BiSSE on 250
-trees from the posterior distribution and found the same pattern as the
-BiSSE analysis on the maximum clade credibility tree (combined mcmc run
-in supp. mat. 20; profiles plot of speciation rates in Fig. S3; boxplot
-of 95\% credibility intervals in Fig. S4). A BiSSE analysis to test
-whether the trait ``feeding on Apocynaceae'' had any effect on increased
-net diversification rates found similar speciation rates for lineages
-feeding on Apocynaceae and other plants (Fig. S5). It has been shown
-that BiSSE performs poorly under certain conditions {[}39{]}. However,
-our data has adequate number of taxa under analysis (more than 300
-tips), adequate speciation bias (between 1.5x and 2.0x), character state
-bias (around 8x) and extinction bias (around 4x) for the analysis of
-Solanaceae hostplants. Thus, BiSSE is expected to produce robust results
-{[}39{]}.
++ Napeogenina + Dircennina + Godyridina) (Fig. 4). The estimated mean
+net diversification rate for taxa that do not feed on Solanaceae was
+\texttt{r = 0.11} while the net diversification rate for the Solanaceae
+feeders was \texttt{r = 0.16} (see Fig. S1 for a boxplot of speciation
+and extinction values for the 95\% credibility intervals). The same
+analysis considering \emph{Vanessa} and \emph{Hypanartia} as
+non-Solanaceae feeders due to dubious records produced the same pattern
+and net diversification rates (Fig. S2). Therefore, the rest of the
+analyses were performed assuming these two genera as Solanaceae feeders.
+We constrained the BiSSE likelihood model to force equal rates of
+speciation for both character states in order to test whether the model
+of different speciation rates is a significantly better explanation for
+the data. A likelihood ratio test found that the model for increased net
+diversification rate for nymphalids feeding on Solanaceae is a
+significantly better explanation than this character state having no
+effect on diversification ($\chi^2 = 12.3; 1 df; p < 0.001$) (Table 3;
+character states available in supp. mat. 17, code in supp. mat. 18, and
+mcmc run in supp. mat. 19). We combined the post-burnin mcmc generations
+from running BiSSE on 250 trees from the random sample of 1000 trees and
+found the same pattern as the BiSSE analysis on the maximum clade
+credibility tree (combined mcmc run in supp. mat. 20; profiles plot of
+speciation rates in Fig. S3; boxplot of 95\% credibility intervals in
+Fig. S4). A BiSSE analysis to test whether the trait ``feeding on
+Apocynaceae'' had any effect on increased net diversification rates
+found similar speciation rates for lineages feeding on Apocynaceae and
+other plants (Fig. S5).
 
@@
 
-\subsection*{Effects of phylogenetic uncertainty on the performance of
-MEDUSA}
+\subsection*{Effects of phylogenetic uncertainty on the performance of MEDUSA}
 
 The MEDUSA method has been used to infer changes in net diversification
-rates along a phylogenetic tree. Since its publication {[}13{]} the
-results of using MEDUSA on a single tree, the maximum clade credibility
-tree, have been used for generation of hypotheses and discussion
-{[}7,40,41{]}. However, we found that MEDUSA estimated different
-diversification shifts and different rates of diversification for
-certain lineages when phylogenetic uncertainty was taken into account by
-using MEDUSA on a random sample of trees from the posterior distribution
-of a Bayesian run. We found that some diversification splits, estimated
-on the Nymphalidae maximum clade credibility tree, were found with a
-very low probability in a random sample of 1000 trees from the posterior
-distribution (Fig. 3, Table 2). We also found that, even though MEDUSA
-could estimate the same diversification splits on two or more trees, the
-estimated net diversification rates could vary widely (Fig. 2). For
+rates in a phylogenetic tree. Since its publication {[}16{]} the results
+of using MEDUSA on a single tree, the maximum clade credibility tree,
+have been used for generation of hypotheses and discussion
+{[}10,43,44{]}. However, different diversification shifts and different
+rates of diversification are found for certain lineages when
+phylogenetic uncertainty was taken into account by using MEDUSA on a
+random sample of trees from the posterior distribution of a Bayesian
+run. We found that some diversification splits, estimated on the
+Nymphalidae maximum clade credibility tree, were found with very low
+probability in the random sample of 1000 trees from the posterior
+distribution (Fig. 3, Table 2). We also found that, even though the
+analyses estimated the same diversification splits on two or more trees,
+the estimated net diversification rates could vary widely (Fig. 2). For
 example, in our Nymphalidae trees, we found that the split for
@@
 The clade Ithomiini and the non-basal danaids are well supported by high
-posterior probability values in Reference 10. Therefore our MEDUSA
-analyses recovered an increase in net diversification rate with
-probability higher than 90\% in the posterior distribution of trees
-(Fig. 3).
-
-\subsection*{Hostplant use and diversification in
-Nymphalidae}
-
-\paragraph*{Ithomiini}
+posterior probability values in Wahlberg et al. (2009) {[}13{]}.
+Therefore our MEDUSA analyses recovered an increase in net
+diversification rate with probability higher than 0.90 in the posterior
+distribution of trees (Fig. 3).
+
+\subsection*{Hostplant use and diversification in Nymphalidae}
+
+\subsubsection*{Ithomiini}
 
 Keith Brown suggested that feeding on Solanaceae was an important event
-in the diversification of Ithomiini butterflies {[}42{]}. Ithomiini
+in the diversification of Ithomiini butterflies {[}45{]}. Ithomiini
 butterflies are exclusively Neotropical and most species feed on
-Solanaceae hostplants during larval stage {[}37{]}. Optimizations of the
+Solanaceae hostplants during larval stage {[}8{]}. Optimizations of the
 evolution of hostplant use on phylogenies evidence a probable shift from
-Apocynaceae to Solanaceae in the ancestor of the tribe {[}37,43{]}.
-Reference 9 found that the Gamma statistics, a LTT plot of an Ithomiini
-phylogeny and the fit of the density-dependent model of diversification
-are consistent with a burst of diversification in Ithomiini following
-the shift from Apocynaceae to Solanaceae.
+Apocynaceae to Solanaceae in the ancestor of the tribe {[}8,46{]}.
+Fordyce (2010) {[}12{]} found that the Gamma statistics, a LTT plot of
+an Ithomiini phylogeny and the fit of the density-dependent model of
+diversification are consistent with a burst of diversification in
+Ithomiini following the shift from Apocynaceae to Solanaceae.
 
@@
 due to the use of Solanaceae plants as hosts during larval stage. For
-this, we used a Bayesian approach {[}44{]} to test whether the trait
+this, we used a Bayesian approach {[}47{]} to test whether the trait
 ``feeding on Solanaceae'' had any effect on the diversification of the
@@
 that the high diversity of Ithomiini is consistent with the
-``escape-and-radiate scenario'' due to a shift onto Solanaceae {[}9{]}
+``escape-and-radiate scenario'' due to a shift onto Solanaceae {[}12{]}
 and radiation scenarios among chemically different lineages of
-Solanaceae plants {[}37,42{]}. According to this theory, the shift from
+Solanaceae plants {[}8,45{]}. According to this theory, the shift from
 Apocynaceae to Solanaceae allowed Ithomiini to invade newly available
@@
 detoxification of ingested food as they have been found in other
-butterfly groups {[}45{]}.
+butterfly groups {[}48{]}.
 
@@
 configurations of the proteins, providing more opportunities for
-hostplant switch {[}46{]}. This might be the reason why the basal
+hostplant switch {[}49{]}. This might be the reason why the basal
 Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and
-restricted to few Solanaceae hosts {[}37{]}, while recent subtribes are
+restricted to few Solanaceae hosts {[}8{]}, while recent subtribes are
 species-rich and have expanded their host range into several Solanaceae
-lineages {[}37{]}. It might be that the switch to feeding in Solanaceae
+lineages {[}8{]}. It might be that the switch to feeding in Solanaceae
 was an important event in the evolutionary history of Ithomiini, but the
@@
 diversity'' and the ``escape-and-radiate'' hypotheses state that insects
-diversify after their hostplants {[}2--4{]}. Reference 45 found strong
-evidence for a model of speciation congruent with Ehrlich and Raven's
-hypothesis in Pieridae butterflies due to, in addition to the
-identification of a key innovation, a burst of diversification in
-glucosinolate-feeding taxa shortly afterwards (with a lag of
-\textasciitilde{}10 MY). According to a recent dated phylogeny of the
-Angiosperms {[}47{]}, the family Solanaceae split from its sister group
-about 59 (49-68) MYA and diversification started (crown group age)
-around 37 (29-47) MYA. Reference 10 give the corresponding ages for
-Ithomiini as 45 (39-53) and 37 (32-43) MYA, respectively. Thus, current
-evidence shows that Solanaceae and Ithomiini might have diversified
-around the same time, during the Late Eocene and Oligocene, and this
-would be congruent with the diffuse cospeciation hypothesis.
-
-The uplift of the Andes was a major tectonic event that underwent higher
-activity during the Oligocene {[}48{]}. This caused climatic changes in
-the region that affected the flora and fauna of the time, which
-coincides with the diversification of modern montane plant and animal
-taxa {[}49{]} including Ithomiini butterflies and Solanaceae hostplants.
-Moreover, all Solanaceae clades currently present in New World
-originated in South America {[}50{]} as well as Ithomiini butterflies
-{[}10{]}. This suggests a process of ``diffuse cospeciation'' of
-Ithomiini and hostplants.
-
-\paragraph{Danaini}
-
-Our MultiMEDUSA approach gave a significant slowdown in net
+diversify after their hostplants {[}2--4{]}. Wheat et al. (2007)
+{[}48{]} found strong evidence for a model of speciation congruent with
+Ehrlich and Raven's hypothesis in Pieridae butterflies due to, in
+addition to the identification of a key innovation, a burst of
+diversification in glucosinolate-feeding taxa shortly afterwards (with a
+lag of circa 10 MY). According to a recent dated phylogeny of
+the Angiosperms {[}50{]}, the family Solanaceae split from its sister
+group about 59 (49--68) MYA and diversification started (crown group
+age) around 37 (29--47) MYA. Wahlberg et al. (2009) {[}13{]} give the
+ages for origin and diversification for Ithomiini at 45 (39--53) and 37
+(32--43) MYA, respectively. Thus, current evidence shows that Solanaceae
+and Ithomiini might have diversified around the same time, during the
+Late Eocene and Oligocene, and this would be congruent with the diffuse
+cospeciation hypothesis.
+
+\subsubsection*{Danaini}
+Our MultiMEDUSA approach showed a significant slowdown in net
 diversification rate in the subtribe Danaina of the Danini. Both Danaina
@@
 
-\paragraph{\emph{Charaxes}}
-
-The genus \emph{Charaxes} contain 193 species distributed in the Old
-World with highest diversity in the Afrotropical region. These
-butterflies are also very strong fliers, but contrary to Danaina, which
-are specialized Apocynaceae feeders, \emph{Charaxes} are known to feed
-on at least 28 plant families in 18 orders {[}29{]} and some species
-appear to be polyphagous {[}52{]}. Reference 53 showed that most of the
-diversification of the genus occurred from the late Oligocene to the
-Miocene when there were drastic global climatic fluctuations, indicating
-that the diversity mainly is driven by climate change. Reference 52
-found that climatic changes during the Pliocene, Pleistocene as well as
-dispersal and vicariance might have been responsible for the high
-diversification of the genus.
-
-\paragraph{Satyrini}\label{satyrini}
+\subsubsection*{Satyrini}
 
@@
 their main hostplant, grasses, during the climatic cooling in the
-Oligocene {[}11{]}. Thus, it is somewhat surprising that part of
+Oligocene {[}14{]}. Thus, it is somewhat surprising that part of
 Satyrini (the subtribes Euptychiina, Satyrina and Pronophilina) were
-found to have accelerated diversification in only 13\% of probabilities
-in the trees from the posterior distribution. Although this can be
-attributed to low phylogenetic signal {[}10{]}, the clade Satyrini is
-very robust {[}10{]} and MEDUSA failed to identify any significant
-accelerated net diversification rate for Satyrini. It appears that the
-radiation of Satyrini as a whole was not remarkably fast and therefore
-not picked up by MEDUSA.
-
-The origin of the tribe Satyrini is not completely clear (originated
-either in the Neotropical or Eastern Palaearctic, Oriental and/or
-Indo-Australian regions {[}27{]} and their radiation involved colonizing
-almost all continents starting from their place of origin. For
-butterflies, a prerequisite for colonizing new areas is that suitable
-hostplants are already present. Therefore, it should be expected that
-the diversification of Satyrinae occurred in a stepwise manner, with
-pulses or bursts of diversification for certain lineages but unlikely
-for the tribe Satyrini as a whole.
-
+found to have accelerated diversification in only 13\% of the trees from
+the posterior distribution. Although this can be attributed to low
+phylogenetic signal {[}13{]}, the clade Satyrini is very robust {[}13{]}
+and MEDUSA failed to identify any significant accelerated net
+diversification rate for Satyrini. It appears that the radiation of
+Satyrini as a whole was not remarkably fast and therefore not picked up
+by MEDUSA. This should be expected if the diversification of Satyrinae
+occurred in a stepwise manner, with pulses or bursts of diversification
+for certain lineages but unlikely for the tribe Satyrini as a whole.
+
+\section*{Conclusions}
 We found that even though MEDUSA estimated several diversification
@@
 
-\subsection*{Acknowledgments}\label{acknowledgments}
-
+\section*{Acknowledgments}
 We are thankful to Mark Cornwall for help with the script to extend
@@
 
-\subsection*{Figure legends}\label{figure-legends}
-
-\texttt{Figure 1.} Results of the MEDUSA analysis run on the maximum
-clade credibility tree from Reference 10. Rate shifts were estimated for
-the following nodes (besides the background rate): 2) Limenitidinae +
+\section*{Figure legends}
+\begin{description}
+ \item {\bf Figure 1. Results of the MEDUSA analysis run on the maximum
+clade credibility tree from Wahlberg et al. (2009)} Rate shifts were estimated for the
+following nodes (besides the background rate): 2) Limenitidinae +
 Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
@@
 \emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
-Satyrina, 14) Mycalesinaa, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
-\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae.
-
-\texttt{Figure 2.} Boxplot of the range of diversification values for
-tips or clades estimated by MEDUSA on the 1000 random trees from the
-posterior distribution of the Nymphalidae phylogeny. The tips or clades
-shown are those present on the maximum clade credibility tree from
-Reference 10.
-
-\texttt{Figure 3.} Results of the MultiMEDUSA analysis on 1000 randomly
-sampled trees from the posterior distribution of the Nymphalidae
-phylogeny. Bars show the probability for nodes of being recovered as
-significant increases or decreases in net diversification rates by
-MultiMEDUSA.
-
-\texttt{Figure 4.} BiSSE analysis of diversification of nymphalids due
-to feeding on Solanaceae hostplants. Speciation and net diversification
-rates are significantly higher in Solanaceae feeders (\$lambda\$1, r1).
-
-\subsection*{Tables}\label{tables}
-
-\texttt{Table 1.} Significant net diversification rate shifts found in
-the MEDUSA analysis of the Nymphalid maximum clade credibility tree.
-Split.Node = node number, Model = preferred diversification model by
-MEDUSA, r = net diversification rate, LnLik.part = log likelihood value.
-
-\texttt{Table 2.} Differences in rates estimated by MEDUSA on the MCC
-tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
+Satyrina, 14) Mycalesina, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
+\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae. Circles on
+nodes indicate the diversification shift number as found by MEDUSA.
+Numbers next to circles indicat the posterior probability values for
+such nodes.
+
+\item {\bf Figure 2. Diversification rates for taxa estimated by MEDUSA on
+the samples of 1000 random trees}
+
+\item {\bf Figure 3. Results of MultiMEDUSA analysis showing the
+probability of specific nodes being characterized by significant shifts
+in diversification rate}
+
+\item {\bf Figure 4. BiSSE analysis of diversification of nymphalids due
+to feeding on Solanaceae hostplants} Speciation and net diversification
+rates are significantly higher in Solanaceae feeders (speciation rate =
+$\lambda1$, net diversification rate = $r1$.
+\end{description}
+
+\section*{Supporting Information Legends}
+\begin{description}
+ \item {\bf Supporting Information S01.nex.} MCC Nymphalidae tree from
+{[}13{]}
+
+ \item {\bf Supporting Information S02.R} R script to run MEDUSA on the
+MCC Nymphalidae tree from {[}13{]}. This script removes the outgroup
+taxa and loads the richness data for the tree terminals.
+
+ \item {\bf Supporting Information S03.csv} Species richness for lineages
+in Nymphalidae.
+
+ \item {\bf Supporting Information S04.txt} Results of running MEDUSA on
+the MCC tree.
+
+ \item {\bf Supporting Information S05 1000 random trees no outgroups.zip} 1000 random trees from {[}13{]}.
+
+ \item {\bf Supporting Information S06.R}?R script to run a MultiMEDUSA
+analysis on 1000 random trees from {[}13{]}.
+
+ \item {\bf Supporting Information S07 multimedusa on 1000 trees.zip} Raw results from the MultiMEDUSA run on the random sample of trees from
+the posterior distribution.
+
+ \item {\bf Supporting Information S08.pdf} Summary of the MulitMEDUSA
+analysis on the 1000 trees from the posterior distribution.
+
+ \item {\bf Supporting Information S09.tree} MCC tree from the 1000
+random trees selected from the posterior distribution.
+
+ \item {\bf Supporting Information S10.R} R script to run a MEDUSA
+analysis on the MCC tree from the 1000 random trees selected from the
+posterior distribution.
+
+ \item {\bf Supporting Information S11.txt} Summary results from a MEDUSA
+run on the MCC tree from the 1000 random trees selected from the
+posterior distribution.
+
+ \item {\bf Supporting Information S12.pdf} Figure for MEDUSA run on MCC
+tree from random 1000 trees.
+
+ \item {\bf Supporting Information S13.txt} Summary results from a
+MultiMEDUSA run on the 1000 random trees selected from the posterior
+distribution.
+
+ \item {\bf Supporting Information S14.pdf} Probability of nodes with
+estimated rates from a MultiMEDUSA run on the 1000 random trees selected
 from the posterior distribution.
 
-\texttt{Table 3.} Likelihood ratio test between the model of increased
-diversification of nymphalids feeding on Solanaceae against a model
-forcing equal speciation rates (no effect on diversification).
-
-\section*{Supp. mat.}\label{supp.-mat.}
-
-\textbf{\texttt{supp mat 01 genus.nex}}: MCC Nymphalidae tree from
-Reference 10.
-
-\textbf{\texttt{supp mat 02 run medusa on mct.R}}: Run MEDUSA on
-MCC tree from Reference 10.
-
-\textbf{\texttt{supp mat 03 richness.nex}}: Species richness for
-lineages in Nymphalidae.
-
-\textbf{\texttt{supp mat 04 MEDUSA results.txt}}: Results of running
-MEDUSA on the MCC tree.
-
-\textbf{\texttt{supp mat 05 1000 random trees no outgroups.nex}}:
-1000 random trees from Reference 10.
-
-\textbf{\texttt{supp mat 06 run multimedusa.R}}: Run MultiMEDUSA on
-1000 random trees from Reference 10.
-
-\textbf{\texttt{supp mat 07 multimedusa on 1000 trees.txt}}: Raw
-results from the MultiMEDUSA run on the random sample of trees from the
-posterior distribution.
-
-\textbf{\texttt{supp mat 08 multimedusa result01.pdf}}: Summary of
-MulitMEDUSA analysis on the 1000 trees from the posterior distribution.
-
-\textbf{\texttt{supp mat 09 mcc tree from random sample of 1000.tree}}:
-MCC tree from the 1000 random trees selected from the posterior
-distribution.
-
-\textbf{\texttt{supp mat 10 run medusa on mct from 1000 trees.R}}:
-Run MEDUSA on MCC tree from the 1000 random trees selected from the
-posterior distribution.
-
-\textbf{\texttt{supp mat 11 results of MEDUSA run on mcc from 1000 trees.txt}}:
-Summary results from a MEDUSA run on the MCC tree from the 1000 random
-trees selected from the posterior distribution.
-
-\textbf{\texttt{supp mat 12 MEDUSA mcc tree from random sample of 1000 trees.pdf}}:
-Figure for MEDUSA run on MCC tree from random 1000 trees.
-
-\textbf{\texttt{supp mat 13 results of Multi-MEDUSA run on 1000 trees.txt}}:
-Summary results from a MultiMEDUSA run on the 1000 random trees selected
-from the posterior distribution.
-
-\textbf{\texttt{supp mat 14 multimedusa result prob nodes.pdf}}:
-Probability of nodes with estimated rates from a MultiMEDUSA run on the
-1000 random trees selected from the posterior distribution.
-
-\textbf{\texttt{supp mat 15 host plant data.csv}}: Hostplants of
-Nymphalidae butterflies recorded from the literature.
-
-\textbf{\texttt{supp mat 16 host plant data references.csv}}:
-References for hostplants data.
-
-\textbf{\texttt{supp mat 17 hostplant states.csv}}: Data matrix with
-character states for hosplant use.
-
-\textbf{\texttt{supp mat 18 bisse source.R}}: R code for running the
+ \item {\bf Supporting Information S15.csv} Hostplants of Nymphalidae
+butterflies recorded from the literature.
+
+ \item {\bf Supporting Information S16.csv} References for hostplants
+data.
+
+ \item {\bf Supporting Information S17.csv} Data matrix with character
+states for hosplant use.
+
+ \item {\bf Supporting Information S18.R} R script for running the BiSSE
+analysis.
+
+ \item {\bf Supporting Information S19.csv} Raw results for the BiSSE
+analysis.
+
+ \item {\bf Supporting Information S20.csv} Raw results for the combined
 BiSSE analysis.
 
-\textbf{\texttt{supp mat 19 bisse mcmc run.csv}}: Raw results for
-the BiSSE analysis.
-
-\textbf{\texttt{supp mat 20 combined bisse.csv}}: Raw results for
-the combined BiSSE analysis.
-
-\texttt{Figure S1.} Boxplot of speciation (lambda) and extinction (mu)
-values for the 95\% credibility intervals of values estimated by BiSSE
-analysis of diversification due to feeding on Solanaceae plants.
-
-\texttt{Figure S2.} BiSSE analysis of diversification of nymphalids due
+ \item {\bf Figure S1.} Boxplot of speciation ($\lambda$) and extinction
+($\mu$) values for the 95\% credibility intervals of values estimated
+by BiSSE analysis of diversification due to feeding on Solanaceae
+plants.
+
+ \item {\bf Figure S2.} BiSSE analysis of diversification of nymphalids due
 to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
@@
 recovered, speciation and net diversification rates are significantly
-higher for Solanaceae feeders (\$lambda\$1, r1).
-
-\texttt{Figure S3.} Net diversification rates of nymphalids feeding on
+higher for Solanaceae feeders $\lambda1$, r1).
+
+ \item {\bf Figure S3.} Net diversification rates of nymphalids feeding on
 Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
@@
 
-\texttt{Figure S5.} BiSSE analysis of diversification of nymphalids due
+ \item {\bf Figure S5.} BiSSE analysis of diversification of nymphalids due
 to feeding on Apocynaceae hostplants. Speciation and net diversification
 rates are similar.
-
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
-
+\end{description}
+
+\section*{References}
 1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
@@
 
-7. Heikkilä M, Kaila L, Mutanen M, Peña C, Wahlberg N (2012) Cretaceous
+7. Joron M, Jiggins C, Papanicolaou A, McMillan W (2006) Heliconius wing
+patterns: An evo-devo model for understanding phenotypic diversity.
+Heredity 97: 157--167.
+
+8. Willmott KR, Freitas AVL (2006) Higher-level phylogeny of the
+Ithomiinae (Lepidoptera: Nymphalidae): classification, patterns of
+larval hostplant colonization and diversification. Cladistics 22:
+297--368.
+doi:\href{http://dx.doi.org/10.1111/j.1096-0031.2006.00108.x}{10.1111/j.1096-0031.2006.00108.x}.
+
+9. Brakefield PM, Beldade P, Zwaan BJ (2009) The african butterfly
+bicyclus anynana: A model for evolutionary genetics and evolutionary
+developmental biology. Cold Spring Harbor Protocols 2009: pdb--emo122.
+
+10. Heikkil{\"a} M, Kaila L, Mutanen M, Pe{\~n}a C, Wahlberg N (2012) Cretaceous
 origin and repeated tertiary diversification of the redefined
@@
 
-8. Elias M, Joron M, Willmott KR, Silva-Brandão KL, Kaiser V, et al.
+11. Elias M, Joron M, Willmott KR, Silva-Brand{\~a}o KL, Kaiser V, et al.
 (2009) Out of the Andes: Patterns of diversification in clearwing
@@
 
-10. Wahlberg N, Leneveu J, Kodandaramaiah U, Peña C, Nylin S, et al.
+13. Wahlberg N, Leneveu J, Kodandaramaiah U, Pe{\~n}a C, Nylin S, et al.
 (2009) Nymphalid butterflies diversify following near demise at the
@@
 
-11. Peña C, Wahlberg N (2008) Prehistorical climate change increased
+14. Pe{\~n}a C, Wahlberg N (2008) Prehistorical climate change increased
 diversification of a group of butterflies. Biol Lett 4: 274--278.
@@
 
-12. Ferrer-Paris JR, S{á}nchez-Mercado A, Viloria ÁL, Donaldson J (2013)
+15. Ferrer-Paris JR, S{\'a}nchez-Mercado A, Viloria {\'A}L, Donaldson J (2013)
 Congruence and Diversity of Butterfly-Host Plant Associations at Higher
@@
 
-13. Alfaro ME, Santini F, Brock C, Alamillo H, Dornburg A, et al. (2009)
+16. Alfaro ME, Santini F, Brock C, Alamillo H, Dornburg A, et al. (2009)
 Nine exceptional radiations plus high turnover explain species diversity
@@
 
-14. Harmon LJ, Rabosky DL, {FitzJohn} RG, Brown JW (2011) turboMEDUSA:
+17. Harmon LJ, Rabosky DL, {FitzJohn} RG, Brown JW (2011) turboMEDUSA:
 MEDUSA: Modeling Evolutionary Diversification Using Stepwise AIC.
 
-15. {FitzJohn} RG (2012) Diversitree: comparative phylogenetic analyses
+18. {FitzJohn} RG (2012) Diversitree: comparative phylogenetic analyses
 of diversification in R. Method Ecol Evol 3: 1084--1092.
@@
 
-16. Lamas G (2004) Checklist: Part 4A. Hesperioidea-Papilionoidea.
+19. Lamas G (2004) Checklist: Part 4A. Hesperioidea-Papilionoidea.
 Gainesville.
 
-17. Matos-Maraví PF, Peña C, Willmott KR, Freitas AV, Wahlberg N (2013)
+20. Matos-Marav{\'i} PF, Pe{\~n}a C, Willmott KR, Freitas AV, Wahlberg N (2013)
 Systematics and evolutionary history of butterflies in the ``Taygetis
@@
 
-18. Brower AVZ, Wahlberg N, Ogawa JR, Boppré M, Vane-Wright RI (2010)
+21. Brower AVZ, Wahlberg N, Ogawa JR, Boppr{\'e} M, Vane-Wright RI (2010)
 Phylogenetic relationships among genera of danaine butterflies
@@
 
-19. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
+22. Kodandaramaiah U, Lees DC, M{\"u}ller CJ, Torres E, Karanth KP, et al.
 (2010) Phylogenetics and biogeography of a spectacular Old World
@@
 
-20. Kodandaramaiah U, Peña C, Braby MF, Grund R, Müller CJ, et al.
+23. Kodandaramaiah U, Pe{\~n}a C, Braby MF, Grund R, M{\"u}ller CJ, et al.
 (2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
@@
 
-24. Peña C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
+27. Pe{\~n}a C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
 (2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
@@
 
-26. Silva-Brandão KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
+29. Silva-Brand{\~a}o KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
 Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
@@
 
-27. Peña C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
+30. Pe{\~n}a C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
 butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
@@
 
-28. Peña C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
+31. Pe{\~n}a C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
 history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
@@
 
-29. Ackery PR (1988) Hostplants and classification: A review of
+32. Ackery PR (1988) Hostplants and classification: A review of
 nymphalid butterflies. Biol J Linn Soc 33: 95--203.
@@
 
-31. Beccaloni G, Viloria Á, Hall S, Robinson G (2008) Catalogue of the
+34. Beccaloni G, Viloria {\'A}, Hall S, Robinson G (2008) Catalogue of the
 hostplants of the Neotropical butterflies. London: The Natural History
@@
 
-40. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
+42. Rabosky DL (2014) Automatic Detection of Key Innovations, Rate Shifts, and
+Diversity-Dependence on Phylogenetic Trees. PLoS ONE 9: e89543. doi:
+\href{http://dx.doi.org/10.1371/journal.pone.0089543}{10.1371/journal.pone.0089543}.
+
+43. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
 bees cut leaves? New insights into the early evolution of bees. P Roy
@@
 
-41. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
+44. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
 clades of Agaricales. P Roy Soc B 279: 2003--2011.
@@
 
-42. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
+45. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
 Interface. Ann Mo Bot Gard 74: 359--397.
 
-43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brandão KL, Whinnett A, et
+46. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand{\~a}o KL, Whinnett A, et
 al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
@@
 
-44. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
+47. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
 trait-dependent speciation and extinction rates from incompletely
@@
 
-45. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
+48. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
 The genetic basis of a plant-insect coevolutionary key innovation. P
@@
 
-46. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
+49. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
 furanocoumarin-metabolizing cytochrome P450 monooxygenases in two
@@
 
-47. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
+50. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
 the angiosperms re-revisited. Am J Bot 97: 1296--1303.
@@
 
-48. Somoza R (1998) Updated Nazca (Farallon) - South America relative
-motions during the last 40 My: implications for mountain building in the
-central Andean region. J S Am Earth Sci 11: 211--215.
-doi:\href{http://dx.doi.org/10.1016/S0895-9811(98)00012-1}{10.1016/S0895-9811(98)00012-1}.
-
-49. Hoorn C, Wesselingh FP, Steege H ter, Bermudez MA, Mora A, et al.
-(2010) Amazonia through time: Andean uplift, climate change, landscape
-evolution, and biodiversity. Science 330: 927--931.
-doi:\href{http://dx.doi.org/10.1126/science.1194585}{10.1126/science.1194585}.
-
-50. Olmstead RG (2013) Phylogeny and biogeography in Solanaceae,
-Verbenaceae and Bignoniaceae: a comparison of continental and
-intercontinental diversification patterns. Bot J Linn Soc 171: 80--102.
-doi:\href{http://dx.doi.org/10.1111/j.1095-8339.2012.01306.x}{10.1111/j.1095-8339.2012.01306.x}.
-
-51. Wahlberg N, Rubinoff D (2011) Vagility across Vanessa (Lepidoptera:
+51. Wahlberg N, Rubinoff D (2011) Vagility across \emph{Vanessa} (Lepidoptera:
 Nymphalidae): mobility in butterfly species does not inhibit the
@@
 
-52. Müller CJ, Wahlberg N, Beheregaray LB (2010) `After Africa': the
-evolutionary history and systematics of the genus Charaxes Ochsenheimer
-(Lepidoptera: Nymphalidae) in the Indo-Pacific region. Biol J Linn Soc
-100: 457--481.
-doi:\href{http://dx.doi.org/10.1111/j.1095-8312.2010.01426.x}{10.1111/j.1095-8312.2010.01426.x}.
-
-53. Aduse-Poku K, Vingerhoedt E, Wahlberg N (2009) Out-of-Africa again:
-a phylogenetic hypothesis of the genus Charaxes (Lepidoptera:
-Nymphalidae) based on 5 gene regions. Mol Phylogenet Evol 53: 463--478.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.06.021}{10.1016/j.ympev.2009.06.021}.
+% Table 1
+\section*{Tables}
+\begin{table}[!h]
+\caption{\bf{Significant net diversification rate shifts found in the MEDUSA analysis of the Nymphalid maximum clade credibility tree.}}
+\begin{tabular}{lccrrcl}
+Shift N$^\circ$ & Split.Node & Model & r          & LnLik.part & AICc     & Taxa                                                       \\
+1               & 399        & yule  & 0.092459   & -1055.957  &          & Nymphalidae (root)                                         \\
+2               & 691        & bd    & 0.054129   & -406.3703  &          & Limenitidinae + Heliconiinae                               \\
+3               & 299        & yule  & 0.311199   & -6.3058    &          & \emph{Ypthima}                                             \\
+4               & 224        & yule  & 0.290989   & -6.2601    &          & \emph{Charaxes}                                            \\
+5               & 750        & yule  & 0.186913   & -147.4146  &          & Oleriina + Ithomiina + Napeogenina + Dircennina + Godyrina \\
+6               & 405        & yule  & 0.116252   & -555.0276  &          & Satyrini                                                   \\
+7               & 495        & yule  & 0.064656   & -124.9143  &          & Coenonymphina                                              \\
+8               & 609        & yule  & 0.240562   & -35.0908   &          & Phyciodina in part                                         \\
+9               & 787        & bd    & 0.042332   & -43.7819   &          & Danaini in part                                            \\
+10              & 231        & yule  & 0.209416   & -4.7955    &          & \emph{Coenonympha}                                         \\
+11              & 478        & yule  & 0.311684   & -9.2218    &          & \emph{Caeruleuptychia} + \emph{Magneuptychia}              \\
+12              & 659        & yule  & 0.219253   & -11.2686   &          & \emph{Callicore} + \emph{Diaethria}                        \\
+13              & 444        & yule  & 0.220615   & -43.7812   &          & Satyrina                                                   \\
+14              & 524        & yule  & 0.190754   & -26.0651   &          & Mycalesina                                                 \\
+15              & 355        & yule  & 0.234041   & -6.2013    &          & \emph{Pedaliodes}                                          \\
+16              & 714        & yule  & 0          & 0          &          & \emph{Dryas} + \emph{Dryadula}                             \\
+17              & 377        & yule  & 0.311671   & -4.1986    &          & \emph{Taenaris}                                            \\
+18              & 688        & yule  & 0.024724   & -17.5256   &          & Pseudergolinae                                             \\
+19              & 583        & yule  & 0          & 0          & 5090.492 & \emph{Anaeomorpha} + \emph{Hypna}
+\end{tabular}
+\begin{flushleft}Split.Node=node number, Model=preferred diversification model by MEDUSA, r=net diversification rate, LnLik.part=log likelihood value.
+\end{flushleft}
+\end{table}
+
+% Table 2
+\begin{table}[!h]
+    \caption{\bf{Differences in rates estimated by MEDUSA on the MCC
+tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
+from the posterior distribution.}}
+\begin{tabular}{lcrr}
+Split.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
+1          & 0.090          & not found                  & 0.000                          \\
+2          & 0.054          & -0.030                     & 0.864                          \\
+3          & 0.311          & 0.219                      & 0.961                          \\
+4          & 0.290          & 0.212                      & 0.996                          \\
+5          & 0.187          & 0.101                      & 0.927                          \\
+6          & 0.116          & 0.039                      & 0.131                          \\
+7          & 0.064          & -0.052                     & 0.195                          \\
+8          & 0.240          & 0.166                      & 0.319                          \\
+9          & 0.042          & -0.049                     & 0.897                          \\
+10         & 0.209          & 0.149                      & 0.619                          \\
+11         & 0.311          & 0.208                      & 0.831                          \\
+12         & 0.219          & 0.135                      & 0.911                          \\
+13         & 0.220          & 0.117                      & 0.276                          \\
+14         & 0.190          & 0.115                      & 0.379                          \\
+15         & 0.234          & 0.127                      & 0.825                          \\
+16         & 0.000          & -0.005                     & 0.659                          \\
+17         & 0.311          & 0.248                      & 0.785                          \\
+18         & 0.024          & -0.064                     & 0.024                          \\
+19         & 0.000          & -0.087                     & 0.381
+\end{tabular}
+\end{table}
+
+% Table 3
+\begin{table}[!h]
+    \caption{\bf{Likelihood ratio test between the model of increased
+diversification of nymphalids feeding on Solanaceae against a model
+forcing equal speciation rates (no effect on diversification).}}
+\begin{tabular}{lcrrlr}
+Df           & lnLik & AIC     & ChiSq  & \multicolumn{1}{r}{p} &         \\
+full         & 6     & -1613.3 & 3238.5 &                       &         \\
+equal.lambda & 5     & -1619.4 & 3248.9 & 12.3                  & 0.00045
+\end{tabular}
+\end{table}
 

### 2015-01-12 response_to_reviewers.tex
comment 15

@@
 
+\textbf{15. Reviewer's Comment:}
+Page 6, sentence beginning “MultiMEDUSA provided mean and standard deviation...”: This sentence states that “some of the changes...had great variation” - the changes themselves did not have variation, rather, the rates showed considerable variation.
+Revise sentence to better identify the object that varied.
+
+\textbf{15. Response:}
+
+\begin{quote}
+\color{blue}
+We modified the sentence to the following:
+
+"MultiMEDUSA provided mean and standard deviation statistics for the
+diversification values on the shifts on the 1000 trees (supp. mat.
+13--14), and found that the estimated net diversification rate
+values had great variation across the posterior distribution of trees."
+\end{quote}
 \end{document}

### 2015-01-12 response_to_reviewers.tex
comment 16

@@
 \end{quote}
+
+\textbf{16. Reviewer's Comment:}
+Page 7, final sentence of results: The test of Apocynaceae feeding is not presented in the Methods section.
+
+\textbf{16. Response:}
+
+\begin{quote}
+\color{blue}
+We added the following sentence to the Methods section:
+
+We also run a BiSSE analysis
+to test whether the trait ``feeding on Apocynaceae'' had any effect on
+net diversification rates found for Nymphalidae lineages feeding on
+Apocynaceae.
+\end{quote}
 \end{document}

### 2015-01-12 MS_revised.tex
comment 15
comment 16

@@
 (\texttt{sampling.f}) in \texttt{diversitree}. 
+We also run a BiSSE analysis
+to test whether the trait ``feeding on Apocynaceae'' had any effect on
+net diversification rates found for Nymphalidae lineages feeding on
+Apocynaceae.
 We also used likelihood ratio tests to evaluate the influence of
@@
 diversification values on the shifts on the 1000 trees (supp. mat.
-13--14), and found that some of the changes in net diversification rate
+13--14), and found that the estimated net diversification rate
 values had great variation across the posterior distribution of trees. A

### 2015-01-12 MS_revised.tex
comment 17

@@
 overall model was set to 7.8 units, as prescribed by MEDUSA. In all
-MEDUSA analyses, the maximum number of inferred diversification splits
+MEDUSA analyses, the maximum number of inferred diversification shifts
 in all trees was 26. The background net diversification rate for
@@
 We also compared the results from MultiMEDUSA (derived from the sample
-of 1000 trees) with the splits found by MEDUSA on the MCC tree derived
+of 1000 trees) with the shifts found by MEDUSA on the MCC tree derived
 from this random sample. In the summary statistics, MultiMEDUSA reports
@@
 the MCC tree are consistently recovered in most of the 1000 trees. Some
-of the splits in the MCC tree are recovered in very few trees, for
+of the shifts in the MCC tree are recovered in very few trees, for
 example the split for the clade Satyrinae is recovered with a
@@
 random sample of trees from the posterior distribution of a Bayesian
-run. We found that some diversification splits, estimated on the
+run. We found that some diversification shifts, estimated on the
 Nymphalidae maximum clade credibility tree, were found with very low
@@
 distribution (Fig. 3, Table 2). We also found that, even though the
-analyses estimated the same diversification splits on two or more trees,
+analyses estimated the same diversification shifts on two or more trees,
 the estimated net diversification rates could vary widely (Fig. 2). For
@@
 In this study, the effect of phylogenetic uncertainty on the inferred
-diversification splits by MEDUSA is amplified because some Nymphalidae
+diversification shifts by MEDUSA is amplified because some Nymphalidae
 taxa appear to be strongly affected by long-branch attraction artifacts
@@
 If there is strong phylogenetic signal for increases or decreases in net
-diversification rates for a node, it is expected that these splits would
+diversification rates for a node, it is expected that these shifts would
 be inferred by MEDUSA in most of the posterior distribution of trees.
@@
 diversification shift (due to a non-existent node). This is the reason
-why MEDUSA estimated net diversification rate splits with a probability
-higher than 0.90 in the sample of trees for only four splits: the genus
+why MEDUSA estimated net diversification rate shifts with a probability
+higher than 0.90 in the sample of trees for only four shifts: the genus
 \emph{Charaxes}, the genus \emph{Ypthima}, part of Ithomiini and the
 clade \emph{Callicore} + \emph{Diaethria} (Fig. 3), while estimating
-splits for other lineages with much lower probability.
+shifts for other lineages with much lower probability.
 
@@
 shifts in the maximum clade credibility tree of Nymphalidae, only a few
-of these splits were found in more than 90\% of the trees from the
+of these shifts were found in more than 90\% of the trees from the
 posterior distribution. In the literature, it is common practice that
-conclusions are based on the splits estimated on the maximum clade
+conclusions are based on the shifts estimated on the maximum clade
 credibility tree. However, by using a MultiMEDUSA approach, we found
-that for this Nymphalidae dataset some of these splits might be greatly
-affected by phylogenetic uncertainty. Moreover, some of these splits can
+that for this Nymphalidae dataset some of these shifts might be greatly
+affected by phylogenetic uncertainty. Moreover, some of these shifts can
 be recovered either as increases or decreases in net diversification
@@
 with no conflicting topology, and very similar estimated ages for nodes
-in order to consistently recover most of the diversification splits on
+in order to consistently recover most of the diversification shifts on
 the posterior distribution of trees that were inferred by MEDUSA on the
@@
 subtribe Danaina. Due to phylogenetic uncertainty, we did not obtain
-strong support for other diversification splits in Nymphalidae. Our
+strong support for other diversification shifts in Nymphalidae. Our
 BiSSE analysis corroborated other studies in that the trait ``feeding on

### 2015-01-12 response_to_reviewers.tex
comment 17

@@
 \end{quote}
+
+\textbf{17. Reviewer's Comment:}
+Page 7, first and third paragraphs of Discussion section (and potentially elsewhere): there are cases where the term “diversification splits” is used, where “diversification shifts” is more appropriate.
+
+\textbf{17. Response:}
+
+\begin{quote}
+\color{blue}
+We changed ``splits'' to ``shifts'' throughout the text.
+\end{quote}
+
+\textbf{18. Reviewer's Comment:}
+Page 7, sentence beginning “For example, in our Nymphalidae trees...”: the authors state “we found that the split...had a variation of r = 0.5234”. Should “variation” instead be “rate”?
+
+\textbf{18. Response:}
+
+\begin{quote}
+\color{blue}
+\end{quote}
+
 \end{document}

### 2015-01-12 MS_revised.tex
comment 17

@@
 of the shifts in the MCC tree are recovered in very few trees, for
-example the split for the clade Satyrinae is recovered with a
+example the shift for the clade Satyrinae is recovered with a
 probability of 0.18 (Fig. 3).
@@
 the estimated net diversification rates could vary widely (Fig. 2). For
-example, in our Nymphalidae trees, we found that the split for
+example, in our Nymphalidae trees, we found that the shift for
 \emph{Magneuptychia} and \emph{Caeruleuptychia} had a variation from
@@
 \begin{tabular}{lccrrcl}
-Shift N$^\circ$ & Split.Node & Model & r          & LnLik.part & AICc     & Taxa                                                       \\
+Shift N$^\circ$ & Shift.Node & Model & r          & LnLik.part & AICc     & Taxa                                                       \\
 1               & 399        & yule  & 0.092459   & -1055.957  &          & Nymphalidae (root)                                         \\
@@
 \end{tabular}
-\begin{flushleft}Split.Node=node number, Model=preferred diversification model by MEDUSA, r=net diversification rate, LnLik.part=log likelihood value.
+\begin{flushleft}Shift.Node=node number, Model=preferred diversification model by MEDUSA, r=net diversification rate, LnLik.part=log likelihood value.
 \end{flushleft}
@@
 \begin{tabular}{lcrr}
-Split.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
+Shift.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
 1          & 0.092          & not found                  & 0.000                          \\

### 2015-01-12 response_to_reviewers.tex
comment 18
comment 19

@@
 \color{blue}
+We corrected the word from ``variation'' to ``rate''.
 \end{quote}
 
+\textbf{19. Reviewer's Comment:}
+Page 7, third paragraph of Discussion: Replace “This is the reason why” with
+“The relatively low phylogenetic support for many nodes in the Nymphalidae tree
+is likely the reason why”.
+
+\textbf{19. Response:}
+
+\begin{quote}
+\color{blue}
+We modified the sentence according to the reviewer's suggestion.
+\end{quote}
 \end{document}

### 2015-01-12 MS_revised.tex
comment 18
comment 19

@@
 example, in our Nymphalidae trees, we found that the shift for
-\emph{Magneuptychia} and \emph{Caeruleuptychia} had a variation from
+\emph{Magneuptychia} and \emph{Caeruleuptychia} had a rate from
 \texttt{r = 0.5234}, higher than the background net diversification
@@
 to be absent in some trees and MEDUSA will be unable to estimate any
-diversification shift (due to a non-existent node). This is the reason
+diversification shift (due to a non-existent node). 
+The relatively low phylogenetic support for many nodes in the Nymphalidae
+tree is likelyl the reason
 why MEDUSA estimated net diversification rate shifts with a probability

### 2015-01-14 response_to_reviewers.tex
comment 20
comment 21

@@
 \end{quote}
+
+\textbf{20. Reviewer's Comment:}
+Page 7, fourth paragraph of Discussion: Omit this paragraph.
+
+\textbf{20. Response:}
+
+\begin{quote}
+\color{blue}
+We removed the paragraph according to the reviewer's suggestion.
+\end{quote}
+
+\textbf{21. Reviewer's Comment:}
+Pages 7-8, fourth paragraph of Discussion: Per previous comment of BAMM analysis, this paragraph should probably be removed.
+If it is not, both shifts detected by BAMM analysis (not just one) should be described and compared to the four identified in MultiMEDUSA (are they at the same nodes or different nodes?).
+
+\textbf{21. Response:}
+
+\begin{quote}
+\color{blue}
+We removed the paragraph according to the reviewer's suggestion.
+\end{quote}
+
 \end{document}

### 2015-01-14 MS_revised.tex
comment 20
comment 21
comment 22

@@
 
-The clade Ithomiini and the non-basal danaids are well supported by high
-posterior probability values in Wahlberg et al. (2009) \cite{wahlberg2009}.
-Therefore our MEDUSA analyses recovered an increase in net
-diversification rate with probability higher than 0.90 in the posterior
-distribution of trees (Fig. 3).
-
 \subsection*{Hostplant use and diversification in
@@
 
-The increase in diversification inferred by MEDUSA occurred after the
+The increase in diversification rate inferred by MEDUSA occurred after the
 probable shift from Apocynaceae to Solanaceae, as the Solanaceae feeders

### 2015-01-14 response_to_reviewers.tex
comment 22
comment 23

@@
 
+\textbf{22. Reviewer's Comment:}
+Page 8, Hostplant use and diversification... section, Ithomiini subsection,
+fifth paragraph, first sentence: insert 'rate': “The increase in diversification
+rate inferred by MEDUSA...”
+
+\textbf{22. Response:}
+
+\begin{quote}
+\color{blue}
+We included the word ``rate'' according to the reviewer's suggestion.
+\end{quote}
+
+\textbf{23. Reviewer's Comment:}
+Page 8, Hostplant use and diversification... section, Ithomiini subsection,
+sixth paragraph, second sentence: replace “might not be too efficient” with 
+“may be inefficient”.
+
+\textbf{23. Response:}
+
+\begin{quote}
+\color{blue}
+We reworded the sentence according to the reviewer's suggestion.
+\end{quote}
+
 \end{document}

### 2015-01-14 MS_revised.tex
comment 23

@@
 several lineages of Ithomiini. We can speculate that older lineages
-exploiting a novel toxic resource \cite{willmott2006, wahlberg2009} might not be too efficient
+exploiting a novel toxic resource \cite{willmott2006, wahlberg2009} 
+may be inefficient
 in metabolizing plant toxins and that younger lineages are able to deal

### 2015-01-14 MS_revised.tex
comment 24

@@
 
-\subsection*{Hostplant use and diversification in
-Nymphalidae}
+\subsection*{Hostplant use and diversification in Nymphalidae}
 
@@
 configurations of the proteins, providing more opportunities for
-hostplant switch \cite{li2003}. This might be the reason why the basal
+hostplant switches \cite{li2003}. This might be the reason why the basal
 Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and

### 2015-01-14 response_to_reviewers.tex
comment 24
comment 25

@@
 
+\textbf{24. Reviewer's Comment:}
+Page 8, Hostplant use and diversification... section, Ithomiini subsection,
+sixth paragraph, third sentence: make final word plural: “switches”.
+
+\textbf{24. Response:}
+
+\begin{quote}
+\color{blue}
+We replaced the word ``switch'' according to the reviewer's suggestion.
+\end{quote}
+
+\textbf{25. Reviewer's Comment:}
+Page 9, first full paragraph, first sentence: replace “hypotheses state” with
+“hypotheses posit”.
+
+\textbf{25. Response:}
+
+\begin{quote}
+\color{blue}
+We replaced the word ``state'' according to the reviewer's suggestion.
+\end{quote}
+
 \end{document}

### 2015-01-14 MS_revised.tex
comment 25

@@
 insects and their hostplants, while the ``resource abundance-dependent
-diversity'' and the ``escape-and-radiate'' hypotheses state that insects
+diversity'' and the ``escape-and-radiate'' hypotheses posit that insects
 diversify after their hostplants \cite{nyman2012, ehrlich1964, janz2011}.

### 2015-01-14 MS_revised.tex
comment 26

@@
 Many of the Danaina are large, strong fliers, highly migratory and
-involved in mimicry rings. Among them is for example the monarch
-(\emph{Danaus plexippus}), probably the most well known of all migratory
-butterflies. The causes for a lower net diversification rate in the
+involved in mimicry rings, including the best-known migratory butterfly,
+the monarch (\emph{Danaus plexippus}).
+The causes for a lower net diversification rate in the
 Danaina remains to be investigated, but their great dispersal power

### 2015-01-14 response_to_reviewers.tex
comment 26
comment 27

@@
 
+\textbf{26. Reviewer's Comment:}
+Page 9, Danaini subsection: Remove the sentence beginning “Among them is for
+example...” and just mention the Monarch at the end of the preceding sentence:
+“...involved in mimicry rings, including the best-known migratory butterfly,
+the monarch (Danaus plexippus).”
+
+\textbf{26. Response:}
+
+\begin{quote}
+\color{blue}
+We reworded the sentence according to the reviewer's suggestion.
+\end{quote}
+
+\textbf{27. Reviewer's Comment:}
+Page 9, Danaini subsection: Revise or remove the last sentence of this 
+paragraph. It is not immediately clear what the age of a species has to do 
+with genetic diversity and speciation (or even what the age of a species 
+really is...).
+
+\textbf{27. Response:}
+
+\begin{quote}
+\color{blue}
+We reworded the sentence according to the reviewer's suggestion:
+
+``It has been found
+in highly vagile species in the nymphalid genus \emph{Vanessa} that
+dispersal has homogenized populations due to gene flow, as 
+vagile species seem to be genetically homogeneous in their
+populations Wahlberg N, Rubinoff D (2011).''
+\end{quote}
+
 \end{document}

### 2015-01-14 MS_revised.tex
comment 27

@@
 in highly vagile species in the nymphalid genus \emph{Vanessa} that
-dispersal has homogenized populations due to gene flow, as old and
-vagile species seem to be genetically homogeneous while younger
-widespread species show higher genetic differentiation in their
+dispersal has homogenized populations due to gene flow, as 
+vagile species seem to be genetically homogeneous in their
 populations \cite{wahlberg2011}.

### 2015-01-14 MS_revised.tex
comment 28

@@
 very robust \cite{wahlberg2009} and MEDUSA failed to identify any significant
-accelerated net diversification rate for Satyrini. It appears that the
+accelerated net diversification rate for Satyrini.
+It appears that the
 radiation of Satyrini as a whole was not remarkably fast and therefore
-not picked up by MEDUSA, although it estimated a diversification shift
+not detected by MEDUSA, although it estimated a diversification shift
 for Satyrini + its sister clade. This should be expected if the

### 2015-01-14 response_to_reviewers.tex
comment 28
comment 29

@@
 
+\textbf{28. Reviewer's Comment:}
+Page 9, Satyrinae subsection, sentence beginning “It appears that the 
+radiation...”: replace “picked up” with “detected”.
+
+\textbf{28. Response:}
+
+\begin{quote}
+\color{blue}
+We replaced the words according to the reviewer's suggestion:
+\end{quote}
+
+\textbf{29. Reviewer's Comment:}
+Page 9, Conclusions section, sentence beginning “Moreover, some of these
+splits...”: replace “splits” with “shifts”.
+
+\textbf{29. Response:}
+\begin{quote}
+\color{blue}
+We replaced the word according to the reviewer's suggestion:
+\end{quote}
+
 \end{document}

### 2015-01-14 MS_revised.tex
comment 30

@@
 for analysis. This means that contradictory conclusions would be made if
-only the maximum clade credibility tree was used for analysis. We
-recommend that all datasets should be analyzed using both approaches,
-MEDUSA and MultiMEDUSA, in order to test whether the results are robust
+only the maximum clade credibility tree was used for analysis.
+We recommend that all datasets should be analyzed using the MultiMEDUSA
+approach in order to test whether the results are robust
 when phylogenetic uncertainty is taken into account.

### 2015-01-14 response_to_reviewers.tex
comment 30
comment 31

@@
 
+\textbf{30. Reviewer's Comment:}
+Page 10, sentence beginning “We recommend that all datasets...”: why are the
+authors recommending using MEDUSA at all, as opposed to MultiMEDUSA?
+
+\textbf{30. Response:}
+\begin{quote}
+\color{blue}
+We reworded the sentence and now we recommend using MultiMEDUSA.
+\end{quote}
+
+\textbf{31. Reviewer's Comment:}
+Page 10, second full paragraph: the results described here (rate increase in 
+Ithomiini \& Charaxes, rate decrease in Danaina) are incongruent with those
+described on page six (four rate increases). This incongruence needs to be 
+addressed.
+
+\textbf{31. Response:}
+\begin{quote}
+\color{blue}
+We revised the paragraph and included the two other increases
+in diversification rate: \emph{Ypthima} and the clade \emph{Callicore} + 
+\emph{Diaethria}
+\end{quote}
+
 \end{document}

### 2015-01-14 MS_revised.tex
comment 31

@@
 distribution of trees found strong support for an increase in net
-diversification rate in the tribe Ithomiini and the genus
-\emph{Charaxes}, and for a decrease in net diversification rate in the
+diversification rate in the tribe Ithomiini, the genera \emph{Ypthima},
+\emph{Charaxes}, the clade \emph{Callicore} + \emph{Diaethria} and for a
+decrease in net diversification rate in the
 subtribe Danaina. Due to phylogenetic uncertainty, we did not obtain
@@
 commenting on the manuscript and giving us the posterior distribution of
-trees, Luke Harmon for commenting on the manuscript and anonymous
+trees, Luke Harmon for commenting on the manuscript, Jeffrey C. Oliver and anonymous
 reviewers for their comments, which greatly improved the manuscript,

### 2015-01-14 response_to_reviewers.tex
comment 32
comment 33

@@
 
+\textbf{32. Reviewer's Comment:}
+Page 10, second full paragraph, second sentence: replace “splits” with “shifts”.
+
+\textbf{32. Response:}
+\begin{quote}
+\color{blue}
+We replaced the word according to the reviewer's suggestion:
+\end{quote}
+
+\textbf{33. Reviewer's Comment:}
+Page 10, second full paragraph, third sentence: 2 changes; (A) replace “the
+trait 'feeding on Solanaceae'” to “Solanaceae-feeding”; (B) add qualifier
+“likely”: “was likely important in the diversification...”
+
+\textbf{33. Response:}
+\begin{quote}
+\color{blue}
+We revised the paragraph according to the reviewer's suggestions.
+\end{quote}
+
 \end{document}

### 2015-01-14 MS_revised.tex
comment 33

@@
 strong support for other diversification shifts in Nymphalidae. Our
-BiSSE analysis corroborated other studies in that the trait ``feeding on
-Solanaceae'', or a codistributed character, was important in the
+BiSSE analysis corroborated other studies in that Solanaceae-feeding,
+or a codistributed character, was likely important in the
 diversification of Ithomiini butterflies. However, by applying MEDUSA we

### 2015-01-14 response_to_reviewers.tex
comment 34
comment 35

@@
 
+\textbf{34. Reviewer's Comment:}
+Page 10, second full paragraph, fifth sentence: replace “the trait 'feeding on
+Apocynaceae'” with “Apocynaceae-feeding”.
+
+\textbf{34. Response:}
+\begin{quote}
+\color{blue}
+We revised the paragraph according to the reviewer's suggestion.
+\end{quote}
+
+\textbf{35. Reviewer's Comment:}
+Page 10, Figure legends, figure 1: Include a description of what nodes 1 and
+19 are as well.
+
+\textbf{35. Response:}
+\begin{quote}
+\color{blue}
+We added the description of nodes 1 and 19 to the legend as suggested by the
+reviewer.
+\end{quote}
+
 \end{document}

### 2015-01-14 MS_revised.tex
comment 34
comment 35

@@
 appeared after the shift from Apocynaceae to Solanaceae. We also found
-that the trait ``feeding on Apocynaceae'' is not responsible for the
+that Apocynaceae-feeding is not responsible for the
 slowdown of diversification in Danaina. Ithomiini and Solanaceae
@@
 clade credibility tree from Wahlberg et al. (2009)} Rate shifts were estimated for the
-following nodes (besides the background rate): 2) Limenitidinae +
+following nodes (besides the background rate): 1) root, 2) Limenitidinae +
 Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
@@
 Satyrina, 14) Mycalesina, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
-\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae. Circles on
+\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae, 19)
+\emph{Anaeomorpha} + \emph{Hypna}. Circles on
 nodes indicate the diversification shift number as found by MEDUSA.

### 2015-01-14 response_to_reviewers.tex
comment 36
comment 37

@@
 
+\textbf{36. Reviewer's Comment:}
+Page 10, Figure legends, figure 2: are these rates estimated in MEDUSA or MultiMEDUSA analyses?
+
+\textbf{36. Response:}
+\begin{quote}
+\color{blue}
+We corrected the legend as the rates were estimated by MultiMEDUSA and not by
+MEDUSA.
+\end{quote}
+
+\textbf{37. Reviewer's Comment:}
+Page 11, Figure legends, figure 4: the legend is missing a closing parenthesis,
+“)”.
+\textbf{37. Response:}
+\begin{quote}
+\color{blue}
+We added the required closing parenthesis.
+\end{quote}
+
 \end{document}

### 2015-01-14 MS_revised.tex
comment 36
comment 37

@@
 
-{\bf Figure 2. Diversification rates for taxa estimated by MEDUSA on
+{\bf Figure 2. Diversification rates for taxa estimated by MultiMEDUSA on
 the samples of 1000 random trees}
@@
 rates are significantly higher in Solanaceae feeders (speciation rate =
-$\lambda1$, net diversification rate = $r1$.
+$\lambda1$, net diversification rate = $r1$).
 

### 2015-01-14 MS_revised_R1.tex
to do diff.tex

--- 2015-01-12 out.tex
@@
-% !TeX encoding: utf-8
 \documentclass[10pt]{article}
@@
 
-% Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2009}
-
 % Remove brackets from numbering in List of References
@@
 
-% Include all macros below. Please limit the use of macros.
-% END MACROS SECTION
+%% Include all macros below. Please limit the use of macros.
+
+%% END MACROS SECTION
+
 \begin{document}
+
 \begin{flushleft}
@@
 ``escape-and-radiate'' hypothesis {[}3{]}, the ``oscillation
-hypothesis'' {[}4,5{]} or ``diffuse cospeciation'' {[}2{]}.
+hypothesis'' {[}4,5{]} or ``diffuse cospeciation'' {[}2{]} and the
+``musical chairs hypothesis'' {[}6{]}.
 
@@
 developing some of the mentioned hypotheses. Nymphalidae contains around
-6000 species {[}6{]}, and several members are considered model organisms
-in evolutionary biology {[}7--9{]}. The family most likely originated
+6000 species {[}7{]}, and several members are considered model organisms
+in evolutionary biology {[}8--10{]}. The family most likely originated
 around 94 MYA in the mid Cretaceous. Diversification of the group began
@@
 appeared shortly after the Cretaceous-Paleogene (K-Pg) boundary
-{[}10{]}. Several studies have used time-calibrated phylogenies and
+{[}11{]}. Several studies have used time-calibrated phylogenies and
 diversification models to reconstruct the evolutionary history of the
 group to identify patterns of accelerated or decelerated diversification
-of some Nymphalidae clades {[}10--13{]}. For example, it has been
+of some Nymphalidae clades {[}11--14{]}. For example, it has been
 suggested that climate change in the Oligocene and the subsequent
 diversification of grasses has led to diversification of the subfamily
-Satyrinae {[}14{]} due to the abundance of grasses over extensive
+Satyrinae {[}15{]} due to the abundance of grasses over extensive
 geographic areas (``resource abundance-dependent diversity dynamics''
-hypothesis). Fordyce (2010) {[}12{]} found increased net diversification
+hypothesis). Fordyce (2010) {[}13{]} found increased net diversification
 rates in some Nymphalidae lineages after a major hostplant shift, which
@@
 diversification pattern of Nymphalidae is congruent with events of
-sudden diversification bursts due to hostplant shift, climatic events or
-shifts to closely related hostplants {[}5,15{]}.
+sudden diversification bursts due to hostplant shift or climatic events
+{[}5,16{]}.
 
 In this study, we used a time-calibrated genus-level phylogenetic
-hypothesis for Nymphalidae butterflies {[}13{]} to investigate patterns
-of diversification. We applied MEDUSA {[}16,17{]}, a recently developed
-statistical method, to study the diversification pattern of Nymphalidae
+hypothesis for Nymphalidae butterflies {[}14{]} to investigate patterns
+of diversification. We applied the statistical method MEDUSA
+{[}17,18{]}, to study the diversification pattern of Nymphalidae
 butterflies. MEDUSA fits likelihood models of diversification into a
@@
 study the effects of phylogenetic uncertainty and by using the extended
-MEDUSA method called MultiMEDUSA {[}16{]}. We also tested whether
+MEDUSA method called MultiMEDUSA {[}17{]}. We also tested whether
 hostplant association dynamics can explain the diversification patterns
@@
 employing the method BiSSE as implemented in the R package
-\texttt{diversitree} {[}18{]}.
+\texttt{diversitree} {[}19{]}.
 
@@
 For analyses, we used the phylogenetic trees from the study of Wahlberg
-et al. (2009) {[}13{]} that were generated using DNA sequence data from
+et al. (2009) {[}14{]} that were generated using DNA sequence data from
 10 gene regions for 398 of the 540 valid genera in Nymphalidae. We
@@
 as a random sample of 1000 trees from their BEAST run after burnin
-{[}13{]}. Their original BEAST run was for 40 million generations. We
+{[}14{]}. Their original BEAST run was for 40 million generations. We
 used a burnin of 25 million generations and took a random sample of 1000
@@
 sources including the specialist-curated lists on
-\url{http://tolweb.org}, Lamas (2004) {[}19{]} and curated lists of
+\url{http://tolweb.org}, Lamas (2004) {[}20{]} and curated lists of
 Global Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}).
@@
 to the closest related genus that was included in Wahlberg et al. (2009)
-{[}13{]} study according to available phylogenetic studies {[}20--31{]},
+{[}14{]} study according to available phylogenetic studies {[}21--32{]},
 taxonomical classification and morphological resemblance when no
@@
 Hostplant data for Nymphalidae species were compiled from several
-sources including {[}32{]}, HOSTS database (\url{http://bit.ly/YI7nwW}),
-{[}33--35{]} and others (supp. mat. 15--16) for a total of 6586
+sources including {[}33{]}, HOSTS database (\url{http://bit.ly/YI7nwW}),
+{[}34--36{]} and others (supp. mat. 15--16) for a total of 6586
 hostplant records, including 428 Nymphalidae genera and 143 plant
@@
 
-We used the statistical software R version 3.0.1 {[}36{]} in combination
-with the APE {[}37{]}, GEIGER {[}38{]}, MEDUSA {[}16{]} and
-\texttt{diversitree} {[}18{]} packages along with our own scripts to
+We used the statistical software R version 3.0.1 {[}37{]} in combination
+with the APE {[}38{]}, GEIGER {[}39{]}, MEDUSA {[}17{]} and
+\texttt{diversitree} {[}19{]} packages along with our own scripts to
 perform the analyses (included as supplementary materials 2, 6, 10). All
 analyses were run on the 1000 random trees from Wahlberg et al. (2009)
-{[}13{]}, on the maximum clade credibility tree derived from these and
-the MCC tree {[}13{]}.
-
-\subsection*{Detecting diversification shifts on phylogenetic
-trees}
+{[}14{]}, on the maximum clade credibility tree derived from these and
+the MCC tree {[}14{]}.
+
+We analyzed our MCC tree in the Bayesian software BAMM in order to
+contrast the results found by MEDUSA and MultiMEDUSA. BAMM (Bayesian
+analysis of macroevolutionary mixtures) {[}40{]} is a software that fits
+several models of speciation and extinction. The best model can be
+selected by comparison of Bayes Factors.
+
+\subsection*{Detecting diversification shifts on phylogenetic trees}
 
 Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
-version 093.4.33 {[}16{]} on the maximum clade credibility tree from
-Wahlberg et al. (2009) {[}13{]}. MEDUSA fits alternative birth-death
+version 093.4.33 {[}17{]} on the maximum clade credibility tree from
+Wahlberg et al. (2009) {[}14{]}. MEDUSA fits alternative birth-death
 likelihood models to a phylogenetic tree in order to estimate changes in
@@
 the rates \texttt{r}: net diversification \texttt{=} speciation
-(\texttt{b}) \texttt{-} extinction (\texttt{d}) and ($\varepsilon$):
+(\texttt{b}) \texttt{-} extinction (\texttt{d}) and (\(\varepsilon\):
 relative extinction \texttt{= d/b}. The AIC scores of the two-parameter
@@
 value. MEDUSA finds the likelihood of the models after taking into
-account branch lengths and number of species per lineage {[}16{]}. Most
+account branch lengths and number of species per lineage {[}17{]}. Most
 studies using MEDUSA only run the method on a single tree, usually the
@@
 therefore selected 1000 random genus-level trees from the posterior
-distribution of the Bayesian run from Wahlberg et al., 2009 {[}13{]}
+distribution of the Bayesian run from Wahlberg et al., 2009 {[}14{]}
 (MultiMEDUSA, supp. mat. 5). We calculated a new MCC tree from the
@@
 net diversification (\texttt{r = b - d}) and relative extinction rate
-($\varepsilon$) \texttt{= d/b}) values across the 1000 trees for the
+(\(\varepsilon\) \texttt{= d/b}) values across the 1000 trees for the
 nodes where changes in diversification tempo occurs. We used the AICc
@@
 
-\subsection*{Estimation of trait-dependent speciation rates}
+\subsection*{Estimation of trait-dependent speciation
+rates}
 
@@
 and performed analyses using the ``binary state speciation and
-extinction'' {[}39{]} Bayesian approach as implemented in the R package
-\texttt{diversitree} {[}18{]}. The Markov Chain Monte Carlo algorithm
+extinction'' {[}41{]} Bayesian approach as implemented in the R package
+\texttt{diversitree} {[}19{]}. The Markov Chain Monte Carlo algorithm
 was run for 10000 generations discarding the first 7500 as burnin. The
-Multiple State Speciation and Extinction approach (MuSSE) {[}18{]} is
+Multiple State Speciation and Extinction approach (MuSSE) {[}19{]} is
 designed to examine the joint effects of two or more traits on
@@
 binary character state has had any effect on increased net
-diversification rate for a clade {[}39{]}. We used our compiled data of
+diversification rate for a clade {[}41{]}. We used our compiled data of
 hostplant use to produce a binary dataset for the character ``feeding on
@@
 1) (supp. mat. 17). The family Solanaceae include the hostplants of the
-diverse Ithomiini butterflies and closest relatives {[}8{]}. Other
+diverse Ithomiini butterflies and closest relatives {[}9{]}. Other
 hostplant shifts were not tested for effect on net diversification rates
@@
 distribution. The records of \emph{Vanessa} and \emph{Hypanartia}
-feeding on Solanaceae {[}33,40{]} might be incorrect as it is unlikely
+feeding on Solanaceae {[}35,42{]} might be incorrect as it is unlikely
 that these species can be feeding on this plant family. We tested
@@
 or presence would not have a significant effect on our results. It has
-been shown that BiSSE performs poorly under certain conditions {[}41{]}.
+been shown that BiSSE performs poorly under certain conditions {[}43{]}.
 However, our data has adequate number of taxa under analysis (more than
@@
 of Solanaceae hostplants. Thus, BiSSE is expected to produce robust
-results {[}41{]}.
+results {[}43{]}.
 
@@
 The MEDUSA analysis on the MCC tree in combination with richness data
-estimated 18 changes in the tempo of diversification in Nymphalidae
-history (Fig. 1; Table 1). The corrected AICc acceptance threshold for
-adding subsequent piecewise birth-death processes to the overall model
-was set to 7.8 units, as prescribed by MEDUSA. In all MEDUSA analyses,
-the maximum number of inferred diversification splits in all trees was
-26. The background net diversification rate for Nymphalidae was
-estimated as \texttt{r = 0.092} lineages per Million of years and the
-AICc score for the best fit model was \texttt{5090.5} (Table 1). MEDUSA
-also estimated that the basic constant birth-death model was not a
-better explanation for our data (AICc \texttt{= 5449.3}).
+estimated 18 significant changes in the tempo of diversification in
+Nymphalidae history (Fig. 1; Table 1). The corrected AICc acceptance
+threshold for adding subsequent piecewise birth-death processes to the
+overall model was set to 7.8 units, as prescribed by MEDUSA. In all
+MEDUSA analyses, the maximum number of inferred diversification splits
+in all trees was 26. The background net diversification rate for
+Nymphalidae was estimated as \texttt{r = 0.092} lineages per Million of
+years and the AICc score for the best fit model was \texttt{5090.5}
+(Table 1). MEDUSA also estimated that the basic constant birth-death
+model was not a better explanation for our data (AICc
+\texttt{= 5449.3}).
 
@@
 
-We contrasted these results with those obtained from analysing the data in
-the software BAMM {[}42{]}. We found
-that the best model (best configuration) estimated 2 diversification shifts in
-our MCC tree (posterior probability = 0.79) for the Ithomiini
- and Satyrini. Bayes factor (BF) values strongly 
-favored this model in comparison with a constant rate model (no diversification
-shifts, BF $>$ 1,884).
-
-\subsection*{Phylogenetic uncertainty in the MultiMEDUSA approach}
+We contrasted these results with those obtained from analyzing the data
+in the software BAMM {[}40{]}. We found that the best model (best
+configuration) estimated 2 diversification shifts in our MCC tree
+(posterior probability = 0.79) for the Ithomiini and Satyrini (Fig. S1).
+Bayes factor (BF) values strongly favored this model in comparison with
+a constant rate model (no diversification shifts, BF \(>\) 1,884).
+
+\subsection*{Phylogenetic uncertainty in the MultiMEDUSA
+approach}
 
@@
 lower median net diversification rates for the diversification shifts
-found by MEDUSA on the MCC tree from {[}13{]} (Table 2). Although the
-diversification pattern found by MEDUSA and MultiMEDUSA was the same,
-the latter consistently estimated lower rates. Furthermore, the shifts
-recovered with low net diversification rate on the MCC were recovered
-with negative net diversification rate by MultiMEDUSA. The background
-diversification and all shifts found by MEDUSA on the 1000 trees are
-provided as an R object in supp. mat. 7.
+found by MEDUSA on the MCC tree derived from the random sample of trees
+(Table 2). Although the diversification pattern found by MEDUSA and
+MultiMEDUSA was the same, the latter consistently estimated lower rates.
+Furthermore, the shifts recovered with low net diversification rate on
+the MCC were recovered with negative net diversification rate by
+MultiMEDUSA. The background diversification and all shifts found by
+MEDUSA on the 1000 trees are provided as an R object in supp. mat. 7.
 
@@
 MultiMEDUSA approach recovered different rates of diversification than
-those found using when the MCC tree alone.
-
-There were four diversification shifts found with more than 0.90 of
-probability in the trees from the random sample (Table 2): (i) net
-diversification rate increase in the genus \emph{Ypthima}; (ii) net
-diversification rate increase in the genus \emph{Charaxes}; (iii) rate
+those found when the MCC tree alone.
+
+There were four diversification shifts found in the trees from the
+random sample (Table 2): (i) net diversification rate increase in the
+genus \emph{Ypthima} (\(r = 0.22\)); (ii) net diversification rate
+increase in the genus \emph{Charaxes} (\(r = 0.21\)); (iii) rate
 increase in Ithomiini subtribes Oleriina + Ithomiina + Napeogenina +
-Dircennina + Godyridina; and (iv) rate increase in \emph{Callicore} +
-\emph{Diaethria}.
+Dircennina + Godyridina (\(r = 0.10\)); and (iv) rate increase in
+\emph{Callicore} + \emph{Diaethria} (\(r = 0.135\)).
 
 MultiMEDUSA provided mean and standard deviation statistics for the
-diversification values found on the shifts on the 1000 trees (supp. mat.
+diversification values on the shifts on the 1000 trees (supp. mat.
 13--14), and found that some of the changes in net diversification rate
@@
 of the splits in the MCC tree are recovered in very few trees, for
-example the split for the clade Satyrini (Euptychiina + Pronophilina +
-Satyrina + Maniolina) is recovered with a probability of 0.18 (Fig. 3).
-
-\subsection*{Estimation of trait-dependent speciation rates}
+example the split for the clade Satyrinae is recovered with a
+probability of 0.18 (Fig. 3).
+
+\subsection*{Estimation of trait-dependent speciation
+rates}
 
@@
 significantly better explanation than this character state having no
-effect on diversification ($\chi^2 = 12.3; 1 df; p < 0.001$) (Table 3;
+effect on diversification (\(\chi^2 = 12.3; 1 df; p < 0.001\)) (Table 3;
 character states available in supp. mat. 17, code in supp. mat. 18, and
@@
 
-\subsection*{Effects of phylogenetic uncertainty on the performance of MEDUSA}
+\subsection*{Effects of phylogenetic uncertainty on the performance of
+MEDUSA}
 
 The MEDUSA method has been used to infer changes in net diversification
-rates in a phylogenetic tree. Since its publication {[}16{]} the results
+rates in a phylogenetic tree. Since its publication {[}17{]} the results
 of using MEDUSA on a single tree, the maximum clade credibility tree,
 have been used for generation of hypotheses and discussion
-{[}10,43,44{]}. However, different diversification shifts and different
+{[}11,44,45{]}. However, different diversification shifts and different
 rates of diversification are found for certain lineages when
@@
 taxa appear to be strongly affected by long-branch attraction artifacts
-{[}30{]}. Thus, the Bayesian runs are expected to recover alternative
+{[}31{]}. Thus, the Bayesian runs are expected to recover alternative
 topologies on the posterior distribution of trees, resulting in low
 support and posterior probability values for the nodes. For example,
-posterior probability values for clades in Satyrini are very low
-{[}13{]}. As a result, MEDUSA inferred a net diversification rate
-increase for part of Satyrini in the maximum clade credibility tree, but
-this was recovered in the MultiMEDUSA analysis in only in 13\% of the
-random sample of trees.
+posterior probability values for clades in Satyrinae are very low
+{[}14{]}. As a result, MEDUSA inferred a net diversification rate
+increase for the Satyrinae (which includes Satyrini and its sister
+clade) in the maximum clade credibility tree, but this was recovered in
+the MultiMEDUSA analysis in only 13\% of the random sample of trees.
 
@@
 The clade Ithomiini and the non-basal danaids are well supported by high
-posterior probability values in Wahlberg et al. (2009) {[}13{]}.
+posterior probability values in Wahlberg et al. (2009) {[}14{]}.
 Therefore our MEDUSA analyses recovered an increase in net
@@
 
-\subsection*{Hostplant use and diversification in Nymphalidae}
+The results from BAMM recovered only two significant shifts in our MCC
+tree as this method seems to be much more conservative than the MEDUSA
+approach. One of these shifts is the Satyrini clade. MEDUSA estimated
+its parent node as significant shift in only a few trees from the
+posterior distribution. Thus, it might be important to take into account
+phylogenetic uncertainty in BAMM as well.
+
+\subsection*{Hostplant use and diversification in
+Nymphalidae}
 
@@
 Keith Brown suggested that feeding on Solanaceae was an important event
-in the diversification of Ithomiini butterflies {[}45{]}. Ithomiini
+in the diversification of Ithomiini butterflies {[}46{]}. Ithomiini
 butterflies are exclusively Neotropical and most species feed on
-Solanaceae hostplants during larval stage {[}8{]}. Optimizations of the
+Solanaceae hostplants during larval stage {[}9{]}. Optimizations of the
 evolution of hostplant use on phylogenies evidence a probable shift from
-Apocynaceae to Solanaceae in the ancestor of the tribe {[}8,46{]}.
-Fordyce (2010) {[}12{]} found that the Gamma statistics, a LTT plot of
+Apocynaceae to Solanaceae in the ancestor of the tribe {[}9,47{]}.
+Fordyce (2010) {[}13{]} found that the Gamma statistics, a LTT plot of
 an Ithomiini phylogeny and the fit of the density-dependent model of
@@
 due to the use of Solanaceae plants as hosts during larval stage. For
-this, we used a Bayesian approach {[}47{]} to test whether the trait
+this, we used a Bayesian approach {[}48{]} to test whether the trait
 ``feeding on Solanaceae'' had any effect on the diversification of the
@@
 trait ``feeding on Solanaceae hostplants'' (Fig. 4). This is in
-agreement with previoues findings using other statistical methods
-{[}12{]}. Due to the fact that Ithomiini are virtually the only
+agreement with previous findings using other statistical methods
+{[}13{]}. Due to the fact that Ithomiini are virtually the only
 nymphalids using Solanaceae as hostplants, it is possible that the trait
 responsible for a higher diversification of Ithomiini might not be the
-hostplant character. As noted by Maddison et al. (2007) {[}39{]}, the
+hostplant character. As noted by Maddison et al. (2007) {[}41{]}, the
 responsible trait might be a codistributed character such as a trait
@@
 that the high diversity of Ithomiini is consistent with the
-``escape-and-radiate scenario'' due to a shift onto Solanaceae {[}12{]}
+``escape-and-radiate scenario'' due to a shift onto Solanaceae {[}13{]}
 and radiation scenarios among chemically different lineages of
-Solanaceae plants {[}8,45{]}. According to this theory, the shift from
+Solanaceae plants {[}9,46{]}. According to this theory, the shift from
 Apocynaceae to Solanaceae allowed Ithomiini to invade newly available
@@
 detoxification of ingested food as they have been found in other
-butterfly groups {[}48{]}.
-
-The increase in diversification inferred by MEDUSA ocurred after the
+butterfly groups {[}49{]}.
+
+The increase in diversification inferred by MEDUSA occurred after the
 probable shift from Apocynaceae to Solanaceae, as the Solanaceae feeders
@@
 Although the Solanaceae genera used by the Ithomiini clades are well
-known {[}8{]}, we do not have any understanding on the physiological
+known {[}9{]}, we do not have any understanding on the physiological
 routes involved in the detoxification of Solanaceae compounds by the
 several lineages of Ithomiini. We can speculate that older lineages
-exploiting a novel toxic resource {[}8,13{]} might not be too efficient
+exploiting a novel toxic resource {[}9,14{]} might not be too efficient
 in metabolizing plant toxins and that younger lineages are able to deal
@@
 configurations of the proteins, providing more opportunities for
-hostplant switch {[}49{]}. This might be the reason why the basal
+hostplant switch {[}50{]}. This might be the reason why the basal
 Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and
-restricted to few Solanaceae hosts {[}8{]}, while recent subtribes are
+restricted to few Solanaceae hosts {[}9{]}, while recent subtribes are
 species-rich and have expanded their host range into several Solanaceae
-lineages {[}8{]}. It might be that the switch to feeding in Solanaceae
+lineages {[}9{]}. It might be that the switch to feeding in Solanaceae
 was an important event in the evolutionary history of Ithomiini, but the
@@
 diversify after their hostplants {[}2--4{]}. Wheat et al. (2007)
-{[}48{]} found strong evidence for a model of speciation congruent with
+{[}49{]} found strong evidence for a model of speciation congruent with
 Ehrlich and Raven's hypothesis in Pieridae butterflies due to, in
@@
 diversification in glucosinolate-feeding taxa shortly afterwards (with a
-lag of circa 10 MY). According to a recent dated phylogeny of
-the Angiosperms {[}50{]}, the family Solanaceae split from its sister
-group about 59 (49--68) MYA and diversification started (crown group
-age) around 37 (29--47) MYA. Wahlberg et al. (2009) {[}13{]} give the
-ages for origin and diversification for Ithomiini at 45 (39--53) and 37
+lag of circa 10 MY). According to a recent dated phylogeny of the
+Angiosperms {[}51{]}, the family Solanaceae split from its sister group
+about 59 (49--68) MYA and diversification started (crown group age)
+around 37 (29--47) MYA. Wahlberg et al. (2009) {[}14{]} give the ages
+for origin and diversification for Ithomiini at 45 (39--53) and 37
 (32--43) MYA, respectively. Thus, current evidence shows that Solanaceae
@@
 widespread species show higher genetic differentiation in their
-populations {[}51{]}.
-
-\subsubsection*{Satyrini}
-
-The diverse tribe Satyrini radiated simultaneously with the radiation of
-their main hostplant, grasses, during the climatic cooling in the
-Oligocene {[}14{]}. Thus, it is somewhat surprising that part of
-Satyrini (the subtribes Euptychiina, Satyrina and Pronophilina) were
-found to have accelerated diversification in only 13\% of the trees from
-the posterior distribution. Although this can be attributed to low
-phylogenetic signal {[}13{]}, the clade Satyrini is very robust {[}13{]}
-and MEDUSA failed to identify any significant accelerated net
-diversification rate for Satyrini. It appears that the radiation of
-Satyrini as a whole was not remarkably fast and therefore not picked up
-by MEDUSA. This should be expected if the diversification of Satyrinae
-occurred in a stepwise manner, with pulses or bursts of diversification
-for certain lineages but unlikely for the tribe Satyrini as a whole.
+populations {[}52{]}.
+
+\subsubsection*{Satyrinae}
+
+Lineages in the diverse family Satyrinae radiated simultaneously with
+the radiation of their main hostplant, grasses, during the climatic
+cooling in the Oligocene {[}15{]}. Thus, it is somewhat surprising that
+part of Satyrinae were found to have accelerated diversification in only
+13\% of the trees from the posterior distribution. Although this can be
+attributed to low phylogenetic signal {[}14{]}, the clade Satyrini is
+very robust {[}14{]} and MEDUSA failed to identify any significant
+accelerated net diversification rate for Satyrini. It appears that the
+radiation of Satyrini as a whole was not remarkably fast and therefore
+not picked up by MEDUSA, although it estimated a diversification shift
+for Satyrini + its sister clade. This should be expected if the
+diversification of Satyrini occurred in a stepwise manner, with pulses
+or bursts of diversification for certain lineages but unlikely for the
+tribe Satyrini as a whole.
 
@@
 diversified near simultaneously, which is in agreement with the diffuse
-co-speciation hypothesis {[}2,4{]}.
+cospeciation hypothesis {[}2,4{]}.
 
@@
 nodes indicate the diversification shift number as found by MEDUSA.
-Numbers next to circles indicat the posterior probability values for
+Numbers next to circles indicate the posterior probability values for
 such nodes.
 
-\item {\bf Figure 2. Diversification rates for taxa estimated by MEDUSA on
+{\bf Figure 2. Diversification rates for taxa estimated by MEDUSA on
 the samples of 1000 random trees}
 
-\item {\bf Figure 3. Results of MultiMEDUSA analysis showing the
+{\bf Figure 3. Results of MultiMEDUSA analysis showing the
 probability of specific nodes being characterized by significant shifts
@@
 $\lambda1$, net diversification rate = $r1$.
-\end{description}
 
 \section*{Supporting Information Legends}
-\begin{description}
- \item {\bf Supporting Information S01.nex.} MCC Nymphalidae tree from
-{[}13{]}
-
- \item {\bf Supporting Information S02.R} R script to run MEDUSA on the
-MCC Nymphalidae tree from {[}13{]}. This script removes the outgroup
+{\textbf{Supporting Information S01.nex.}} MCC Nymphalidae tree from
+{[}14{]}
+
+{\textbf{Supporting Information S02.R}} R script to run MEDUSA on the
+MCC Nymphalidae tree from {[}14{]}. This script removes the outgroup
 taxa and loads the richness data for the tree terminals.
 
- \item {\bf Supporting Information S03.csv} Species richness for lineages
+{\textbf{Supporting Information S03.csv}} Species richness for lineages
 in Nymphalidae.
 
- \item {\bf Supporting Information S04.txt} Results of running MEDUSA on
+{\textbf{Supporting Information S04.txt}} Results of running MEDUSA on
 the MCC tree.
 
- \item {\bf Supporting Information S05 1000 random trees no outgroups.zip} 1000 random trees from {[}13{]}.
-
- \item {\bf Supporting Information S06.R}?R script to run a MultiMEDUSA
-analysis on 1000 random trees from {[}13{]}.
-
- \item {\bf Supporting Information S07 multimedusa on 1000 trees.zip} Raw results from the MultiMEDUSA run on the random sample of trees from
-the posterior distribution.
-
- \item {\bf Supporting Information S08.pdf} Summary of the MulitMEDUSA
+{\textbf{Supporting Information S05 1000 random trees no outgroups.zip}}
+1000 random trees from {[}14{]}.
+
+{\textbf{Supporting Information S06.R}}~R script to run a MultiMEDUSA
+analysis on 1000 random trees from {[}14{]}.
+
+{\textbf{Supporting Information S07 multimedusa on 1000 trees.zip}} Raw
+results from the MultiMEDUSA run on the random sample of trees from the
+posterior distribution.
+
+{\textbf{Supporting Information S08.pdf}} Summary of the MultiMEDUSA
 analysis on the 1000 trees from the posterior distribution.
 
- \item {\bf Supporting Information S09.tree} MCC tree from the 1000
-random trees selected from the posterior distribution.
-
- \item {\bf Supporting Information S10.R} R script to run a MEDUSA
+{\textbf{Supporting Information S09.tree}} MCC tree from the 1000 random
+trees selected from the posterior distribution.
+
+{\textbf{Supporting Information S10.R}} R script to run a MEDUSA
 analysis on the MCC tree from the 1000 random trees selected from the
@@
 
- \item {\bf Supporting Information S12.pdf} Figure for MEDUSA run on MCC
+{\textbf{Supporting Information S12.pdf}} Figure for MEDUSA run on MCC
 tree from random 1000 trees.
 
- \item {\bf Supporting Information S13.txt} Summary results from a
+{\textbf{Supporting Information S13.txt}} Summary results from a
 MultiMEDUSA run on the 1000 random trees selected from the posterior
@@
 
- \item {\bf Supporting Information S15.csv} Hostplants of Nymphalidae
+{\textbf{Supporting Information S15.csv}} Hostplants of Nymphalidae
 butterflies recorded from the literature.
 
- \item {\bf Supporting Information S16.csv} References for hostplants
+{\textbf{Supporting Information S16.csv}} References for hostplants
 data.
 
- \item {\bf Supporting Information S17.csv} Data matrix with character
+{\textbf{Supporting Information S17.csv}} Data matrix with character
 states for hosplant use.
 
- \item {\bf Supporting Information S18.R} R script for running the BiSSE
+{\textbf{Supporting Information S18.R}} R script for running the BiSSE
 analysis.
 
- \item {\bf Supporting Information S19.csv} Raw results for the BiSSE
+{\textbf{Supporting Information S19.csv}} Raw results for the BiSSE
 analysis.
 
- \item {\bf Supporting Information S20.csv} Raw results for the combined
+{\textbf{Supporting Information S20.csv}} Raw results for the combined
 BiSSE analysis.
 
- \item {\bf Figure S1.} Boxplot of speciation ($\lambda$) and extinction
-($\mu$) values for the 95\% credibility intervals of values estimated
+{\textbf{Figure S1.}} Best diversification model as estimated by BAMM.
+Two diversification shifts were recovered (Ithomiini and Satyrini). The
+MEDUSA analysis found a diversification shifts in Satyrinae (the parent
+node of Satyini).
+
+{\textbf{Figure S2.}} Boxplot of speciation (\(\lambda\)) and extinction
+(\(\mu\)) values for the 95\% credibility intervals of values estimated
 by BiSSE analysis of diversification due to feeding on Solanaceae
@@
 
- \item {\bf Figure S2.} BiSSE analysis of diversification of nymphalids due
-to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
+{\textbf{Figure S3.}} BiSSE analysis of diversification of nymphalids
+due to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
 \emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
 recovered, speciation and net diversification rates are significantly
-higher for Solanaceae feeders $\lambda1$, r1).
-
- \item {\bf Figure S3.} Net diversification rates of nymphalids feeding on
+higher for Solanaceae feeders \(\lambda1\), r1).
+
+{\textbf{Figure S4.}} Net diversification rates of nymphalids feeding on
 Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
@@
 
- \item {\bf Figure S4.} Boxplot of speciation and extinction values for the
-95\% credibility intervals of values estimated by BiSSE analysis of
+{\textbf{Figure S5.}} Boxplot of speciation and extinction values for
+the 95\% credibility intervals of values estimated by BiSSE analysis of
 diversification due to feeding on Solanaceae plants on the combined
@@
 
- \item {\bf Figure S5.} BiSSE analysis of diversification of nymphalids due
-to feeding on Apocynaceae hostplants. Speciation and net diversification
-rates are similar.
-\end{description}
-
-\section*{References}
-1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
-adaptive zones: Has phytophagy promoted insect diversification? Am Nat
-132: 107--128.
-doi:\href{http://dx.doi.org/10.1086/284840}{10.1086/284840}.
-
-2. Nyman T, Linder HP, Pe{\~n}a C, Malm T, Wahlberg N (2012) Climate-driven
-diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
-889--898.
-doi:\href{http://dx.doi.org/10.1111/j.1461-0248.2012.01782.x}{10.1111/j.1461-0248.2012.01782.x}.
-
-3. Ehrlich PR, Raven PH (1964) Butterflies and plants: A study in
-coevolution. Evolution 18: 586--608.
-doi:\href{http://dx.doi.org/10.2307/2406212}{10.2307/2406212}.
-
-4. Janz N (2011) Ehrlich and Raven Revisited: Mechanisms Underlying
-Codiversification of Plants and Enemies. Annu Rev Ecol Evol S 42:
-71--89.
-doi:\href{http://dx.doi.org/10.1146/annurev-ecolsys-102710-145024}{10.1146/annurev-ecolsys-102710-145024}.
-
-5. Nylin S, Slove J, Janz N (2014) Host plant utilization, host range
-oscillations and diversification in Nymphalid butterflies: A
-phylogenetic investigation. Evolution 68: 105--124.
-doi:\href{http://dx.doi.org/10.1111/evo.12227}{10.1111/evo.12227}.
-
-6. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
-al. (2011) Order Lepidoptera. In: Animal biodiversity: An outline of
-higher-level classification and survey of taxonomic richness. Zootaxa.
-pp. 212--221.
-
-7. Joron M, Jiggins C, Papanicolaou A, McMillan W (2006) Heliconius wing
-patterns: An evo-devo model for understanding phenotypic diversity.
-Heredity 97: 157--167.
-
-8. Willmott KR, Freitas AVL (2006) Higher-level phylogeny of the
-Ithomiinae (Lepidoptera: Nymphalidae): classification, patterns of
-larval hostplant colonization and diversification. Cladistics 22:
-297--368.
-doi:\href{http://dx.doi.org/10.1111/j.1096-0031.2006.00108.x}{10.1111/j.1096-0031.2006.00108.x}.
-
-9. Brakefield PM, Beldade P, Zwaan BJ (2009) The african butterfly
-bicyclus anynana: A model for evolutionary genetics and evolutionary
-developmental biology. Cold Spring Harbor Protocols 2009: pdb--emo122.
-
-10. Heikkil{\"a} M, Kaila L, Mutanen M, Pe{\~n}a C, Wahlberg N (2012) Cretaceous
-origin and repeated tertiary diversification of the redefined
-butterflies. P Roy Soc B 279: 1093--1099.
-doi:\href{http://dx.doi.org/10.1098/rspb.2011.1430}{10.1098/rspb.2011.1430}.
-
-11. Elias M, Joron M, Willmott KR, Silva-Brand{\~a}o KL, Kaiser V, et al.
-(2009) Out of the Andes: Patterns of diversification in clearwing
-butterflies. Mol Ecol 18: 1716--1729.
-doi:\href{http://dx.doi.org/10.1111/j.1365-294X.2009.04149.x}{10.1111/j.1365-294X.2009.04149.x}.
-
-12. Fordyce JA (2010) Host shifts and evolutionary radiations of
-butterflies. P Roy Soc B 277: 3735--3743.
-doi:\href{http://dx.doi.org/10.1098/rspb.2010.0211}{10.1098/rspb.2010.0211}.
-
-13. Wahlberg N, Leneveu J, Kodandaramaiah U, Pe{\~n}a C, Nylin S, et al.
-(2009) Nymphalid butterflies diversify following near demise at the
-Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302.
-doi:\href{http://dx.doi.org/10.1098/rspb.2009.1303}{10.1098/rspb.2009.1303}.
-
-14. Pe{\~n}a C, Wahlberg N (2008) Prehistorical climate change increased
-diversification of a group of butterflies. Biol Lett 4: 274--278.
-doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.
-
-15. Ferrer-Paris JR, S{\'a}nchez-Mercado A, Viloria {\'A}L, Donaldson J (2013)
-Congruence and Diversity of Butterfly-Host Plant Associations at Higher
-Taxonomic Levels. PloS one 8: e63570.
-doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.
-
-16. Alfaro ME, Santini F, Brock C, Alamillo H, Dornburg A, et al. (2009)
-Nine exceptional radiations plus high turnover explain species diversity
-in jawed vertebrates. P Natl Acad Sci USA 106: 13410--13414.
-doi:\href{http://dx.doi.org/10.1073/pnas.0811087106}{10.1073/pnas.0811087106}.
-
-17. Harmon LJ, Rabosky DL, {FitzJohn} RG, Brown JW (2011) turboMEDUSA:
-MEDUSA: Modeling Evolutionary Diversification Using Stepwise AIC.
-
-18. {FitzJohn} RG (2012) Diversitree: comparative phylogenetic analyses
-of diversification in R. Method Ecol Evol 3: 1084--1092.
-doi:\href{http://dx.doi.org/10.1111/j.2041-210X.2012.00234.x}{10.1111/j.2041-210X.2012.00234.x}.
-
-19. Lamas G (2004) Checklist: Part 4A. Hesperioidea-Papilionoidea.
-Gainesville.
-
-20. Matos-Marav{\'i} PF, Pe{\~n}a C, Willmott KR, Freitas AV, Wahlberg N (2013)
-Systematics and evolutionary history of butterflies in the ``Taygetis
-clade'' (Nymphalidae: Satyrinae: Euptychiina): Towards a better
-understanding of Neotropical biogeography. Mol Phylogenet Evol 66:
-54--68.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2012.09.005}{10.1016/j.ympev.2012.09.005}.
-
-21. Brower AVZ, Wahlberg N, Ogawa JR, Boppr{\'e} M, Vane-Wright RI (2010)
-Phylogenetic relationships among genera of danaine butterflies
-(Lepidoptera: Nymphalidae) as implied by morphology and DNA sequences.
-Syst Biodivers 8: 75--89.
-doi:\href{http://dx.doi.org/10.1080/14772001003626814}{10.1080/14772001003626814}.
-
-22. Kodandaramaiah U, Lees DC, M{\"u}ller CJ, Torres E, Karanth KP, et al.
-(2010) Phylogenetics and biogeography of a spectacular Old World
-radiation of butterflies: the subtribe Mycalesina (Lepidoptera:
-Nymphalidae: Satyrini). BMC Evol Biol 10: 172.
-doi:\href{http://dx.doi.org/10.1186/1471-2148-10-172}{10.1186/1471-2148-10-172}.
-
-23. Kodandaramaiah U, Pe{\~n}a C, Braby MF, Grund R, M{\"u}ller CJ, et al.
-(2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
-problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.08.012}{10.1016/j.ympev.2009.08.012}.
-
-24. Ortiz-Acevedo E, Willmott KR (2013) Molecular systematics of the
-butterfly tribe Preponini (Nymphalidae: Charaxinae). Syst Entomol 38:
-440--449.
-doi:\href{http://dx.doi.org/10.1111/syen.12008}{10.1111/syen.12008}.
-
-25. De-Silva DL, Day JJ, Elias M, Willmott KR, Whinnett A, et al. (2010)
-Molecular phylogenetics of the Neotropical butterfly subtribe Oleriina
-(Nymphalidae: Danainae: Ithomiini). Mol Phylogenet Evol 55: 1032--1041.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2010.01.010}{10.1016/j.ympev.2010.01.010}.
-
-26. Freitas AVL, Brown Jr. KS (2004) Phylogeny of the Nymphalidae
-(Lepidoptera). Syst Biol 53: 363--383.
-doi:\href{http://dx.doi.org/10.1080/10635150490445670}{10.1080/10635150490445670}.
-
-27. Pe{\~n}a C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
-(2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
-Nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2006.02.007}{10.1016/j.ympev.2006.02.007}.
-
-28. Penz CM (1999) Higher level phylogeny for the passion-vine
-butterflies (Nymphalidae, Heliconiinae) based on early stage and adult
-morphology. Zool J Linn Soc 127: 277--344.
-doi:\href{http://dx.doi.org/10.1111/j.1096-3642.1999.tb00680.x}{10.1111/j.1096-3642.1999.tb00680.x}.
-
-29. Silva-Brand{\~a}o KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
-Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
-tribe Acraeini (Lepidoptera, Nymphalidae, Heliconiinae) and the
-evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2007.11.024}{10.1016/j.ympev.2007.11.024}.
-
-30. Pe{\~n}a C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
-butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
-methods. Zool J Linn Soc 161: 64--87.
-doi:\href{http://dx.doi.org/10.1111/j.1096-3642.2009.00627.x}{10.1111/j.1096-3642.2009.00627.x}.
-
-31. Pe{\~n}a C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
-history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
-Satyrinae). Zool Scr 39: 243--258.
-doi:\href{http://dx.doi.org/10.1111/j.1463-6409.2010.00421.x}{10.1111/j.1463-6409.2010.00421.x}.
-
-32. Ackery PR (1988) Hostplants and classification: A review of
-nymphalid butterflies. Biol J Linn Soc 33: 95--203.
-doi:\href{http://dx.doi.org/10.1111/j.1095-8312.1988.tb00446.x}{10.1111/j.1095-8312.1988.tb00446.x}.
-
-33. Dyer LA, Gentry GL (2002) Caterpillars and parasitoids of a tropical
-lowland wet forest. Available: \url{http://www.caterpillars.org}.
-Accessed 17 April 2013.
-
-34. Beccaloni G, Viloria {\'A}, Hall S, Robinson G (2008) Catalogue of the
-hostplants of the Neotropical butterflies. London: The Natural History
-Museum.
-
-35. Janzen DH, Hallwachs W (2009) Dynamic database for an inventory of
-the macrocaterpillar fauna, and its food plants and parasitoids, of Area
-de Conservacion Guanacaste (ACG), northwestern Costa Rica. Available:
-\url{http://janzen.sas.upenn.edu}. Accessed 17 April 2013.
-
-36. Core Team R (2013) R: A Language and Environment for Statistical
-Computing. Vienna, Austria: R Foundation for Statistical Computing.
-Available: \url{http://www.r-project.org}.
-
-37. Popescu A-A, Huber KT, Paradis E (2012) ape 3.0: New tools for
-distance-based phylogenetics and evolutionary analysis in R.
-Bioinformatics 28: 1536--1537.
-doi:\href{http://dx.doi.org/10.1093/bioinformatics/bts184}{10.1093/bioinformatics/bts184}.
-
-38. Harmon LJ, Weir JT, Brock CD, Glor RE, Challenger W (2008) GEIGER:
-investigating evolutionary radiations. Bioinformatics 24: 129--131.
-doi:\href{http://dx.doi.org/10.1093/bioinformatics/btm538}{10.1093/bioinformatics/btm538}.
-
-39. Maddison WP, Midford PE, Otto SP (2007) Estimating a binary
-character's effect on speciation and extinction. Syst Biol 56: 701--710.
-doi:\href{http://dx.doi.org/10.1080/10635150701607033}{10.1080/10635150701607033}.
-
-40. Scott JA (1986) The Butterflies of North America. Stanford
-University Press.
-
-41. Davis MP, Midford PE, Maddison W (2013) Exploring power and
-parameter estimation of the BiSSE method for analyzing species
-diversification. BMC Evol Biol 13: 38.
-doi:\href{http://dx.doi.org/10.1186/1471-2148-13-38}{10.1186/1471-2148-13-38}.
-
-42. Rabosky DL (2014) Automatic Detection of Key Innovations, Rate Shifts, and
-Diversity-Dependence on Phylogenetic Trees. PLoS ONE 9: e89543. doi:
-\href{http://dx.doi.org/10.1371/journal.pone.0089543}{10.1371/journal.pone.0089543}.
-
-43. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
-bees cut leaves? New insights into the early evolution of bees. P Roy
-Soc B 278: 3593--3600.
-doi:\href{http://dx.doi.org/10.1098/rspb.2011.0365}{10.1098/rspb.2011.0365}.
-
-44. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
-clades of Agaricales. P Roy Soc B 279: 2003--2011.
-doi:\href{http://dx.doi.org/10.1098/rspb.2011.2428}{10.1098/rspb.2011.2428}.
-
-45. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
-Interface. Ann Mo Bot Gard 74: 359--397.
-
-46. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand{\~a}o KL, Whinnett A, et
-al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
-Nymphalidae) inferred from one mitochondrial and two nuclear gene
-regions. Syst Entomol 31: 288--301.
-doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2006.00321.x}{10.1111/j.1365-3113.2006.00321.x}.
-
-47. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
-trait-dependent speciation and extinction rates from incompletely
-resolved phylogenies. Syst Biol 58: 595--611.
-doi:\href{http://dx.doi.org/10.1093/sysbio/syp067}{10.1093/sysbio/syp067}.
-
-48. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
-The genetic basis of a plant-insect coevolutionary key innovation. P
-Natl Acad Sci USA 104: 20427--20431.
-doi:\href{http://dx.doi.org/10.1073/pnas.0706229104}{10.1073/pnas.0706229104}.
-
-49. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
-furanocoumarin-metabolizing cytochrome P450 monooxygenases in two
-papilionids: Specificity and substrate encounter rate. P Natl Acad Sci
-USA 100: 14593--14598.
-doi:\href{http://dx.doi.org/10.1073/pnas.1934643100}{10.1073/pnas.1934643100}.
-
-50. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
-the angiosperms re-revisited. Am J Bot 97: 1296--1303.
-doi:\href{http://dx.doi.org/10.3732/ajb.0900346}{10.3732/ajb.0900346}.
-
-51. Wahlberg N, Rubinoff D (2011) Vagility across \emph{Vanessa} (Lepidoptera:
-Nymphalidae): mobility in butterfly species does not inhibit the
-formation and persistence of isolated sister taxa. Syst Entomol 36:
-362--370.
-doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2010.00566.x}{10.1111/j.1365-3113.2010.00566.x}.
+{\textbf{Figure S6.}} BiSSE analysis of diversification of nymphalids
+due to feeding on Apocynaceae hostplants. Speciation and net
+diversification rates are similar.
 
@@
     \caption{\bf{Differences in rates estimated by MEDUSA on the MCC
-tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
-from the posterior distribution.}}
+tree from the sample of trees from the posterior distribution and the
+MultiMEDUSA approach. Shift consistently recovered across the
+sample of trees in bold face.}}
 \begin{tabular}{lcrr}
 Split.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
-1          & 0.090          & not found                  & 0.000                          \\
-2          & 0.054          & -0.030                     & 0.864                          \\
-3          & 0.311          & 0.219                      & 0.961                          \\
-4          & 0.290          & 0.212                      & 0.996                          \\
-5          & 0.187          & 0.101                      & 0.927                          \\
-6          & 0.116          & 0.039                      & 0.131                          \\
-7          & 0.064          & -0.052                     & 0.195                          \\
-8          & 0.240          & 0.166                      & 0.319                          \\
+1          & 0.092          & not found                  & 0.000                          \\
+2          & 0.055          & -0.030                     & 0.864                          \\
+\bf{3}     & 0.184          & \bf{0.219}                 & \bf{0.961}                     \\
+\bf{4}     & 0.166          & \bf{0.212}                 & \bf{0.996}                     \\
+\bf{5}     & 0.111          & \bf{0.101}                 & \bf{0.927}                     \\
+6          & 0.119          & 0.039                      & 0.131                          \\
+7          & 0.066          & -0.052                     & 0.195                          \\
+8          & 0.232          & 0.166                      & 0.319                          \\
 9          & 0.042          & -0.049                     & 0.897                          \\
-10         & 0.209          & 0.149                      & 0.619                          \\
+10         & 0.058          & 0.149                      & 0.619                          \\
 11         & 0.311          & 0.208                      & 0.831                          \\
-12         & 0.219          & 0.135                      & 0.911                          \\
-13         & 0.220          & 0.117                      & 0.276                          \\
-14         & 0.190          & 0.115                      & 0.379                          \\
-15         & 0.234          & 0.127                      & 0.825                          \\
-16         & 0.000          & -0.005                     & 0.659                          \\
-17         & 0.311          & 0.248                      & 0.785                          \\
-18         & 0.024          & -0.064                     & 0.024                          \\
-19         & 0.000          & -0.087                     & 0.381
+\bf{12}    & 0.219          & \bf{0.135}                 & \bf{0.911}                     \\
+13         & 0.082          & 0.117                      & 0.276                          \\
+14         & 0.099          & 0.115                      & 0.379                          \\
+15         & 0.113          & 0.127                      & 0.825                          \\
+16         & 0.222          & -0.005                     & 0.659                          \\
+17         & 0.243          & 0.248                      & 0.785                          \\
+18         & 0.192          & -0.064                     & 0.024                          \\
+19         & 0.064          & -0.087                     & 0.381                         
 \end{tabular}
@@
 
+\pagebreak
+\section*{References}
+
+1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
+adaptive zones: Has phytophagy promoted insect diversification? Am Nat
+132: 107--128.
+doi:\href{http://dx.doi.org/10.1086/284840}{10.1086/284840}.
+
+2. Nyman T, Linder HP, Pe\~na C, Malm T, Wahlberg N (2012) Climate-driven
+diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
+889--898.
+doi:\href{http://dx.doi.org/10.1111/j.1461-0248.2012.01782.x}{10.1111/j.1461-0248.2012.01782.x}.
+
+3. Ehrlich PR, Raven PH (1964) Butterflies and plants: A study in
+coevolution. Evolution 18: 586--608.
+doi:\href{http://dx.doi.org/10.2307/2406212}{10.2307/2406212}.
+
+4. Janz N (2011) Ehrlich and Raven Revisited: Mechanisms Underlying
+Codiversification of Plants and Enemies. Annu Rev Ecol Evol S 42:
+71--89.
+doi:\href{http://dx.doi.org/10.1146/annurev-ecolsys-102710-145024}{10.1146/annurev-ecolsys-102710-145024}.
+
+5. Nylin S, Slove J, Janz N (2014) Host plant utilization, host range
+oscillations and diversification in Nymphalid butterflies: A
+phylogenetic investigation. Evolution 68: 105--124.
+doi:\href{http://dx.doi.org/10.1111/evo.12227}{10.1111/evo.12227}.
+
+6. Hardy NB, Otto SP (2014) Specialization and generalization in the
+diversification of phytophagous insects: Tests of the musical chairs and
+oscillation hypotheses. P Roy Soc B 281: 20132960.
+
+7. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
+al. (2011) Order Lepidoptera. In: Animal biodiversity: An outline of
+higher-level classification and survey of taxonomic richness. Zootaxa.
+pp. 212--221.
+
+8. Joron M, Jiggins C, Papanicolaou A, McMillan W (2006) Heliconius wing
+patterns: An evo-devo model for understanding phenotypic diversity.
+Heredity 97: 157--167.
+
+9. Willmott KR, Freitas AVL (2006) Higher-level phylogeny of the
+Ithomiinae (Lepidoptera: Nymphalidae): classification, patterns of
+larval hostplant colonization and diversification. Cladistics 22:
+297--368.
+doi:\href{http://dx.doi.org/10.1111/j.1096-0031.2006.00108.x}{10.1111/j.1096-0031.2006.00108.x}.
+
+10. Brakefield PM, Beldade P, Zwaan BJ (2009) The african butterfly
+Bicyclus anynana: A model for evolutionary genetics and evolutionary
+developmental biology. Cold Spring Harb Prot 2009: pdb--emo122.
+
+11. Heikkil\"a M, Kaila L, Mutanen M, Pe\~na C, Wahlberg N (2012) Cretaceous
+origin and repeated tertiary diversification of the redefined
+butterflies. P Roy Soc B 279: 1093--1099.
+doi:\href{http://dx.doi.org/10.1098/rspb.2011.1430}{10.1098/rspb.2011.1430}.
+
+12. Elias M, Joron M, Willmott KR, Silva-Brand\~ao KL, Kaiser V, et al.
+(2009) Out of the Andes: Patterns of diversification in clearwing
+butterflies. Mol Ecol 18: 1716--1729.
+doi:\href{http://dx.doi.org/10.1111/j.1365-294X.2009.04149.x}{10.1111/j.1365-294X.2009.04149.x}.
+
+13. Fordyce JA (2010) Host shifts and evolutionary radiations of
+butterflies. P Roy Soc B 277: 3735--3743.
+doi:\href{http://dx.doi.org/10.1098/rspb.2010.0211}{10.1098/rspb.2010.0211}.
+
+14. Wahlberg N, Leneveu J, Kodandaramaiah U, Pe\~na C, Nylin S, et al.
+(2009) Nymphalid butterflies diversify following near demise at the
+Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302.
+doi:\href{http://dx.doi.org/10.1098/rspb.2009.1303}{10.1098/rspb.2009.1303}.
+
+15. Pe\~na C, Wahlberg N (2008) Prehistorical climate change increased
+diversification of a group of butterflies. Biol Lett 4: 274--278.
+doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.
+
+16. Ferrer-Paris JR, S{\'a}nchez-Mercado A, Viloria \'AL, Donaldson J (2013)
+Congruence and Diversity of Butterfly-Host Plant Associations at Higher
+Taxonomic Levels. PLoS ONE 8: e63570.
+doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.
+
+17. Alfaro ME, Santini F, Brock C, Alamillo H, Dornburg A, et al. (2009)
+Nine exceptional radiations plus high turnover explain species diversity
+in jawed vertebrates. P Natl Acad Sci USA 106: 13410--13414.
+doi:\href{http://dx.doi.org/10.1073/pnas.0811087106}{10.1073/pnas.0811087106}.
+
+18. Harmon LJ, Rabosky DL, {FitzJohn} RG, Brown JW (2011) turboMEDUSA:
+MEDUSA: Modeling Evolutionary Diversification Using Stepwise AIC.
+Available: \url{https://github.com/josephwb/turboMEDUSA}.
+
+19. {FitzJohn} RG (2012) Diversitree: comparative phylogenetic analyses
+of diversification in R. Method Ecol Evol 3: 1084--1092.
+doi:\href{http://dx.doi.org/10.1111/j.2041-210X.2012.00234.x}{10.1111/j.2041-210X.2012.00234.x}.
+
+20. Lamas G (2004) Checklist: Part 4A. Hesperioidea-Papilionoidea.
+Gainesville: Association for Tropical Lepidoptera; Scientific
+Publishers.
+
+21. Matos-Marav\'i PF, Pe\~na C, Willmott KR, Freitas AV, Wahlberg N (2013)
+Systematics and evolutionary history of butterflies in the ``Taygetis
+clade'' (Nymphalidae: Satyrinae: Euptychiina): Towards a better
+understanding of Neotropical biogeography. Mol Phylogenet Evol 66:
+54--68.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2012.09.005}{10.1016/j.ympev.2012.09.005}.
+
+22. Brower AVZ, Wahlberg N, Ogawa JR, Boppr\'e M, Vane-Wright RI (2010)
+Phylogenetic relationships among genera of danaine butterflies
+(Lepidoptera: Nymphalidae) as implied by morphology and DNA sequences.
+Syst Biodivers 8: 75--89.
+doi:\href{http://dx.doi.org/10.1080/14772001003626814}{10.1080/14772001003626814}.
+
+23. Kodandaramaiah U, Lees DC, M\"uller CJ, Torres E, Karanth KP, et al.
+(2010) Phylogenetics and biogeography of a spectacular Old World
+radiation of butterflies: the subtribe Mycalesina (Lepidoptera:
+Nymphalidae: Satyrini). BMC Evol Biol 10: 172.
+doi:\href{http://dx.doi.org/10.1186/1471-2148-10-172}{10.1186/1471-2148-10-172}.
+
+24. Kodandaramaiah U, Pe\~na C, Braby MF, Grund R, M\"uller CJ, et al.
+(2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
+problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.08.012}{10.1016/j.ympev.2009.08.012}.
+
+25. Ortiz-Acevedo E, Willmott KR (2013) Molecular systematics of the
+butterfly tribe Preponini (Nymphalidae: Charaxinae). Syst Entomol 38:
+440--449.
+doi:\href{http://dx.doi.org/10.1111/syen.12008}{10.1111/syen.12008}.
+
+26. De-Silva DL, Day JJ, Elias M, Willmott KR, Whinnett A, et al. (2010)
+Molecular phylogenetics of the Neotropical butterfly subtribe Oleriina
+(Nymphalidae: Danainae: Ithomiini). Mol Phylogenet Evol 55: 1032--1041.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2010.01.010}{10.1016/j.ympev.2010.01.010}.
+
+27. Freitas AVL, Brown Jr. KS (2004) Phylogeny of the Nymphalidae
+(Lepidoptera). Syst Biol 53: 363--383.
+doi:\href{http://dx.doi.org/10.1080/10635150490445670}{10.1080/10635150490445670}.
+
+28. Pe\~na C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
+(2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
+Nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2006.02.007}{10.1016/j.ympev.2006.02.007}.
+
+29. Penz CM (1999) Higher level phylogeny for the passion-vine
+butterflies (Nymphalidae, Heliconiinae) based on early stage and adult
+morphology. Zool J Linn Soc 127: 277--344.
+doi:\href{http://dx.doi.org/10.1111/j.1096-3642.1999.tb00680.x}{10.1111/j.1096-3642.1999.tb00680.x}.
+
+30. Silva-Brand\~ao KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
+Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
+tribe Acraeini (Lepidoptera, Nymphalidae, Heliconiinae) and the
+evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
+doi:\href{http://dx.doi.org/10.1016/j.ympev.2007.11.024}{10.1016/j.ympev.2007.11.024}.
+
+31. Pe\~na C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
+butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
+methods. Zool J Linn Soc 161: 64--87.
+doi:\href{http://dx.doi.org/10.1111/j.1096-3642.2009.00627.x}{10.1111/j.1096-3642.2009.00627.x}.
+
+32. Pe\~na C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
+history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
+Satyrinae). Zool Scr 39: 243--258.
+doi:\href{http://dx.doi.org/10.1111/j.1463-6409.2010.00421.x}{10.1111/j.1463-6409.2010.00421.x}.
+
+33. Ackery PR (1988) Hostplants and classification: A review of
+nymphalid butterflies. Biol J Linn Soc 33: 95--203.
+doi:\href{http://dx.doi.org/10.1111/j.1095-8312.1988.tb00446.x}{10.1111/j.1095-8312.1988.tb00446.x}.
+
+34. Dyer LA, Gentry GL (2002) Caterpillars and parasitoids of a tropical
+lowland wet forest. Available: \url{http://www.caterpillars.org}.
+Accessed 17 April 2013.
+
+35. Beccaloni G, Viloria \'A, Hall S, Robinson G (2008) Catalogue of the
+hostplants of the Neotropical butterflies. London: The Natural History
+Museum.
+
+36. Janzen DH, Hallwachs W (2009) Dynamic database for an inventory of
+the macrocaterpillar fauna, and its food plants and parasitoids, of Area
+de Conservacion Guanacaste (ACG), northwestern Costa Rica. Available:
+\url{http://janzen.sas.upenn.edu}. Accessed 17 April 2013.
+
+37. Core Team R (2013) R: A Language and Environment for Statistical
+Computing. Vienna, Austria: R Foundation for Statistical Computing.
+Available: \url{http://www.r-project.org}.
+
+38. Popescu A-A, Huber KT, Paradis E (2012) ape 3.0: New tools for
+distance-based phylogenetics and evolutionary analysis in R.
+Bioinformatics 28: 1536--1537.
+doi:\href{http://dx.doi.org/10.1093/bioinformatics/bts184}{10.1093/bioinformatics/bts184}.
+
+39. Harmon LJ, Weir JT, Brock CD, Glor RE, Challenger W (2008) GEIGER:
+investigating evolutionary radiations. Bioinformatics 24: 129--131.
+doi:\href{http://dx.doi.org/10.1093/bioinformatics/btm538}{10.1093/bioinformatics/btm538}.
+
+40. Rabosky DL (2014) Automatic detection of key innovations, rate
+shifts, and diversity-dependence on phylogenetic trees. PLoS ONE 9:
+e89543. Available:
+\url{http://dx.doi.org/10.1371\%2Fjournal.pone.0089543}.
+
+41. Maddison WP, Midford PE, Otto SP (2007) Estimating a binary
+character's effect on speciation and extinction. Syst Biol 56: 701--710.
+doi:\href{http://dx.doi.org/10.1080/10635150701607033}{10.1080/10635150701607033}.
+
+42. Scott JA (1986) The Butterflies of North America. Stanford
+University Press.
+
+43. Davis MP, Midford PE, Maddison W (2013) Exploring power and
+parameter estimation of the BiSSE method for analyzing species
+diversification. BMC Evol Biol 13: 38.
+doi:\href{http://dx.doi.org/10.1186/1471-2148-13-38}{10.1186/1471-2148-13-38}.
+
+44. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
+bees cut leaves? New insights into the early evolution of bees. P Roy
+Soc B 278: 3593--3600.
+doi:\href{http://dx.doi.org/10.1098/rspb.2011.0365}{10.1098/rspb.2011.0365}.
+
+45. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
+clades of Agaricales. P Roy Soc B 279: 2003--2011.
+doi:\href{http://dx.doi.org/10.1098/rspb.2011.2428}{10.1098/rspb.2011.2428}.
+
+46. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
+Interface. Ann Mo Bot Gard 74: 359--397.
+
+47. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand\~ao KL, Whinnett A, et
+al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
+Nymphalidae) inferred from one mitochondrial and two nuclear gene
+regions. Syst Entomol 31: 288--301.
+doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2006.00321.x}{10.1111/j.1365-3113.2006.00321.x}.
+
+48. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
+trait-dependent speciation and extinction rates from incompletely
+resolved phylogenies. Syst Biol 58: 595--611.
+doi:\href{http://dx.doi.org/10.1093/sysbio/syp067}{10.1093/sysbio/syp067}.
+
+49. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
+The genetic basis of a plant-insect coevolutionary key innovation. P
+Natl Acad Sci USA 104: 20427--20431.
+doi:\href{http://dx.doi.org/10.1073/pnas.0706229104}{10.1073/pnas.0706229104}.
+
+50. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
+furanocoumarin-metabolizing cytochrome P450 monooxygenases in two
+papilionids: Specificity and substrate encounter rate. P Natl Acad Sci
+USA 100: 14593--14598.
+doi:\href{http://dx.doi.org/10.1073/pnas.1934643100}{10.1073/pnas.1934643100}.
+
+51. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
+the angiosperms re-revisited. Am J Bot 97: 1296--1303.
+doi:\href{http://dx.doi.org/10.3732/ajb.0900346}{10.3732/ajb.0900346}.
+
+52. Wahlberg N, Rubinoff D (2011) Vagility across Vanessa (Lepidoptera:
+Nymphalidae): mobility in butterfly species does not inhibit the
+formation and persistence of isolated sister taxa. Syst Entomol 36:
+362--370.
+doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2010.00566.x}{10.1111/j.1365-3113.2010.00566.x}.
+
 \end{document}

### 2015-01-14 MS_orig.tex
ready for resubmittion
revision 2

@@
 {\Large
-\title{Diversity dynamics in Nymphalidae butterflies: Effect of
-phylogenetic uncertainty on diversification rate shift estimates}
+    \title{Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates}
 }
@@
 Hostplant shifts have been invoked to be responsible for a great part of
-the biodiversity of herbivorous insects {[}1{]}. The study of the
+the biodiversity of herbivorous insects \cite{mitter1988}. The study of the
 evolution of hostplant use has spawned several theories explaining the
-evolutionary interactions between plants and insects {[}2{]}. The
-``escape-and-radiate'' hypothesis {[}3{]}, the ``oscillation
-hypothesis'' {[}4,5{]} or ``diffuse cospeciation'' {[}2{]}.
+evolutionary interactions between plants and insects \cite{nyman2012}. The
+``escape-and-radiate'' hypothesis \cite{ehrlich1964}, the ``oscillation
+hypothesis'' \cite{janz2011, nylin2014} or ``diffuse cospeciation'' \cite{nyman2012} and the
+``musical chairs hypothesis'' \cite{hardy2014}.
 
@@
 developing some of the mentioned hypotheses. Nymphalidae contains around
-6000 species {[}6{]}, and is the largest family within the true
-butterflies. The family most likely originated around 94 MYA in the mid
-Cretaceous. Diversification of the group began in the Late Cretaceous
-and most major radiations (current subfamilies) appeared shortly after
-the Cretaceous-Paleogene (K-Pg) boundary {[}7{]}. Several studies have
-used time-calibrated phylogenies and diversification models to
-reconstruct the evolutionary history of the group to identify patterns
-of accelerated or decelerated diversification of some Nymphalidae clades
-{[}7--10{]}. For example, it has been suggested that climate change in
-the Oligocene and the subsequent diversification of grasses has led to
-diversification of the subfamily Satyrinae {[}11{]} due to the abundance
-of grasses over extensive geographic areas (``resource
-abundance-dependent diversity dynamics'' hypothesis). Reference 9 found
-increased net diversification rates in some Nymphalidae lineages after a
-major hostplant shift, which appears to be in agreement with Reference 3
-``escape-and-radiate'' model of diversification.
+6000 species \cite{van_nieukerken2011}, and several members are considered model organisms
+in evolutionary biology \cite{joron2006, willmott2006, brakefield2009}.
+The family most likely originated
+around 94 MYA in the mid Cretaceous. Diversification of the group began
+in the Late Cretaceous and most major radiations (current subfamilies)
+appeared shortly after the Cretaceous-Paleogene (K-Pg) boundary
+\cite{heikkila2012}. Several studies have used time-calibrated phylogenies and
+diversification models to reconstruct the evolutionary history of the
+group to identify patterns of accelerated or decelerated diversification
+of some Nymphalidae clades \cite{heikkila2012, elias2009, fordyce2010, wahlberg2009}.
+For example, it has been
+suggested that climate change in the Oligocene and the subsequent
+diversification of grasses has led to diversification of the subfamily
+Satyrinae \cite{pena2008} due to the abundance of grasses over extensive
+geographic areas (``resource abundance-dependent diversity dynamics''
+hypothesis): Fordyce (2010) \cite{fordyce2010} found increased net diversification
+rates in some Nymphalidae lineages after a major hostplant shift, which
+appears to be in agreement with the ``escape-and-radiate'' model of
+diversification \cite{ehrlich1964}.
 
@@
 diversification pattern of Nymphalidae is congruent with events of
-sudden diversification bursts due to hostplant shift, climatic events or
-shifts to closely related hostplants {[}5,12{]}.
+sudden diversification bursts due to hostplant shift or climatic events
+\cite{nylin2014, ferrer2013}.
 
 In this study, we used a time-calibrated genus-level phylogenetic
-hypothesis for Nymphalidae butterflies {[}10{]} to investigate patterns
-of diversification. We applied MEDUSA {[}13,14{]}, a recently developed
-statistical method, to study the diversification pattern of Nymphalidae
+hypothesis for Nymphalidae butterflies \cite{wahlberg2009} to investigate patterns
+of diversification. We applied the statistical method MEDUSA
+\cite{alfaro2009, harmon2011}, to study the diversification pattern of Nymphalidae
 butterflies. MEDUSA fits likelihood models of diversification into a
@@
 study the effects of phylogenetic uncertainty and by using the extended
-MEDUSA method called MultiMEDUSA {[}13{]}. We also tested whether
+MEDUSA method called MultiMEDUSA \cite{alfaro2009}. We also tested whether
 hostplant association dynamics can explain the diversification patterns
@@
 employing the method BiSSE as implemented in the R package
-\texttt{diversitree} {[}15{]}.
+\texttt{diversitree} \cite{fitzjohn2012}.
 
-\section*{Methods}
+\section*{Materials and Methods}
 
@@
 
-For analyses, we used the phylogenetic trees from the study of Reference
-10 that were generated using DNA sequence data from 10 gene regions for
-398 of the 540 valid genera in Nymphalidae. We employed Reference 10
-maximum clade credibility tree (MCC tree) (Fig. 1) as well as a random
-sample of 1000 trees from their BEAST run after burnin. Their original
-BEAST run was for 40 million generations. We used a burnin of 25 million
-generations and took a random sample of 1000 trees using Burntrees
-v.0.1.9 \url{http://www.abc.se/~nylander/} (supp. mat. 5) in order to
-correct for phylogenetic uncertainty when performing the diversification
-analyses.
+For analyses, we used the phylogenetic trees from the study of Wahlberg
+et al. (2009) \cite{wahlberg2009} that were generated using DNA sequence data from
+10 gene regions for 398 of the 540 valid genera in Nymphalidae. We
+employed the maximum clade credibility tree (MCC tree) (Fig. 1) as well
+as a random sample of 1000 trees from their BEAST run after burnin
+\cite{wahlberg2009}. Their original BEAST run was for 40 million generations. We
+used a burnin of 25 million generations and took a random sample of 1000
+trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (supp.
+mat. 5) in order to correct for phylogenetic uncertainty when performing
+the diversification analyses.
 
@@
 sources including the specialist-curated lists on
-\url{http://tolweb.org}, Reference 16 and curated lists of Global
-Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}). We
-assigned the species numbers of genera not included in the phylogeny to
-the closest related genus that was included in Reference 10 study
-according to available phylogenetic studies {[}17--28{]}, taxonomical
-classification and morphological resemblance when no phylogenies were
-available (supp. mat. 3).
+\url{http://tolweb.org}, Lamas (2004) \cite{lamas2004} and curated lists of
+Global Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}).
+We assigned the species numbers of genera not included in the phylogeny
+to the closest related genus that was included in Wahlberg et al. (2009)
+\cite{wahlberg2009} study according to available phylogenetic studies 
+\cite{matos2013, brower2010, kodandaramaiah2010, kodandaramaiah2010a,
+ortiz2013, desilva2010, freitas2004, pena2006, penz1999, silva2008,
+pena2011, pena2010},
+taxonomical classification and morphological resemblance when no
+phylogenies were available (supp. mat. 3).
 
 Hostplant data for Nymphalidae species were compiled from several
-sources including Reference 29, HOSTS database
-(\url{http://bit.ly/YI7nwW}), Reference 30, Reference 31, Reference 32
-and others (supp. mat. 15--16) for a total of 6586 hostplant records,
-including 428 Nymphalidae genera and 143 plant families and 1070 plant
-genera. It was not possible to find any hostplant data for 35 butterfly
-genera.
+sources including \cite{ackery1988}, HOSTS database (\url{http://bit.ly/YI7nwW}),
+\cite{dyer2002, beccaloni2008, janzen2009} and others (supp. mat. 15--16) for a total of 6586
+hostplant records, including 428 Nymphalidae genera and 143 plant
+families and 1070 plant genera. It was not possible to find any
+hostplant data for 35 butterfly genera.
 
-\subsection*{Analyses of
-Diversification}
+\subsection*{Analyses of Diversification}
 
-We used the statistical software R version 3.0.1 {[}33{]} in combination
-with the APE {[}34{]}, GEIGER {[}35{]}, MEDUSA {[}13{]} and
-\texttt{diversitree} {[}15{]} packages along with our own scripts to
+We used the statistical software R version 3.0.1 \cite{r2013} in combination
+with the APE \cite{popescu2012}, GEIGER \cite{harmon2008}, MEDUSA
+\cite{alfaro2009} and
+\texttt{diversitree} \cite{fitzjohn2012} packages along with our own scripts to
 perform the analyses (included as supplementary materials 2, 6, 10). All
-analyses were run on the 1000 random trees from Reference 10, on the
-maximum clade credibility tree derived from these and the MCC tree from
-Reference 10.
+analyses were run on the 1000 random trees from Wahlberg et al. (2009)
+\cite{wahlberg2009}, on the maximum clade credibility tree derived from these and
+the MCC tree \cite{wahlberg2009}.
 
-\subsection*{Detecting diversification shifts on phylogenetic
-trees}
+We analyzed our MCC tree in the Bayesian software BAMM in order to
+contrast the results found by MEDUSA and MultiMEDUSA. BAMM (Bayesian
+analysis of macroevolutionary mixtures) \cite{rabosky2014} is a software that fits
+several models of speciation and extinction. The best model can be
+selected by comparison of Bayes Factors.
+
+\subsection*{Detecting diversification shifts on phylogenetic trees}
 
 Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
-version 093.4.33 {[}13{]} on the maximum clade credibility tree from
-Reference 10. MEDUSA fits alternative birth-death likelihood models to a
-phylogenetic tree in order to estimate changes in net diversification
-rates along branches. MEDUSA estimates likelihood and AIC scores for the
-simplest birth-death model, with two parameters, the rates \texttt{r}:
-net diversification \texttt{=} speciation (\texttt{b}) \texttt{-}
-extinction (\texttt{d}) and \(\varepsilon\): relative extinction
-\texttt{= d/b}. The AIC scores of the two-parameter model are then
-compared with incrementally more complex models until the addition of
-parameters do not improve the AIC scores beyond a cutoff value. MEDUSA
-finds the likelihood of the models after taking into account branch
-lengths and number of species per lineage {[}13{]}. Most studies using
-MEDUSA only run the method on a single tree, usually the maximum clade
-credibility tree, which makes the assumption that this tree is correct.
-We wanted to study the effects of phylogenetic uncertainty on estimation
-of net diversification rate shifts and therefore selected 1000 random
-genus-level trees from the posterior distribution (MultiMEDUSA, supp.
-mat. 5). We calculated a new MCC tree from the selection of trees, ran
-MEDUSA and MultiMEDUSA on the selection of 1000 trees and summarized the
-estimated changes in net diversification rates for nodes across all
-trees. Patterns of change in net diversification rates are considered
-significant if they are found at the same node in at least 90\% of the
-trees. We also expected to find similar values of net diversification
-(\texttt{r = b - d}) and relative extinction rate (\(\varepsilon\)
-\texttt{= d/b}) values across the 1000 trees for the nodes where changes
-in diversification tempo occurs. We used the AICc threshold of 7.8
-units, as estimated by MEDUSA, as the limit for a significantly better
-fit to select among increasingly complex alternative models.
+version 093.4.33 \cite{alfaro2009} on the maximum clade credibility tree from
+Wahlberg et al. (2009) \cite{wahlberg2009}. MEDUSA fits alternative birth-death
+likelihood models to a phylogenetic tree in order to estimate changes in
+net diversification rates along branches. MEDUSA estimates likelihood
+and AIC scores for the simplest birth-death model, with two parameters,
+the rates \texttt{r}: net diversification \texttt{=} speciation
+(\texttt{b}) \texttt{-} extinction (\texttt{d}) and (\(\varepsilon\):
+relative extinction \texttt{= d/b}. The AIC scores of the two-parameter
+model are then compared with incrementally more complex models until the
+addition of parameters do not improve the AIC scores beyond a cutoff
+value. MEDUSA finds the likelihood of the models after taking into
+account branch lengths and number of species per lineage \cite{alfaro2009}. Most
+studies using MEDUSA only run the method on a single tree, usually the
+maximum clade credibility tree, which makes the assumption that this
+tree is correct. We wanted to study the effects of phylogenetic
+uncertainty on estimation of net diversification rate shifts and
+therefore selected 1000 random genus-level trees from the posterior
+distribution of the Bayesian run from Wahlberg et al., 2009 \cite{wahlberg2009}
+(MultiMEDUSA, supp. mat. 5). We calculated a new MCC tree from the
+selection of trees, ran MEDUSA and MultiMEDUSA on the selection of 1000
+trees and summarized the estimated changes in net diversification rates
+for nodes across all trees. Patterns of change in net diversification
+rates are considered significant if they are found at the same node in
+at least 90\% of the trees. We also expected to find similar values of
+net diversification (\texttt{r = b - d}) and relative extinction rate
+(\(\varepsilon\) \texttt{= d/b}) values across the 1000 trees for the
+nodes where changes in diversification tempo occurs. We used the AICc
+threshold of 7.8 units, as estimated by MEDUSA, as the limit for a
+significantly better fit to select among increasingly complex
+alternative models.
 
@@
 diversification in the clade Ithomiini, we tested whether this pattern
-can be explained by increase in birth-rate due to hostplant use and
-performed analyses with the ``binary state speciation and extinction''
-{[}36{]} Bayesian approach as implemented in the R package
-\texttt{diversitree} {[}15{]}. MuSSE {[}15{]} is designed to examine the
-joint effects of two or more traits on speciation. Because most of
-Nymphalidae butterflies are restricted to use one plant family as
-hostplant, the character states can be coded as presence/absence, for
-which the BiSSE analysis is better suited. BiSSE was designed to test
-whether a binary character state has had any effect on increased net
-diversification rate for a clade {[}36{]}. We used our compiled data of
+can be explained by an increase in the birth-rate due to hostplant use
+and performed analyses using the ``binary state speciation and
+extinction'' \cite{maddison2007} Bayesian approach as implemented in the R package
+\texttt{diversitree} \cite{fitzjohn2012}. The Markov Chain Monte Carlo algorithm
+was run for 10000 generations discarding the first 7500 as burnin. The
+Multiple State Speciation and Extinction approach (MuSSE) \cite{fitzjohn2012} is
+designed to examine the joint effects of two or more traits on
+speciation. Because most of Nymphalidae butterflies are restricted to
+use one plant family as hostplant (232 genera use only one plant family
+versus 176 that use more than one; based on our data in supp. mat. 15),
+the character states can be coded as presence/absence, for which the
+BiSSE analysis is better suited. BiSSE was designed to test whether a
+binary character state has had any effect on increased net
+diversification rate for a clade \cite{maddison2007}. We used our compiled data of
 hostplant use to produce a binary dataset for the character ``feeding on
 the plant family Solanaceae'' with two states (absence = 0; presence =
-1) (supp. mat. 17), which is the main hostplants of the diverse
-Ithomiini butterflies and closest relatives {[}37{]}. Other hostplant
-shifts were not tested for effect on net diversification rates due to
-the low support for diversification shifts found in the MEDUSA analyses.
-We analyzed the data using BiSSE employing the Markov Chain Monte Carlo
-algorithm on the maximum clade credibility tree, taking into account
-missing taxa by using the parameter ``sampling factor''
+1) (supp. mat. 17). The family Solanaceae include the hostplants of the
+diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. Other
+hostplant shifts were not tested for effect on net diversification rates
+due to the low support for diversification shifts found in the MEDUSA
+analyses. We analyzed the data using BiSSE employing the Markov Chain
+Monte Carlo algorithm on the maximum clade credibility tree, taking into
+account missing taxa by using the parameter ``sampling factor''
 (\texttt{sampling.f}) in \texttt{diversitree}. We also used constrained
@@
 hypothesis (no effect). The analyses were run across a sample of 250
-trees from the posterior distribution. The records of \emph{Vanessa} and
-\emph{Hypanartia} feeding on Solanaceae {[}30,38{]} might be incorrect
-as it is unlikely that these species can be feeding on this plant
-family. We tested whether coding these two genera as not feeding on
-Solanaceae affected our results.
+trees from the selected 1000 random trees from the posterior
+distribution. The records of \emph{Vanessa} and \emph{Hypanartia}
+feeding on Solanaceae \cite{beccaloni2008, scott1986} might be incorrect as it is unlikely
+that these species can be feeding on this plant family. We tested
+whether coding these two genera as not feeding on Solanaceae affected
+our results. We expected that coding these two genera as either absence
+or presence would not have a significant effect on our results. It has
+been shown that BiSSE performs poorly under certain conditions \cite{davis2013}.
+However, our data has adequate number of taxa under analysis (more than
+300 tips), adequate speciation bias (between 1.5x and 2.0x), character
+state bias (around 8x) and extinction bias (around 4x) for the analysis
+of Solanaceae hostplants. Thus, BiSSE is expected to produce robust
+results \cite{davis2013}.
 
@@
 The MEDUSA analysis on the MCC tree in combination with richness data
-estimated 18 changes in the tempo of diversification in Nymphalidae
-history (Fig. 1; Table 1). The corrected AICc acceptance threshold for
-adding subsequent piecewise birth-death processes to the overall model
-was set to 7.8 units, as prescribed by MEDUSA. In all MEDUSA analyses,
-the maximum number of inferred diversification splits in all trees was
-26. The background net diversification rate for Nymphalidae was
-estimated as \texttt{r = 0.092} lineages per Million of years and the
-AICc score for the best fit model was \texttt{5090.5} (Table 1). MEDUSA
-also estimated that the basic constant birth--death model was not a
-better explanation for our data (AICc \texttt{= 5449.3}).
+estimated 18 significant changes in the tempo of diversification in
+Nymphalidae history (Fig. 1; Table 1). The corrected AICc acceptance
+threshold for adding subsequent piecewise birth-death processes to the
+overall model was set to 7.8 units, as prescribed by MEDUSA. In all
+MEDUSA analyses, the maximum number of inferred diversification splits
+in all trees was 26. The background net diversification rate for
+Nymphalidae was estimated as \texttt{r = 0.092} lineages per Million of
+years and the AICc score for the best fit model was \texttt{5090.5}
+(Table 1). MEDUSA also estimated that the basic constant birth-death
+model was not a better explanation for our data (AICc
+\texttt{= 5449.3}).
 
@@
 
+We contrasted these results with those obtained from analyzing the data
+in the software BAMM \cite{rabosky20140}. We found that the best model (best
+configuration) estimated 2 diversification shifts in our MCC tree
+(posterior probability = 0.79) for the Ithomiini and Satyrini (Fig. S1).
+Bayes factor (BF) values strongly favored this model in comparison with
+a constant rate model (no diversification shifts, BF \(>\) 1,884).
+
 \subsection*{Phylogenetic uncertainty in the MultiMEDUSA
@@
 
-We tested MEDUSA to find out whether taking into account the
-phylogenetic signal from a random sample of 1000 trees from the
-posterior distribution can return similar estimates of diversification
-to the values obtained from the MCC tree. We run MultiMEDUSA on a random
+We used MEDUSA to find out whether taking into account the phylogenetic
+signal from the random sample of 1000 trees from the posterior
+distribution can return similar estimates of diversification to the
+values obtained from the MCC tree. We ran MultiMEDUSA on the random
 sample of 1000 trees (supp. mat. 5) from the posterior distribution and
@@
 lower median net diversification rates for the diversification shifts
-found by MEDUSA on the MCC tree from Reference 10 (Table 2). Although
-the diversification pattern found by MEDUSA and MultiMEDUSA was the
-same, the latter consistently estimated lower rates. Thus, the shifts
-recovered with low net diversification rate on the MCC were recovered
-with negative net diversification rate by MultiMEDUSA. The background
-diversification and all shifts found by MEDUSA on the 1000 trees are
-provided as an R object in supp. mat. 7).
+found by MEDUSA on the MCC tree derived from the random sample of trees
+(Table 2). Although the diversification pattern found by MEDUSA and
+MultiMEDUSA was the same, the latter consistently estimated lower rates.
+Furthermore, the shifts recovered with low net diversification rate on
+the MCC were recovered with negative net diversification rate by
+MultiMEDUSA. The background diversification and all shifts found by
+MEDUSA on the 1000 trees are provided as an R object in supp. mat. 7.
 
@@
 picked up by MEDUSA and a node shift was found most of the time, then
-the \texttt{sum.prop} should be close to 1. For example the node 625
-consists of the clade \texttt{Charaxes} + \texttt{Polyura} and it was
-found in only 256 trees, however MultiMEDUSA was consistenly able to
-find a diversification shift for that node and the \texttt{sum.prop}
-value is 0.996.
+the \texttt{sum.prop} should be close to 1. For example, the
+\texttt{Charaxes} + \texttt{Polyura} clade was found in only 256 trees,
+however MultiMEDUSA was consistently able to find a diversification
+shift for that node and the \texttt{sum.prop} value is 0.996.
 
-We found different rates for diversification shifts even for the nodes
-that appear in most of the sample of 1000 trees (Table 2) as many of the
-nodes also found in the MCC tree had a probability of being recovered
-higher than 0.90.
+For the diversification shifts found in both the MCC tree and most of
+the samples of 1000 trees (frequency more than 90\%; Table 2), the
+MultiMEDUSA approach recovered different rates of diversification than
+those found when the MCC tree alone.
 
-There were four diversification shifts found with more than 90\% of
-probability in the trees from the random sample (Table 2): (i) net
-diversification rate increase in the genus \emph{Ypthima}; (ii) net
-diversification rate increase in the genus \emph{Charaxes}; (iii) rate
+There were four diversification shifts found in the trees from the
+random sample (Table 2): (i) net diversification rate increase in the
+genus \emph{Ypthima} (\(r = 0.22\)); (ii) net diversification rate
+increase in the genus \emph{Charaxes} (\(r = 0.21\)); (iii) rate
 increase in Ithomiini subtribes Oleriina + Ithomiina + Napeogenina +
-Dircennina + Godyridina; and (iv) rate increase in \emph{Callicore} +
-\emph{Diaethria}.
+Dircennina + Godyridina (\(r = 0.10\)); and (iv) rate increase in
+\emph{Callicore} + \emph{Diaethria} (\(r = 0.135\)).
 
 MultiMEDUSA provided mean and standard deviation statistics for the
-diversification values found on the shifts on the 1000 trees (supp. mat.
+diversification values on the shifts on the 1000 trees (supp. mat.
 13--14), and found that some of the changes in net diversification rate
@@
 of the splits in the MCC tree are recovered in very few trees, for
-example the split for the clade Satyrini (Euptychiina + Pronophilina +
-Satyrina + Maniolina) is recovered with a probability of 18\% (Fig. 3).
+example the split for the clade Satyrinae is recovered with a
+probability of 0.18 (Fig. 3).
 
@@
 
-The MEDUSA analyses taking into account phylogenetic uncertainty
+The MEDUSA analyses, taking into account phylogenetic uncertainty,
 estimated a net diversification rate increase in part of the clade
@@
 net net diversification rate on part of Ithomiini (Oleriina + Ithomiina
-+ Napeogenina + Dircennina + Godyridina) (Fig. 4). The Markov Chain
-Monte Carlo algorithm was run for 10000 generations discarding the first
-7500 as burnin. The estimated mean net diversification rate for taxa
-that do not feed on Solanaceae was \texttt{r = 0.11} while the net
-diversification rate for the Solanaceae feeders was \texttt{r = 0.16}
-(see Fig. S1 for a boxplot of speciation and extinction values for the
-95\% credibility intervals). The same analysis considering
-\emph{Vanessa} and \emph{Hypanartia} as non-Solanaceae feeders due to
-dubious records produced the same pattern and net diversification rates
-(Fig. S2). Therefore, the rest of the analyses were performed assuming
-these two genera as Solanaceae feeders. We constrained the BiSSE
-likelihood model to force equal rates of speciation for both character
-states in order to test whether the model of different speciation rates
-is a significantly better explanation for the data. A likelihood ratio
-test found that the model for increased net diversification rate for
-nymphalids feeding on Solanaceae is a significantly better explanation
-than this character state having no effect on diversification
-(\texttt{p \textless{} 0.001}) (Table 3; character states available in
-supp. mat. 17, code in supp. mat. 18, and mcmc run in supp. mat. 19). We
-combined the post-burnin mcmc generations from running BiSSE on 250
-trees from the posterior distribution and found the same pattern as the
-BiSSE analysis on the maximum clade credibility tree (combined mcmc run
-in supp. mat. 20; profiles plot of speciation rates in Fig. S3; boxplot
-of 95\% credibility intervals in Fig. S4). A BiSSE analysis to test
-whether the trait ``feeding on Apocynaceae'' had any effect on increased
-net diversification rates found similar speciation rates for lineages
-feeding on Apocynaceae and other plants (Fig. S5). It has been shown
-that BiSSE performs poorly under certain conditions {[}39{]}. However,
-our data has adequate number of taxa under analysis (more than 300
-tips), adequate speciation bias (between 1.5x and 2.0x), character state
-bias (around 8x) and extinction bias (around 4x) for the analysis of
-Solanaceae hostplants. Thus, BiSSE is expected to produce robust results
-{[}39{]}.
++ Napeogenina + Dircennina + Godyridina) (Fig. 4). The estimated mean
+net diversification rate for taxa that do not feed on Solanaceae was
+\texttt{r = 0.11} while the net diversification rate for the Solanaceae
+feeders was \texttt{r = 0.16} (see Fig. S2 for a boxplot of speciation
+and extinction values for the 95\% credibility intervals). The same
+analysis considering \emph{Vanessa} and \emph{Hypanartia} as
+non-Solanaceae feeders due to dubious records produced the same pattern
+and net diversification rates (Fig. S3). Therefore, the rest of the
+analyses were performed assuming these two genera as Solanaceae feeders.
+We constrained the BiSSE likelihood model to force equal rates of
+speciation for both character states in order to test whether the model
+of different speciation rates is a significantly better explanation for
+the data. A likelihood ratio test found that the model for increased net
+diversification rate for nymphalids feeding on Solanaceae is a
+significantly better explanation than this character state having no
+effect on diversification (\(\chi^2 = 12.3; 1 df; p < 0.001\)) (Table 3;
+character states available in supp. mat. 17, code in supp. mat. 18, and
+mcmc run in supp. mat. 19). We combined the post-burnin mcmc generations
+from running BiSSE on 250 trees from the random sample of 1000 trees and
+found the same pattern as the BiSSE analysis on the maximum clade
+credibility tree (combined mcmc run in supp. mat. 20; profiles plot of
+speciation rates in Fig. S4; boxplot of 95\% credibility intervals in
+Fig. S5). A BiSSE analysis to test whether the trait ``feeding on
+Apocynaceae'' had any effect on increased net diversification rates
+found similar speciation rates for lineages feeding on Apocynaceae and
+other plants (Fig. S6).
 
@@
 The MEDUSA method has been used to infer changes in net diversification
-rates along a phylogenetic tree. Since its publication {[}13{]} the
-results of using MEDUSA on a single tree, the maximum clade credibility
-tree, have been used for generation of hypotheses and discussion
-{[}7,40,41{]}. However, we found that MEDUSA estimated different
-diversification shifts and different rates of diversification for
-certain lineages when phylogenetic uncertainty was taken into account by
-using MEDUSA on a random sample of trees from the posterior distribution
-of a Bayesian run. We found that some diversification splits, estimated
-on the Nymphalidae maximum clade credibility tree, were found with a
-very low probability in a random sample of 1000 trees from the posterior
-distribution (Fig. 3, Table 2). We also found that, even though MEDUSA
-could estimate the same diversification splits on two or more trees, the
-estimated net diversification rates could vary widely (Fig. 2). For
+rates in a phylogenetic tree. Since its publication \cite{alfaro2009} the results
+of using MEDUSA on a single tree, the maximum clade credibility tree,
+have been used for generation of hypotheses and discussion
+\cite{heikkila2012, litman2011, ryberg2012}. However, different diversification shifts and different
+rates of diversification are found for certain lineages when
+phylogenetic uncertainty was taken into account by using MEDUSA on a
+random sample of trees from the posterior distribution of a Bayesian
+run. We found that some diversification splits, estimated on the
+Nymphalidae maximum clade credibility tree, were found with very low
+probability in the random sample of 1000 trees from the posterior
+distribution (Fig. 3, Table 2). We also found that, even though the
+analyses estimated the same diversification splits on two or more trees,
+the estimated net diversification rates could vary widely (Fig. 2). For
 example, in our Nymphalidae trees, we found that the split for
@@
 taxa appear to be strongly affected by long-branch attraction artifacts
-{[}27{]}. Thus, the Bayesian runs are expected to recover alternative
+\cite{pena2011}. Thus, the Bayesian runs are expected to recover alternative
 topologies on the posterior distribution of trees, resulting in low
 support and posterior probability values for the nodes. For example,
-posterior probability values for clades in Satyrini are very low
-{[}10{]}. As a result, MEDUSA inferred a net diversification rate
-increase for part of Satyrini in the maximum clade credibility tree, but
-this was recovered only in 13\% of probabilities in the MultiMEDUSA
-analysis on the random sample of trees.
+posterior probability values for clades in Satyrinae are very low
+\cite{wahlberg2009}. As a result, MEDUSA inferred a net diversification rate
+increase for the Satyrinae (which includes Satyrini and its sister
+clade) in the maximum clade credibility tree, but this was recovered in
+the MultiMEDUSA analysis in only 13\% of the random sample of trees.
 
@@
 The clade Ithomiini and the non-basal danaids are well supported by high
-posterior probability values in Reference 10. Therefore our MEDUSA
-analyses recovered an increase in net diversification rate with
-probability higher than 90\% in the posterior distribution of trees
-(Fig. 3).
+posterior probability values in Wahlberg et al. (2009) \cite{wahlberg2009}.
+Therefore our MEDUSA analyses recovered an increase in net
+diversification rate with probability higher than 0.90 in the posterior
+distribution of trees (Fig. 3).
+
+The results from BAMM recovered only two significant shifts in our MCC
+tree as this method seems to be much more conservative than the MEDUSA
+approach. One of these shifts is the Satyrini clade. MEDUSA estimated
+its parent node as significant shift in only a few trees from the
+posterior distribution. Thus, it might be important to take into account
+phylogenetic uncertainty in BAMM as well.
 
@@
 
-\paragraph*{Ithomiini}
+\subsubsection*{Ithomiini}
 
 Keith Brown suggested that feeding on Solanaceae was an important event
-in the diversification of Ithomiini butterflies {[}42{]}. Ithomiini
+in the diversification of Ithomiini butterflies \cite{brown1987}. Ithomiini
 butterflies are exclusively Neotropical and most species feed on
-Solanaceae hostplants during larval stage {[}37{]}. Optimizations of the
+Solanaceae hostplants during larval stage \cite{willmott2006}. Optimizations of the
 evolution of hostplant use on phylogenies evidence a probable shift from
-Apocynaceae to Solanaceae in the ancestor of the tribe {[}37,43{]}.
-Reference 9 found that the Gamma statistics, a LTT plot of an Ithomiini
-phylogeny and the fit of the density-dependent model of diversification
-are consistent with a burst of diversification in Ithomiini following
-the shift from Apocynaceae to Solanaceae.
+Apocynaceae to Solanaceae in the ancestor of the tribe \cite{willmott2006, brower2006}.
+Fordyce (2010) \cite{fordyce2010} found that the Gamma statistics, a LTT plot of
+an Ithomiini phylogeny and the fit of the density-dependent model of
+diversification are consistent with a burst of diversification in
+Ithomiini following the shift from Apocynaceae to Solanaceae.
 
@@
 due to the use of Solanaceae plants as hosts during larval stage. For
-this, we used a Bayesian approach {[}44{]} to test whether the trait
+this, we used a Bayesian approach \cite{fitzjohn2009} to test whether the trait
 ``feeding on Solanaceae'' had any effect on the diversification of the
@@
 trait ``feeding on Solanaceae hostplants'' (Fig. 4). This is in
-agreement with the findings of Reference 9 using other statistical
-methods. Due to the fact that Ithomiini are virtually the only
+agreement with previous findings using other statistical methods
+\cite{fordyce2010}. Due to the fact that Ithomiini are virtually the only
 nymphalids using Solanaceae as hostplants, it is possible that the trait
 responsible for a higher diversification of Ithomiini might not be the
-hostplant character. As noted by Reference 36, the responsible trait
-might be a codistributed character such as a trait related to the
-ability to digest secondary metabolites.
+hostplant character. As noted by Maddison et al. (2007) \cite{maddison2007}, the
+responsible trait might be a codistributed character such as a trait
+related to the ability to digest secondary metabolites.
 
@@
 that the high diversity of Ithomiini is consistent with the
-``escape-and-radiate scenario'' due to a shift onto Solanaceae {[}9{]}
+``escape-and-radiate scenario'' due to a shift onto Solanaceae
+\cite{fordyce2010}
 and radiation scenarios among chemically different lineages of
-Solanaceae plants {[}37,42{]}. According to this theory, the shift from
+Solanaceae plants \cite{willmott2006, brown1987}. According to this theory, the shift from
 Apocynaceae to Solanaceae allowed Ithomiini to invade newly available
@@
 detoxification of ingested food as they have been found in other
-butterfly groups {[}45{]}.
+butterfly groups \cite{wheat2007}.
 
-The increase in diversification inferred by MEDUSA ocurred after the
+The increase in diversification inferred by MEDUSA occurred after the
 probable shift from Apocynaceae to Solanaceae, as the Solanaceae feeders
@@
 Although the Solanaceae genera used by the Ithomiini clades are well
-known {[}37{]}, we do not have any understanding on the physiological
+known \cite{willmott2006}, we do not have any understanding on the physiological
 routes involved in the detoxification of Solanaceae compounds by the
 several lineages of Ithomiini. We can speculate that older lineages
-exploiting a novel toxic resource {[}10,37{]} might not be too efficient
+exploiting a novel toxic resource \cite{willmott2006, wahlberg2009}  might not be too efficient
 in metabolizing plant toxins and that younger lineages are able to deal
@@
 configurations of the proteins, providing more opportunities for
-hostplant switch {[}46{]}. This might be the reason why the basal
+hostplant switch \cite{li2003}. This might be the reason why the basal
 Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and
-restricted to few Solanaceae hosts {[}37{]}, while recent subtribes are
+restricted to few Solanaceae hosts \cite{willmott2006}, while recent subtribes are
 species-rich and have expanded their host range into several Solanaceae
-lineages {[}37{]}. It might be that the switch to feeding in Solanaceae
+lineages \cite{willmott2006}. It might be that the switch to feeding in Solanaceae
 was an important event in the evolutionary history of Ithomiini, but the
@@
 diversity'' and the ``escape-and-radiate'' hypotheses state that insects
-diversify after their hostplants {[}2--4{]}. Reference 45 found strong
-evidence for a model of speciation congruent with Ehrlich and Raven's
-hypothesis in Pieridae butterflies due to, in addition to the
-identification of a key innovation, a burst of diversification in
-glucosinolate-feeding taxa shortly afterwards (with a lag of
-\textasciitilde{}10 MY). According to a recent dated phylogeny of the
-Angiosperms {[}47{]}, the family Solanaceae split from its sister group
-about 59 (49-68) MYA and diversification started (crown group age)
-around 37 (29-47) MYA. Reference 10 give the corresponding ages for
-Ithomiini as 45 (39-53) and 37 (32-43) MYA, respectively. Thus, current
-evidence shows that Solanaceae and Ithomiini might have diversified
-around the same time, during the Late Eocene and Oligocene, and this
-would be congruent with the diffuse cospeciation hypothesis.
+diversify after their hostplants \cite{nyman2012, ehrlich1964, janz2011}. Wheat et al. (2007)
+\cite{wheat2007} found strong evidence for a model of speciation congruent with
+Ehrlich and Raven's hypothesis in Pieridae butterflies due to, in
+addition to the identification of a key innovation, a burst of
+diversification in glucosinolate-feeding taxa shortly afterwards (with a
+lag of circa 10 MY). According to a recent dated phylogeny of the
+Angiosperms \cite{bell2010}, the family Solanaceae split from its sister group
+about 59 (49--68) MYA and diversification started (crown group age)
+around 37 (29--47) MYA. Wahlberg et al. (2009) \cite{wahlberg2009} give the ages
+for origin and diversification for Ithomiini at 45 (39--53) and 37
+(32--43) MYA, respectively. Thus, current evidence shows that Solanaceae
+and Ithomiini might have diversified around the same time, during the
+Late Eocene and Oligocene, and this would be congruent with the diffuse
+cospeciation hypothesis.
 
-The uplift of the Andes was a major tectonic event that underwent higher
-activity during the Oligocene {[}48{]}. This caused climatic changes in
-the region that affected the flora and fauna of the time, which
-coincides with the diversification of modern montane plant and animal
-taxa {[}49{]} including Ithomiini butterflies and Solanaceae hostplants.
-Moreover, all Solanaceae clades currently present in New World
-originated in South America {[}50{]} as well as Ithomiini butterflies
-{[}10{]}. This suggests a process of ``diffuse cospeciation'' of
-Ithomiini and hostplants.
-
-\paragraph{Danaini}
-
-Our MultiMEDUSA approach gave a significant slowdown in net
+\subsubsection*{Danaini}
+Our MultiMEDUSA approach showed a significant slowdown in net
 diversification rate in the subtribe Danaina of the Danini. Both Danaina
@@
 widespread species show higher genetic differentiation in their
-populations {[}51{]}.
+populations \cite{wahlberg2011}.
 
-\paragraph{\emph{Charaxes}}
+\subsubsection*{Satyrinae}
 
-The genus \emph{Charaxes} contain 193 species distributed in the Old
-World with highest diversity in the Afrotropical region. These
-butterflies are also very strong fliers, but contrary to Danaina, which
-are specialized Apocynaceae feeders, \emph{Charaxes} are known to feed
-on at least 28 plant families in 18 orders {[}29{]} and some species
-appear to be polyphagous {[}52{]}. Reference 53 showed that most of the
-diversification of the genus occurred from the late Oligocene to the
-Miocene when there were drastic global climatic fluctuations, indicating
-that the diversity mainly is driven by climate change. Reference 52
-found that climatic changes during the Pliocene, Pleistocene as well as
-dispersal and vicariance might have been responsible for the high
-diversification of the genus.
+Lineages in the diverse family Satyrinae radiated simultaneously with
+the radiation of their main hostplant, grasses, during the climatic
+cooling in the Oligocene \cite{pena2008}. Thus, it is somewhat surprising that
+part of Satyrinae were found to have accelerated diversification in only
+13\% of the trees from the posterior distribution. Although this can be
+attributed to low phylogenetic signal \cite{wahlberg2009}, the clade Satyrini is
+very robust \cite{wahlberg2009} and MEDUSA failed to identify any significant
+accelerated net diversification rate for Satyrini.
+It appears that the
+radiation of Satyrini as a whole was not remarkably fast and therefore
+not picked up by MEDUSA, although it estimated a diversification shift
+for Satyrini + its sister clade. This should be expected if the
+diversification of Satyrini occurred in a stepwise manner, with pulses
+or bursts of diversification for certain lineages but unlikely for the
+tribe Satyrini as a whole.
 
-\paragraph{Satyrini}\label{satyrini}
-
-The diverse tribe Satyrini radiated simultaneously with the radiation of
-their main hostplant, grasses, during the climatic cooling in the
-Oligocene {[}11{]}. Thus, it is somewhat surprising that part of
-Satyrini (the subtribes Euptychiina, Satyrina and Pronophilina) were
-found to have accelerated diversification in only 13\% of probabilities
-in the trees from the posterior distribution. Although this can be
-attributed to low phylogenetic signal {[}10{]}, the clade Satyrini is
-very robust {[}10{]} and MEDUSA failed to identify any significant
-accelerated net diversification rate for Satyrini. It appears that the
-radiation of Satyrini as a whole was not remarkably fast and therefore
-not picked up by MEDUSA.
-
-The origin of the tribe Satyrini is not completely clear (originated
-either in the Neotropical or Eastern Palaearctic, Oriental and/or
-Indo-Australian regions {[}27{]} and their radiation involved colonizing
-almost all continents starting from their place of origin. For
-butterflies, a prerequisite for colonizing new areas is that suitable
-hostplants are already present. Therefore, it should be expected that
-the diversification of Satyrinae occurred in a stepwise manner, with
-pulses or bursts of diversification for certain lineages but unlikely
-for the tribe Satyrini as a whole.
-
+\section*{Conclusions}
 We found that even though MEDUSA estimated several diversification
@@
 diversified near simultaneously, which is in agreement with the diffuse
-co-speciation hypothesis {[}2,4{]}.
+cospeciation hypothesis \cite{nyman2012, janz2011}.
 
-\subsection*{Acknowledgments}\label{acknowledgments}
-
+\section*{Acknowledgments}
 We are thankful to Mark Cornwall for help with the script to extend
@@
 
-\subsection*{Figure legends}\label{figure-legends}
-
-\texttt{Figure 1.} Results of the MEDUSA analysis run on the maximum
-clade credibility tree from Reference 10. Rate shifts were estimated for
-the following nodes (besides the background rate): 2) Limenitidinae +
+\section*{Figure legends}
+{\bf Figure 1. Results of the MEDUSA analysis run on the maximum
+clade credibility tree from Wahlberg et al. (2009)} Rate shifts were estimated for the
+following nodes (besides the background rate): 2) Limenitidinae +
 Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
-part, 6) Satyrini, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
+part, 6) Satyrinae, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
 part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
 \emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
-Satyrina, 14) Mycalesinaa, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
-\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae.
+Satyrina, 14) Mycalesina, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
+\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae. Circles on
+nodes indicate the diversification shift number as found by MEDUSA.
+Numbers next to circles indicate the posterior probability values for
+such nodes.
 
-\texttt{Figure 2.} Boxplot of the range of diversification values for
-tips or clades estimated by MEDUSA on the 1000 random trees from the
-posterior distribution of the Nymphalidae phylogeny. The tips or clades
-shown are those present on the maximum clade credibility tree from
-Reference 10.
+{\bf Figure 2. Diversification rates for taxa estimated by MEDUSA on
+the samples of 1000 random trees}
 
-\texttt{Figure 3.} Results of the MultiMEDUSA analysis on 1000 randomly
-sampled trees from the posterior distribution of the Nymphalidae
-phylogeny. Bars show the probability for nodes of being recovered as
-significant increases or decreases in net diversification rates by
-MultiMEDUSA.
+{\bf Figure 3. Results of MultiMEDUSA analysis showing the
+probability of specific nodes being characterized by significant shifts
+in diversification rate}
 
-\texttt{Figure 4.} BiSSE analysis of diversification of nymphalids due
-to feeding on Solanaceae hostplants. Speciation and net diversification
-rates are significantly higher in Solanaceae feeders (\$lambda\$1, r1).
+{\bf Figure 4. BiSSE analysis of diversification of nymphalids due
+to feeding on Solanaceae hostplants} Speciation and net diversification
+rates are significantly higher in Solanaceae feeders (speciation rate =
+$\lambda1$, net diversification rate = $r1$.
 
-\subsection*{Tables}\label{tables}
+\section*{Supporting Information Legends}
+{\textbf{Supporting Information S01.nex.}} MCC Nymphalidae tree from
+\cite{wahlberg2009}
 
-\texttt{Table 1.} Significant net diversification rate shifts found in
-the MEDUSA analysis of the Nymphalid maximum clade credibility tree.
-Split.Node = node number, Model = preferred diversification model by
-MEDUSA, r = net diversification rate, LnLik.part = log likelihood value.
+{\textbf{Supporting Information S02.R}} R script to run MEDUSA on the
+MCC Nymphalidae tree from \cite{wahlberg2009}. This script removes the outgroup
+taxa and loads the richness data for the tree terminals.
 
-\texttt{Table 2.} Differences in rates estimated by MEDUSA on the MCC
-tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
-from the posterior distribution.
+{\textbf{Supporting Information S03.csv}} Species richness for lineages
+in Nymphalidae.
 
-\texttt{Table 3.} Likelihood ratio test between the model of increased
-diversification of nymphalids feeding on Solanaceae against a model
-forcing equal speciation rates (no effect on diversification).
+{\textbf{Supporting Information S04.txt}} Results of running MEDUSA on
+the MCC tree.
 
-\section*{Supp. mat.}\label{supp.-mat.}
+{\textbf{Supporting Information S05 1000 random trees no outgroups.zip}}
+1000 random trees from \cite{wahlberg2009}.
 
-\textbf{\texttt{supp mat 01 genus.nex}}: MCC Nymphalidae tree from
-Reference 10.
+{\textbf{Supporting Information S06.R}}~R script to run a MultiMEDUSA
+analysis on 1000 random trees from \cite{wahlberg2009}.
 
-\textbf{\texttt{supp mat 02 run medusa on mct.R}}: Run MEDUSA on
-MCC tree from Reference 10.
-
-\textbf{\texttt{supp mat 03 richness.nex}}: Species richness for
-lineages in Nymphalidae.
-
-\textbf{\texttt{supp mat 04 MEDUSA results.txt}}: Results of running
-MEDUSA on the MCC tree.
-
-\textbf{\texttt{supp mat 05 1000 random trees no outgroups.nex}}:
-1000 random trees from Reference 10.
-
-\textbf{\texttt{supp mat 06 run multimedusa.R}}: Run MultiMEDUSA on
-1000 random trees from Reference 10.
-
-\textbf{\texttt{supp mat 07 multimedusa on 1000 trees.txt}}: Raw
+{\textbf{Supporting Information S07 multimedusa on 1000 trees.zip}} Raw
 results from the MultiMEDUSA run on the random sample of trees from the
@@
 
-\textbf{\texttt{supp mat 08 multimedusa result01.pdf}}: Summary of
-MulitMEDUSA analysis on the 1000 trees from the posterior distribution.
+{\textbf{Supporting Information S08.pdf}} Summary of the MultiMEDUSA
+analysis on the 1000 trees from the posterior distribution.
 
-\textbf{\texttt{supp mat 09 mcc tree from random sample of 1000.tree}}:
-MCC tree from the 1000 random trees selected from the posterior
+{\textbf{Supporting Information S09.tree}} MCC tree from the 1000 random
+trees selected from the posterior distribution.
+
+{\textbf{Supporting Information S10.R}} R script to run a MEDUSA
+analysis on the MCC tree from the 1000 random trees selected from the
+posterior distribution.
+
+{\textbf{Supporting Information S11.txt}} Summary results from a MEDUSA
+run on the MCC tree from the 1000 random trees selected from the
+posterior distribution.
+
+{\textbf{Supporting Information S12.pdf}} Figure for MEDUSA run on MCC
+tree from random 1000 trees.
+
+{\textbf{Supporting Information S13.txt}} Summary results from a
+MultiMEDUSA run on the 1000 random trees selected from the posterior
 distribution.
 
-\textbf{\texttt{supp mat 10 run medusa on mct from 1000 trees.R}}:
-Run MEDUSA on MCC tree from the 1000 random trees selected from the
-posterior distribution.
-
-\textbf{\texttt{supp mat 11 results of MEDUSA run on mcc from 1000 trees.txt}}:
-Summary results from a MEDUSA run on the MCC tree from the 1000 random
-trees selected from the posterior distribution.
-
-\textbf{\texttt{supp mat 12 MEDUSA mcc tree from random sample of 1000 trees.pdf}}:
-Figure for MEDUSA run on MCC tree from random 1000 trees.
-
-\textbf{\texttt{supp mat 13 results of Multi-MEDUSA run on 1000 trees.txt}}:
-Summary results from a MultiMEDUSA run on the 1000 random trees selected
+{\textbf{Supporting Information S14.pdf}} Probability of nodes with
+estimated rates from a MultiMEDUSA run on the 1000 random trees selected
 from the posterior distribution.
 
-\textbf{\texttt{supp mat 14 multimedusa result prob nodes.pdf}}:
-Probability of nodes with estimated rates from a MultiMEDUSA run on the
-1000 random trees selected from the posterior distribution.
+{\textbf{Supporting Information S15.csv}} Hostplants of Nymphalidae
+butterflies recorded from the literature.
 
-\textbf{\texttt{supp mat 15 host plant data.csv}}: Hostplants of
-Nymphalidae butterflies recorded from the literature.
+{\textbf{Supporting Information S16.csv}} References for hostplants
+data.
 
-\textbf{\texttt{supp mat 16 host plant data references.csv}}:
-References for hostplants data.
+{\textbf{Supporting Information S17.csv}} Data matrix with character
+states for hosplant use.
 
-\textbf{\texttt{supp mat 17 hostplant states.csv}}: Data matrix with
-character states for hosplant use.
+{\textbf{Supporting Information S18.R}} R script for running the BiSSE
+analysis.
 
-\textbf{\texttt{supp mat 18 bisse source.R}}: R code for running the
+{\textbf{Supporting Information S19.csv}} Raw results for the BiSSE
+analysis.
+
+{\textbf{Supporting Information S20.csv}} Raw results for the combined
 BiSSE analysis.
 
-\textbf{\texttt{supp mat 19 bisse mcmc run.csv}}: Raw results for
-the BiSSE analysis.
+{\textbf{Figure S1.}} Best diversification model as estimated by BAMM.
+Two diversification shifts were recovered (Ithomiini and Satyrini). The
+MEDUSA analysis found a diversification shifts in Satyrinae (the parent
+node of Satyini).
 
-\textbf{\texttt{supp mat 20 combined bisse.csv}}: Raw results for
-the combined BiSSE analysis.
+{\textbf{Figure S2.}} Boxplot of speciation (\(\lambda\)) and extinction
+(\(\mu\)) values for the 95\% credibility intervals of values estimated
+by BiSSE analysis of diversification due to feeding on Solanaceae
+plants.
 
-\texttt{Figure S1.} Boxplot of speciation (lambda) and extinction (mu)
-values for the 95\% credibility intervals of values estimated by BiSSE
-analysis of diversification due to feeding on Solanaceae plants.
-
-\texttt{Figure S2.} BiSSE analysis of diversification of nymphalids due
-to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
+{\textbf{Figure S3.}} BiSSE analysis of diversification of nymphalids
+due to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
 \emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
 recovered, speciation and net diversification rates are significantly
-higher for Solanaceae feeders (\$lambda\$1, r1).
+higher for Solanaceae feeders \(\lambda1\), r1).
 
-\texttt{Figure S3.} Net diversification rates of nymphalids feeding on
+{\textbf{Figure S4.}} Net diversification rates of nymphalids feeding on
 Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
@@
 
-\texttt{Figure S4.} Boxplot of speciation and extinction values for the
-95\% credibility intervals of values estimated by BiSSE analysis of
+{\textbf{Figure S5.}} Boxplot of speciation and extinction values for
+the 95\% credibility intervals of values estimated by BiSSE analysis of
 diversification due to feeding on Solanaceae plants on the combined
@@
 
-\texttt{Figure S5.} BiSSE analysis of diversification of nymphalids due
-to feeding on Apocynaceae hostplants. Speciation and net diversification
-rates are similar.
+{\textbf{Figure S6.}} BiSSE analysis of diversification of nymphalids
+due to feeding on Apocynaceae hostplants. Speciation and net
+diversification rates are similar.
 
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
+% Table 1
+\section*{Tables}
+\begin{table}[!h]
+\caption{\bf{Significant net diversification rate shifts found in the MEDUSA analysis of the Nymphalid maximum clade credibility tree.}}
+\begin{tabular}{lccrrcl}
+Shift N$^\circ$ & Split.Node & Model & r          & LnLik.part & AICc     & Taxa                                                       \\
+1               & 399        & yule  & 0.092459   & -1055.957  &          & Nymphalidae (root)                                         \\
+2               & 691        & bd    & 0.054129   & -406.3703  &          & Limenitidinae + Heliconiinae                               \\
+3               & 299        & yule  & 0.311199   & -6.3058    &          & \emph{Ypthima}                                             \\
+4               & 224        & yule  & 0.290989   & -6.2601    &          & \emph{Charaxes}                                            \\
+5               & 750        & yule  & 0.186913   & -147.4146  &          & Oleriina + Ithomiina + Napeogenina + Dircennina + Godyrina \\
+6               & 405        & yule  & 0.116252   & -555.0276  &          & Satyrinae                                                   \\
+7               & 495        & yule  & 0.064656   & -124.9143  &          & Coenonymphina                                              \\
+8               & 609        & yule  & 0.240562   & -35.0908   &          & Phyciodina in part                                         \\
+9               & 787        & bd    & 0.042332   & -43.7819   &          & Danaini in part                                            \\
+10              & 231        & yule  & 0.209416   & -4.7955    &          & \emph{Coenonympha}                                         \\
+11              & 478        & yule  & 0.311684   & -9.2218    &          & \emph{Caeruleuptychia} + \emph{Magneuptychia}              \\
+12              & 659        & yule  & 0.219253   & -11.2686   &          & \emph{Callicore} + \emph{Diaethria}                        \\
+13              & 444        & yule  & 0.220615   & -43.7812   &          & Satyrina                                                   \\
+14              & 524        & yule  & 0.190754   & -26.0651   &          & Mycalesina                                                 \\
+15              & 355        & yule  & 0.234041   & -6.2013    &          & \emph{Pedaliodes}                                          \\
+16              & 714        & yule  & 0          & 0          &          & \emph{Dryas} + \emph{Dryadula}                             \\
+17              & 377        & yule  & 0.311671   & -4.1986    &          & \emph{Taenaris}                                            \\
+18              & 688        & yule  & 0.024724   & -17.5256   &          & Pseudergolinae                                             \\
+19              & 583        & yule  & 0          & 0          & 5090.492 & \emph{Anaeomorpha} + \emph{Hypna}                                       
+\end{tabular}
+\begin{flushleft}Split.Node=node number, Model=preferred diversification model by MEDUSA, r=net diversification rate, LnLik.part=log likelihood value.
+\end{flushleft}
+\end{table}
 
-1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
-adaptive zones: Has phytophagy promoted insect diversification? Am Nat
-132: 107--128.
-doi:\href{http://dx.doi.org/10.1086/284840}{10.1086/284840}.
+% Table 2
+\begin{table}[!h]
+    \caption{\bf{Differences in rates estimated by MEDUSA on the MCC
+tree from the sample of trees from the posterior distribution and the
+MultiMEDUSA approach. Shift consistently recovered across the
+sample of trees in bold face.}}
+\begin{tabular}{lcrr}
+Split.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
+1          & 0.092          & not found                  & 0.000                          \\
+2          & 0.055          & -0.030                     & 0.864                          \\
+\bf{3}     & 0.184          & \bf{0.219}                 & \bf{0.961}                     \\
+\bf{4}     & 0.166          & \bf{0.212}                 & \bf{0.996}                     \\
+\bf{5}     & 0.111          & \bf{0.101}                 & \bf{0.927}                     \\
+6          & 0.119          & 0.039                      & 0.131                          \\
+7          & 0.066          & -0.052                     & 0.195                          \\
+8          & 0.232          & 0.166                      & 0.319                          \\
+9          & 0.042          & -0.049                     & 0.897                          \\
+10         & 0.058          & 0.149                      & 0.619                          \\
+11         & 0.311          & 0.208                      & 0.831                          \\
+\bf{12}    & 0.219          & \bf{0.135}                 & \bf{0.911}                     \\
+13         & 0.082          & 0.117                      & 0.276                          \\
+14         & 0.099          & 0.115                      & 0.379                          \\
+15         & 0.113          & 0.127                      & 0.825                          \\
+16         & 0.222          & -0.005                     & 0.659                          \\
+17         & 0.243          & 0.248                      & 0.785                          \\
+18         & 0.192          & -0.064                     & 0.024                          \\
+19         & 0.064          & -0.087                     & 0.381                         
+\end{tabular}
+\end{table}
 
-2. Nyman T, Linder HP, Peña C, Malm T, Wahlberg N (2012) Climate-driven
-diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
-889--898.
-doi:\href{http://dx.doi.org/10.1111/j.1461-0248.2012.01782.x}{10.1111/j.1461-0248.2012.01782.x}.
+% Table 3
+\begin{table}[!h]
+    \caption{\bf{Likelihood ratio test between the model of increased
+diversification of nymphalids feeding on Solanaceae against a model
+forcing equal speciation rates (no effect on diversification).}}
+\begin{tabular}{lcrrlr}
+Df           & lnLik & AIC     & ChiSq  & \multicolumn{1}{r}{p} &         \\
+full         & 6     & -1613.3 & 3238.5 &                       &         \\
+equal.lambda & 5     & -1619.4 & 3248.9 & 12.3                  & 0.00045
+\end{tabular}
+\end{table}
 
-3. Ehrlich PR, Raven PH (1964) Butterflies and plants: A study in
-coevolution. Evolution 18: 586--608.
-doi:\href{http://dx.doi.org/10.2307/2406212}{10.2307/2406212}.
-
-4. Janz N (2011) Ehrlich and Raven Revisited: Mechanisms Underlying
-Codiversification of Plants and Enemies. Annu Rev Ecol Evol S 42:
-71--89.
-doi:\href{http://dx.doi.org/10.1146/annurev-ecolsys-102710-145024}{10.1146/annurev-ecolsys-102710-145024}.
-
-5. Nylin S, Slove J, Janz N (2014) Host plant utilization, host range
-oscillations and diversification in Nymphalid butterflies: A
-phylogenetic investigation. Evolution 68: 105--124.
-doi:\href{http://dx.doi.org/10.1111/evo.12227}{10.1111/evo.12227}.
-
-6. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
-al. (2011) Order Lepidoptera. In: Animal biodiversity: An outline of
-higher-level classification and survey of taxonomic richness. Zootaxa.
-pp. 212--221.
-
-7. Heikkilä M, Kaila L, Mutanen M, Peña C, Wahlberg N (2012) Cretaceous
-origin and repeated tertiary diversification of the redefined
-butterflies. P Roy Soc B 279: 1093--1099.
-doi:\href{http://dx.doi.org/10.1098/rspb.2011.1430}{10.1098/rspb.2011.1430}.
-
-8. Elias M, Joron M, Willmott KR, Silva-Brandão KL, Kaiser V, et al.
-(2009) Out of the Andes: Patterns of diversification in clearwing
-butterflies. Mol Ecol 18: 1716--1729.
-doi:\href{http://dx.doi.org/10.1111/j.1365-294X.2009.04149.x}{10.1111/j.1365-294X.2009.04149.x}.
-
-9. Fordyce JA (2010) Host shifts and evolutionary radiations of
-butterflies. P Roy Soc B 277: 3735--3743.
-doi:\href{http://dx.doi.org/10.1098/rspb.2010.0211}{10.1098/rspb.2010.0211}.
-
-10. Wahlberg N, Leneveu J, Kodandaramaiah U, Peña C, Nylin S, et al.
-(2009) Nymphalid butterflies diversify following near demise at the
-Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302.
-doi:\href{http://dx.doi.org/10.1098/rspb.2009.1303}{10.1098/rspb.2009.1303}.
-
-11. Peña C, Wahlberg N (2008) Prehistorical climate change increased
-diversification of a group of butterflies. Biol Lett 4: 274--278.
-doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.
-
-12. Ferrer-Paris JR, S{á}nchez-Mercado A, Viloria ÁL, Donaldson J (2013)
-Congruence and Diversity of Butterfly-Host Plant Associations at Higher
-Taxonomic Levels. PloS one 8: e63570.
-doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.
-
-13. Alfaro ME, Santini F, Brock C, Alamillo H, Dornburg A, et al. (2009)
-Nine exceptional radiations plus high turnover explain species diversity
-in jawed vertebrates. P Natl Acad Sci USA 106: 13410--13414.
-doi:\href{http://dx.doi.org/10.1073/pnas.0811087106}{10.1073/pnas.0811087106}.
-
-14. Harmon LJ, Rabosky DL, {FitzJohn} RG, Brown JW (2011) turboMEDUSA:
-MEDUSA: Modeling Evolutionary Diversification Using Stepwise AIC.
-
-15. {FitzJohn} RG (2012) Diversitree: comparative phylogenetic analyses
-of diversification in R. Method Ecol Evol 3: 1084--1092.
-doi:\href{http://dx.doi.org/10.1111/j.2041-210X.2012.00234.x}{10.1111/j.2041-210X.2012.00234.x}.
-
-16. Lamas G (2004) Checklist: Part 4A. Hesperioidea-Papilionoidea.
-Gainesville.
-
-17. Matos-Maraví PF, Peña C, Willmott KR, Freitas AV, Wahlberg N (2013)
-Systematics and evolutionary history of butterflies in the ``Taygetis
-clade'' (Nymphalidae: Satyrinae: Euptychiina): Towards a better
-understanding of Neotropical biogeography. Mol Phylogenet Evol 66:
-54--68.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2012.09.005}{10.1016/j.ympev.2012.09.005}.
-
-18. Brower AVZ, Wahlberg N, Ogawa JR, Boppré M, Vane-Wright RI (2010)
-Phylogenetic relationships among genera of danaine butterflies
-(Lepidoptera: Nymphalidae) as implied by morphology and DNA sequences.
-Syst Biodivers 8: 75--89.
-doi:\href{http://dx.doi.org/10.1080/14772001003626814}{10.1080/14772001003626814}.
-
-19. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
-(2010) Phylogenetics and biogeography of a spectacular Old World
-radiation of butterflies: the subtribe Mycalesina (Lepidoptera:
-Nymphalidae: Satyrini). BMC Evol Biol 10: 172.
-doi:\href{http://dx.doi.org/10.1186/1471-2148-10-172}{10.1186/1471-2148-10-172}.
-
-20. Kodandaramaiah U, Peña C, Braby MF, Grund R, Müller CJ, et al.
-(2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
-problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.08.012}{10.1016/j.ympev.2009.08.012}.
-
-21. Ortiz-Acevedo E, Willmott KR (2013) Molecular systematics of the
-butterfly tribe Preponini (Nymphalidae: Charaxinae). Syst Entomol 38:
-440--449.
-doi:\href{http://dx.doi.org/10.1111/syen.12008}{10.1111/syen.12008}.
-
-22. De-Silva DL, Day JJ, Elias M, Willmott KR, Whinnett A, et al. (2010)
-Molecular phylogenetics of the Neotropical butterfly subtribe Oleriina
-(Nymphalidae: Danainae: Ithomiini). Mol Phylogenet Evol 55: 1032--1041.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2010.01.010}{10.1016/j.ympev.2010.01.010}.
-
-23. Freitas AVL, Brown Jr. KS (2004) Phylogeny of the Nymphalidae
-(Lepidoptera). Syst Biol 53: 363--383.
-doi:\href{http://dx.doi.org/10.1080/10635150490445670}{10.1080/10635150490445670}.
-
-24. Peña C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
-(2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
-Nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2006.02.007}{10.1016/j.ympev.2006.02.007}.
-
-25. Penz CM (1999) Higher level phylogeny for the passion-vine
-butterflies (Nymphalidae, Heliconiinae) based on early stage and adult
-morphology. Zool J Linn Soc 127: 277--344.
-doi:\href{http://dx.doi.org/10.1111/j.1096-3642.1999.tb00680.x}{10.1111/j.1096-3642.1999.tb00680.x}.
-
-26. Silva-Brandão KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
-Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
-tribe Acraeini (Lepidoptera, Nymphalidae, Heliconiinae) and the
-evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2007.11.024}{10.1016/j.ympev.2007.11.024}.
-
-27. Peña C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
-butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
-methods. Zool J Linn Soc 161: 64--87.
-doi:\href{http://dx.doi.org/10.1111/j.1096-3642.2009.00627.x}{10.1111/j.1096-3642.2009.00627.x}.
-
-28. Peña C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
-history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
-Satyrinae). Zool Scr 39: 243--258.
-doi:\href{http://dx.doi.org/10.1111/j.1463-6409.2010.00421.x}{10.1111/j.1463-6409.2010.00421.x}.
-
-29. Ackery PR (1988) Hostplants and classification: A review of
-nymphalid butterflies. Biol J Linn Soc 33: 95--203.
-doi:\href{http://dx.doi.org/10.1111/j.1095-8312.1988.tb00446.x}{10.1111/j.1095-8312.1988.tb00446.x}.
-
-30. Dyer LA, Gentry GL (2002) Caterpillars and parasitoids of a tropical
-lowland wet forest. Available: \url{http://www.caterpillars.org}.
-Accessed 17 April 2013.
-
-31. Beccaloni G, Viloria Á, Hall S, Robinson G (2008) Catalogue of the
-hostplants of the Neotropical butterflies. London: The Natural History
-Museum.
-
-32. Janzen DH, Hallwachs W (2009) Dynamic database for an inventory of
-the macrocaterpillar fauna, and its food plants and parasitoids, of Area
-de Conservacion Guanacaste (ACG), northwestern Costa Rica. Available:
-\url{http://janzen.sas.upenn.edu}. Accessed 17 April 2013.
-
-33. Core Team R (2013) R: A Language and Environment for Statistical
-Computing. Vienna, Austria: R Foundation for Statistical Computing.
-Available: \url{http://www.r-project.org}.
-
-34. Popescu A-A, Huber KT, Paradis E (2012) ape 3.0: New tools for
-distance-based phylogenetics and evolutionary analysis in R.
-Bioinformatics 28: 1536--1537.
-doi:\href{http://dx.doi.org/10.1093/bioinformatics/bts184}{10.1093/bioinformatics/bts184}.
-
-35. Harmon LJ, Weir JT, Brock CD, Glor RE, Challenger W (2008) GEIGER:
-investigating evolutionary radiations. Bioinformatics 24: 129--131.
-doi:\href{http://dx.doi.org/10.1093/bioinformatics/btm538}{10.1093/bioinformatics/btm538}.
-
-36. Maddison WP, Midford PE, Otto SP (2007) Estimating a binary
-character's effect on speciation and extinction. Syst Biol 56: 701--710.
-doi:\href{http://dx.doi.org/10.1080/10635150701607033}{10.1080/10635150701607033}.
-
-37. Willmott KR, Freitas AVL (2006) Higher-level phylogeny of the
-Ithomiinae (Lepidoptera: Nymphalidae): classification, patterns of
-larval hostplant colonization and diversification. Cladistics 22:
-297--368.
-doi:\href{http://dx.doi.org/10.1111/j.1096-0031.2006.00108.x}{10.1111/j.1096-0031.2006.00108.x}.
-
-38. Scott JA (1986) The Butterflies of North America. Stanford
-University Press.
-
-39. Davis MP, Midford PE, Maddison W (2013) Exploring power and
-parameter estimation of the BiSSE method for analyzing species
-diversification. BMC Evol Biol 13: 38.
-doi:\href{http://dx.doi.org/10.1186/1471-2148-13-38}{10.1186/1471-2148-13-38}.
-
-40. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
-bees cut leaves? New insights into the early evolution of bees. P Roy
-Soc B 278: 3593--3600.
-doi:\href{http://dx.doi.org/10.1098/rspb.2011.0365}{10.1098/rspb.2011.0365}.
-
-41. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
-clades of Agaricales. P Roy Soc B 279: 2003--2011.
-doi:\href{http://dx.doi.org/10.1098/rspb.2011.2428}{10.1098/rspb.2011.2428}.
-
-42. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
-Interface. Ann Mo Bot Gard 74: 359--397.
-
-43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brandão KL, Whinnett A, et
-al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
-Nymphalidae) inferred from one mitochondrial and two nuclear gene
-regions. Syst Entomol 31: 288--301.
-doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2006.00321.x}{10.1111/j.1365-3113.2006.00321.x}.
-
-44. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
-trait-dependent speciation and extinction rates from incompletely
-resolved phylogenies. Syst Biol 58: 595--611.
-doi:\href{http://dx.doi.org/10.1093/sysbio/syp067}{10.1093/sysbio/syp067}.
-
-45. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
-The genetic basis of a plant-insect coevolutionary key innovation. P
-Natl Acad Sci USA 104: 20427--20431.
-doi:\href{http://dx.doi.org/10.1073/pnas.0706229104}{10.1073/pnas.0706229104}.
-
-46. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
-furanocoumarin-metabolizing cytochrome P450 monooxygenases in two
-papilionids: Specificity and substrate encounter rate. P Natl Acad Sci
-USA 100: 14593--14598.
-doi:\href{http://dx.doi.org/10.1073/pnas.1934643100}{10.1073/pnas.1934643100}.
-
-47. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
-the angiosperms re-revisited. Am J Bot 97: 1296--1303.
-doi:\href{http://dx.doi.org/10.3732/ajb.0900346}{10.3732/ajb.0900346}.
-
-48. Somoza R (1998) Updated Nazca (Farallon) - South America relative
-motions during the last 40 My: implications for mountain building in the
-central Andean region. J S Am Earth Sci 11: 211--215.
-doi:\href{http://dx.doi.org/10.1016/S0895-9811(98)00012-1}{10.1016/S0895-9811(98)00012-1}.
-
-49. Hoorn C, Wesselingh FP, Steege H ter, Bermudez MA, Mora A, et al.
-(2010) Amazonia through time: Andean uplift, climate change, landscape
-evolution, and biodiversity. Science 330: 927--931.
-doi:\href{http://dx.doi.org/10.1126/science.1194585}{10.1126/science.1194585}.
-
-50. Olmstead RG (2013) Phylogeny and biogeography in Solanaceae,
-Verbenaceae and Bignoniaceae: a comparison of continental and
-intercontinental diversification patterns. Bot J Linn Soc 171: 80--102.
-doi:\href{http://dx.doi.org/10.1111/j.1095-8339.2012.01306.x}{10.1111/j.1095-8339.2012.01306.x}.
-
-51. Wahlberg N, Rubinoff D (2011) Vagility across Vanessa (Lepidoptera:
-Nymphalidae): mobility in butterfly species does not inhibit the
-formation and persistence of isolated sister taxa. Syst Entomol 36:
-362--370.
-doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2010.00566.x}{10.1111/j.1365-3113.2010.00566.x}.
-
-52. Müller CJ, Wahlberg N, Beheregaray LB (2010) `After Africa': the
-evolutionary history and systematics of the genus Charaxes Ochsenheimer
-(Lepidoptera: Nymphalidae) in the Indo-Pacific region. Biol J Linn Soc
-100: 457--481.
-doi:\href{http://dx.doi.org/10.1111/j.1095-8312.2010.01426.x}{10.1111/j.1095-8312.2010.01426.x}.
-
-53. Aduse-Poku K, Vingerhoedt E, Wahlberg N (2009) Out-of-Africa again:
-a phylogenetic hypothesis of the genus Charaxes (Lepidoptera:
-Nymphalidae) based on 5 gene regions. Mol Phylogenet Evol 53: 463--478.
-doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.06.021}{10.1016/j.ympev.2009.06.021}.
+\pagebreak
+\bibliography{refs}{}
+\bibliographystyle{plos2009}
 

### 2015-01-14 MS_revised.tex
revision 2

@@
 evolution of hostplant use has spawned several theories explaining the
-evolutionary interactions between plants and insects \cite{nyman2012}. The
+evolutionary interactions between plants and insects \cite{nyman2012}: the
 ``escape-and-radiate'' hypothesis \cite{ehrlich1964}, the ``oscillation
@@
 The relatively low phylogenetic support for many nodes in the Nymphalidae
-tree is likelyl the reason
+tree is likely the reason
 why MEDUSA estimated net diversification rate shifts with a probability
@@
 dispersal has homogenized populations due to gene flow, as 
-vagile species seem to be genetically homogeneous in their
+vagile species seem to be genetically homogeneous among
 populations \cite{wahlberg2011}.
@@
 distribution of trees found strong support for an increase in net
-diversification rate in the tribe Ithomiini, the genera \emph{Ypthima},
-\emph{Charaxes}, the clade \emph{Callicore} + \emph{Diaethria} and for a
+diversification rate in the tribe Ithomiini, the genus \emph{Ypthima},
+genus \emph{Charaxes}, the clade \emph{Callicore} + \emph{Diaethria} and for a
 decrease in net diversification rate in the

### 2015-01-22 MS_revised.tex
fix supp figures in text
reformat supp materials in text

@@
-\documentclass[10pt]{article}
-\usepackage{amsmath}
-\usepackage{amssymb}
+% Template for PLoS
+% Version 3.0 December 2014
+%
+% To compile to pdf, run:
+% latex plos.template
+% bibtex plos.template
+% latex plos.template
+% latex plos.template
+% dvipdf plos.template
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
+% the final text of your manuscript.
+%
+% There are no restrictions on package use within the LaTeX files except that 
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - tabs/spacing/line breaks within cells to alter layout or alignment
+% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
+% - colors, shading, or graphic objects
+% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
+%
+% Please be sure to include all portions of an equation in the math environment.
+%
+% Do not include text that is not math in the math environment. For example, CO2 will be CO\textsubscript{2}.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% Use Unicode characters when possible
+\usepackage[utf8]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% fixltx2e package for \textsubscript
+\usepackage{fixltx2e}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
+
+% cite package, to clean up citations in the main text. Do not remove.
 \usepackage{cite}
-\usepackage{hyperref}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
 \usepackage{microtype}
@@
 
+% rotating package for sideways tables
+\usepackage{rotating}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
 % Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
 
-% Bold the 'Figure #' in the caption and separate it with a period
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
 % Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2009}
 
@@
 
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
 \pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+\lhead{\includegraphics[natwidth=1.3in,natheight=0.4in]{PLOSlogo.png}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\sf PLOS}
 
-%% Include all macros below. Please limit the use of macros.
+%% Include all macros below
+
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
 
@@
 \begin{document}
+\vspace*{0.35in}
 
+% Title must be 150 characters or less
 \begin{flushleft}
 {\Large
-    \title{Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates}
+\textbf\newline{Article Title - Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates}
 }
-\maketitle
-Carlos Pe\~na$^{1,\ast}$,
-Marianne Espeland$^{2}$
+\newline
+% Insert Author names, affiliations and corresponding author email.
 \\
-\bf{1} Carlos Pe\~na Laboratory of Genetics, Department of Biology, University of Turku, Turku, Finland. Email: \href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}
+Carlos Pe\~na\textsuperscript{1,\ast},
+Marianne Espeland\textsuperscript{2,\textcurrency}
 \\
-\bf{2} Marianne Espeland Museum of Comparative Zoology and Department
-of Organismic and Evolutionary Biology, Harvard University, Cambridge,
-USA. Email:
-\href{mailto:marianne.espeland@gmail.com}{\nolinkurl{marianne.espeland@gmail.com}}.
-Current address: McGuire Center for Lepidoptera and Biodiversity,
+\bf{1} Laboratory of Genetics, Department of Biology, University of Turku, Turku, Finland
+\\
+\bf{2} Museum of Comparative Zoology and Department of Organismic and Evolutionary Biology, Harvard University, Cambridge, USA
+\\
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+
+% Current address notes
+\textcurrency Current address: McGuire Center for Lepidoptera and Biodiversity,
 Florida Museum of Natural History, University of Florida, Gainesville,
 FL, US.
-\\
-$\ast$ \textbf{Corresponding author:}
-\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}
-Telephone: +358 417063065 Fax: +358 2 333 6680
+% \textcurrency b Insert current address of second author with an address update
+% \textcurrency c Insert current address of third author with an address update
 
+% Deceased author note
+
+% Group/Consortium Author Note
+
+* E-mail: mycalesis@gmail.com
 \end{flushleft}
 
-\subsubsection*{Running title: Diversity dynamics in Nymphalidae
+\subsection*{Running title: Diversity dynamics in Nymphalidae
 butterflies}
 
+% Please keep the abstract below 300 words
 \section*{Abstract}
-
 The species rich butterfly family Nymphalidae has been used to study
@@
 
+\linenumbers
+
 \section*{Introduction}
-
 Hostplant shifts have been invoked to be responsible for a great part of
@@
 
+% You may title this section "Methods" or "Models". 
+% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
+% authors may use "Analysis" 
 \section*{Materials and Methods}
-
 \subsection*{Data}
@@
 hostplant data for 35 butterfly genera.
+\begin{figure}[h]
+\caption{{\bf Figure Title first bold sentence Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}
+Figure Caption Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
+\label{fig1}
+\end{figure}
 
@@
 \texttt{r = 0.11} while the net diversification rate for the Solanaceae
-feeders was \texttt{r = 0.16} (see Fig. S2 for a boxplot of speciation
+feeders was \texttt{r = 0.16} (see Fig. S1 for a boxplot of speciation
 and extinction values for the 95\% credibility intervals). The same
@@
 non-Solanaceae feeders due to dubious records produced the same pattern
-and net diversification rates (Fig. S3). Therefore, the rest of the
+and net diversification rates (Fig. S2). Therefore, the rest of the
 analyses were performed assuming these two genera as Solanaceae feeders.
@@
 credibility tree (combined mcmc run in supp. mat. 20; profiles plot of
-speciation rates in Fig. S4; boxplot of 95\% credibility intervals in
-Fig. S5). A BiSSE analysis to test whether the trait ``feeding on
+speciation rates in Fig. S3; boxplot of 95\% credibility intervals in
+Fig. S4). A BiSSE analysis to test whether the trait ``feeding on
 Apocynaceae'' had any effect on increased net diversification rates
 found similar speciation rates for lineages feeding on Apocynaceae and
-other plants (Fig. S6).
+other plants (Fig. S5).
 
@@
 
-\subsubsection*{Ithomiini}
+\paragraph*{Ithomiini}
 
@@
 
-\subsubsection*{Danaini}
+\paragraph*{Danaini}
 Our MultiMEDUSA approach showed a significant slowdown in net
@@
 
-\subsubsection*{Satyrinae}
+\paragraph*{Satyrinae}
 
@@
 
-\section*{Acknowledgments}
-We are thankful to Mark Cornwall for help with the script to extend
-MEDUSA to include phylogenetic uncertainty, Niklas Wahlberg for
-commenting on the manuscript and giving us the posterior distribution of
-trees, Luke Harmon for commenting on the manuscript, Jeffrey C. Oliver and anonymous
-reviewers for their comments, which greatly improved the manuscript,
-Jessica Slove Davidson and Niklas Janz for access to their hostplant
-data. The study was supported by a Kone Foundation grant (awarded to
-Niklas Wahlberg), Finland (C. Pe\~na) and the Research Council of Norway
-(grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
-Science Ltd. (Finland) for the allocation of computational resources.
-
 \section*{Figure legends}
@@
 \section*{Supporting Information Legends}
-{\textbf{Supporting Information S01.nex.}} MCC Nymphalidae tree from
-\cite{wahlberg2009}
+\subsection*{S1 Dataset}
+\label{S1_Dataset}
+{\bf MCC Nymphalidae tree from \cite{wahlberg2009}}
 
-{\textbf{Supporting Information S02.R}} R script to run MEDUSA on the
-MCC Nymphalidae tree from \cite{wahlberg2009}. This script removes the outgroup
+\subsection*{S2 Supporting Information}
+\label{S2_Supporting_Information}
+{\bf R script to run MEDUSA on the MCC Nymphalidae tree from
+\cite{wahlberg2009}}. This script removes the outgroup
 taxa and loads the richness data for the tree terminals.
 
-{\textbf{Supporting Information S03.csv}} Species richness for lineages
-in Nymphalidae.
+\subsection*{S3 Dataset}
+\label{S3_Dataset}
+{\bf Species richness for lineages in Nymphalidae}.
 
-{\textbf{Supporting Information S04.txt}} Results of running MEDUSA on
-the MCC tree.
+\subsection*{S4 Text}
+\label{S4_Text}
+{\bf Results of running MEDUSA on the MCC tree}.
 
-{\textbf{Supporting Information S05 1000 random trees no outgroups.zip}}
-1000 random trees from \cite{wahlberg2009}.
+\subsection*{S5 File}
+\label{S5_File}
+{\bf 1000 random trees from \cite{wahlberg2009}}.
 
-{\textbf{Supporting Information S06.R}}~R script to run a MultiMEDUSA
-analysis on 1000 random trees from \cite{wahlberg2009}.
+\subsection*{S6 Supporting Information}
+\label{S6_Supporting_Information}
+{\bf R script to run a MultiMEDUSA analysis on 1000 random trees from
+\cite{wahlberg2009}}.
 
-{\textbf{Supporting Information S07 multimedusa on 1000 trees.zip}} Raw
-results from the MultiMEDUSA run on the random sample of trees from the
-posterior distribution.
+\subsection*{S7 File}
+\label{S7_File}
+{\bf Raw results from the MultiMEDUSA run on the random sample of trees from 
+the posterior distribution}.
 
-{\textbf{Supporting Information S08.pdf}} Summary of the MultiMEDUSA
-analysis on the 1000 trees from the posterior distribution.
+\subsection*{S8 Figure}
+\label{S8_Figure}
+{\bf Summary of the MultiMEDUSA
+analysis on the 1000 trees from the posterior distribution}.
 
-{\textbf{Supporting Information S09.tree}} MCC tree from the 1000 random
-trees selected from the posterior distribution.
+\subsection*{S9 Supporting Information}
+\label{S9_Supporting_Information}
+{\bf MCC tree from the 1000 random
+trees selected from the posterior distribution}.
 
-{\textbf{Supporting Information S10.R}} R script to run a MEDUSA
+\subsection*{S10 Supporting Information}
+\label{S10_Supporting_Information}
+{\bf R script to run a MEDUSA
 analysis on the MCC tree from the 1000 random trees selected from the
-posterior distribution.
+posterior distribution}.
 
-{\textbf{Supporting Information S11.txt}} Summary results from a MEDUSA
+\subsection*{S11 Text}
+\label{S11_Text}
+{\bf Summary results from a MEDUSA
 run on the MCC tree from the 1000 random trees selected from the
-posterior distribution.
+posterior distribution}.
 
-{\textbf{Supporting Information S12.pdf}} Figure for MEDUSA run on MCC
-tree from random 1000 trees.
+\subsection*{S12 Figure}
+\label{S12_Figure}
+{\bf Figure for MEDUSA run on MCC tree from random 1000 trees}.
 
-{\textbf{Supporting Information S13.txt}} Summary results from a
+\subsection*{S13 Text}
+\label{S13_Text}
+{\bf Summary results from a
 MultiMEDUSA run on the 1000 random trees selected from the posterior
-distribution.
+distribution}.
 
-{\textbf{Supporting Information S14.pdf}} Probability of nodes with
+\subsection*{S14 Figure}
+\label{S14_Figure}
+{\bf Probability of nodes with
 estimated rates from a MultiMEDUSA run on the 1000 random trees selected
-from the posterior distribution.
+from the posterior distribution}.
 
-{\textbf{Supporting Information S15.csv}} Hostplants of Nymphalidae
-butterflies recorded from the literature.
+\subsection*{S15 Dataset}
+\label{S15_Dataset}
+{\bf Hostplants of Nymphalidae
+butterflies recorded from the literature}.
 
-{\textbf{Supporting Information S16.csv}} References for hostplants
-data.
+\subsection*{S16 Dataset}
+\label{S16_Dataset}
+{\bf References for hostplants data}.
 
-{\textbf{Supporting Information S17.csv}} Data matrix with character
-states for hosplant use.
+\subsection*{S17 Dataset}
+\label{S17_Dataset}
+{\bf Data matrix with character states for hosplant use}.
 
-{\textbf{Supporting Information S18.R}} R script for running the BiSSE
-analysis.
+\subsection*{S18 Supporting Information}
+\label{S18_Supporting_Information}
+{\bf R script for running the BiSSE analysis}.
 
-{\textbf{Supporting Information S19.csv}} Raw results for the BiSSE
-analysis.
+\subsection*{S19 Supporting Information}
+\label{S19_Supporting_Information}
+{\bf Raw results for the BiSSE analysis}.
 
-{\textbf{Supporting Information S20.csv}} Raw results for the combined
-BiSSE analysis.
+\subsection*{S20 Supporting Information}
+\label{S20_Supporting_Information}
+{\bf Raw results for the combined BiSSE analysis}.
 
@@
 diversification rates are similar.
+
+\section*{Acknowledgments}
+We are thankful to Mark Cornwall for help with the script to extend
+MEDUSA to include phylogenetic uncertainty, Niklas Wahlberg for
+commenting on the manuscript and giving us the posterior distribution of
+trees, Luke Harmon for commenting on the manuscript, Jeffrey C. Oliver and anonymous
+reviewers for their comments, which greatly improved the manuscript,
+Jessica Slove Davidson and Niklas Janz for access to their hostplant
+data. The study was supported by a Kone Foundation grant (awarded to
+Niklas Wahlberg), Finland (C. Pe\~na) and the Research Council of Norway
+(grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
+Science Ltd. (Finland) for the allocation of computational resources.
+
+\nolinenumbers
 

### 2015-01-22 MS_revised.tex
add logo and new bib style
reorganize sup mat

@@
 \\
-Carlos Pe\~na\textsuperscript{1,\ast},
+Carlos Pe\~na\textsuperscript{1,*},
 Marianne Espeland\textsuperscript{2,\textcurrency}
@@
 used a burnin of 25 million generations and took a random sample of 1000
-trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (supp.
-mat. 5) in order to account for phylogenetic uncertainty when performing
+trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (S6 File)
+in order to account for phylogenetic uncertainty when performing
 the diversification analyses.
@@
 taxonomical classification and morphological resemblance when no
-phylogenies were available (supp. mat. 3).
+phylogenies were available (S7 Dataset).
 
@@
 sources including \cite{ackery1988}, HOSTS database (\url{http://bit.ly/YI7nwW}),
-\cite{dyer2002, beccaloni2008, janzen2009} and others (supp. mat. 15--16) for a total of 6586
+\cite{dyer2002, beccaloni2008, janzen2009} and others (S8--S9 Datasets) for a
+total of 6586
 hostplant records, including 428 Nymphalidae genera and 143 plant
@@
 hostplant data for 35 butterfly genera.
+
 \begin{figure}[h]
@@
 \texttt{diversitree} \cite{fitzjohn2012} packages along with our own scripts to
-perform the analyses (included as supplementary materials 2, 6, 10). All
+perform the analyses (included as S10--S12 Supporting Information). All
 analyses were run on the 1000 random trees from Wahlberg et al. (2009)
@@
 distribution of the Bayesian run from Wahlberg et al., 2009 \cite{wahlberg2009}
-(MultiMEDUSA, supp. mat. 5). We calculated a new MCC tree from the
+(MultiMEDUSA, S6 File). We calculated a new MCC tree from the
 selection of trees, ran MEDUSA and MultiMEDUSA on the selection of 1000
@@
 
-\subsection*{Estimation of trait-dependent speciation
-rates}
+\subsection*{Estimation of trait-dependent speciation rates}
 
@@
 use one plant family as hostplant (232 genera use only one plant family
-versus 176 that use more than one; based on our data in supp. mat. 15),
+versus 176 that use more than one; based on our data in S8 Dataset),
 the character states can be coded as presence/absence, for which the
@@
 binary character state has had any effect on increased net
-diversification rate for a clade \cite{maddison2007}. We used our compiled data of
-hostplant use to produce a binary dataset for the character ``feeding on
+diversification rate for a clade \cite{maddison2007}. We used our compiled data
+of hostplant use to produce a binary dataset for the character ``feeding on
 the plant family Solanaceae'' with two states (absence = 0; presence =
-1) (supp. mat. 17). The family Solanaceae include the hostplants of the
+1) (S13 Dataset). The family Solanaceae include the hostplants of the
 diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. 
@@
 values obtained from the MCC tree. We ran MultiMEDUSA on the random
-sample of 1000 trees (supp. mat. 5) from the posterior distribution and
+sample of 1000 trees (S6 File) from the posterior distribution and
 compared the results with a MEDUSA analysis on the MCC tree derived from
-this sample (supp. mat. 9).
+this sample (S14 Supporting Information).
 
@@
 MultiMEDUSA. The background diversification and all shifts found by
-MEDUSA on the 1000 trees are provided as an R object in supp. mat. 7.
+MEDUSA on the 1000 trees are provided as an R object in S15 File.
 
@@
 MultiMEDUSA provided mean and standard deviation statistics for the
-diversification values on the shifts on the 1000 trees (supp. mat.
-13--14), and found that the estimated net diversification rate
+diversification values on the shifts on the 1000 trees 
+(S16--S17 Text and Figure), and found that the estimated net diversification rate
 values had great variation across the posterior distribution of trees. A
@@
 \texttt{r = 0.11} while the net diversification rate for the Solanaceae
-feeders was \texttt{r = 0.16} (see Fig. S1 for a boxplot of speciation
+feeders was \texttt{r = 0.16} (see S1 Fig. for a boxplot of speciation
 and extinction values for the 95\% credibility intervals). The same
@@
 non-Solanaceae feeders due to dubious records produced the same pattern
-and net diversification rates (Fig. S2). Therefore, the rest of the
+and net diversification rates (S2 Fig.). Therefore, the rest of the
 analyses were performed assuming these two genera as Solanaceae feeders.
@@
 effect on diversification (\(\chi^2 = 12.3; 1 df; p < 0.001\)) (Table 3;
-character states available in supp. mat. 17, code in supp. mat. 18, and
-mcmc run in supp. mat. 19). We combined the post-burnin mcmc generations
+character states available in S13 Dataset, code in S18 Supporting Information,
+and mcmc run in S19 Supporting Information). We combined the post-burnin mcmc generations
 from running BiSSE on 250 trees from the random sample of 1000 trees and
 found the same pattern as the BiSSE analysis on the maximum clade
-credibility tree (combined mcmc run in supp. mat. 20; profiles plot of
-speciation rates in Fig. S3; boxplot of 95\% credibility intervals in
-Fig. S4). A BiSSE analysis to test whether the trait ``feeding on
+credibility tree (combined mcmc run in S20 Supporting Information; profiles
+plot of speciation rates in S3 Fig.; boxplot of 95\% credibility intervals in
+S4 Fig.). A BiSSE analysis to test whether the trait ``feeding on
 Apocynaceae'' had any effect on increased net diversification rates
 found similar speciation rates for lineages feeding on Apocynaceae and
-other plants (Fig. S5).
+other plants (S5 Fig.).
 
@@
 have been used for generation of hypotheses and discussion
-\cite{heikkila2012, litman2011, ryberg2012}. However, different diversification shifts and different
+\cite{heikkila2012, litman2011, ryberg2012}. However, different diversification
+shifts and different
 rates of diversification are found for certain lineages when
@@
 \section*{Supporting Information Legends}
-\subsection*{S1 Dataset}
-\label{S1_Dataset}
-{\bf MCC Nymphalidae tree from \cite{wahlberg2009}}
-
-\subsection*{S2 Supporting Information}
-\label{S2_Supporting_Information}
+\subsection*{S1 Fig.}
+\label{S1_Fig.}
+{\bf Boxplot of speciation (\(\lambda\)) and extinction
+(\(\mu\)) values for the 95\% credibility intervals of values estimated
+by BiSSE analysis of diversification due to feeding on Solanaceae
+plants}.
+
+\subsection*{S2 Fig.}
+\label{S2_Fig.}
+{\bf BiSSE analysis of diversification of nymphalids
+due to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
+\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
+recovered, speciation and net diversification rates are significantly
+higher for Solanaceae feeders \(\lambda1\), r1)}.
+
+\subsection*{S3 Fig.}
+\label{S3_Fig.}
+{\bf Net diversification rates of nymphalids feeding on
+Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
+the 1000 trees from the posterior distribution}.
+
+\subsection*{S4 Fig.}
+\label{S4_Fig.}
+{\bf Boxplot of speciation and extinction values for
+the 95\% credibility intervals of values estimated by BiSSE analysis of
+diversification due to feeding on Solanaceae plants on the combined
+post-burnin runs on 1000 trees from the posterior distribution.}.
+
+\subsection*{S5 Fig.}
+\label{S5_Fig.}
+{\bf BiSSE analysis of diversification of nymphalids
+due to feeding on Apocynaceae hostplants}. Speciation and net
+diversification rates are similar.
+
+\subsection*{S6 File}
+\label{S6_File}
+{\bf 1000 random trees from \cite{wahlberg2009}}.
+
+\subsection*{S7 Dataset}
+\label{S7_Dataset}
+{\bf Species richness for lineages in Nymphalidae}.
+
+\subsection*{S8Dataset}
+\label{S8_Dataset}
+{\bf Hostplants of Nymphalidae
+butterflies recorded from the literature}.
+
+\subsection*{S9 Dataset}
+\label{S9_Dataset}
+{\bf References for hostplants data}.
+
+\subsection*{S10 Supporting Information}
+\label{S10_Supporting_Information}
 {\bf R script to run MEDUSA on the MCC Nymphalidae tree from
@@
 
-\subsection*{S3 Dataset}
-\label{S3_Dataset}
-{\bf Species richness for lineages in Nymphalidae}.
-
-\subsection*{S4 Text}
-\label{S4_Text}
-{\bf Results of running MEDUSA on the MCC tree}.
-
-\subsection*{S5 File}
-\label{S5_File}
-{\bf 1000 random trees from \cite{wahlberg2009}}.
-
-\subsection*{S6 Supporting Information}
-\label{S6_Supporting_Information}
+\subsection*{S11 Supporting Information}
+\label{S11_Supporting_Information}
 {\bf R script to run a MultiMEDUSA analysis on 1000 random trees from
@@
 
-\subsection*{S7 File}
-\label{S7_File}
-{\bf Raw results from the MultiMEDUSA run on the random sample of trees from 
-the posterior distribution}.
-
-\subsection*{S8 Figure}
-\label{S8_Figure}
-{\bf Summary of the MultiMEDUSA
-analysis on the 1000 trees from the posterior distribution}.
-
-\subsection*{S9 Supporting Information}
-\label{S9_Supporting_Information}
-{\bf MCC tree from the 1000 random
-trees selected from the posterior distribution}.
-
-\subsection*{S10 Supporting Information}
-\label{S10_Supporting_Information}
+\subsection*{S12 Supporting Information}
+\label{S12_Supporting_Information}
 {\bf R script to run a MEDUSA
@@
 
-\subsection*{S11 Text}
-\label{S11_Text}
+\subsection*{S13 Dataset}
+\label{S13_Dataset}
+{\bf Data matrix with character states for hosplant use}.
+
+\subsection*{S14 Supporting Information}
+\label{S14_Supporting_Information}
+{\bf MCC tree from the 1000 random
+trees selected from the posterior distribution}.
+
+\subsection*{S15 File}
+\label{S15_File}
+{\bf Raw results from the MultiMEDUSA run on the random sample of trees from 
+the posterior distribution}.
+
+\subsection*{S16 Text}
+\label{S16_Text}
+{\bf Summary results from a
+MultiMEDUSA run on the 1000 random trees selected from the posterior
+distribution}.
+
+\subsection*{S17 Figure}
+\label{S17_Figure}
+{\bf Probability of nodes with
+estimated rates from a MultiMEDUSA run on the 1000 random trees selected
+from the posterior distribution}.
+
+\subsection*{S18 Supporting Information}
+\label{S18_Supporting_Information}
+{\bf R script for running the BiSSE analysis}.
+
+\subsection*{S19 Supporting Information}
+\label{S19_Supporting_Information}
+{\bf Raw results for the BiSSE analysis}.
+
+\subsection*{S20 Supporting Information}
+\label{S20_Supporting_Information}
+{\bf Raw results for the combined BiSSE analysis}.
+
+\subsection*{S21 Dataset}
+\label{S21_Dataset}
+{\bf MCC Nymphalidae tree from \cite{wahlberg2009}}
+
+\subsection*{S22 Text}
+\label{S22_Text}
+{\bf Results of running MEDUSA on the MCC tree}.
+
+\subsection*{S23 Figure}
+\label{S23_Figure}
+{\bf Summary of the MultiMEDUSA
+analysis on the 1000 trees from the posterior distribution}.
+
+\subsection*{S24 Text}
+\label{S24_Text}
 {\bf Summary results from a MEDUSA
@@
 
-\subsection*{S12 Figure}
-\label{S12_Figure}
+\subsection*{S25 Figure}
+\label{S25_Figure}
 {\bf Figure for MEDUSA run on MCC tree from random 1000 trees}.
-
-\subsection*{S13 Text}
-\label{S13_Text}
-{\bf Summary results from a
-MultiMEDUSA run on the 1000 random trees selected from the posterior
-distribution}.
-
-\subsection*{S14 Figure}
-\label{S14_Figure}
-{\bf Probability of nodes with
-estimated rates from a MultiMEDUSA run on the 1000 random trees selected
-from the posterior distribution}.
-
-\subsection*{S15 Dataset}
-\label{S15_Dataset}
-{\bf Hostplants of Nymphalidae
-butterflies recorded from the literature}.
-
-\subsection*{S16 Dataset}
-\label{S16_Dataset}
-{\bf References for hostplants data}.
-
-\subsection*{S17 Dataset}
-\label{S17_Dataset}
-{\bf Data matrix with character states for hosplant use}.
-
-\subsection*{S18 Supporting Information}
-\label{S18_Supporting_Information}
-{\bf R script for running the BiSSE analysis}.
-
-\subsection*{S19 Supporting Information}
-\label{S19_Supporting_Information}
-{\bf Raw results for the BiSSE analysis}.
-
-\subsection*{S20 Supporting Information}
-\label{S20_Supporting_Information}
-{\bf Raw results for the combined BiSSE analysis}.
-
-{\textbf{Figure S1.}} Boxplot of speciation (\(\lambda\)) and extinction
-(\(\mu\)) values for the 95\% credibility intervals of values estimated
-by BiSSE analysis of diversification due to feeding on Solanaceae
-plants.
-
-{\textbf{Figure S2.}} BiSSE analysis of diversification of nymphalids
-due to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
-\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
-recovered, speciation and net diversification rates are significantly
-higher for Solanaceae feeders \(\lambda1\), r1).
-
-{\textbf{Figure S3.}} Net diversification rates of nymphalids feeding on
-Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
-the 1000 trees from the posterior distribution.
-
-{\textbf{Figure S4.}} Boxplot of speciation and extinction values for
-the 95\% credibility intervals of values estimated by BiSSE analysis of
-diversification due to feeding on Solanaceae plants on the combined
-post-burnin runs on 1000 trees from the posterior distribution.
-
-{\textbf{Figure S5.}} BiSSE analysis of diversification of nymphalids
-due to feeding on Apocynaceae hostplants. Speciation and net
-diversification rates are similar.
 
@@
 \bibliography{refs}{}
-\bibliographystyle{plos2009}
 

### 2015-01-22 ancillary/Supporting_Information_S10.R

@@ -1,17 +1,0 @@
-# run MEDUSA 0.93-4-33
-
-library("MEDUSA")
-library("multicore")
-phy <- read.nexus("ancillary/Supporting_Information_S09.tree");
-phy <- ladderize(phy)
-
-
-
-richness <- read.csv("ancillary/Supporting_Information_S03.csv")
-
-res <- MEDUSA(phy, richness, stop="threshold", model="mixed", modelLimit=25, mc=TRUE)
-
-
-pdf(file="ancillary/supp_mat_12_MEDUSA_mcc_tree_from_random_sample_of_1000_trees.pdf", width=9, height=23)
-summ <- medusaSummary(res, cex=0.3)
-dev.off()

### 2015-01-22 MS_revised.tex
moved figure legends

@@
 
+\begin{figure}[h]
+\caption{{\bf Results of the MEDUSA analysis run on the maximum
+clade credibility tree from Wahlberg et al. (2009)}. Rate shifts were estimated for the
+following nodes (besides the background rate): 1) root, 2) Limenitidinae +
+Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
+part, 6) Satyrinae, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
+part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
+\emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
+Satyrina, 14) Mycalesina, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
+\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae, 19)
+\emph{Anaeomorpha} + \emph{Hypna}. Circles on
+nodes indicate the diversification shift number as found by MEDUSA.
+Numbers next to circles indicate the posterior probability values for
+such nodes.}
+\label{fig1}
+\end{figure}
+
 Species richness data for Nymphalidae genera were compiled from several
@@
 hostplant data for 35 butterfly genera.
-
-\begin{figure}[h]
-\caption{{\bf Figure Title first bold sentence Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}
-Figure Caption Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
-\label{fig1}
-\end{figure}
 
@@
 
+\begin{figure}[h]
+\caption{{\bf Diversification rates for taxa estimated by MultiMEDUSA on
+the samples of 1000 random trees}}
+\label{fig2}
+\end{figure}
+
 It is also evident that not all the diversification shifts estimated on
@@
 probability of 0.18 (Fig. 3).
+
+\begin{figure}[h]
+\caption{{\bf Results of MultiMEDUSA analysis showing the
+probability of specific nodes being characterized by significant shifts
+in diversification rate}}
+\label{fig3}
+\end{figure}
 
@@
 
+\begin{figure}[h]
+\caption{{\bf BiSSE analysis of diversification of nymphalids due
+to feeding on Solanaceae hostplants}.
+Speciation and net diversification
+rates are significantly higher in Solanaceae feeders (speciation rate =
+$\lambda1$, net diversification rate = $r1$}
+\label{fig4}
+\end{figure}
+
 \section*{Discussion}
@@
 cospeciation hypothesis \cite{nyman2012, janz2011}.
-
-\section*{Figure legends}
-{\bf Figure 1. Results of the MEDUSA analysis run on the maximum
-clade credibility tree from Wahlberg et al. (2009)} Rate shifts were estimated for the
-following nodes (besides the background rate): 1) root, 2) Limenitidinae +
-Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
-part, 6) Satyrinae, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
-part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
-\emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
-Satyrina, 14) Mycalesina, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
-\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae, 19)
-\emph{Anaeomorpha} + \emph{Hypna}. Circles on
-nodes indicate the diversification shift number as found by MEDUSA.
-Numbers next to circles indicate the posterior probability values for
-such nodes.
-
-{\bf Figure 2. Diversification rates for taxa estimated by MultiMEDUSA on
-the samples of 1000 random trees}
-
-{\bf Figure 3. Results of MultiMEDUSA analysis showing the
-probability of specific nodes being characterized by significant shifts
-in diversification rate}
-
-{\bf Figure 4. BiSSE analysis of diversification of nymphalids due
-to feeding on Solanaceae hostplants} Speciation and net diversification
-rates are significantly higher in Solanaceae feeders (speciation rate =
-$\lambda1$, net diversification rate = $r1$).
 

### 2015-01-22 MS_revised.tex
move tables to text
table 1

@@
 
+% Table 1
+\begin{table}[!ht]
+\begin{adjustwidth}{-2.25in}{0in}
+\caption{{\bf Significant net diversification rate shifts found in the MEDUSA analysis of the Nymphalid maximum clade credibility tree.}}
+\begin{tabular}{|l|c|c|r|r|c|l|}
+\hline
+Shift N$^\circ$ & Shift.Node & Model & r          & LnLik.part & AICc     & Taxa                                                       \\ \hline
+1               & 399        & yule  & 0.092459   & -1055.957  &          & Nymphalidae (root)                                         \\ \hline
+2               & 691        & bd    & 0.054129   & -406.3703  &          & Limenitidinae + Heliconiinae                               \\ \hline
+3               & 299        & yule  & 0.311199   & -6.3058    &          & \emph{Ypthima}                                             \\ \hline
+4               & 224        & yule  & 0.290989   & -6.2601    &          & \emph{Charaxes}                                            \\ \hline
+5               & 750        & yule  & 0.186913   & -147.4146  &          & Oleriina + Ithomiina + Napeogenina + Dircennina + Godyrina \\ \hline
+6               & 405        & yule  & 0.116252   & -555.0276  &          & Satyrinae                                                  \\ \hline
+7               & 495        & yule  & 0.064656   & -124.9143  &          & Coenonymphina                                              \\ \hline
+8               & 609        & yule  & 0.240562   & -35.0908   &          & Phyciodina in part                                         \\ \hline
+9               & 787        & bd    & 0.042332   & -43.7819   &          & Danaini in part                                            \\ \hline
+10              & 231        & yule  & 0.209416   & -4.7955    &          & \emph{Coenonympha}                                         \\ \hline
+11              & 478        & yule  & 0.311684   & -9.2218    &          & \emph{Caeruleuptychia} + \emph{Magneuptychia}              \\ \hline
+12              & 659        & yule  & 0.219253   & -11.2686   &          & \emph{Callicore} + \emph{Diaethria}                        \\ \hline
+13              & 444        & yule  & 0.220615   & -43.7812   &          & Satyrina                                                   \\ \hline
+14              & 524        & yule  & 0.190754   & -26.0651   &          & Mycalesina                                                 \\ \hline
+15              & 355        & yule  & 0.234041   & -6.2013    &          & \emph{Pedaliodes}                                          \\ \hline
+16              & 714        & yule  & 0          & 0          &          & \emph{Dryas} + \emph{Dryadula}                             \\ \hline
+17              & 377        & yule  & 0.311671   & -4.1986    &          & \emph{Taenaris}                                            \\ \hline
+18              & 688        & yule  & 0.024724   & -17.5256   &          & Pseudergolinae                                             \\ \hline
+19              & 583        & yule  & 0          & 0          & 5090.492 & \emph{Anaeomorpha} + \emph{Hypna}                          \\ \hline
+\end{tabular}
+\begin{flushleft}Shift.Node=node number, Model=preferred diversification model by MEDUSA, r=net diversification rate, LnLik.part=log likelihood value.
+\end{flushleft}
+\label{table1}
+\end{adjustwidth}
+\end{table}
+
 Some of the 18 changes in diversification correspond to rate increases
@@
 MEDUSA on the 1000 trees are provided as an R object in S15 File.
+
+% Table 2
+\begin{table}[!h]
+    \caption{\bf{Differences in rates estimated by MEDUSA on the MCC
+tree from the sample of trees from the posterior distribution and the
+MultiMEDUSA approach. Shift consistently recovered across the
+sample of trees in bold face.}}
+\begin{tabular}{lcrr}
+Shift.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
+1          & 0.092          & not found                  & 0.000                          \\
+2          & 0.055          & -0.030                     & 0.864                          \\
+\bf{3}     & 0.184          & \bf{0.219}                 & \bf{0.961}                     \\
+\bf{4}     & 0.166          & \bf{0.212}                 & \bf{0.996}                     \\
+\bf{5}     & 0.111          & \bf{0.101}                 & \bf{0.927}                     \\
+6          & 0.119          & 0.039                      & 0.131                          \\
+7          & 0.066          & -0.052                     & 0.195                          \\
+8          & 0.232          & 0.166                      & 0.319                          \\
+9          & 0.042          & -0.049                     & 0.897                          \\
+10         & 0.058          & 0.149                      & 0.619                          \\
+11         & 0.311          & 0.208                      & 0.831                          \\
+\bf{12}    & 0.219          & \bf{0.135}                 & \bf{0.911}                     \\
+13         & 0.082          & 0.117                      & 0.276                          \\
+14         & 0.099          & 0.115                      & 0.379                          \\
+15         & 0.113          & 0.127                      & 0.825                          \\
+16         & 0.222          & -0.005                     & 0.659                          \\
+17         & 0.243          & 0.248                      & 0.785                          \\
+18         & 0.192          & -0.064                     & 0.024                          \\
+19         & 0.064          & -0.087                     & 0.381                         
+\end{tabular}
+\end{table}
 
@@
 
+% Table 3
+\begin{table}[!h]
+    \caption{\bf{Likelihood ratio test between the model of increased
+diversification of nymphalids feeding on Solanaceae against a model
+forcing equal speciation rates (no effect on diversification).}}
+\begin{tabular}{lcrrlr}
+Df           & lnLik & AIC     & ChiSq  & \multicolumn{1}{r}{p} &         \\
+full         & 6     & -1613.3 & 3238.5 &                       &         \\
+equal.lambda & 5     & -1619.4 & 3248.9 & 12.3                  & 0.00045
+\end{tabular}
+\end{table}
+
 \begin{figure}[h]
@@
 
-% Table 1
-\section*{Tables}
-\begin{table}[!h]
-\caption{\bf{Significant net diversification rate shifts found in the MEDUSA analysis of the Nymphalid maximum clade credibility tree.}}
-\begin{tabular}{lccrrcl}
-Shift N$^\circ$ & Shift.Node & Model & r          & LnLik.part & AICc     & Taxa                                                       \\
-1               & 399        & yule  & 0.092459   & -1055.957  &          & Nymphalidae (root)                                         \\
-2               & 691        & bd    & 0.054129   & -406.3703  &          & Limenitidinae + Heliconiinae                               \\
-3               & 299        & yule  & 0.311199   & -6.3058    &          & \emph{Ypthima}                                             \\
-4               & 224        & yule  & 0.290989   & -6.2601    &          & \emph{Charaxes}                                            \\
-5               & 750        & yule  & 0.186913   & -147.4146  &          & Oleriina + Ithomiina + Napeogenina + Dircennina + Godyrina \\
-6               & 405        & yule  & 0.116252   & -555.0276  &          & Satyrinae                                                   \\
-7               & 495        & yule  & 0.064656   & -124.9143  &          & Coenonymphina                                              \\
-8               & 609        & yule  & 0.240562   & -35.0908   &          & Phyciodina in part                                         \\
-9               & 787        & bd    & 0.042332   & -43.7819   &          & Danaini in part                                            \\
-10              & 231        & yule  & 0.209416   & -4.7955    &          & \emph{Coenonympha}                                         \\
-11              & 478        & yule  & 0.311684   & -9.2218    &          & \emph{Caeruleuptychia} + \emph{Magneuptychia}              \\
-12              & 659        & yule  & 0.219253   & -11.2686   &          & \emph{Callicore} + \emph{Diaethria}                        \\
-13              & 444        & yule  & 0.220615   & -43.7812   &          & Satyrina                                                   \\
-14              & 524        & yule  & 0.190754   & -26.0651   &          & Mycalesina                                                 \\
-15              & 355        & yule  & 0.234041   & -6.2013    &          & \emph{Pedaliodes}                                          \\
-16              & 714        & yule  & 0          & 0          &          & \emph{Dryas} + \emph{Dryadula}                             \\
-17              & 377        & yule  & 0.311671   & -4.1986    &          & \emph{Taenaris}                                            \\
-18              & 688        & yule  & 0.024724   & -17.5256   &          & Pseudergolinae                                             \\
-19              & 583        & yule  & 0          & 0          & 5090.492 & \emph{Anaeomorpha} + \emph{Hypna}                                       
-\end{tabular}
-\begin{flushleft}Shift.Node=node number, Model=preferred diversification model by MEDUSA, r=net diversification rate, LnLik.part=log likelihood value.
-\end{flushleft}
-\end{table}
-
-% Table 2
-\begin{table}[!h]
-    \caption{\bf{Differences in rates estimated by MEDUSA on the MCC
-tree from the sample of trees from the posterior distribution and the
-MultiMEDUSA approach. Shift consistently recovered across the
-sample of trees in bold face.}}
-\begin{tabular}{lcrr}
-Shift.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
-1          & 0.092          & not found                  & 0.000                          \\
-2          & 0.055          & -0.030                     & 0.864                          \\
-\bf{3}     & 0.184          & \bf{0.219}                 & \bf{0.961}                     \\
-\bf{4}     & 0.166          & \bf{0.212}                 & \bf{0.996}                     \\
-\bf{5}     & 0.111          & \bf{0.101}                 & \bf{0.927}                     \\
-6          & 0.119          & 0.039                      & 0.131                          \\
-7          & 0.066          & -0.052                     & 0.195                          \\
-8          & 0.232          & 0.166                      & 0.319                          \\
-9          & 0.042          & -0.049                     & 0.897                          \\
-10         & 0.058          & 0.149                      & 0.619                          \\
-11         & 0.311          & 0.208                      & 0.831                          \\
-\bf{12}    & 0.219          & \bf{0.135}                 & \bf{0.911}                     \\
-13         & 0.082          & 0.117                      & 0.276                          \\
-14         & 0.099          & 0.115                      & 0.379                          \\
-15         & 0.113          & 0.127                      & 0.825                          \\
-16         & 0.222          & -0.005                     & 0.659                          \\
-17         & 0.243          & 0.248                      & 0.785                          \\
-18         & 0.192          & -0.064                     & 0.024                          \\
-19         & 0.064          & -0.087                     & 0.381                         
-\end{tabular}
-\end{table}
-
-% Table 3
-\begin{table}[!h]
-    \caption{\bf{Likelihood ratio test between the model of increased
-diversification of nymphalids feeding on Solanaceae against a model
-forcing equal speciation rates (no effect on diversification).}}
-\begin{tabular}{lcrrlr}
-Df           & lnLik & AIC     & ChiSq  & \multicolumn{1}{r}{p} &         \\
-full         & 6     & -1613.3 & 3238.5 &                       &         \\
-equal.lambda & 5     & -1619.4 & 3248.9 & 12.3                  & 0.00045
-\end{tabular}
-\end{table}
-
-\pagebreak
 \bibliography{refs}{}

### 2015-01-22 MS_revised.tex
table 2

@@
 % Table 2
-\begin{table}[!h]
-    \caption{\bf{Differences in rates estimated by MEDUSA on the MCC
+\begin{table}[!ht]
+\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+\caption{{\bf Differences in rates estimated by MEDUSA on the MCC
 tree from the sample of trees from the posterior distribution and the
@@
 sample of trees in bold face.}}
-\begin{tabular}{lcrr}
-Shift.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
-1          & 0.092          & not found                  & 0.000                          \\
-2          & 0.055          & -0.030                     & 0.864                          \\
-\bf{3}     & 0.184          & \bf{0.219}                 & \bf{0.961}                     \\
-\bf{4}     & 0.166          & \bf{0.212}                 & \bf{0.996}                     \\
-\bf{5}     & 0.111          & \bf{0.101}                 & \bf{0.927}                     \\
-6          & 0.119          & 0.039                      & 0.131                          \\
-7          & 0.066          & -0.052                     & 0.195                          \\
-8          & 0.232          & 0.166                      & 0.319                          \\
-9          & 0.042          & -0.049                     & 0.897                          \\
-10         & 0.058          & 0.149                      & 0.619                          \\
-11         & 0.311          & 0.208                      & 0.831                          \\
-\bf{12}    & 0.219          & \bf{0.135}                 & \bf{0.911}                     \\
-13         & 0.082          & 0.117                      & 0.276                          \\
-14         & 0.099          & 0.115                      & 0.379                          \\
-15         & 0.113          & 0.127                      & 0.825                          \\
-16         & 0.222          & -0.005                     & 0.659                          \\
-17         & 0.243          & 0.248                      & 0.785                          \\
-18         & 0.192          & -0.064                     & 0.024                          \\
-19         & 0.064          & -0.087                     & 0.381                         
+\begin{tabular}{|l|c|r|r|}
+    \hline
+Shift.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\ \hline
+1          & 0.092          & not found                  & 0.000                          \\ \hline
+2          & 0.055          & -0.030                     & 0.864                          \\ \hline
+\bf{3}     & 0.184          & \bf{0.219}                 & \bf{0.961}                     \\ \hline
+\bf{4}     & 0.166          & \bf{0.212}                 & \bf{0.996}                     \\ \hline
+\bf{5}     & 0.111          & \bf{0.101}                 & \bf{0.927}                     \\ \hline
+6          & 0.119          & 0.039                      & 0.131                          \\ \hline
+7          & 0.066          & -0.052                     & 0.195                          \\ \hline
+8          & 0.232          & 0.166                      & 0.319                          \\ \hline
+9          & 0.042          & -0.049                     & 0.897                          \\ \hline
+10         & 0.058          & 0.149                      & 0.619                          \\ \hline
+11         & 0.311          & 0.208                      & 0.831                          \\ \hline
+\bf{12}    & 0.219          & \bf{0.135}                 & \bf{0.911}                     \\ \hline
+13         & 0.082          & 0.117                      & 0.276                          \\ \hline
+14         & 0.099          & 0.115                      & 0.379                          \\ \hline
+15         & 0.113          & 0.127                      & 0.825                          \\ \hline
+16         & 0.222          & -0.005                     & 0.659                          \\ \hline
+17         & 0.243          & 0.248                      & 0.785                          \\ \hline
+18         & 0.192          & -0.064                     & 0.024                          \\ \hline
+19         & 0.064          & -0.087                     & 0.381                          \\ \hline
 \end{tabular}
+\begin{flushleft}
+\end{flushleft}
+\label{table2}
+\end{adjustwidth}
 \end{table}

### 2015-01-22 MS_revised.tex
table 3

@@
 % Table 3
-\begin{table}[!h]
-    \caption{\bf{Likelihood ratio test between the model of increased
+\begin{table}[!ht]
+% \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+    \caption{{\bf Likelihood ratio test between the model of increased
 diversification of nymphalids feeding on Solanaceae against a model
 forcing equal speciation rates (no effect on diversification).}}
-\begin{tabular}{lcrrlr}
-Df           & lnLik & AIC     & ChiSq  & \multicolumn{1}{r}{p} &         \\
-full         & 6     & -1613.3 & 3238.5 &                       &         \\
-equal.lambda & 5     & -1619.4 & 3248.9 & 12.3                  & 0.00045
+\begin{tabular}{|l|c|r|r|l|r|}
+    \hline
+Df           & lnLik & AIC     & ChiSq  & \multicolumn{1}{r}{p} &         \\ \hline
+full         & 6     & -1613.3 & 3238.5 &                       &         \\ \hline
+equal.lambda & 5     & -1619.4 & 3248.9 & 12.3                  & 0.00045 \\ \hline
 \end{tabular}
+\begin{flushleft}
+\end{flushleft}
+\label{table3}
+% \end{adjustwidth}
 \end{table}

### 2015-02-17 MS_revised.tex
submitted
ref to figures

@@
 % Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2009}
+\bibliographystyle{plos2015}
 
@@
 Carlos Pe\~na\textsuperscript{1,*},
-Marianne Espeland\textsuperscript{2,\textcurrency}
+Marianne Espeland\textsuperscript{2,3}
 \\
@@
 % Current address notes
-\textcurrency Current address: McGuire Center for Lepidoptera and Biodiversity,
+\bf{3} Current address: McGuire Center for Lepidoptera and Biodiversity,
 Florida Museum of Natural History, University of Florida, Gainesville,
@@
 \end{flushleft}
-
-\subsection*{Running title: Diversity dynamics in Nymphalidae
-butterflies}
 
@@
 
-\textbf{Keywords:} diversification analysis, MEDUSA, BiSSE, speciation
-rate, insect-hostplant dynamics
-
 \linenumbers
@@
 10 gene regions for 398 of the 540 valid genera in Nymphalidae. We
-employed the maximum clade credibility tree (MCC tree) (Fig. 1) as well
+employed the maximum clade credibility tree (MCC tree) (Fig.~\ref{fig1}) as well
 as a random sample of 1000 trees from their BEAST run after burnin
@@
 sources including \cite{ackery1988}, HOSTS database (\url{http://bit.ly/YI7nwW}),
-\cite{dyer2002, beccaloni2008, janzen2009} and others (S8--S9 Datasets) for a
+\cite{dyer2002, beccaloni2008, janzen2009} and others (S8 and S9 Dataset) for a
 total of 6586
@@
 \texttt{diversitree} \cite{fitzjohn2012} packages along with our own scripts to
-perform the analyses (included as S10--S12 Supporting Information). All
+perform the analyses (included as S10 and S11 File). All
 analyses were run on the 1000 random trees from Wahlberg et al. (2009)
@@
 estimated 18 significant changes in the tempo of diversification in
-Nymphalidae history (Fig. 1; Table 1). The corrected AICc acceptance
+Nymphalidae history (Fig.~\ref{fig1}; Table 1). The corrected AICc acceptance
 threshold for adding subsequent piecewise birth-death processes to the
@@
 diversification values on the shifts on the 1000 trees 
-(S16--S17 Text and Figure), and found that the estimated net diversification rate
+(S16 Text and S17 Figure), and found that the estimated net diversification rate
 values had great variation across the posterior distribution of trees. A
@@
 increased or slowed diversification pace depending on the tree used for
-analysis (Fig. 2). This variation is especially wide for the clade
+analysis (Fig.~\ref{fig2}). This variation is especially wide for the clade
 formed by the genera \emph{Magneuptychia} and \emph{Caeruleuptychia}
@@
 the nodes for \emph{Ypthima}, \emph{Charaxes} and \emph{Coenonympha}
-(Fig. 2). The net diversification rates estimates for the clades
+(Fig.~\ref{fig2}). The net diversification rates estimates for the clades
 (Oleriina + Ithomiina + Napeogenina + Dircennina + Godyridina),
 Limenitidinae + Heliconiinae and Pseudergolinae are relatively
-consistent across the 1000 trees (Fig. 2).
+consistent across the 1000 trees (Fig.~\ref{fig2}).
 
@@
 example the shift for the clade Satyrinae is recovered with a
-probability of 0.18 (Fig. 3).
+probability of 0.18 (Fig.~\ref{fig3}).
 
@@
 net net diversification rate on part of Ithomiini (Oleriina + Ithomiina
-+ Napeogenina + Dircennina + Godyridina) (Fig. 4). The estimated mean
++ Napeogenina + Dircennina + Godyridina) (Fig.~\ref{fig4}). The estimated mean
 net diversification rate for taxa that do not feed on Solanaceae was
 \texttt{r = 0.11} while the net diversification rate for the Solanaceae
-feeders was \texttt{r = 0.16} (see S1 Fig. for a boxplot of speciation
+feeders was \texttt{r = 0.16} (see S1 Figure for a boxplot of speciation
 and extinction values for the 95\% credibility intervals). The same
@@
 non-Solanaceae feeders due to dubious records produced the same pattern
-and net diversification rates (S2 Fig.). Therefore, the rest of the
+and net diversification rates (S2 Figure). Therefore, the rest of the
 analyses were performed assuming these two genera as Solanaceae feeders.
@@
 credibility tree (combined mcmc run in S20 Supporting Information; profiles
-plot of speciation rates in S3 Fig.; boxplot of 95\% credibility intervals in
-S4 Fig.). A BiSSE analysis to test whether the trait ``feeding on
+plot of speciation rates in S3 Figure; boxplot of 95\% credibility intervals in
+S4 Figure). A BiSSE analysis to test whether the trait ``feeding on
 Apocynaceae'' had any effect on increased net diversification rates
 found similar speciation rates for lineages feeding on Apocynaceae and
-other plants (S5 Fig.).
+other plants (S5 Figure).
 
@@
 rates are significantly higher in Solanaceae feeders (speciation rate =
-$\lambda1$, net diversification rate = $r1$}
+$\lambda1$, net diversification rate = $r1$)}
 \label{fig4}
@@
 probability in the random sample of 1000 trees from the posterior
-distribution (Fig. 3, Table 2). We also found that, even though the
+distribution (Fig.~\ref{fig3}, Table 2). We also found that, even though the
 analyses estimated the same diversification shifts on two or more trees,
-the estimated net diversification rates could vary widely (Fig. 2). For
+the estimated net diversification rates could vary widely (Fig.~\ref{fig2}). For
 example, in our Nymphalidae trees, we found that the shift for
@@
 \emph{Charaxes}, the genus \emph{Ypthima}, part of Ithomiini and the
-clade \emph{Callicore} + \emph{Diaethria} (Fig. 3), while estimating
+clade \emph{Callicore} + \emph{Diaethria} (Fig.~\ref{fig3}), while estimating
 shifts for other lineages with much lower probability.
@@
 diversification rate for Ithomiini taxa, which can be attributed to the
-trait ``feeding on Solanaceae hostplants'' (Fig. 4). This is in
+trait ``feeding on Solanaceae hostplants'' (Fig.~\ref{fig4}). This is in
 agreement with previous findings using other statistical methods
@@
 in the subtribes Melinaeina and Mechanitina are not included in the
-diversification shift (shift number 5 in Fig. 1). The apparent
+diversification shift (shift number 5 in Fig.~\ref{fig1}). The apparent
 conflicting results from MEDUSA and BiSSE can be explained by the low
@@
 \section*{Supporting Information Legends}
-\subsection*{S1 Fig.}
-\label{S1_Fig.}
-{\bf Boxplot of speciation (\(\lambda\)) and extinction
+\textbf{S1 Figure.
+\bf Boxplot of speciation (\(\lambda\)) and extinction
 (\(\mu\)) values for the 95\% credibility intervals of values estimated
 by BiSSE analysis of diversification due to feeding on Solanaceae
-plants}.
-
-\subsection*{S2 Fig.}
-\label{S2_Fig.}
-{\bf BiSSE analysis of diversification of nymphalids
+plants.}
+
+\textbf{S2 Figure.
+BiSSE analysis of diversification of nymphalids
 due to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
-\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
+\emph{Hypanartia} as non-Solanaceae feeders}. The same pattern is
 recovered, speciation and net diversification rates are significantly
-higher for Solanaceae feeders \(\lambda1\), r1)}.
-
-\subsection*{S3 Fig.}
-\label{S3_Fig.}
-{\bf Net diversification rates of nymphalids feeding on
+higher for Solanaceae feeders \(\lambda1\), r1).
+
+\textbf{S3 Figure.
+Net diversification rates of nymphalids feeding on
 Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
@@
 
-\subsection*{S4 Fig.}
-\label{S4_Fig.}
-{\bf Boxplot of speciation and extinction values for
+\textbf{S4 Figure.
+Boxplot of speciation and extinction values for
 the 95\% credibility intervals of values estimated by BiSSE analysis of
 diversification due to feeding on Solanaceae plants on the combined
-post-burnin runs on 1000 trees from the posterior distribution.}.
-
-\subsection*{S5 Fig.}
-\label{S5_Fig.}
-{\bf BiSSE analysis of diversification of nymphalids
+post-burnin runs on 1000 trees from the posterior distribution.}
+
+\textbf{S5 Figure.
+BiSSE analysis of diversification of nymphalids
 due to feeding on Apocynaceae hostplants}. Speciation and net
@@
 
-\subsection*{S6 File}
-\label{S6_File}
-{\bf 1000 random trees from \cite{wahlberg2009}}.
-
-\subsection*{S7 Dataset}
-\label{S7_Dataset}
-{\bf Species richness for lineages in Nymphalidae}.
-
-\subsection*{S8Dataset}
-\label{S8_Dataset}
-{\bf Hostplants of Nymphalidae
+\textbf{S6 File.
+1000 random trees from \cite{wahlberg2009}}.
+
+\textbf{S7 Dataset.
+Species richness for lineages in Nymphalidae}.
+
+\textbf{S8 Dataset.
+Hostplants of Nymphalidae
 butterflies recorded from the literature}.
 
-\subsection*{S9 Dataset}
-\label{S9_Dataset}
-{\bf References for hostplants data}.
-
-\subsection*{S10 Supporting Information}
-\label{S10_Supporting_Information}
-{\bf R script to run MEDUSA on the MCC Nymphalidae tree from
+\textbf{S9 Dataset.
+References for hostplants data}.
+
+\textbf{S10 File.
+R script to run MEDUSA on the MCC Nymphalidae tree from
 \cite{wahlberg2009}}. This script removes the outgroup
@@
 
-\subsection*{S11 Supporting Information}
-\label{S11_Supporting_Information}
-{\bf R script to run a MultiMEDUSA analysis on 1000 random trees from
+\textbf{S11 File.
+R script to run a MultiMEDUSA analysis on 1000 random trees from
 \cite{wahlberg2009}}.
 
-\subsection*{S12 Supporting Information}
-\label{S12_Supporting_Information}
-{\bf R script to run a MEDUSA
+\textbf{S12 Supporting Information.
+R script to run a MEDUSA
 analysis on the MCC tree from the 1000 random trees selected from the
@@
 
-\subsection*{S13 Dataset}
-\label{S13_Dataset}
-{\bf Data matrix with character states for hosplant use}.
-
-\subsection*{S14 Supporting Information}
-\label{S14_Supporting_Information}
-{\bf MCC tree from the 1000 random
+\textbf{S13 Dataset.
+Data matrix with character states for hosplant use}.
+
+\textbf{S14 Supporting Information.
+MCC tree from the 1000 random
 trees selected from the posterior distribution}.
 
-\subsection*{S15 File}
-\label{S15_File}
-{\bf Raw results from the MultiMEDUSA run on the random sample of trees from 
+\textbf{S15 File.
+Raw results from the MultiMEDUSA run on the random sample of trees from 
 the posterior distribution}.
 
-\subsection*{S16 Text}
-\label{S16_Text}
-{\bf Summary results from a
+\textbf{S16 Text.
+Summary results from a
 MultiMEDUSA run on the 1000 random trees selected from the posterior
@@
 
-\subsection*{S17 Figure}
-\label{S17_Figure}
-{\bf Probability of nodes with
+\textbf{S17 Figure.
+Probability of nodes with
 estimated rates from a MultiMEDUSA run on the 1000 random trees selected
@@
 
-\subsection*{S18 Supporting Information}
-\label{S18_Supporting_Information}
-{\bf R script for running the BiSSE analysis}.
-
-\subsection*{S19 Supporting Information}
-\label{S19_Supporting_Information}
-{\bf Raw results for the BiSSE analysis}.
-
-\subsection*{S20 Supporting Information}
-\label{S20_Supporting_Information}
-{\bf Raw results for the combined BiSSE analysis}.
-
-\subsection*{S21 Dataset}
-\label{S21_Dataset}
-{\bf MCC Nymphalidae tree from \cite{wahlberg2009}}
-
-\subsection*{S22 Text}
-\label{S22_Text}
-{\bf Results of running MEDUSA on the MCC tree}.
-
-\subsection*{S23 Figure}
-\label{S23_Figure}
-{\bf Summary of the MultiMEDUSA
+\textbf{S18 Supporting Information.
+R script for running the BiSSE analysis}.
+
+\textbf{S19 Supporting Information.
+Raw results for the BiSSE analysis}.
+
+\textbf{S20 Supporting Information.
+Raw results for the combined BiSSE analysis}.
+
+\textbf{S21 Dataset.
+MCC Nymphalidae tree from \cite{wahlberg2009}}
+
+\textbf{S22 Text.
+Results of running MEDUSA on the MCC tree}.
+
+\textbf{S23 Figure.
+Summary of the MultiMEDUSA
 analysis on the 1000 trees from the posterior distribution}.
 
-\subsection*{S24 Text}
-\label{S24_Text}
-{\bf Summary results from a MEDUSA
+\textbf{S24 Text.
+Summary results from a MEDUSA
 run on the MCC tree from the 1000 random trees selected from the
@@
 
-\subsection*{S25 Figure}
-\label{S25_Figure}
-{\bf Figure for MEDUSA run on MCC tree from random 1000 trees}.
+\textbf{S25 Figure.
+Figure for MEDUSA run on MCC tree from random 1000 trees}.
 

### 2015-02-17 MS_revised.tex
ref to tables

@@
 estimated 18 significant changes in the tempo of diversification in
-Nymphalidae history (Fig.~\ref{fig1}; Table 1). The corrected AICc acceptance
+Nymphalidae history (Fig.~\ref{fig1}; Table~\ref{table1}). The corrected AICc acceptance
 threshold for adding subsequent piecewise birth-death processes to the
@@
 years and the AICc score for the best fit model was \texttt{5090.5}
-(Table 1). MEDUSA also estimated that the basic constant birth-death
+(Table~\ref{table1}). MEDUSA also estimated that the basic constant birth-death
 model was not a better explanation for our data (AICc
@@
 Pseudergolinae (\texttt{r = 0.024}) and Coenonymphina
-(\texttt{r = 0.065}) (Table 1).
+(\texttt{r = 0.065}) (Table~\ref{table1}).
 
@@
 found by MEDUSA on the MCC tree derived from the random sample of trees
-(Table 2). Although the diversification pattern found by MEDUSA and
+(Table~\ref{table2}). Although the diversification pattern found by MEDUSA and
 MultiMEDUSA was the same, the latter consistently estimated lower rates.
@@
 For the diversification shifts found in both the MCC tree and most of
-the samples of 1000 trees (frequency more than 90\%; Table 2), the
+the samples of 1000 trees (frequency more than 90\%; Table~\ref{table2}), the
 MultiMEDUSA approach recovered different rates of diversification than
@@
 There were four net diversification rate increases found in the trees from the
-random sample (Table 2): (i) the
+random sample (Table~\ref{table2}): (i) the
 genus \emph{Ypthima} (\(r = 0.22\)); (ii) the genus \emph{Charaxes}
@@
 significantly better explanation than this character state having no
-effect on diversification (\(\chi^2 = 12.3; 1 df; p < 0.001\)) (Table 3;
+effect on diversification (\(\chi^2 = 12.3; 1 df; p < 0.001\))
+(Table~\ref{table3};
 character states available in S13 Dataset, code in S18 Supporting Information,
@@
 probability in the random sample of 1000 trees from the posterior
-distribution (Fig.~\ref{fig3}, Table 2). We also found that, even though the
+distribution (Fig.~\ref{fig3}, Table~\ref{table2}). We also found that, even though the
 analyses estimated the same diversification shifts on two or more trees,

### 2015-02-17 MS_revised.tex
refix captions supp mat

@@
 
-\section*{Supporting Information Legends}
-\textbf{S1 Figure.
-\bf Boxplot of speciation (\(\lambda\)) and extinction
+\section*{Supporting Information}
+
+\subsection*{S1 Figure}
+\label{S1_Figure}
+{\bf Boxplot of speciation (\(\lambda\)) and extinction
 (\(\mu\)) values for the 95\% credibility intervals of values estimated
@@
 
-\textbf{S2 Figure.
-BiSSE analysis of diversification of nymphalids
+\subsection*{S2 Figure}
+\label{S2_Figure}
+{\bf BiSSE analysis of diversification of nymphalids
 due to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
@@
 
-\textbf{S3 Figure.
-Net diversification rates of nymphalids feeding on
+\subsection*{S3 Figure}
+\label{S3_Figure}
+{\bf Net diversification rates of nymphalids feeding on
 Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
@@
 
-\textbf{S4 Figure.
-Boxplot of speciation and extinction values for
+\subsection*{S4 Figure}
+\label{S4_Figure}
+{\bf Boxplot of speciation and extinction values for
 the 95\% credibility intervals of values estimated by BiSSE analysis of
@@
 
-\textbf{S5 Figure.
-BiSSE analysis of diversification of nymphalids
+\subsection*{S5 Figure}
+\label{S5_Figure}
+{\bf BiSSE analysis of diversification of nymphalids
 due to feeding on Apocynaceae hostplants}. Speciation and net
@@
 
-\textbf{S6 File.
-1000 random trees from \cite{wahlberg2009}}.
-
-\textbf{S7 Dataset.
-Species richness for lineages in Nymphalidae}.
-
-\textbf{S8 Dataset.
-Hostplants of Nymphalidae
+\subsection*{S6 File}
+\label{S6_File}
+{\bf 1000 random trees from \cite{wahlberg2009}}.
+
+\subsection*{S7 Dataset}
+\label{S7_Dataset}
+{\bf Species richness for lineages in Nymphalidae}.
+
+\subsection*{S8 Dataset}
+\label{S8_Dataset}
+{\bf Hostplants of Nymphalidae
 butterflies recorded from the literature}.
 
-\textbf{S9 Dataset.
-References for hostplants data}.
-
-\textbf{S10 File.
-R script to run MEDUSA on the MCC Nymphalidae tree from
+\subsection*{S9 Dataset}
+\label{S9_Dataset}
+{\bf References for hostplants data}.
+
+\subsection*{S10 File}
+\label{S10_File}
+{\bf R script to run MEDUSA on the MCC Nymphalidae tree from
 \cite{wahlberg2009}}. This script removes the outgroup
@@
 
-\textbf{S11 File.
-R script to run a MultiMEDUSA analysis on 1000 random trees from
+\subsection*{S11 File}
+\label{S11_File}
+{\bf R script to run a MultiMEDUSA analysis on 1000 random trees from
 \cite{wahlberg2009}}.
 
-\textbf{S12 Supporting Information.
-R script to run a MEDUSA
+\subsection*{S12 Supporting Information}
+\label{S12_Supporting_Information}
+{\bf R script to run a MEDUSA
 analysis on the MCC tree from the 1000 random trees selected from the
@@
 
-\textbf{S13 Dataset.
-Data matrix with character states for hosplant use}.
-
-\textbf{S14 Supporting Information.
-MCC tree from the 1000 random
+\subsection*{S13 Dataset}
+\label{S13_Dataset}
+{\bf Data matrix with character states for hosplant use}.
+
+\subsection*{S14 Supporting Information}
+\label{S14_Supporting_Information}
+{\bf MCC tree from the 1000 random
 trees selected from the posterior distribution}.
 
-\textbf{S15 File.
-Raw results from the MultiMEDUSA run on the random sample of trees from 
+\subsection*{S15 File}
+\label{S15_File}
+{\bf Raw results from the MultiMEDUSA run on the random sample of trees from 
 the posterior distribution}.
 
-\textbf{S16 Text.
-Summary results from a
+\subsection*{S16 Text}
+\label{S16_Text}
+{\bf Summary results from a
 MultiMEDUSA run on the 1000 random trees selected from the posterior
@@
 
-\textbf{S17 Figure.
-Probability of nodes with
+\subsection*{S17 Figure}
+\label{S17_Figure}
+{\bf Probability of nodes with
 estimated rates from a MultiMEDUSA run on the 1000 random trees selected
@@
 
-\textbf{S18 Supporting Information.
-R script for running the BiSSE analysis}.
-
-\textbf{S19 Supporting Information.
-Raw results for the BiSSE analysis}.
-
-\textbf{S20 Supporting Information.
-Raw results for the combined BiSSE analysis}.
-
-\textbf{S21 Dataset.
-MCC Nymphalidae tree from \cite{wahlberg2009}}
-
-\textbf{S22 Text.
-Results of running MEDUSA on the MCC tree}.
-
-\textbf{S23 Figure.
-Summary of the MultiMEDUSA
+\subsection*{S18 supporting information}
+\label{S18_Supporting_Information}
+{\bf R script for running the BiSSE analysis}.
+
+\subsection*{S19 Supporting Information}
+\label{S19_Supporting_Information}
+{\bf Raw results for the BiSSE analysis}.
+
+\subsection*{S20 Supporting Information}
+\label{S20_Supporting_Information}
+{\bf Raw results for the combined BiSSE analysis}.
+
+\subsection*{S21 Dataset}
+\label{S21_Dataset}
+{\bf MCC Nymphalidae tree from \cite{wahlberg2009}}
+
+\subsection*{S22 Text}
+\label{S22_Text}
+{\bf Results of running MEDUSA on the MCC tree}.
+
+\subsection*{S23 Figure}
+\label{S23_Figure}
+{\bf Summary of the MultiMEDUSA
 analysis on the 1000 trees from the posterior distribution}.
 
-\textbf{S24 Text.
-Summary results from a MEDUSA
+\subsection*{S24 Text}
+\label{S24_Text}
+{\bf Summary results from a MEDUSA
 run on the MCC tree from the 1000 random trees selected from the
@@
 
-\textbf{S25 Figure.
-Figure for MEDUSA run on MCC tree from random 1000 trees}.
+\subsection*{S25 Figure}
+\label{S25_Figure}
+{\bf Figure for MEDUSA run on MCC tree from random 1000 trees}.
 

### 2015-02-21 MS_revised.tex
supp figs

@@
 diversification values on the shifts on the 1000 trees 
-(S16 Text and S17 Figure), and found that the estimated net diversification rate
+(S16 Text and S17 Fig.), and found that the estimated net diversification rate
 values had great variation across the posterior distribution of trees. A
@@
 \texttt{r = 0.11} while the net diversification rate for the Solanaceae
-feeders was \texttt{r = 0.16} (see S1 Figure for a boxplot of speciation
+feeders was \texttt{r = 0.16} (see S1 Fig. for a boxplot of speciation
 and extinction values for the 95\% credibility intervals). The same
@@
 non-Solanaceae feeders due to dubious records produced the same pattern
-and net diversification rates (S2 Figure). Therefore, the rest of the
+and net diversification rates (S2 Fig.). Therefore, the rest of the
 analyses were performed assuming these two genera as Solanaceae feeders.
@@
 credibility tree (combined mcmc run in S20 Supporting Information; profiles
-plot of speciation rates in S3 Figure; boxplot of 95\% credibility intervals in
-S4 Figure). A BiSSE analysis to test whether the trait ``feeding on
+plot of speciation rates in S3 Fig.; boxplot of 95\% credibility intervals in
+S4 Fig.). A BiSSE analysis to test whether the trait ``feeding on
 Apocynaceae'' had any effect on increased net diversification rates
 found similar speciation rates for lineages feeding on Apocynaceae and
-other plants (S5 Figure).
+other plants (S5 Fig.).
 

### 2015-02-21 MS_revised.tex
supp captions
revised captions

@@
 Carlos Pe\~na\textsuperscript{1,*},
-Marianne Espeland\textsuperscript{2,3}
+Marianne Espeland\textsuperscript{2,3,4}
 \\
@@
 \\
-\bf{2} Museum of Comparative Zoology and Department of Organismic and Evolutionary Biology, Harvard University, Cambridge, USA
+\bf{2} Museum of Comparative Zoology, Harvard University, Cambridge, USA
+\bf{3} Department of Organismic and Evolutionary Biology, Harvard University, Cambridge, USA
 \\
@@
 % Current address notes
-\bf{3} Current address: McGuire Center for Lepidoptera and Biodiversity,
+\bf{4} Current address: McGuire Center for Lepidoptera and Biodiversity,
 Florida Museum of Natural History, University of Florida, Gainesville,
@@
 
-\subsection*{S1 Figure}
+\subsection*{S1 Fig.}
 \label{S1_Figure}
@@
 
-\subsection*{S2 Figure}
+\subsection*{S2 Fig.}
 \label{S2_Figure}
@@
 
-\subsection*{S3 Figure}
+\subsection*{S3 Fig.}
 \label{S3_Figure}
@@
 
-\subsection*{S4 Figure}
+\subsection*{S4 Fig.}
 \label{S4_Figure}
@@
 
-\subsection*{S5 Figure}
+\subsection*{S5 Fig.}
 \label{S5_Figure}
@@
 
-\subsection*{S17 Figure}
+\subsection*{S17 Fig.}
 \label{S17_Figure}
@@
 
-\subsection*{S18 supporting information}
+\subsection*{S18 Supporting Information}
 \label{S18_Supporting_Information}
@@
 
-\subsection*{S23 Figure}
+\subsection*{S23 Fig.}
 \label{S23_Figure}
@@
 
-\subsection*{S25 Figure}
+\subsection*{S25 Fig.}
 \label{S25_Figure}

### 2015-03-12 MS_revised.tex
fixed order of supp mat in text

@@
 used a burnin of 25 million generations and took a random sample of 1000
-trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (S6 File)
+trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (S9 File)
 in order to account for phylogenetic uncertainty when performing
@@
 taxonomical classification and morphological resemblance when no
-phylogenies were available (S7 Dataset).
+phylogenies were available (S13 Dataset).
 
@@
 sources including \cite{ackery1988}, HOSTS database (\url{http://bit.ly/YI7nwW}),
-\cite{dyer2002, beccaloni2008, janzen2009} and others (S8 and S9 Dataset) for a
+\cite{dyer2002, beccaloni2008, janzen2009} and others 
+(S14 and S15 Dataset) for a
 total of 6586
@@
 distribution of the Bayesian run from Wahlberg et al., 2009 \cite{wahlberg2009}
-(MultiMEDUSA, S6 File). We calculated a new MCC tree from the
+(MultiMEDUSA, S9 File). We calculated a new MCC tree from the
 selection of trees, ran MEDUSA and MultiMEDUSA on the selection of 1000
@@
 use one plant family as hostplant (232 genera use only one plant family
-versus 176 that use more than one; based on our data in S8 Dataset),
+versus 176 that use more than one; based on our data in S14 Dataset),
 the character states can be coded as presence/absence, for which the
@@
 the plant family Solanaceae'' with two states (absence = 0; presence =
-1) (S13 Dataset). The family Solanaceae include the hostplants of the
+1) (S16 Dataset). The family Solanaceae include the hostplants of the
 diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. 
@@
 values obtained from the MCC tree. We ran MultiMEDUSA on the random
-sample of 1000 trees (S6 File) from the posterior distribution and
+sample of 1000 trees (S9 File) from the posterior distribution and
 compared the results with a MEDUSA analysis on the MCC tree derived from
-this sample (S14 Supporting Information).
+this sample (S19 Supporting Information).
 
@@
 MultiMEDUSA. The background diversification and all shifts found by
-MEDUSA on the 1000 trees are provided as an R object in S15 File.
+MEDUSA on the 1000 trees are provided as an R object in S12 File.
 
@@
 diversification values on the shifts on the 1000 trees 
-(S16 Text and S17 Fig.), and found that the estimated net diversification rate
+(S6 Fig. and S23 Text), and found that the estimated net diversification rate
 values had great variation across the posterior distribution of trees. A
@@
 positive effect of the character state ``feeding on Solanaceae'' on the
-net net diversification rate on part of Ithomiini (Oleriina + Ithomiina
+net diversification rate on part of Ithomiini (Oleriina + Ithomiina
 + Napeogenina + Dircennina + Godyridina) (Fig.~\ref{fig4}). The estimated mean
@@
 (Table~\ref{table3};
-character states available in S13 Dataset, code in S18 Supporting Information,
-and mcmc run in S19 Supporting Information). We combined the post-burnin mcmc generations
+character states available in S16 Dataset, code in S20 Supporting Information,
+and mcmc run in S21 Supporting Information). We combined the post-burnin mcmc generations
 from running BiSSE on 250 trees from the random sample of 1000 trees and
 found the same pattern as the BiSSE analysis on the maximum clade
-credibility tree (combined mcmc run in S20 Supporting Information; profiles
+credibility tree (combined mcmc run in S22 Supporting Information; profiles
 plot of speciation rates in S3 Fig.; boxplot of 95\% credibility intervals in
@@
 
-\subsection*{S6 File}
-\label{S6_File}
+\subsection*{S6 Fig.}
+#\subsection*{S17 Fig.}
+\label{S6_Figure}
+{\bf Probability of nodes with
+estimated rates from a MultiMEDUSA run on the 1000 random trees selected
+from the posterior distribution}.
+
+\subsection*{S7 Fig.}
+#\subsection*{S23 Fig.}
+\label{S7_Figure}
+{\bf Summary of the MultiMEDUSA
+analysis on the 1000 trees from the posterior distribution}.
+
+\subsection*{S8 Fig.}
+#\subsection*{S25 Fig.}
+\label{S8_Figure}
+{\bf Figure for MEDUSA run on MCC tree from random 1000 trees}.
+
+\subsection*{S9 File}
+#\subsection*{S6 File}
+\label{S9_File}
 {\bf 1000 random trees from \cite{wahlberg2009}}.
-
-\subsection*{S7 Dataset}
-\label{S7_Dataset}
-{\bf Species richness for lineages in Nymphalidae}.
-
-\subsection*{S8 Dataset}
-\label{S8_Dataset}
-{\bf Hostplants of Nymphalidae
-butterflies recorded from the literature}.
-
-\subsection*{S9 Dataset}
-\label{S9_Dataset}
-{\bf References for hostplants data}.
 
@@
 
-\subsection*{S12 Supporting Information}
-\label{S12_Supporting_Information}
+\subsection*{S12 File}
+#\subsection*{S15 File}
+\label{S12_File}
+{\bf Raw results from the MultiMEDUSA run on the random sample of trees from 
+the posterior distribution}.
+
+\subsection*{S13 Dataset}
+#\subsection*{S7 Dataset}
+\label{S13_Dataset}
+{\bf Species richness for lineages in Nymphalidae}.
+
+\subsection*{S14 Dataset}
+#\subsection*{S8 Dataset}
+\label{S14_Dataset}
+{\bf Hostplants of Nymphalidae
+butterflies recorded from the literature}.
+
+\subsection*{S15 Dataset}
+#\subsection*{S9 Dataset}
+\label{S15_Dataset}
+{\bf References for hostplants data}.
+
+\subsection*{S16 Dataset}
+#\subsection*{S13 Dataset}
+\label{S16_Dataset}
+{\bf Data matrix with character states for hosplant use}.
+
+\subsection*{S17 Dataset}
+#\subsection*{S21 Dataset}
+\label{S17_Dataset}
+{\bf MCC Nymphalidae tree from \cite{wahlberg2009}}
+
+\subsection*{S18 Supporting Information}
+#\subsection*{S12 Supporting Information}
+\label{S18_Supporting_Information}
 {\bf R script to run a MEDUSA
@@
 
-\subsection*{S13 Dataset}
-\label{S13_Dataset}
-{\bf Data matrix with character states for hosplant use}.
-
-\subsection*{S14 Supporting Information}
-\label{S14_Supporting_Information}
+\subsection*{S19 Supporting Information}
+#\subsection*{S14 Supporting Information}
+\label{S19_Supporting_Information}
 {\bf MCC tree from the 1000 random
@@
 
-\subsection*{S15 File}
-\label{S15_File}
-{\bf Raw results from the MultiMEDUSA run on the random sample of trees from 
-the posterior distribution}.
-
-\subsection*{S16 Text}
-\label{S16_Text}
+\subsection*{S20 Supporting Information}
+#\subsection*{S18 Supporting Information}
+\label{S20_Supporting_Information}
+{\bf R script for running the BiSSE analysis}.
+
+\subsection*{S21 Supporting Information}
+#\subsection*{S19 Supporting Information}
+\label{S21_Supporting_Information}
+{\bf Raw results for the BiSSE analysis}.
+
+\subsection*{S22 Supporting Information}
+#\subsection*{S20 Supporting Information}
+\label{S22_Supporting_Information}
+{\bf Raw results for the combined BiSSE analysis}.
+
+\subsection*{S23 Text}
+#\subsection*{S16 Text}
+\label{S23_Text}
 {\bf Summary results from a
@@
 
-\subsection*{S17 Fig.}
-\label{S17_Figure}
-{\bf Probability of nodes with
-estimated rates from a MultiMEDUSA run on the 1000 random trees selected
-from the posterior distribution}.
-
-\subsection*{S18 Supporting Information}
-\label{S18_Supporting_Information}
-{\bf R script for running the BiSSE analysis}.
-
-\subsection*{S19 Supporting Information}
-\label{S19_Supporting_Information}
-{\bf Raw results for the BiSSE analysis}.
-
-\subsection*{S20 Supporting Information}
-\label{S20_Supporting_Information}
-{\bf Raw results for the combined BiSSE analysis}.
-
-\subsection*{S21 Dataset}
-\label{S21_Dataset}
-{\bf MCC Nymphalidae tree from \cite{wahlberg2009}}
-
-\subsection*{S22 Text}
-\label{S22_Text}
+\subsection*{S24 Text}
+#\subsection*{S22 Text}
+\label{S24_Text}
 {\bf Results of running MEDUSA on the MCC tree}.
 
-\subsection*{S23 Fig.}
-\label{S23_Figure}
-{\bf Summary of the MultiMEDUSA
-analysis on the 1000 trees from the posterior distribution}.
-
-\subsection*{S24 Text}
-\label{S24_Text}
+\subsection*{S25 Text}
+\label{S25_Text}
 {\bf Summary results from a MEDUSA
@@
 posterior distribution}.
-
-\subsection*{S25 Fig.}
-\label{S25_Figure}
-{\bf Figure for MEDUSA run on MCC tree from random 1000 trees}.
 

### 2015-03-16 MS_revised.tex
cleanned supp info in ms
rename sup mat in tex file

@@
 used a burnin of 25 million generations and took a random sample of 1000
-trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (S9 File)
+trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (S1 File)
 in order to account for phylogenetic uncertainty when performing
@@
 taxonomical classification and morphological resemblance when no
-phylogenies were available (S13 Dataset).
+phylogenies were available (S1 Dataset).
 
@@
 \cite{dyer2002, beccaloni2008, janzen2009} and others 
-(S14 and S15 Dataset) for a
+(S2 and S3 Dataset) for a
 total of 6586
@@
 \texttt{diversitree} \cite{fitzjohn2012} packages along with our own scripts to
-perform the analyses (included as S10 and S11 File). All
+perform the analyses (included as S2 and S3 File). All
 analyses were run on the 1000 random trees from Wahlberg et al. (2009)
@@
 distribution of the Bayesian run from Wahlberg et al., 2009 \cite{wahlberg2009}
-(MultiMEDUSA, S9 File). We calculated a new MCC tree from the
+(MultiMEDUSA, S1 File). We calculated a new MCC tree from the
 selection of trees, ran MEDUSA and MultiMEDUSA on the selection of 1000
@@
 use one plant family as hostplant (232 genera use only one plant family
-versus 176 that use more than one; based on our data in S14 Dataset),
+versus 176 that use more than one; based on our data in S2 Dataset),
 the character states can be coded as presence/absence, for which the
@@
 the plant family Solanaceae'' with two states (absence = 0; presence =
-1) (S16 Dataset). The family Solanaceae include the hostplants of the
+1) (S4 Dataset). The family Solanaceae include the hostplants of the
 diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. 
@@
 values obtained from the MCC tree. We ran MultiMEDUSA on the random
-sample of 1000 trees (S9 File) from the posterior distribution and
+sample of 1000 trees (S1 File) from the posterior distribution and
 compared the results with a MEDUSA analysis on the MCC tree derived from
-this sample (S19 Supporting Information).
+this sample (S2 Supporting Information).
 
@@
 MultiMEDUSA. The background diversification and all shifts found by
-MEDUSA on the 1000 trees are provided as an R object in S12 File.
+MEDUSA on the 1000 trees are provided as an R object in S4 File.
 
@@
 diversification values on the shifts on the 1000 trees 
-(S6 Fig. and S23 Text), and found that the estimated net diversification rate
+(S6 Fig. and S1 Text), and found that the estimated net diversification rate
 values had great variation across the posterior distribution of trees. A
@@
 (Table~\ref{table3};
-character states available in S16 Dataset, code in S20 Supporting Information,
-and mcmc run in S21 Supporting Information). We combined the post-burnin mcmc generations
+character states available in S4 Dataset, code in S3 Supporting Information,
+and mcmc run in S4 Supporting Information). We combined the post-burnin mcmc generations
 from running BiSSE on 250 trees from the random sample of 1000 trees and
 found the same pattern as the BiSSE analysis on the maximum clade
-credibility tree (combined mcmc run in S22 Supporting Information; profiles
+credibility tree (combined mcmc run in S5 Supporting Information; profiles
 plot of speciation rates in S3 Fig.; boxplot of 95\% credibility intervals in
@@
 \subsection*{S6 Fig.}
-#\subsection*{S17 Fig.}
 \label{S6_Figure}
@@
 \subsection*{S7 Fig.}
-#\subsection*{S23 Fig.}
 \label{S7_Figure}
@@
 \subsection*{S8 Fig.}
-#\subsection*{S25 Fig.}
 \label{S8_Figure}
@@
 
-\subsection*{S9 File}
-#\subsection*{S6 File}
-\label{S9_File}
+\subsection*{S1 File}
+\label{S1_File}
 {\bf 1000 random trees from \cite{wahlberg2009}}.
 
-\subsection*{S10 File}
-\label{S10_File}
+\subsection*{S2 File}
+\label{S2_File}
 {\bf R script to run MEDUSA on the MCC Nymphalidae tree from
@@
 
-\subsection*{S11 File}
-\label{S11_File}
+\subsection*{S3 File}
+\label{S3_File}
 {\bf R script to run a MultiMEDUSA analysis on 1000 random trees from
@@
 
-\subsection*{S12 File}
-#\subsection*{S15 File}
-\label{S12_File}
+\subsection*{S4 File}
+\label{S4_File}
 {\bf Raw results from the MultiMEDUSA run on the random sample of trees from 
@@
 
-\subsection*{S13 Dataset}
-#\subsection*{S7 Dataset}
-\label{S13_Dataset}
+\subsection*{S1 Dataset}
+\label{S1_Dataset}
 {\bf Species richness for lineages in Nymphalidae}.
 
-\subsection*{S14 Dataset}
-#\subsection*{S8 Dataset}
-\label{S14_Dataset}
+\subsection*{S2 Dataset}
+\label{S2_Dataset}
 {\bf Hostplants of Nymphalidae
@@
 
-\subsection*{S15 Dataset}
-#\subsection*{S9 Dataset}
-\label{S15_Dataset}
+\subsection*{S3 Dataset}
+\label{S3_Dataset}
 {\bf References for hostplants data}.
 
-\subsection*{S16 Dataset}
-#\subsection*{S13 Dataset}
-\label{S16_Dataset}
+\subsection*{S4 Dataset}
+\label{S4_Dataset}
 {\bf Data matrix with character states for hosplant use}.
 
-\subsection*{S17 Dataset}
-#\subsection*{S21 Dataset}
-\label{S17_Dataset}
+\subsection*{S5 Dataset}
+\label{S5_Dataset}
 {\bf MCC Nymphalidae tree from \cite{wahlberg2009}}
 
-\subsection*{S18 Supporting Information}
-#\subsection*{S12 Supporting Information}
-\label{S18_Supporting_Information}
+\subsection*{S1 Supporting Information}
+\label{S1_Supporting_Information}
 {\bf R script to run a MEDUSA
@@
 
-\subsection*{S19 Supporting Information}
-#\subsection*{S14 Supporting Information}
-\label{S19_Supporting_Information}
+\subsection*{S2 Supporting Information}
+\label{S2_Supporting_Information}
 {\bf MCC tree from the 1000 random
@@
 
-\subsection*{S20 Supporting Information}
-#\subsection*{S18 Supporting Information}
-\label{S20_Supporting_Information}
+\subsection*{S3 Supporting Information}
+\label{S3_Supporting_Information}
 {\bf R script for running the BiSSE analysis}.
 
-\subsection*{S21 Supporting Information}
-#\subsection*{S19 Supporting Information}
-\label{S21_Supporting_Information}
+\subsection*{S4 Supporting Information}
+\label{S4_Supporting_Information}
 {\bf Raw results for the BiSSE analysis}.
 
-\subsection*{S22 Supporting Information}
-#\subsection*{S20 Supporting Information}
-\label{S22_Supporting_Information}
+\subsection*{S5 Supporting Information}
+\label{S5_Supporting_Information}
 {\bf Raw results for the combined BiSSE analysis}.
 
-\subsection*{S23 Text}
-#\subsection*{S16 Text}
-\label{S23_Text}
+\subsection*{S1 Text}
+\label{S1_Text}
 {\bf Summary results from a
@@
 
-\subsection*{S24 Text}
-#\subsection*{S22 Text}
-\label{S24_Text}
+\subsection*{S2 Text}
+\label{S2_Text}
 {\bf Results of running MEDUSA on the MCC tree}.
 
-\subsection*{S25 Text}
-\label{S25_Text}
+\subsection*{S3 Text}
+\label{S3_Text}
 {\bf Summary results from a MEDUSA
